### 2019-05-11 unimodular-covers.tex
first version

\documentclass{amsart}

\textheight=22.5cm
\topmargin=0.5cm

\usepackage{amssymb}
\usepackage{graphicx,color}
\usepackage{amsmath}
\usepackage{url}
\usepackage[colorlinks=true]{hyperref}

\usepackage{cleveref}
\Crefname{counterexample}{counterexample}{counterexamples}
\Crefname{counterexample}{Counterexample}{Counterexamples}
\Crefname{conjecture}{conjecture}{conjectures}
\Crefname{conjecture}{Conjecture}{Conjectures}

\usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{matrix}

\newcommand{\limites}{}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{example}[theorem]{Example}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{question}[theorem]{Question}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}

% Macros in general

\newcommand{\F}{ \ensuremath{\mathbb{F}}}
\newcommand{\N}{ \ensuremath{\mathbb{N}}}
\newcommand{\Z}{ \ensuremath{\mathbb{Z}}}
\newcommand{\C}{ \ensuremath{\mathbb{C}}}
\newcommand{\R}{ \ensuremath{\mathbb{R}}}

\newcommand{\bb}{{\mathbf{b}}}
\newcommand{\cc}{{\mathbf{c}}}
\newcommand{\ttt}{\mathbf{t}}
\newcommand{\GL}{{GL}_r (K)}
\newcommand{\Gin}{\ensuremath{\mathrm{Gin}}}
\newcommand{\gin}{\ensuremath{\mathrm{gin}}}
\newcommand{\GIN}{\ensuremath{\mathrm{GIN}}}
\newcommand{\Gins}{\ensuremath{\mathrm{Gins}}}
\newcommand{\gins}{\ensuremath{\mathrm{gins}}}
\newcommand{\Lex}{\ensuremath{\mathrm{Lex}}}
\newcommand{\dele}[1]{\ensuremath{\Delta^e({#1})}}
\newcommand{\dels}[1]{\ensuremath{\Delta^s({#1})}}
\newcommand{\delc}[1]{\ensuremath{\Delta^c({#1})}}
\newcommand{\Shift}{\ensuremath{\mathrm{Shift}}}

\renewcommand{\int}{\operatorname{int}}
\newcommand{\width}{\operatorname{width}}
\newcommand{\length}{\operatorname{length}}

\newcommand{\cone}{\ensuremath{\mathrm{cone}}\hspace{1pt}}
\newcommand{\conv}{\ensuremath{\mathrm{conv}}\hspace{1pt}}
\newcommand{\cayley}{\ensuremath{\mathrm{Cay}}\hspace{1pt}}
\newcommand{\rank}{\ensuremath{\mathrm{rank}}\hspace{1pt}}
\newcommand{\reg}{\ensuremath{\mathrm{reg}}\hspace{1pt}}
\newcommand{\depth}{\ensuremath{\mathrm{depth}}\hspace{1pt}}

\newcommand{\Image}{\ensuremath{\mathrm{Im}}\hspace{1pt}}

\usepackage[colorinlistoftodos]{todonotes}

\newcommand{\giulia}[1]{\todo[size=\tiny,color=blue!30]{#1 \\ \hfill --- G.}}
\newcommand{\Giulia}[1]{\todo[size=\tiny,inline,color=blue!30]{#1 \\ \hfill --- G.}}

\newcommand{\paco}[1]{\todo[size=\tiny,color=green!30]{#1 \\ \hfill --- P.}}
\newcommand{\Paco}[1]{\todo[size=\tiny,inline,color=green!30]{#1 \\ \hfill --- P.}}

\date{\today}

\author[G.~Codenotti]{Giulia Codenotti}
\author[F.~Santos]{Francisco Santos}

\address[G.~Codenotti]
{
Institut f\"ur Mathematik, Freie Universit\"at Berlin, Germany
}
\email{giulia.codenotti@fu-berlin.de}

\address[F.~Santos]
{
Department of Mathematics, Statistics and Computer Science, University of Cantabria, Spain
}
\email{francisco.santos@unican.es}

\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230 and, while they were in residence at the Mathematical Sciences Research Institute in Berkeley, California during the Fall 2017 semester, by the Clay Institute and the National Science Foundation (Grant No. DMS-1440140).
 Work of F. Santos is also supported by project MTM2017-83750-P of the Spanish Ministry of Science (AEI/FEDER, UE)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Unimodular covers of 3-dimensional parallelepipeds and Cayley sums}

\begin{document}

\begin{abstract}
We show that the following classes of three-dimensional lattice polytopes have unimodular covers: the class of all parallelepipeds, the class of all centrally symmetric polytopes, and the class of all three-dimensional Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$. This improves results of Beck et al. (2018) and Haase et al. (2008) where the last two classes were shown to be IDP.
\end{abstract}

\maketitle

\section{Introduction}

\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, Gaku Liu, Johannes Hofscheier, Christian Haase, .... et al.

\section{Parallelepipeds (and centrally symmetric polytopes)}

\section{nef Calyley polytopes}

\begin{thebibliography}{99}

\end{thebibliography}

\end{document}

### 2019-05-13 unimodular-covers.tex
wrote draft of parallelepipeds section

@@
 
+A lattice polytope $P\subset \R^d$ has the \emph{integer decomposition property} if for every positive integer $n$, every lattice point $p \in nP\cap \Z^d$ can be written as a sum of $n$ lattice points in $P$.
+We abbreviate this by saying that ``$P$ is IDP''.
+Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (normality of toric varieties) and it falls into a hierarchy of some ten properties each stronger than the previous one.
+See, e.g., \cite[Sect.
+1.2.5]{HPPS-survey}, \cite[p.
+2097f]{mfo2004}, \cite[p.
+2313ff]{mfo2007}.
+\paco{cite Bruns-Gubeladze book?}
+Let us here only mention that
+\[
+P \text{ has a unimodular triangulation}\Rightarrow
+P \text{ has a unimodular cover}\Rightarrow
+P \text{ is IDP.}
+\]
+Remember that a \emph{unimodular triangulation} is a triangulation of $P$ into unimodular simplices, and a \emph{unimodular cover} is a collection of unimodular simplices whose union equals $P$.
+
+\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions}
+Related to this, a lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice.
+
+Oda~\cite{Oda1997} posed several questions/conjectures regarding IDP polytopes, among which are the following two:
+\begin{question}
+\begin{enumerate}
+\item Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
+\[
+(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d
+\]
+\paco{perhaps we can call this formula ``being mixed IDP''. The property of ``$N(Q)$ refines $N(P)$'' is sometimes called ``$P$ is a Minkowski summand of $Q$''.}
+
+\item Suppose $P$ is smooth. Then $P$ is IDP.
+\end{enumerate}
+\end{question}
+
+Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes. In particular:
+\begin{enumerate}
+\item Haase et al.~\cite{HNPS2008} proved the mixed IDP property for $2$-polytopes $P$ and $Q$ when $N(Q) \le N(P)$.
+
+\item Beck et al.~\cite{BHHHJKM2018} proved IDP for smooth centrally symmetric $3$-polytopes.
+\end{enumerate}
+
+We improve both of these results to the level of unimodular covers.
+
+The structure of the paper is as follows: 
+%after introducing the necessary tools and concepts in \Cref{sec:prelim},  
+\Cref{sec:parallelepipeds} is devoted to $3$-dimensional lattice parallelepipeds. 
+Observe that lattice parallelotopes in arbitrary dimension are trivially IDP. Here we show:
+
+\begin{theorem}
+\label{thm:parallelepipeds}
+Every $3$-dimensional lattice parallelepiped has a unimodular cover.
+\end{theorem}
+
+This, in turn, implies our improvement of Beck. et al.'s result on smooth c.s. 3-polytopes:
+
+\begin{corollary}
+\label{coro:3cs}
+Every smooth centrally symmetric lattice $3$-polytope has a unimodular cover. 
+\end{corollary}
+
+\begin{proof}
+Let $P$ be a smooth centrally symmetric lattice $3$-polytope.
+Beck et al.~\cite{BHHHJKM2018} show that $P$ is IDP by covering it with parallelepipeds and unimodular simplices.
+Once we know parallelepipeds have unimodular covers (\Cref{thm:parallelepipeds}), we conclude $P$ has a unimodular cover too.
+\end{proof}
+
+Then in \Cref{sec:cayley} we look at Cayley sums and show:
+
+\begin{theorem}
+\label{thm:cayley}
+Let $P$ and $Q$ be lattice polygons with the normal fan of $Q$ refining that of $P$. Then the Cayley sum $\cayley(P,Q)$ has a unimodular cover.
+\end{theorem}
+
+\begin{corollary}
+\label{coro:3cs}
+Every smooth $3$-dimensional prismatoid has a unimodular cover.
+\end{corollary}
+
+\Paco{This corollary can be made stronger: it works for every $3$-prismatoid 
+$P= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\}$ with $k\ge 2$ and such that $P\cap\{z=i\}$ is a lattice polytope for every $i=0,\dots, k$.}
+
+We believe that the $3$-polytopes in these four statements all have unimodular triangulations, but we do not have a proof.
+
 \subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, Gaku Liu, Johannes Hofscheier, Christian Haase, .... et al.
 
-\section{Parallelepipeds (and centrally symmetric polytopes)}
+%\section{Preliminaries}
+%\label{sec:prelim}
 
-\section{nef Calyley polytopes}
+\section{Parallelepipeds}
+\label{sec:parallelepipeds}
+
+The main tool is what I call the “parallelepiped circumscribed to a given tetrahedron, defined as follows:
+
+\begin{definition}
+Definition: Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$.
+Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and let $S$ be the convex hull of the $p_i$’s and the $q_i$’s.
+$S$ is called the \emph{parallelepiped circumscribed} to $T$.
+\end{definition}
+\paco{define the ``corner tetrahedra''}
+
+The idea to have in mind is that of the regular tetrahedron inscribed in a cube. Modulo an affine transformation this is exactly the situation of $T$ and $S$.
+
+Some properties of this construction:
+
+1) $S$ is ``half a fundamental domain’’ for the affine lattice $\Lambda_P$ generated by $p_1,…,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
+Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it.
+(Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).
+
+\begin{lemma}
+\label{lemma:corner}
+Let $T=\conv\{p_1,p_2,p_3,p_4\}$ be lattice tetrahedron and $C(T)$ be the parallelepiped circumscribed to $T$.
+Let $T_1, T_2,T_3$ and $T_4$ be the corresponding corner tetrahedra in $C(T)$.
+Then, every $T_i$ contains at least one lattice point different from $\{p_1,\dots,p_4\}$.
+\end{lemma}
+
+\begin{proof}
+%2) As a corollary, if $T$ is a non-unimodular lattice tetrahedron with respect to a certain lattice $\Lambda$ then $S$ contains some additional lattice point $p_5$.
+Proof: since $T$ is not unimodular, there is a lattice point $x \in \Lambda \setminus \Lambda_P$.
+Then either $x + \Lambda_P$ or $x-\Lambda_P$ contains the desired lattice point.
+%
+%3) Even more so, if $T$ is a non-unimodular   **empty**  lattice tetrahedron then each of the four ``corner tetrahedra’’ in $S$ (the tetrahedra $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$) contains a lattice point $x \in \Lambda \setminus \Lambda_P$.
+Proof:
+This requires using the classification of empty tetrahedra, which can be rephrased as follows: let $T$ be an empty tetrahedron and consider the tiling of 3-space by copies of $T$, its opposite, and octahedra.
+(Think of the tiling of $\R^3$ by hyperplanes of the form $\pm x \pm y \pm z = even$).
+Then, all lattice points lie in parallelograms that divide the octahedra in two.
+Moreover, the lattice points in each of the parallelograms are a discrete cyclic group of the torus obtained gluing opposite sides of it.
+The key property now is that one fourth of the fundamental parallelogram goes to each of the corner tetrahedra (where by “one fourth” I mean the triangle obtained joining the center of the parallelogram of one of the edges).
+And, by the classification of empty tetrahedra, each of these fourths of parallelogram necessarily contains at least a lattice point of $\Lambda \setminus \Lambda_P$.
+(This last assertion is essentially Lemma 3.2(2) in my paper with Monica https://arxiv.org/abs/1409.6701).
+\end{proof}
+
+\begin{lemma}
+\label{lemma:3<4}
+Let $P$ be a lattice parallelepiped and let $T\subset P$ be an empty tetrahedron. Then, at least one of the four corner tetrahedra in $C(T)$ is fully contained in $P$.
+\end{lemma}
+
+\begin{proof}
+\paco{fill in details...}
+Each of the three pairs of parallel facets of $P$ cuts out at most one of the four new vertices in $C(T)$.
+\end{proof}
+
+\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
+\paco{rewrite}
+4) For every point $p_5 \in S \setminus T$ we have that $\conv(T\cup \{p_5\})$ can be triangulated into three tetrahedra of volume strictly smaller than that of $T$.
+Proof: $S$ is naturally divided into five tetrahedra: $T$ plus the four ``corner tetrahedra’’ $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$.
+Suppose wlog that $p_5 \in \conv(p_1,p_2,p_3, q_4)$.
+Then $\conv(T\cup \{p_5\})$ can be triangulated into
+\[
+\conv(p_1,p_2,p_4, p_5),
+\conv(p_1,p_3,p_4, p_5),
+\conv(p_2,p_3,p_4, p_5).
+\]
+
+Putting these properties together we get the following: Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if: ``for every empty tetrahedron $T$ contained in $P$ at least one of the other four vertices of the circumscribed parallelepiped $S$ is also in $P$’'.
+
+Theorem: Every lattice polytope with the circumscribed parallelepiped property has a unimodular cover.
+
+Proof: triangulate $P$ into empty simplices and iteratively apply properties (3) and (4) to them.
+This allows to cover every non-unimudular lattice simplex contained in $P$ with lattice simplices of smaller volume, until you cover by unimodular ones.
+QED
+\end{proof}
+
+\begin{remark}
+\label{rem:octahedra_and_tri-prisms}
+Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is
+contained in $P$’'. Then, $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.
+
+In turn, our proof of \Cref{lemma:3<4} relies on the fact that parallelepipeds have only three (pairs of) normal vectors.
+The proof fails if $P$ has four of them, as the following non-IDP octahedron and triangular prism show:
+\[
+\text{ write the octa and tri prism}.
+\]
+In both cases, the only lattice points are the six vertices and the origin. The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
+\end{remark}
+
+So, the questions to look at are:
+
+1) Does every lattice $3$-parallelepiped have the circumscribed parallelepiped property? I think yes.
+
+2) Does the Cayley embedding of a $P$ and a $Q$ such that one normal fan refines the other have the circumscribed parallelepiped property?
+Probably not in general, but perhaps yes in some cases; probably yes if one of them is a Minkowski summand of the other.
+
+2’) Even if the answer to (2) is negative, the following can be enough to prove unimodular cover: find a triangulation of $Cayley(P,Q)$ in which all the tetrahedra with an edge in $P$ and an edge in $Q$ have the
+
+3) The long shot: does every smooth 3-polytope have the circumscribed parallelepiped property?
+
+Paco
+
+\section{nef Cayley polytopes}
+\label{sec:cayley}
 

### 2019-05-13 unimodular-covers.tex
wrote draft of cayley section

@@
 
-So, the questions to look at are:
+\begin{question}
+Does every smooth 3-polytope have the circumscribed parallelepiped property? By the above remark, if the answer is yes then every smooth $3$-polytope has a unimodular cover.
+\end{question}
 
-1) Does every lattice $3$-parallelepiped have the circumscribed parallelepiped property? I think yes.
+\section{nef Cayley sums}
+\label{sec:cayley}
 
-2) Does the Cayley embedding of a $P$ and a $Q$ such that one normal fan refines the other have the circumscribed parallelepiped property?
-Probably not in general, but perhaps yes in some cases; probably yes if one of them is a Minkowski summand of the other.
+Throughout this section let $P$ and $Q$ be two lattice polytopes in $\R^d$.
+We do not require them to be full-dimensional, but we assume their Minkowski sum is.
+Remember that the Minkowski sum $P+Q$ of $P$ and $Q$ is defined as:
+\[
+P + Q := \{ p+q \in \R^d: p\in P, q \in Q\}.
+\]
+The \emph{Cayley sum} of $P$ and $Q$ is defined as
+\[
+\cayley (P,Q) = \conv( P\times\{0\} \cup Q\times \{1\}) \in \R^{d+1}.
+\]
+The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ is a $(d+1)$-dimensional.
 
-2’) Even if the answer to (2) is negative, the following can be enough to prove unimodular cover: find a triangulation of $Cayley(P,Q)$ in which all the tetrahedra with an edge in $P$ and an edge in $Q$ have the
+We introduce the following two concepts.
 
-3) The long shot: does every smooth 3-polytope have the circumscribed parallelepiped property?
+\begin{definition}
+In the above conditions, we say that $Q$ \emph{is nef for $P$} if the normal fan of $Q$ refines that of $P$. If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.
 
-Paco
+We say that the pair $(P,Q)$ is \emph{mixed IDP} if 
+\paco{``mixed IDP'' is not the right property, since this is weaker than IDP}
+\[
+(P\cap \Z^d) + (Q\cap \Z^d) = (P+Q)\cap \Z^d.
+\]
+\end{definition}
 
-\section{nef Cayley polytopes}
-\label{sec:cayley}
+In this language, Problems 3 and 4 i \cite{Oda1997} translate to
+
+\begin{question}
+If $Q$ is nef for $P$, is the pair $(P,Q)$ mixed IDP?
+\end{question}
+
+In dimension two, the answer is always yes by \cite{HNPS2008}.
+In dimension three it is easy to construct counter-examples (e.g., let $P=Q$ be the empty simplex of volume two), but the case where $P$ is smooth is open, and attributed as a conjecture of Oda.
+
+\begin{proposition}
+If $\cayley(P,Q)$ is IDP then the pair $(P,Q)$ is mixed IDP.
+\end{proposition}
+
+The so-called \emph{Cayley Trick} is the isomorphism
+\[
+2\cayley(P, Q) \cap (\R^d\times \{1\}) \cong P+Q,
+\]
+which induces the following canonical bijections:
+\begin{align*}
+\text{polyhedral subdivisions of $\cayley(P,Q)$} &\leftrightarrow \text{mixed subdivisions of $P + Q$}\\
+\text{triangulations of $\cayley(P,Q)$} &\leftrightarrow \text{fine mixed subdivisions of $P + Q$}\\
+\text{unimodular simplices in $\cayley(P,Q)$} &\leftrightarrow \text{unimodular prod-simplices in $P + Q$}.
+\end{align*}
+See \cite{DLRS2010} for more details on the Cayley Trick and on triangulations and polyhedral subdivisions of polytopes.
+In fact these bijections can be taken as definitions of the objects in the right-hand sides. In particular, 
+we call \emph{prod-simplices} in $P+Q$ the Minkowski sums $T_1+T_2$ where $T_1\subset P$ and $T_2\subset Q$ are simplices with complementary affine spans.
+A prod-simplex is \emph{unimodular} if the edge vectors from a vertex of $T_1$ and from a vertex of $T_2$ form a unimodular basis.
+
+%As a consequence, we also have that:
+%\begin{align*}
+%\text{$\cayley(P,Q)$ has a unimodular cover} &\leftrightarrow \text{$P+Q$ can be covered by unimodular prod-simplices}.
+%\end{align*}
+
+Our proof of \Cref{thm:cayley} is based on the following result:
+
+\begin{lemma}
+\label{lemma:cayley}
+Let $P$ and $Q$ be two lattice polygons with $Q$ nef with respect to $P$.
+Let $p_1p_2 \subset P$ and $q_1q_2\subset Q$ be two primitive and non-parallel segments, and assume that the prod-simplex (parallelogram)  $p_1p_2 + q_1q_2$ is not unimodular.
+Then,
+at least one of the strips
+\paco{add picture for these strips, and perhaps for the proof}
+\[
+(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle) \cap P
+\qquad \text{and} \qquad
+(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle) \cap Q
+\]
+has a lattice point in its interior.
+\end{lemma}
+
+\begin{proof}
+To DO
+\end{proof}
+
+\begin{proof}[Proof of \Cref{thm:cayley}]
+TO DO 
+\end{proof}
 

### 2019-05-13 unimodular-covers.tex
finished first full version (modulo some proofs and pictures)

@@
 \newcommand{\conv}{\ensuremath{\mathrm{conv}}\hspace{1pt}}
-\newcommand{\cayley}{\ensuremath{\mathrm{Cay}}\hspace{1pt}}
-\newcommand{\rank}{\ensuremath{\mathrm{rank}}\hspace{1pt}}
-\newcommand{\reg}{\ensuremath{\mathrm{reg}}\hspace{1pt}}
-\newcommand{\depth}{\ensuremath{\mathrm{depth}}\hspace{1pt}}
+\newcommand{\cayley}{\operatorname{Cay}}
+\newcommand{\rank}{\operatorname{rank}}
+\newcommand{\reg}{\operatorname{reg}}
+\newcommand{\depth}{\operatorname{depth}}
 
@@
 
-\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230 and, while they were in residence at the Mathematical Sciences Research Institute in Berkeley, California during the Fall 2017 semester, by the Clay Institute and the National Science Foundation (Grant No.
-DMS-1440140).
+\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
 Work of F. Santos is also supported by project MTM2017-83750-P of the Spanish Ministry of Science (AEI/FEDER, UE)}
@@
 We abbreviate this by saying that ``$P$ is IDP''.
-Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (normality of toric varieties) and it falls into a hierarchy of some ten properties each stronger than the previous one.
+Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (normality of toric varieties) and it falls into a hierarchy of several properties each stronger than the previous one.
 See, e.g., \cite[Sect.
@@
 2097f]{mfo2004}, \cite[p.
-2313ff]{mfo2007}.
+2313f]{mfo2007}.
 \paco{cite Bruns-Gubeladze book?}
@@
 \paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions}
-Related to this, a lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice.
-
-Oda~\cite{Oda1997} posed several questions/conjectures regarding IDP polytopes, among which are the following two:
-\begin{question}
+A lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice. 
+%
+Oda~\cite{Oda1997} posed several questions regarding IDP polytopes which are now considered conjectures (REFERENCES). 
+Among them:
+\begin{conjecture}
+\label{conj:Oda}
 \begin{enumerate}
 \item Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
-\[
-(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d
-\]
+\begin{align*}
+\label{eq:mixedIDP}
+(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
+\end{align*}
 \paco{perhaps we can call this formula ``being mixed IDP''. The property of ``$N(Q)$ refines $N(P)$'' is sometimes called ``$P$ is a Minkowski summand of $Q$''.}
 
-\item Suppose $P$ is smooth. Then $P$ is IDP.
+\item Every smooth lattice polytope is IDP.
 \end{enumerate}
-\end{question}
+\end{conjecture}
 
-Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes. In particular:
-\begin{enumerate}
-\item Haase et al.~\cite{HNPS2008} proved the mixed IDP property for $2$-polytopes $P$ and $Q$ when $N(Q) \le N(P)$.
+Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes.
 
-\item Beck et al.~\cite{BHHHJKM2018} proved IDP for smooth centrally symmetric $3$-polytopes.
-\end{enumerate}
+For example,  very recently
+Beck et al.~\cite{BHHHJKM2019} proved that all smooth centrally symmetric $3$-polytopes are IDP.
+More precisely, they show that any such polytope can be covered by lattice 
+parallelepipeds and unimodular simplices, both of which are trivially IDP.
 
-We improve both of these results to the level of unimodular covers.
-
-The structure of the paper is as follows: 
-%after introducing the necessary tools and concepts in \Cref{sec:prelim},  
-\Cref{sec:parallelepipeds} is devoted to $3$-dimensional lattice parallelepipeds. 
-Observe that lattice parallelotopes in arbitrary dimension are trivially IDP. Here we show:
+Partially motivated by this result in \Cref{sec:parallelepipeds} we show:
 
@@
 
-This, in turn, implies our improvement of Beck. et al.'s result on smooth c.s. 3-polytopes:
+As a consequence:
 
@@
 Every smooth centrally symmetric lattice $3$-polytope has a unimodular cover. 
+\qed
 \end{corollary}
 
-\begin{proof}
-Let $P$ be a smooth centrally symmetric lattice $3$-polytope.
-Beck et al.~\cite{BHHHJKM2018} show that $P$ is IDP by covering it with parallelepipeds and unimodular simplices.
+\begin{remark}
+Higher dimensional parallelotopes (affine images of cubes) are also IDP, but we do not know whether 
+they have unimodular covers, or whether $3$-dimensional ones have unimodular triangulations.
+\end{remark}
+
+%\begin{proof}
+%Let $P$ be a smooth centrally symmetric lattice $3$-polytope.
+%Beck et al.~\cite{BHHHJKM2019} show that $P$ is IDP by covering it with parallelepipeds and unimodular simplices.
 Once we know parallelepipeds have unimodular covers (\Cref{thm:parallelepipeds}), we conclude $P$ has a unimodular cover too.
-\end{proof}
+%\end{proof}
 
-Then in \Cref{sec:cayley} we look at Cayley sums and show:
+Another case of Oda's questions that has been solved is the two-dimensional case of \Cref{conj:Oda}(1), with three different proofs by
+Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth. 
+An alternative approach to \Cref{conj:Oda}(1) is via Cayley sums. As we note in \Cref{sec:cayley} (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
+\[
+\cayley(P,Q) \text{ is IDP} 
+\Rightarrow
+(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
+\]
+In \Cref{sec:cayley} we show the following even stronger result, in the two-dimensional case:
 
@@
 
+As an application of this, we prove \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
+
 \begin{corollary}
-\label{coro:3cs}
-Every smooth $3$-dimensional prismatoid has a unimodular cover.
+\label{coro:prismatoid}
+Every smooth $3$-dimensional lattice prismatoid has a unimodular cover.
 \end{corollary}
-
-\Paco{This corollary can be made stronger: it works for every $3$-prismatoid 
-$P= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\}$ with $k\ge 2$ and such that $P\cap\{z=i\}$ is a lattice polytope for every $i=0,\dots, k$.}
 
@@
 
-The main tool is what I call the “parallelepiped circumscribed to a given tetrahedron, defined as follows:
+The main tool is what I call the parallelepiped circumscribed to a given tetrahedron, defined as follows:
 
@@
 Definition: Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$.
-Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and let $S$ be the convex hull of the $p_i$’s and the $q_i$’s.
+Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and let $S$ be the convex hull of the $p_i$s and the $q_i$s.
 $S$ is called the \emph{parallelepiped circumscribed} to $T$.
@@
 
-1) $S$ is ``half a fundamental domain’’ for the affine lattice $\Lambda_P$ generated by $p_1,…,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
+1) $S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
 Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it.
@@
 %
-%3) Even more so, if $T$ is a non-unimodular   **empty**  lattice tetrahedron then each of the four ``corner tetrahedra’’ in $S$ (the tetrahedra $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$) contains a lattice point $x \in \Lambda \setminus \Lambda_P$.
+%3) Even more so, if $T$ is a non-unimodular   **empty**  lattice tetrahedron then each of the four ``corner tetrahedraï¿½ï¿½F in $S$ (the tetrahedra $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$) contains a lattice point $x \in \Lambda \setminus \Lambda_P$.
 Proof:
@@
 Moreover, the lattice points in each of the parallelograms are a discrete cyclic group of the torus obtained gluing opposite sides of it.
-The key property now is that one fourth of the fundamental parallelogram goes to each of the corner tetrahedra (where by “one fourth” I mean the triangle obtained joining the center of the parallelogram of one of the edges).
+The key property now is that one fourth of the fundamental parallelogram goes to each of the corner tetrahedra (where by 'one fourth' I mean the triangle obtained joining the center of the parallelogram of one of the edges).
 And, by the classification of empty tetrahedra, each of these fourths of parallelogram necessarily contains at least a lattice point of $\Lambda \setminus \Lambda_P$.
@@
 4) For every point $p_5 \in S \setminus T$ we have that $\conv(T\cup \{p_5\})$ can be triangulated into three tetrahedra of volume strictly smaller than that of $T$.
-Proof: $S$ is naturally divided into five tetrahedra: $T$ plus the four ``corner tetrahedra’’ $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$.
+Proof: $S$ is naturally divided into five tetrahedra: $T$ plus the four ``corner tetrahedra'' $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$.
 Suppose wlog that $p_5 \in \conv(p_1,p_2,p_3, q_4)$.
@@
 
-Putting these properties together we get the following: Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if: ``for every empty tetrahedron $T$ contained in $P$ at least one of the other four vertices of the circumscribed parallelepiped $S$ is also in $P$’'.
+Putting these properties together we get the following: Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if: ``for every empty tetrahedron $T$ contained in $P$ at least one of the other four vertices of the circumscribed parallelepiped $S$ is also in $P$''.
 
@@
 Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is
-contained in $P$’'. Then, $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.
+contained in $P$''. Then, $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.
 
@@
 \begin{proposition}
+\label{prop:mixedIDP}
 If $\cayley(P,Q)$ is IDP then the pair $(P,Q)$ is mixed IDP.
@@
 
+\Cref{coro:prismatoid} is a particular case of the following situation. 
+Let $Q_1$ and $Q_2$ be two polygons and consider the prismatoid 
+\[
+P= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\},
+\]
+with $k\ge 2$. 
+%Assume that $P\cap(\R^2\times\{i\})$ is a lattice polytope for every $i=0,\dots, k$. This is equivalent to assuming that $P\cap(\R^2\times\{1\})$ is a lattice polytope:
+
+\begin{corollary}
+If $P\cap(\R^2\times\{1\})$ is a lattice polygon then $P$ has a unimodular cover.
+\end{corollary}
+
+\begin{proof}
+\paco{perhaps add details}
+The condition that $P\cap(\R^2\times\{1\})$ is a lattice polygon implies the same for $P\cap(\R^2\times\{i\})$, for every $i$.
+Then, each slice
+\[
+P \cap (\R^2\times[i,i+1])
+\]
+is a nef Cayley polytope, hence it has a unimodular cover by \Cref{thm:cayley}.
+\end{proof}
+
 \begin{thebibliography}{99}
+
+\bibitem{BHHHJKM2019}
+Matthias Beck, Christian Haase, Akihiro Higashitani, Johannes Hofscheier, Katharina Jochemko, Lukas Katth\"an, Mateusz Micha{\l}ek,
+Smooth Centrally Symmetric Polytopes in Dimension 3 are IDP,
+\emph{Ann.~Combin.}, in press, 2019.
+\url{https://doi.org/10.1007/s00026-019-00418-x}
+
+\bibitem{DLRS2010}
+J. A. De Loera, J. Rambau, F. Santos
+Triangulations: Structures for Algorithms and Applications, 539 pp.
+Algorithms and Computation in Mathematics, Vol. 25, Springer-Verlag. 
+ISBN: 978-3-642-12970-4
+
+\bibitem{HPPS-survey}
+Christian Haase, Andreas Paffenholz, Lindsay C. Piechnik, Francisco Santos. Existence of unimodular triangulations - positive results. 
+%Preprint May 2014, updated December 2017, 89 pages. 
+ \emph{Mem. Amer. Math. Soc.}, to appear.
+ Available as arXiv preprint \href{https://arxiv.org/abs/1405.1687}{arXiv:1405.16878}
+
+\bibitem{HNPS2008}
+Christian Haase, Benjamin Nill, Andreas Paffenholz, and Francisco Santos, Lattice points in Minkowski sums, 
+Electron J. Combin., 15 (2008), no. 1, Note 11, 5 pp.
+
+\bibitem{mfo2004}
+Mini-workshop: Ehrhart Quasipolynomials: Algebra, Combinatorics, and Geometry, Oberwolfach Rep.
+1 (2004), no.
+3, 2071Ð2101, Abstracts from the mini-workshop held August 15Ð21, 2004, Organized by Jes\'us A.~De Loera and Christian Haase, Oberwolfach Reports.
+Vol.
+1, no.
+3.
+MR MR2144157
+
+\bibitem{mfo2007}
+Mini-workshop: Projective normality of smooth toric varieties, Oberwolfach Rep.
+4 (2007), no.
+39/2007, Abstracts from the mini-workshop held August 12Ð18, 2007.
+Organized by Christian Haase, Takayuki Hibi and Diane MacLagan.
+
+\bibitem{Oda1997}
+Tadao Oda. Problems on Minkowski sums of convex lattice polytopes. 
+Abstract submitted at the Oberwolfach Conference ÒCombinatorial Convexity and Algebraic GeometryÓ 26.10--01.11, 1997.
+Available as arXiv preprint \href{https://arxiv.org/abs/0812.1418}{arXiv:0812.1418}, 2008.
 

### 2019-05-13 unimodular-covers.tex
some edits

@@
 \begin{enumerate}
-\item Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
+\item (Problems 3 and 4 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
 \begin{align*}
@@
 
-\begin{remark}
-Higher dimensional parallelotopes (affine images of cubes) are also IDP, but we do not know whether 
-they have unimodular covers, or whether $3$-dimensional ones have unimodular triangulations.
-\end{remark}
+\begin{question}
+Do $3$-dimensional parallelepipeds have unimodular triangulations?
+\end{question}
+
+\begin{question}
+Higher dimensional parallelotopes (affine images of cubes) are IDP. Do they have unimodular covers? 
+\end{question}
 
@@
 Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth. 
+In dimension three however the statement fails without the smoothness assumptin, as shown by letting $P=Q$ be the empty simplex of volume two.
+
 An alternative approach to \Cref{conj:Oda}(1) is via Cayley sums. As we note in \Cref{sec:cayley} (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
@@
 
-In this language, Problems 3 and 4 i \cite{Oda1997} translate to
-
-\begin{question}
-If $Q$ is nef for $P$, is the pair $(P,Q)$ mixed IDP?
-\end{question}
-
-In dimension two, the answer is always yes by \cite{HNPS2008}.
-In dimension three it is easy to construct counter-examples (e.g., let $P=Q$ be the empty simplex of volume two), but the case where $P$ is smooth is open, and attributed as a conjecture of Oda.
+In this language, \Cref{conj:Oda}(1) translates to:
+\emph{If $Q$ is nef for $P$ and $P$ is smooth then the pair $(P,Q)$ mixed IDP.}
 

### 2019-05-14 unimodular-covers.tex
proof of lemma:corner

@@
 1.2.5]{HPPS-survey}, \cite[p.
-2097f]{mfo2004}, \cite[p.
-2313f]{mfo2007}.
+2097]{mfo2004}, \cite[p.
+2313]{mfo2007}.
 \paco{cite Bruns-Gubeladze book?}
@@
 \begin{enumerate}
-\item (Problems 3 and 4 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
+\item (Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
 \begin{align*}
@@
 
-\item Every smooth lattice polytope is IDP.
+\item (Related to problems 2 and 5 in \cite{Oda1997})
+Every smooth lattice polytope is IDP.
 \end{enumerate}
@@
 \label{lemma:corner}
-Let $T=\conv\{p_1,p_2,p_3,p_4\}$ be lattice tetrahedron and $C(T)$ be the parallelepiped circumscribed to $T$.
+Let $T=\conv\{p_1,p_2,p_3,p_4\}$ be an empty lattice tetrahedron that is not unimodular and $C(T)$ be the parallelepiped circumscribed to $T$.
 Let $T_1, T_2,T_3$ and $T_4$ be the corresponding corner tetrahedra in $C(T)$.
@@
 \begin{proof}
+By White's classification of empty tetrahedra (\cite{White1964}, see also, e.~g.~\cite[Sect.~4.1]{HPPS-survey}), there is no loss of generality in assuming $T=\conv(p_1,p_2,p_3,p_4)$ with
+\[
+p_1=(0,0,0), \quad
+p_2=(1,0,0), \quad
+p_3=(0,0,1), \quad
+p_4=(a,b,1).
+\]
+where $b\ge 2$ is the (normalized) volume of $T$, and $a\in \{1,\dots,b-1\}$ satisfies $\gcd(a,b)=1$. This gives 
+\paco{picture of the quadrilaterals $p_1q_4p_2q_3$ and $q_2p_3q_1p_4$ would be good.}
+\begin{align*}
+q_1=\left(\frac{1+a}2,\frac{b}2,1\right), &&
+q_2=\left(\frac{a-1}2,\frac{b}2,1\right), \quad\\
+q_3=\left(\frac{1+a}2,\frac{b}2,0\right), &&
+q_4=\left(\frac{a-1}2,-\frac{b}2,0\right).
+\end{align*}
+
+Then, the inequalities $b\ge 1+a$ and $b\ge 2$ imply:
+\[
+(1,1,0)\in \conv(p_1p_2q_3) \subset T_4, \quad
+(0,-1,0)\in \conv(p_1p_2q_4) \subset T_3.
+\]
+\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
+Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.
+Hence, its translate $\conv(q_2p_3q_1p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$.
+By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
+%
 %2) As a corollary, if $T$ is a non-unimodular lattice tetrahedron with respect to a certain lattice $\Lambda$ then $S$ contains some additional lattice point $p_5$.
@@
 %3) Even more so, if $T$ is a non-unimodular   **empty**  lattice tetrahedron then each of the four ``corner tetrahedraï¿½ï¿½F in $S$ (the tetrahedra $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$) contains a lattice point $x \in \Lambda \setminus \Lambda_P$.
-Proof:
-This requires using the classification of empty tetrahedra, which can be rephrased as follows: let $T$ be an empty tetrahedron and consider the tiling of 3-space by copies of $T$, its opposite, and octahedra.
+%Proof:
+%This requires using the classification of empty tetrahedra, which can be rephrased as follows: let $T$ be an empty tetrahedron and consider the tiling of 3-space by copies of $T$, its opposite, and octahedra.
 (Think of the tiling of $\R^3$ by hyperplanes of the form $\pm x \pm y \pm z = even$).
@@
 
+\bibitem{White1964}
+G.~K.~White.
+Lattice tetrahedra.
+\emph{Canadian J.~Math.} 16 (1964), 389--396.
+
 \end{thebibliography}

### 2019-05-24 unimodular-covers.tex
added proofs in section 2

@@
 
-The main tool is what I call the parallelepiped circumscribed to a given tetrahedron, defined as follows:
+The main tool is what we call the parallelepiped circumscribed to a given tetrahedron, defined as follows:
 
 \begin{definition}
-Definition: Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$.
-Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and let $S$ be the convex hull of the $p_i$s and the $q_i$s.
-$S$ is called the \emph{parallelepiped circumscribed} to $T$.
+Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$.
+Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and define $C(T)=\conv(p_i,q_i: i\in[4])$.
+$C(T)$ is a parallelepiped whose faces are of the form $\conv(p_i, p_j, q_k, q_l)$ for all $\{i,j,k,l\}=[4]$.
+We call $C(T)$ the \emph{parallelepiped circumscribed} to $T$.
+
+For each $i \in [4]$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=[4]$; we call these $T_i$ the \emph{corner tetrahedra} of $C(T)$. These along with $T$ form a triangulation of $C(T)$.
 \end{definition}
@@
 The idea to have in mind is that of the regular tetrahedron inscribed in a cube. Modulo an affine transformation this is exactly the situation of $T$ and $S$.
+\giulia{we should add a picture of the standard cube}
 
@@
 By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
-%
-%2) As a corollary, if $T$ is a non-unimodular lattice tetrahedron with respect to a certain lattice $\Lambda$ then $S$ contains some additional lattice point $p_5$.
-Proof: since $T$ is not unimodular, there is a lattice point $x \in \Lambda \setminus \Lambda_P$.
-Then either $x + \Lambda_P$ or $x-\Lambda_P$ contains the desired lattice point.
-%
-%3) Even more so, if $T$ is a non-unimodular   **empty**  lattice tetrahedron then each of the four ``corner tetrahedraï¿½ï¿½F in $S$ (the tetrahedra $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$) contains a lattice point $x \in \Lambda \setminus \Lambda_P$.
-%Proof:
-%This requires using the classification of empty tetrahedra, which can be rephrased as follows: let $T$ be an empty tetrahedron and consider the tiling of 3-space by copies of $T$, its opposite, and octahedra.
-(Think of the tiling of $\R^3$ by hyperplanes of the form $\pm x \pm y \pm z = even$).
-Then, all lattice points lie in parallelograms that divide the octahedra in two.
-Moreover, the lattice points in each of the parallelograms are a discrete cyclic group of the torus obtained gluing opposite sides of it.
-The key property now is that one fourth of the fundamental parallelogram goes to each of the corner tetrahedra (where by 'one fourth' I mean the triangle obtained joining the center of the parallelogram of one of the edges).
-And, by the classification of empty tetrahedra, each of these fourths of parallelogram necessarily contains at least a lattice point of $\Lambda \setminus \Lambda_P$.
-(This last assertion is essentially Lemma 3.2(2) in my paper with Monica https://arxiv.org/abs/1409.6701).
 \end{proof}
@@
 \label{lemma:3<4}
-Let $P$ be a lattice parallelepiped and let $T\subset P$ be an empty tetrahedron. Then, at least one of the four corner tetrahedra in $C(T)$ is fully contained in $P$.
+Let $P$ be a lattice parallelepiped and let $T\subset P$ be an empty tetrahedron.
+Then, at least one of the four corner tetrahedra $T_i$ of the circumscribed parallelogram $C(T)$ is fully contained in $P$.
 \end{lemma}
@@
 \begin{proof}
-\paco{fill in details...}
-Each of the three pairs of parallel facets of $P$ cuts out at most one of the four new vertices in $C(T)$.
+We denote the vertices of $T$ by $p_1, p_2, p_3, p_4$ and by $q_1, q_2, q_3, q_4$ the vertices of $C(T)$ not in $T$, as before.
+Let $f\in (\R^3)^*$ be any functional.
+Fix $i \in [4]$; observe that if $f(q_i) > f( p_j)$ for all $j\neq i$, then
+\begin{gather*}
+f(q_i) \overset{(i)}{>} f(q_j) \overset{(ii)}{>} f(p_i) \\
+f(q_i) \overset{(iii)}{>} f(p_j) \overset{(iv)}{>} f(p_i) \\
+f(p_j)\overset{(v)}{>}f(q_k)
+\end{gather*}
+for all $j \neq i$, $k \notin \{j, i\}$.
+
+Indeed, for any $\{i,j,k,l\}=[4]$ we have
+
+\begin{align*}
+f(q_i) > f( p_j) \Leftrightarrow \frac12 f(p_l+&p_k -p_j-p_i) > 0  \Leftrightarrow f(q_j) > f(p_i)\\
+& \Updownarrow \\
+ f(q_l) < f(p_k) \Leftrightarrow \frac12 f(-p_l-&p_k +p_j+p_i) < 0 \Leftrightarrow f(q_k) < f(p_l),
+\end{align*}
+which shows the equivalence of $(i)$, $(iii)$ and $(v)$; note that $(iii)$ is our assumption.
+Inequalities $(ii)$ and $(iv)$ follow by taking the sum of two inequalities of the form $\frac12 f(p_l+p_k -p_j-p_i) > 0$.
+
+Using this observation, we can show that each of pairs of parallel facets of $P$ cuts out at most one of the $q_i$s:
+if $q_i$ lies outside of $P$, it must violate a facet-defining inequality $f(x) \leq a$.Thus $f(q_i) > a \geq f(p_j)$, and by our observation, it follows that all remaining $q_j$ satisfy the inequality $f(x) \leq a$.
+Since $P$ is a parallelogram, there is a facet with inequality $f(x) \geq b$; by what is said above, the minimum value of $f(x)$ is attained at $p_i$, which is in $P$.
+Thus $f(q_j) > f(p_i) \geq b$, that is, the inequality is satisfied by all the $q_j$.
+
+Since at most one of the $q_i$ does not satisfy the inequalities corresponding to a pairs of parallel facets of $P$, and there are only three pairs of facets, one of $q_i$s must satisfy all the inequalities and therefore be inside of $P$.
 \end{proof}
+\giulia{here it may also be nice to add a picture}
 
 \begin{proof}[Proof of \Cref{thm:parallelepipeds}]
-\paco{rewrite}
-4) For every point $p_5 \in S \setminus T$ we have that $\conv(T\cup \{p_5\})$ can be triangulated into three tetrahedra of volume strictly smaller than that of $T$.
-Proof: $S$ is naturally divided into five tetrahedra: $T$ plus the four ``corner tetrahedra'' $\conv(p_i,p_j,p_k,q_l\}$ where $\{i,j,k,l\}= \{1,2,3,4\}$.
-Suppose wlog that $p_5 \in \conv(p_1,p_2,p_3, q_4)$.
+Let $P \subseteq \R^3$ be a parallelopiped.
+It can be triangulated into empty lattice tetrahedra, simply by taking a triangulation which uses all lattice points as vertices.
+If all the tetrahedra in the triangulation are unimodular we are done.
+Otherwise, suppose $T$ is be empty tetrahedron in the triangulation which is not unimodular.
+\Cref{lemma:3<4} guarantees that one of the corner tetrahedra $T_i$ of the parallelepiped $C(T)$ circumscribed to $T$ is contained in $P$.
+Without loss of generality, suppose $T_4 = \conv(p_1, p_2, p_3,q_4)$ is in $P$.
+By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point, which we denote by $l$.
+
 Then $\conv(T\cup \{p_5\})$ can be triangulated into 
 \[
-\conv(p_1,p_2,p_4, p_5),
-\conv(p_1,p_3,p_4, p_5),
-\conv(p_2,p_3,p_4, p_5).
+\conv(p_1,p_2,p_4, l),
+\conv(p_1,p_3,p_4, l),
+\conv(p_2,p_3,p_4, l).
 \]
 
-Putting these properties together we get the following: Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if: ``for every empty tetrahedron $T$ contained in $P$ at least one of the other four vertices of the circumscribed parallelepiped $S$ is also in $P$''.
-
-Theorem: Every lattice polytope with the circumscribed parallelepiped property has a unimodular cover.
-
-Proof: triangulate $P$ into empty simplices and iteratively apply properties (3) and (4) to them.
-This allows to cover every non-unimudular lattice simplex contained in $P$ with lattice simplices of smaller volume, until you cover by unimodular ones.
-QED
+Since these all have volume strictly smaller than $T$ \giulia{not convinced this is so obvious or true for any $l$: the volume of $\conv(T \cup q_4)$ is 3/2 times that of $T$, and if two of the three have very small volume...
+But I think (perhaps check in coordinates) we can choose $l$ wisely.}, this allows us to cover $T$ with lattice simplices of smaller volume recursively, until it is covered by unimodular ones.
 \end{proof}
@@
 Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is
-contained in $P$''. Then, $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.
+contained in $P$''. Then $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.
 
@@
 1 (2004), no.
-3, 2071Ð2101, Abstracts from the mini-workshop held August 15Ð21, 2004, Organized by Jes\'us A.~De Loera and Christian Haase, Oberwolfach Reports.
+3, 2071-2101, Abstracts from the mini-workshop held August 15-21, 2004, Organized by Jes\'us A.~De Loera and Christian Haase, Oberwolfach Reports.
 Vol.
@@
 4 (2007), no.
-39/2007, Abstracts from the mini-workshop held August 12Ð18, 2007.
+39/2007, Abstracts from the mini-workshop held August 12-18, 2007.
 Organized by Christian Haase, Takayuki Hibi and Diane MacLagan.
@@
 Tadao Oda. Problems on Minkowski sums of convex lattice polytopes. 
-Abstract submitted at the Oberwolfach Conference ÒCombinatorial Convexity and Algebraic GeometryÓ 26.10--01.11, 1997.
+Abstract submitted at the Oberwolfach Conference ''Combinatorial Convexity and Algebraic Geometry'' 26.10--01.11, 1997.
 Available as arXiv preprint \href{https://arxiv.org/abs/0812.1418}{arXiv:0812.1418}, 2008.

### 2019-05-26 unimodular-covers.tex
typos
finished rough draft of section 2

@@
 \newcommand{\GL}{{GL}_r (K)}
-\newcommand{\Gin}{\ensuremath{\mathrm{Gin}}}
-\newcommand{\gin}{\ensuremath{\mathrm{gin}}}
-\newcommand{\GIN}{\ensuremath{\mathrm{GIN}}}
-\newcommand{\Gins}{\ensuremath{\mathrm{Gins}}}
-\newcommand{\gins}{\ensuremath{\mathrm{gins}}}
-\newcommand{\Lex}{\ensuremath{\mathrm{Lex}}}
-\newcommand{\dele}[1]{\ensuremath{\Delta^e({#1})}}
-\newcommand{\dels}[1]{\ensuremath{\Delta^s({#1})}}
-\newcommand{\delc}[1]{\ensuremath{\Delta^c({#1})}}
-\newcommand{\Shift}{\ensuremath{\mathrm{Shift}}}
 
@@
 \newcommand{\width}{\operatorname{width}}
+\newcommand{\Vol}{\operatorname{Vol}}
 \newcommand{\length}{\operatorname{length}}
@@
 \newcommand{\reg}{\operatorname{reg}}
-\newcommand{\depth}{\operatorname{depth}}
 
@@
 \begin{definition}
-Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$.
-Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and define $C(T)=\conv(p_i,q_i: i\in[4])$.
-$C(T)$ is a parallelepiped whose faces are of the form $\conv(p_i, p_j, q_k, q_l)$ for all $\{i,j,k,l\}=[4]$.
-We call $C(T)$ the \emph{parallelepiped circumscribed} to $T$.
+Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$. Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and define
+
+\[
+C(T)=\conv(p_i,q_i: i\in[4]).
+\]
+
+$C(T)$ is a parallelepiped whose faces are of the form $\conv(p_i, p_j, q_k, q_l)$ for all $\{i,j,k,l\}=[4]$, which we call the \emph{parallelepiped circumscribed} to $T$.
 
@@
 
-Some properties of this construction:
-
-1) $S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
+\Giulia{not sure if we need this remark:
+$S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
 Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it.
 (Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).
+}
 
@@
 q_3=\left(\frac{1+a}2,\frac{b}2,0\right), &&
-q_4=\left(\frac{a-1}2,-\frac{b}2,0\right).
+q_4=\left(\frac{1-a}2,-\frac{b}2,0\right).
 \end{align*}
@@
 By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point, which we denote by $l$.
+Observe that from the proof of \Cref{lemma:corner} we know more: $l$ lies in the face $\conv(p_1, p_2, q_4)$.\giulia{we need this for the volume computation below.
+perhaps it should be made explicit in the statemente of the lemma...}
 
-Then $\conv(T\cup \{p_5\})$ can be triangulated into 
+Then $S=\conv(T\cup \{l\})$ can be triangulated in two different ways: $S=T \cup T'_4 = S_1 \cup S_2 \cup S_3$, where $T'_4 = \conv(p_1, p_2, p_3, l) \subseteq T_4$ and
 \[
-\conv(p_1,p_2,p_4, l),
-\conv(p_1,p_3,p_4, l),
-\conv(p_2,p_3,p_4, l).
+S_1= \conv(p_2,p_3,p_4, l),
+S_2=\conv(p_1,p_3,p_4, l),
+S_3=\conv(p_1,p_2,p_4, l).
 \]
 
-Since these all have volume strictly smaller than $T$ \giulia{not convinced this is so obvious or true for any $l$: the volume of $\conv(T \cup q_4)$ is 3/2 times that of $T$, and if two of the three have very small volume...
-But I think (perhaps check in coordinates) we can choose $l$ wisely.}, this allows us to cover $T$ with lattice simplices of smaller volume recursively, until it is covered by unimodular ones.
+Each of the tetrahedra $S_i$ has lattice volume strictly smaller than $T$: indeed, $\Vol(S_3)=\Vol(T'_4)$, since these tetrahedra have a common face $\conv(p_1, p_2, l)$ and the remaining vertex is at distance one from this face in both.
+Now $T'_4$ is contained in $T_4$, and we know that each corner tetrahedron $T_i$ has volume half that of $T$.
+The remaining two tetrahedra $S_1$ and $S_2$ must therefore have volume summing to that of $T$, and therefore each has volume strictly less than $T$.
+
+This allows us to cover $T$ with lattice simplices of smaller volume recursively, until it is covered by unimodular ones.
 \end{proof}
@@
 The proof fails if $P$ has four of them, as the following non-IDP octahedron and triangular prism show:
+
+\begin{example}
+\label{ex:non-IDP}
+The following lattice octahedron is not IDP:
 \[
-\text{ write the octa and tri prism}.
+Q= \conv((0,1,1),(1,0,1),(1,1,0),(0,-1,-1),(-1,0,-1),(-1,-1,0))
 \]
-In both cases, the only lattice points are the six vertices and the origin. The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
+
+The following lattice triangular prism is not IDP:
+\[
+P=\conv((0,1,1),(1,0,1),(1,1,0),(-1,0,0),(0,-1,0),(0,0,-1)).
+\]
+Indeed, in both polytopes the only lattice points are the six vertices and the origin. The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
+\giulia{is it necessary to be more explicit, for example show that these are the only lattice points?}
+\end{example}
+
 \end{remark}

### 2019-05-29 unimodular-covers.tex
pf of thm 1.6 in section 3

@@
 \newcommand{\R}{ \ensuremath{\mathbb{R}}}
+
+\newcommand{\T}{ \ensuremath{\mathcal{T}}}
 
@@
 \end{lemma}
+\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $\{p_1, p_2\} + \{q_1, q_2\}$}
 
@@
 \end{proof}
+]
+\begin{proof}[Proof of \Cref{thm:cayley}]
+Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$.
+Observe that any simplex in $T$ has vertices in $Px\{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$.
+Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$).
+\giulia{is it usually denoted by dimension or cardinality?
+that is, (2,0) or (3,1)?} A triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained by simply unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$).
+This leaves simplices of type $(2,2)$.
+In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can I be without being confusing?}.
 
-\begin{proof}[Proof of \Cref{thm:cayley}]
-TO DO 
+Let $S=\{p_1, p_2\} + \{q_1, q_2\}$ be such a parallelogram.
+By \Cref{lemma:cayley} we know that there is a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $S$, w.l.o.g.
+suppose it is $p+q_1$.
+Then we can cover $S$ by the triangle $\{p_1, p_2, p\} + q_1$ and parallelograms $\{p_1, p\}  + \{q_1, q_2\}$ and $\{p_2, p\} + \{q_1, q_2\}$.
+As above, the triangle can be unimodularly triangulated, while the volume of the two parallelograms sums up to the volume of $S$, and therefore each has volume strictly smaller than $S$.
+By iterating this procedure, we can cover $S$ by unimodular cells.
 \end{proof}

### 2019-06-10 unimodular-covers.tex
started proof of lemma 3.3

@@
 \end{lemma}
-\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $\{p_1, p_2\} + \{q_1, q_2\}$}
+\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in the interior of $\{p_1, p_2\} + \{q_1, q_2\}$ (or symmetric for $Q$)}
 
 \begin{proof}
-To DO
+Suppose by contraddiction that there is no lattice point as described in the lemma.
+Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
+The same holds for $Q$ and its corresponding strip.
+Thus $P$ has two edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $e_1$ and $e_2$.
+The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.
+
+Since the normal fan of $Q$ refines that of $P$, $Q$ will contain edges $e_1'$ and $e_2'$ parallel to $e_1$ and $e_2$ respectively.
+These edges of $Q$ are all distinct and appear alternating in $Q$, that is, they appear in the order $f_1, e_1', f_2, e_2'$.
+Indeed, consider the quadrilateral $S$ with vertices $q_1, p_1+l, q_2, p_2+l$, where $l$ is a lattice point chosen so that the segments $q_1q_2$ and $p_1+l,p_2+l$ intersect in their interior; we can further require that $\conv( q_1, p_1+l, q_2)$ is unimodular \giulia{not sure if we use it}.
+$S$ lies inside the $Q$-strip, while its translate $S-l$ lies inside the $P$-strip.
+Thus in order to have no lattice points in the strips, we must have that the edge $f_i$ separates the vertex $p_i +l$ of $S$ from the remaining three, and  the same for the edge $e_i$ and vertex $q_i-l$ of $S-l$.
+This is equivalent to saying that the normals to the edges $f_1, e_1, f_2, e_2$ lie in different open cones of the normal fan of $S$, in this cyclic order.
+
+Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensively}.
+Since these segments do not contain lattice points, they must have lattice length stricly less than $1$ (or the width wrt $q$ is less than $k$, the index of $p,q$), which means that the smallest width in direction of $p_1 p_2$ between a "consecutive" vertices of $f_1$ and of $f_2$ must also be strictly less than $k$.
+
+Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
 \end{proof}

### 2019-06-13 unimodular-covers.tex
added Christian's comment

@@
 \end{corollary}
+
+\Paco{another corollary, suggested by Christian Haase: the $k$-th dilation of an emptyy simplex is very easy to decompose into slices of width one all of which fit into our description ($P$ and $Q$ are either pararllel parallelograms, one of which perhaps degenerates to a segment).
+In particular, we recover as a corollary that the $k$-th dilation of every empty tetrahedron (and, a fortiori, the $k$-th dilation of every lattice $3$-polytope) has a unimodular cover for every $k\ge 2$}
 

### 2019-06-13 unimodular-covers.tex
sketch of  proof of lemma 3.3

--- 2019-06-10 unimodular-covers.tex
@@
 $S$ lies inside the $Q$-strip, while its translate $S-l$ lies inside the $P$-strip.
-Thus in order to have no lattice points in the strips, we must have that the edge $f_i$ separates the vertex $p_i +l$ of $S$ from the remaining three, and  the same for the edge $e_i$ and vertex $q_i-l$ of $S-l$.
+Thus in order to have no lattice points in the strips, we must have that the edge $f_i$ separates the vertex $p_i +l$ of $S$ from the remaining three, and  the same for the edge $e_i +l$ and vertex $q_i$ of $S$.
 This is equivalent to saying that the normals to the edges $f_1, e_1, f_2, e_2$ lie in different open cones of the normal fan of $S$, in this cyclic order.
 
-Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensively}.
+- both $f_1$ or $f_2$ must meet the line through $q_1$ and $q_2$, else they would not cut out the lattice point they need to. same goes for $e_1$ and $e_2$ and the line through $p_1$ and $p_2$.
+
+claim: the segments $e_1 + l$ and $e_2 + l$ respectively intersect both $f_1$ and $f_2$.
+
+claim: if we tilt on of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.
+
+%Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensibly}.
 Since these segments do not contain lattice points, they must have lattice length stricly less than $1$ (or the width wrt $q$ is less than $k$, the index of $p,q$), which means that the smallest width in direction of $p_1 p_2$ between a "consecutive" vertices of $f_1$ and of $f_2$ must also be strictly less than $k$.
 
-Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
+%Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
 \end{proof}

### 2019-06-14 unimodular-covers.tex
drawn attempt of counter-example

@@
 \usetikzlibrary{matrix}
+\usepackage{tkz-euclide}
 
@@
 
+\begin{figure}[htb]
+  \centering
+\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+  .8]
+%coordinate axes
+   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%        \tkzLabelY[orig=false,label options={font=\tiny}]
+        \tkzGrid
+%       \tkzDrawX
+%       \tkzDrawY
+
+%hexagon
+  \node[mynode] (A) at (0,-1) {}; 
+%  \node[mynode] (B) at (1,1) {};  
+  \node[mynode] (C) at (2,3) {};  
+  \node[mynode] (D) at (0,1) {};  
+%  \node[mynode] (E) at (-1,-1) {};  
+  \node[mynode] (F) at (-2,-3) {};
+
+%other points
+%  \node[mynode] () at (0,-1) {};
+
+  \draw[black] (A) -- (C) -- (D) -- (F) -- (A);
+  \draw[black, thick] (A) -- (D);
+\end{tikzpicture}
+%
+\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+  .8]
+%coordinate axes
+   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%        \tkzLabelY[orig=false,label options={font=\tiny}]
+        \tkzGrid
+%        \tkzDrawX
+%        \tkzDrawY
+
+%hexagon
+  \node[mynode] (A) at (1,0) {}; 
+  \node[mynode] (B) at (2,1) {};  
+  \node[mynode] (C) at (3,3) {};  
+  \node[mynode] (D) at (-1,0) {};  
+  \node[mynode] (E) at (-2,-1) {};  
+  \node[mynode] (F) at (-3,-3) {};
+
+  \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
+  \draw[black, thick] (A) -- (D);
+\end{tikzpicture}
+
+\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
+
+\caption{Partial counter-example to \Cref{lemma:cayley}.
+It is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one).
+%
+With respect to the lattice of index two it is not a counter-example either, because two vertices of $Q$ are not in that lattice.
+}
+\label{fig:width3}
+\end{figure}
+
 \begin{lemma}
@@
 \end{proof}
-]
+
 \begin{proof}[Proof of \Cref{thm:cayley}]

### 2019-06-26 unimodular-covers.tex
added proof of offending lemma for index two

@@
 \newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{claim}[theorem]{Claim}
 
@@
 \begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  .8]
+  .7]
 %coordinate axes
@@
 
+  \node[mynode] (p1) at (-1,0) {}; 
+  \node[mynode] (p2) at (1,0) {};
+
   \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
-  \draw[black, thick] (A) -- (D);
+  \draw[black, thick] (p1) -- (p2);
 \end{tikzpicture}
@@
 \caption{Partial counter-example to \Cref{lemma:cayley}.
-It is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one).
+With respect to the standard lattice it is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one).
 %
@@
 (p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle) \cap P
-\qquad \text{and} \qquad
+\qquad \text{or} \qquad
 (q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle) \cap Q
@@
 \begin{proof}
-Suppose by contraddiction that there is no lattice point as described in the lemma.
+Suppose by contradiction that there is no lattice point as described in the lemma.
 Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
@@
 
+\appendix
+\section{Attempt at the proof of \Cref{lemma:cayley}}
+
+At some point I'll assume volume two for the parallelogram $p_1p_2 + q_1q_2$, but for the time being we don't.
+
+Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
+\[
+p_1=(0,-\alpha), \quad
+p_2=(0,1-\alpha), \quad
+q_1=(-\beta,0), \quad
+q_2=(1-\beta,0).
+\] 
+The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
+
+Suppose by contradiction that there is no lattice point as described in the lemma.
+Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
+The same holds for $Q$ and its corresponding strip.
+Thus $P$ has two primitive edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
+The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges  $b, t$ (with $b$ below and $t$ above).
+To make things more explicit, we assume the following coordinates for the edge-vectors:
+\[
+\vec l = (l_1,l_2), \quad
+\vec r= (r_1,r_2), \quad
+\vec b= (b_1,b_2), \quad
+\vec t= (t_1,t_2).
+\]
+Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
+
+Then:
+
+\begin{claim}
+$l_2, r_2, b_1, t_1 >1$.
+\end{claim}
+
+\begin{proof}
+$l_2, r_2, b_1, t_1 \ge 1$ follows from the fact that each edge crosses the corresponding strip.
+If one of the inequalities is not strict then one of the strips contains half of the square $p_1p_2+q_1q_2$, hence it contains a lattice point in its interior.
+\end{proof}
+
+\begin{claim}
+$b_2$ and $t_2$ are non-zero and have the same sign. W.l.o.g.~we assume they are positive.
+\end{claim}
+
+\begin{proof}
+If, say, $b_2 \le 0 \le t_2$.
+Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$).
+This implies the vertical distance between any two points in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
+\end{proof}
+
+\begin{claim}
+Assume w.l.o.g.~that $b$ has greater slope than $t$. Then, 
+\begin{enumerate}
+\item The intersection of $Q$ with any vertical line to the right of $q_2$ has length strictly smaller than one.
+\item $r_1>0$.
+\end{enumerate}
+\end{claim}
+
+\begin{proof}
+The assumption implies that the intersections of $t$ and $b$ with the vertical line $\{x=1-\beta\}$ through $q_2$ are less than distance one apart (consider a vertical lattice line intersecting $q_1q_2$ and two consecutive lattice points in it, one above and one below the horizontal axis).
+This in turn implies part (a).
+For (b), if $r_1$ was $\le 0$ it would be impossible to fit a translated copy of $r$ in the right side of $Q$, because of (a).
+\end{proof}
+
+As a conclusion of the last two claims we have that $b$, $t$ and $r$ have positive slope.
+
+Starting here we assume index 2.
+
+\begin{claim}
+The slopes of $t$ and $b$ are strictly smaller than one and that of $r$ strictly greater than one.
+\end{claim}
+
+\begin{proof}
+Otherwise the points $(0,1/2)$, $(0,-1/2)$, $(1/2,0)$, respectively, would be in the corresponding band.
+\end{proof}
+
+We are now ready to show a contradiction.
+Let $u$ and $v$ be the end-points of the translated copy of $r$ in $Q$, with $\vec r=v-u$.
+Let $d$ be the line of slope $1$ through $u$, which is a lattice line.
+Let $d'$ be the next parallel lattice line above $d$.
+By Claim A.3.1, the intersection of $d'$ with the vertical line through $u$ is above $Q$.
+Since the slope of $t$ is less than one, the same happens for every point of $d'$ to the right of $u$.
+In particular, $v$ must be below $d'$ and hence it must be on or below $d$.
+This contradicts the fact that the slope of $r$ is strictly greater than one.
+QED
+
 \end{document}

### 2019-07-02 unimodular-covers.tex
expanding proofs in appendix

@@
 \begin{proof}
-Suppose by contradiction that there is no lattice point as described in the lemma.
+%Suppose by contradiction that there is no lattice point as described in the lemma.
 Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
@@
 The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.
+
+Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
+\[
+p_1=(0,-\alpha), \quad
+p_2=(0,1-\alpha), \quad
+q_1=(-\beta,0), \quad
+q_2=(1-\beta,0).
+\] 
+The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
+
+Suppose by contradiction that there is no lattice point as described in the lemma.
+Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
+The same holds for $Q$ and its corresponding strip.
+Thus $P$ has two primitive edges which each have one vertex on each side of the strip $p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
+The same holds for $Q$ and the strip $q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle$, and we call the edges  $b, t$ (with $b$ below and $t$ above).
+To make things more explicit, we assume the following coordinates for the edge-vectors:
+\[
+\vec l = (l_1,l_2), \quad
+\vec r= (r_1,r_2), \quad
+\vec b= (b_1,b_2), \quad
+\vec t= (t_1,t_2).
+\]
+Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
 
@@
 
-claim: if we tilt on of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.
+claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.
 
@@
 $l_2, r_2, b_1, t_1 \ge 1$ follows from the fact that each edge crosses the corresponding strip. 
-If one of the inequalities is not strict then one of the strips contains half of the square $p_1p_2+q_1q_2$, hence it contains a lattice point in its interior.
+If one of the inequalities is not strict, then one edge must have vertices on the corresponding strip.
+Then one of $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle) \cap P$ or $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle) \cap Q$
+contains a triangle with one edge a translation of $p_1p_2$ and another edge a translation of $q_1q_2$.
+Such a triangle must contain a lattice point other than its vertices, and since $p_1p_2$ and $q_1q_2$ are primitive, the lattice point must be in the interior of the strip.
 \end{proof}
@@
 Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$).
-This implies the vertical distance between any two points in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
+This implies the width w.r.t to $f_q$, the functional constant on $q_1q_2$, of any segment with vertices in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
 \end{proof}
@@
 This in turn implies part (a).
+
 For (b), if $r_1$ was $\le 0$ it would be impossible to fit a translated copy of $r$ in the right side of $Q$, because of (a).

### 2019-07-03 unimodular-covers.tex
edited intro

@@
 We show that the following classes of three-dimensional lattice polytopes have unimodular covers: the class of all parallelepipeds, the class of all centrally symmetric polytopes, and the class of all three-dimensional Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$.
-This improves results of Beck et al.
-(2018) and Haase et al.
-(2008) where the last two classes were shown to be IDP.
+This improves results of Beck et al.~(2018) and Haase et al.~(2008) where the last two classes were shown to be IDP.
 \end{abstract}
@@
 We abbreviate this by saying that ``$P$ is IDP''.
-Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (normality of toric varieties) and it falls into a hierarchy of several properties each stronger than the previous one.
-See, e.g., \cite[Sect.
+Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (normality of toric varieties).
+It falls into a hierarchy of several properties each stronger than the previous one; see, e.g., \cite[Section 2.D]{BGbook}, \cite[Sect.
 1.2.5]{HPPS-survey}, \cite[p.
@@
 2313]{mfo2007}.
-\paco{cite Bruns-Gubeladze book?}
 Let us here only mention that
@@
 
-\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions}
+\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions?}
 A lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice. 
 %
-Oda~\cite{Oda1997} posed several questions regarding IDP polytopes which are now considered conjectures (REFERENCES). 
+Oda~\cite{Oda1997} posed several questions regarding IDP polytopes which are now considered conjectures~\cite{HNPS2008,mfo2007}. 
 Among them:
@@
 Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes. 
-
+%
 For example,  very recently
@@
 
-Partially motivated by this result in \Cref{sec:parallelepipeds} we show:
+In \Cref{sec:parallelepipeds} we show the following stronger statement:
 
@@
 
-\begin{question}
-Do $3$-dimensional parallelepipeds have unimodular triangulations?
-\end{question}
-
-\begin{question}
-Higher dimensional parallelotopes (affine images of cubes) are IDP. Do they have unimodular covers? 
-\end{question}
-
 %\begin{proof}
@@
 Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth. 
-In dimension three however the statement fails without the smoothness assumptin, as shown by letting $P=Q$ be the empty simplex of volume two.
+In dimension three however the statement fails without the smoothness assumption, as shown by letting $P=Q$ be any non-unimodular empty tetrahedron.
 
-An alternative approach to \Cref{conj:Oda}(1) is via Cayley sums. As we note in \Cref{sec:cayley} (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
+An alternative approach to \Cref{conj:Oda}(1) is via Cayley sums, that we discuss in  \Cref{sec:cayley}. 
+As we note there (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
 \[
@@
 \]
-In \Cref{sec:cayley} we show the following even stronger result, in the two-dimensional case:
+In particular, the following statement that we prove in \Cref{sec:cayley} is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:
 
@@
 
-As an application of this, we prove \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
+This not only implies \Cref{conj:Oda}(1) in dimension three, but also \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
 
@@
 \Paco{another corollary, suggested by Christian Haase: the $k$-th dilation of an emptyy simplex is very easy to decompose into slices of width one all of which fit into our description ($P$ and $Q$ are either pararllel parallelograms, one of which perhaps degenerates to a segment).
-In particular, we recover as a corollary that the $k$-th dilation of every empty tetrahedron (and, a fortiori, the $k$-th dilation of every lattice $3$-polytope) has a unimodular cover for every $k\ge 2$}
+In particular, we recover as a corollary that the $k$-th dilation of every empty tetrahedron (and, a fortiori, the $k$-th dilation of every lattice $3$-polytope) has a unimodular cover for every $k\ge 2$
+
+EXCEPT FOR THIS TO BE TRUE WE NEED TO ALLOW $Q$ TO DEGENERATE TO AN EDGE IN \Cref{hm:cayley}. CAN WE?
+}
 
@@
 
-\Cref{coro:prismatoid} is a particular case of the following situation. 
-Let $Q_1$ and $Q_2$ be two polygons and consider the prismatoid 
+We now turn our attention to \emph{prismatoids}, defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets.
+Since we are only interested in lattice polytopes, we here use \emph{prismatoid} meaning any $d$-polytope of the form
 \[
-P= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\},
+\conv(Q_1\times\{0\} \cup Q_2 \times \{k\},
+\]
+with $k\in \Z_{>0}$ and $Q_1,Q_2$ being lattice $(d-1)$-polytopes. This generalizes Cayley sums, which are the case $k=1$.
+
+\begin{proposition}
+\label{prop:prismatoid}
+Let $Q_1$ and $Q_2$ be two lattice polygons and consider the prismatoid 
+\[
+P:= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\},
 \]
@@
 %Assume that $P\cap(\R^2\times\{i\})$ is a lattice polytope for every $i=0,\dots, k$. This is equivalent to assuming that $P\cap(\R^2\times\{1\})$ is a lattice polytope:
-
-\begin{corollary}
 If $P\cap(\R^2\times\{1\})$ is a lattice polygon then $P$ has a unimodular cover.
-\end{corollary}
+\end{proposition}
 
@@
 
+\Cref{coro:prismatoid} follows from \Cref{prop:prismatoid} since smoothness of a prismatoid implies $P\cap(\R^2\times\{i\})$ to be a lattice polytope for every $i$.
+
+both smooth prismatoids and dilations of lattice polytopse of width 1
+
+is a particular case of the following situation.
+
 \begin{thebibliography}{99}
@@
 
+\bibitem{BGbook}
+Winfried Bruns and Joseph Gubeladze, \emph{Polytopes, rings, and k-theory}, Monographs in Mathematics,
+  Springer-Verlag, 2009, XIV, 461 p. 52 illus.
+
  \bibitem{DLRS2010}
@@
 ISBN: 978-3-642-12970-4
+
+\bibitem{Fakhruddin}
+Najmuddin Fakhruddin.
+\newblock Multiplication maps of linear systems on smooth projective toric
+  surfaces, arXiv:0208.5178, August 2002.
 
@@
 Available as arXiv preprint \href{https://arxiv.org/abs/0812.1418}{arXiv:0812.1418}, 2008.
+
+\bibitem{Ogata}
+Shoetsu Ogata.
+\newblock {Multiplication maps of complete linear systems on projective toric
+  surfaces.}
+\newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006.
 

### 2019-07-04 unimodular-covers.tex
put back two questions I accidentally removed
worked on the proof in the appendix

--- 2019-07-02 unimodular-covers.tex
@@
 \newcommand{\width}{\operatorname{width}}
+\newcommand{\ind}{\operatorname{index}}
 \newcommand{\Vol}{\operatorname{Vol}}
@@
 
-At some point I'll assume volume two for the parallelogram $p_1p_2 + q_1q_2$, but for the time being we don't.
+%At some point I'll assume volume two for the parallelogram $p_1p_2 + q_1q_2$, but for the time being we don't.
 
-Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
+Let us fix coordinates in which $p=\conv(p_1,p_2)$ and $q= \conv(q_1,q_2)$ are a vertical and horizontal segment of length one intersecting at the origin.
+That is, there are numbers $\alpha, \beta \in (0,1)$ such that
 \[
@@
 q_2=(1-\beta,0).
+\] \giulia{perhaps the explicit coordinates are superfluous now}
+The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
+We can further assume (since translating the polytopes by lattice vectors will not result in any loss of generality) that $p_1$ is on the consecutive lattice line parallel to the line $q$ lies on.
+That is, if $f_q$ is the primitive lattice functional which is zero on $q$, then $f_q(p_1)=-1$.
+Equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
+Let $f_p$ be the functional constant on $p$.
+We observe then that
+\[
+\width_{f_q}(p + \overrightarrow{q})=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\overrightarrow{p}) = \ind(p, q) =:i.
 \]
-The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
+
+Observe that $p_1$ is the unique lattice point in the strip $q +\overrightarrow{p}$ on the line $f_q(x)=-1$.
+Indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p$ and $q$ have index $1$, which is a contradiction.
+The unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.
 
@@
 The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges  $b, t$ (with $b$ below and $t$ above).
+Observe that one of $l$ and $r$ can coincide with $p$, if this is an edge of $P$, and similarly one of $t,b$ might be $q$, if this is an edge of $Q$.
 To make things more explicit, we assume the following coordinates for the edge-vectors:
@@
 
-\begin{claim}
-$l_2, r_2, b_1, t_1 >1$.
+%\begin{claim}
+%$l_2, r_2, b_1, t_1 >1$.
+%\end{claim}
+
+\begin{claim}\giulia{I think we need to treat separately the case where we deal with edges (these are easier, in any case, but I couldn't find a unification)}
+The following inequalities hold, provided the segment in question is not an edge of the corresponding polytope:
+\begin{align*}
+\width_{f_q}(l) > i, \\
+\width_{f_q}(r) > i, \\
+\width_{f_p}(t) > i, \\
+\width_{f_p}(b) > i.
+\end{align*}
 \end{claim}
@@
 \begin{proof}
-$l_2, r_2, b_1, t_1 \ge 1$ follows from the fact that each edge crosses the corresponding strip. 
-If one of the inequalities is not strict, then one edge must have vertices on the corresponding strip.
-Then one of $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle) \cap P$ or $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle) \cap Q$
-contains a triangle with one edge a translation of $p_1p_2$ and another edge a translation of $q_1q_2$.
+The inequality $\geq i$ follows in each case from the fact that each segment crosses the corresponding strip, the width of which (w.r.t the appropriate functional) is $i$, as observed above. 
+If one of the inequalities were not strict, then the corresponding segment, say $l$, must have vertices on the corresponding strip.
+Then the strip $(p + \langle \overrightarrow{q}\rangle) \cap P$ contains a triangle with one edge a translation of $p$ and another edge a translation of $q$.
 Such a triangle must contain a lattice point other than its vertices, and since $p_1p_2$ and $q_1q_2$ are primitive, the lattice point must be in the interior of the strip.
@@
 
+%\begin{claim}
+%$b_2$ and $t_2$ are non-zero and have the same sign. W.l.o.g.~we assume they are positive.
+%\end{claim}
+%
+%\begin{proof}
+%If, say, $b_2 \le 0 \le t_2$.
+Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$).
+This implies the width w.r.t to $f_q$, the functional constant on $q_1q_2$, of any segment with vertices in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
+%\end{proof}
+
 \begin{claim}
-$b_2$ and $t_2$ are non-zero and have the same sign. W.l.o.g.~we assume they are positive.
+The slopes of $b$ and $t$ must either be both strictly larger than the slope of $q$ or both be strictly smaller. W.l.o.g.~we assume they are larger.
 \end{claim}
@@
 \begin{proof}
-If, say, $b_2 \le 0 \le t_2$.
-Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$).
-This implies the width w.r.t to $f_q$, the functional constant on $q_1q_2$, of any segment with vertices in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
+If, say, $b$ has positive slope and $t$ negative.
+Then since $t$ intersects the vertical line $x=0$ spanned by $p$ below $p_2$, and $b$ intersects it above $p_1$, the slopes of $b$ and $t$ imply that  $Q \cap \{x\ge 0\}$ is contained in the strip $-1<f_q(x)<i-1$, of width $i$.
+This cannot contain a translated copy of $r$, since $\width_{f_q}(r) >i$ .
 \end{proof}
@@
 \begin{claim}
+\label{positive slopes}
 Assume w.l.o.g.~that $b$ has greater slope than $t$. Then, 
 \begin{enumerate}
-\item The intersection of $Q$ with any vertical line to the right of $q_2$ has length strictly smaller than one.
-\item $r_1>0$.
+\item The intersection of $Q$ with any vertical line to the right of $p$ has width w.r.t. $f_q$ strictly smaller than $i$.
+\item $r$ also has positive slope.
 \end{enumerate}
@@
 \begin{proof}
-The assumption implies that the intersections of $t$ and $b$ with the vertical line $\{x=1-\beta\}$ through $q_2$ are less than distance one apart (consider a vertical lattice line intersecting $q_1q_2$ and two consecutive lattice points in it, one above and one below the horizontal axis).
-This in turn implies part (a).
+\giulia{translate this into the width language!}
+Since $t$ and $b$ both intersect $p$, these intersections form a vertical segment which has width w.r.t $f_q$ less than $i$, the width of $p$.
+The assumptions on the slopes implies that this is true for any vertical segment with endpoints in $b$ and $t$ to the right of this one, which is part (a).
 
-For (b), if $r_1$ was $\le 0$ it would be impossible to fit a translated copy of $r$ in the right side of $Q$, because of (a).
+For (b), if $r$ had negative slope, it would be impossible to fit a translated copy of $r$ in the right side of $Q$, because of (a): it would lie \giulia{finish, with picture}
 \end{proof}
@@
 
-Starting here we assume index 2.
-
-\begin{claim}
-The slopes of $t$ and $b$ are strictly smaller than one and that of $r$ strictly greater than one.
-\end{claim}
-
-\begin{proof}
-Otherwise the points $(0,1/2)$, $(0,-1/2)$, $(1/2,0)$, respectively, would be in the corresponding band.
-\end{proof}
-
-We are now ready to show a contradiction.
+%Starting here we assume index 2.
+%
+%\begin{claim}
+%The slopes of $t$ and $b$ are strictly smaller than one and that of $r$ strictly greater than one.
+%\end{claim}
+%
+%\begin{proof}
+%Otherwise the points $(0,1/2)$, $(0,-1/2)$, $(1/2,0)$, respectively, would be in the corresponding band.
+%\end{proof}
+%
+%We are now ready to show a contradiction.
 Let $u$ and $v$ be the end-points of the translated copy of $r$ in $Q$, with $\vec r=v-u$.
@@
 
+\begin{claim}
+The slopes of $t$ and $b$ are strictly smaller than the slope of $\conv(p_1, q_2)$ and that of $r$ (and $l$) is strictly greater than it.
+\end{claim}
+
+\begin{proof}
+Since $b$ and $t$ must respectively separate $p_1$ and $q_1+q_2-p_1$ from the remaining vertices of the parallelogram $\conv(q_1, p_1, q_2, q_1+q_2-p_1)$, they must respectively intersect its (parallel) edges $\conv(p_1, q_2)$ and $\conv(q_1, q_1+q_2-p_1)$.
+This means their slopes must be strictly less than that of these segments.
+The same argument can be applied to $l$ and $r$ if one considers the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-p_2)$, yielding that the slopes of $l$ and $r$ are strictly larger than that of $\conv(p_1, q_2)$.
+\end{proof}
+
+We are now ready to show a contradiction.
+Since $Q$ has normal fan refining that of $P$, it must have an edge which is a translated copy of $r$.
+Let $u$ and $v$ be its endpoints.
+Now consider the lattice line $d$ through $u$ parallel to  $\conv(p_1, q_2)$.
+Let $d'$ be the next parallel lattice line above $d$.
+The width w.r.t $f_q$ of the vertical segment starting at $u$ with other endpoint $v'$ on $d'$ is $i$, since this vertical segment is a translation of $p$.
+Thus $v'$ must lie outside $Q$, by claim \ref{positive slopes}.
+That is, $v'$ lies above the line spanned by $t$.
+Since the slope of $t$ is less than the slope of $d$, the same happens for every point of $d'$ to the right of $u$.
+In particular, $v$ must be below $d'$ and hence it must be on or below $d$.
+This contradicts the fact that the slope of $r$ is strictly greater than the slope of $d$.
+
 \end{document}

### 2019-07-04 unimodular-covers.tex

@@
 \end{corollary}
+
+These results leave open the following important questions:
 

### 2019-07-04 unimodular-covers.tex
made proof of Lemma a new section. Put old stuff in the appendix

@@
 \maketitle
+
+\tableofcontents
 
@@
 
-Our proof of \Cref{thm:cayley} is based on the following result:
-
-\begin{figure}[htb]
-  \centering
-\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  .7]
-%coordinate axes
-   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%        \tkzLabelY[orig=false,label options={font=\tiny}]
-        \tkzGrid
-%       \tkzDrawX
-%       \tkzDrawY
-
-%hexagon
-  \node[mynode] (A) at (0,-1) {}; 
-%  \node[mynode] (B) at (1,1) {};  
-  \node[mynode] (C) at (2,3) {};  
-  \node[mynode] (D) at (0,1) {};  
-%  \node[mynode] (E) at (-1,-1) {};  
-  \node[mynode] (F) at (-2,-3) {};
-
-%other points
-%  \node[mynode] () at (0,-1) {};
-
-  \draw[black] (A) -- (C) -- (D) -- (F) -- (A);
-  \draw[black, thick] (A) -- (D);
-\end{tikzpicture}
-%
-\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  .7]
-%coordinate axes
-   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%        \tkzLabelY[orig=false,label options={font=\tiny}]
-        \tkzGrid
-%        \tkzDrawX
-%        \tkzDrawY
-
-%hexagon
-  \node[mynode] (A) at (1,0) {}; 
-  \node[mynode] (B) at (2,1) {};  
-  \node[mynode] (C) at (3,3) {};  
-  \node[mynode] (D) at (-1,0) {};  
-  \node[mynode] (E) at (-2,-1) {};  
-  \node[mynode] (F) at (-3,-3) {};
-
-  \node[mynode] (p1) at (-1,0) {}; 
-  \node[mynode] (p2) at (1,0) {};
-
-  \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
-  \draw[black, thick] (p1) -- (p2);
-\end{tikzpicture}
-
-\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
-
-\caption{Partial counter-example to \Cref{lemma:cayley}.
-With respect to the standard lattice it is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one).
-%
-With respect to the lattice of index two it is not a counter-example either, because two vertices of $Q$ are not in that lattice.
-}
-\label{fig:width3}
-\end{figure}
+Our proof of \Cref{thm:cayley} is based on the following result, the proof of which we postpone to \Cref{sec:the_lemma}:
 
@@
 \giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in the interior of $\{p_1, p_2\} + \{q_1, q_2\}$ (or symmetric for $Q$)}
-
-\begin{proof}
-%Suppose by contradiction that there is no lattice point as described in the lemma.
-Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
-The same holds for $Q$ and its corresponding strip.
-Thus $P$ has two edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $e_1$ and $e_2$.
-The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.
-
-Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
-\[
-p_1=(0,-\alpha), \quad
-p_2=(0,1-\alpha), \quad
-q_1=(-\beta,0), \quad
-q_2=(1-\beta,0).
-\] 
-The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
-
-Suppose by contradiction that there is no lattice point as described in the lemma.
-Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
-The same holds for $Q$ and its corresponding strip.
-Thus $P$ has two primitive edges which each have one vertex on each side of the strip $p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
-The same holds for $Q$ and the strip $q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle$, and we call the edges  $b, t$ (with $b$ below and $t$ above).
-To make things more explicit, we assume the following coordinates for the edge-vectors:
-\[
-\vec l = (l_1,l_2), \quad
-\vec r= (r_1,r_2), \quad
-\vec b= (b_1,b_2), \quad
-\vec t= (t_1,t_2).
-\]
-Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
-
-Since the normal fan of $Q$ refines that of $P$, $Q$ will contain edges $e_1'$ and $e_2'$ parallel to $e_1$ and $e_2$ respectively.
-These edges of $Q$ are all distinct and appear alternating in $Q$, that is, they appear in the order $f_1, e_1', f_2, e_2'$.
-Indeed, consider the quadrilateral $S$ with vertices $q_1, p_1+l, q_2, p_2+l$, where $l$ is a lattice point chosen so that the segments $q_1q_2$ and $p_1+l,p_2+l$ intersect in their interior; we can further require that $\conv( q_1, p_1+l, q_2)$ is unimodular \giulia{not sure if we use it}.
-$S$ lies inside the $Q$-strip, while its translate $S-l$ lies inside the $P$-strip.
-Thus in order to have no lattice points in the strips, we must have that the edge $f_i$ separates the vertex $p_i +l$ of $S$ from the remaining three, and  the same for the edge $e_i +l$ and vertex $q_i$ of $S$.
-This is equivalent to saying that the normals to the edges $f_1, e_1, f_2, e_2$ lie in different open cones of the normal fan of $S$, in this cyclic order.
-
-- both $f_1$ or $f_2$ must meet the line through $q_1$ and $q_2$, else they would not cut out the lattice point they need to. same goes for $e_1$ and $e_2$ and the line through $p_1$ and $p_2$.
-
-claim: the segments $e_1 + l$ and $e_2 + l$ respectively intersect both $f_1$ and $f_2$.
-
-claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.
-
-%Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensibly}.
-Since these segments do not contain lattice points, they must have lattice length stricly less than $1$ (or the width wrt $q$ is less than $k$, the index of $p,q$), which means that the smallest width in direction of $p_1 p_2$ between a "consecutive" vertices of $f_1$ and of $f_2$ must also be strictly less than $k$.
-
-%Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
-\end{proof}
 
@@
 \[
-\conv(Q_1\times\{0\} \cup Q_2 \times \{k\},
+\conv(Q_1\times\{0\} \cup Q_2 \times \{k\}),
 \]
@@
 
-both smooth prismatoids and dilations of lattice polytopse of width 1
-
-is a particular case of the following situation.
-
-\begin{thebibliography}{99}
-
-\bibitem{BHHHJKM2019}
-Matthias Beck, Christian Haase, Akihiro Higashitani, Johannes Hofscheier, Katharina Jochemko, Lukas Katth\"an, Mateusz Micha{\l}ek,
-Smooth Centrally Symmetric Polytopes in Dimension 3 are IDP,
-\emph{Ann.~Combin.}, in press, 2019.
-\url{https://doi.org/10.1007/s00026-019-00418-x}
-
-\bibitem{BGbook}
-Winfried Bruns and Joseph Gubeladze, \emph{Polytopes, rings, and k-theory}, Monographs in Mathematics,
-  Springer-Verlag, 2009, XIV, 461 p. 52 illus.
-
- \bibitem{DLRS2010}
-J. A. De Loera, J. Rambau, F. Santos
-Triangulations: Structures for Algorithms and Applications, 539 pp.
-Algorithms and Computation in Mathematics, Vol. 25, Springer-Verlag. 
-ISBN: 978-3-642-12970-4
-
-\bibitem{Fakhruddin}
-Najmuddin Fakhruddin.
-\newblock Multiplication maps of linear systems on smooth projective toric
-  surfaces, arXiv:0208.5178, August 2002.
-
-  \bibitem{HPPS-survey}
-Christian Haase, Andreas Paffenholz, Lindsay C. Piechnik, Francisco Santos. Existence of unimodular triangulations - positive results. 
-%Preprint May 2014, updated December 2017, 89 pages. 
- \emph{Mem. Amer. Math. Soc.}, to appear.
- Available as arXiv preprint \href{https://arxiv.org/abs/1405.1687}{arXiv:1405.16878}
-
-\bibitem{HNPS2008}
-Christian Haase, Benjamin Nill, Andreas Paffenholz, and Francisco Santos, Lattice points in Minkowski sums, 
-Electron J. Combin., 15 (2008), no. 1, Note 11, 5 pp.
-
-\bibitem{mfo2004}
-Mini-workshop: Ehrhart Quasipolynomials: Algebra, Combinatorics, and Geometry, Oberwolfach Rep.
-1 (2004), no.
-3, 2071-2101, Abstracts from the mini-workshop held August 15-21, 2004, Organized by Jes\'us A.~De Loera and Christian Haase, Oberwolfach Reports.
-Vol.
-1, no.
-3.
-MR MR2144157
-
-\bibitem{mfo2007}
-Mini-workshop: Projective normality of smooth toric varieties, Oberwolfach Rep.
-4 (2007), no.
-39/2007, Abstracts from the mini-workshop held August 12-18, 2007.
-Organized by Christian Haase, Takayuki Hibi and Diane MacLagan.
-
-\bibitem{Oda1997}
-Tadao Oda. Problems on Minkowski sums of convex lattice polytopes. 
-Abstract submitted at the Oberwolfach Conference ''Combinatorial Convexity and Algebraic Geometry'' 26.10--01.11, 1997.
-Available as arXiv preprint \href{https://arxiv.org/abs/0812.1418}{arXiv:0812.1418}, 2008.
-
-\bibitem{Ogata}
-Shoetsu Ogata.
-\newblock {Multiplication maps of complete linear systems on projective toric
-  surfaces.}
-\newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006.
-
-\bibitem{White1964}
-G.~K.~White.
-Lattice tetrahedra.
-\emph{Canadian J.~Math.} 16 (1964), 389--396.
-
-\end{thebibliography}
-
-\appendix
-\section{Attempt at the proof of \Cref{lemma:cayley}}
+\section{Proof of \Cref{lemma:cayley}}
+\label{sec:the_lemma}
 
@@
 
+\begin{thebibliography}{99}
+
+\bibitem{BHHHJKM2019}
+Matthias Beck, Christian Haase, Akihiro Higashitani, Johannes Hofscheier, Katharina Jochemko, Lukas Katth\"an, Mateusz Micha{\l}ek,
+Smooth Centrally Symmetric Polytopes in Dimension 3 are IDP,
+\emph{Ann.~Combin.}, in press, 2019.
+\url{https://doi.org/10.1007/s00026-019-00418-x}
+
+\bibitem{BGbook}
+Winfried Bruns and Joseph Gubeladze, \emph{Polytopes, rings, and k-theory}, Monographs in Mathematics,
+  Springer-Verlag, 2009, XIV, 461 p. 52 illus.
+
+ \bibitem{DLRS2010}
+J. A. De Loera, J. Rambau, F. Santos
+Triangulations: Structures for Algorithms and Applications, 539 pp.
+Algorithms and Computation in Mathematics, Vol. 25, Springer-Verlag. 
+ISBN: 978-3-642-12970-4
+
+\bibitem{Fakhruddin}
+Najmuddin Fakhruddin.
+\newblock Multiplication maps of linear systems on smooth projective toric
+  surfaces, arXiv:0208.5178, August 2002.
+
+  \bibitem{HPPS-survey}
+Christian Haase, Andreas Paffenholz, Lindsay C. Piechnik, Francisco Santos. Existence of unimodular triangulations - positive results. 
+%Preprint May 2014, updated December 2017, 89 pages. 
+ \emph{Mem. Amer. Math. Soc.}, to appear.
+ Available as arXiv preprint \href{https://arxiv.org/abs/1405.1687}{arXiv:1405.16878}
+
+\bibitem{HNPS2008}
+Christian Haase, Benjamin Nill, Andreas Paffenholz, and Francisco Santos, Lattice points in Minkowski sums, 
+Electron J. Combin., 15 (2008), no. 1, Note 11, 5 pp.
+
+\bibitem{mfo2004}
+Mini-workshop: Ehrhart Quasipolynomials: Algebra, Combinatorics, and Geometry, Oberwolfach Rep.
+1 (2004), no.
+3, 2071-2101, Abstracts from the mini-workshop held August 15-21, 2004, Organized by Jes\'us A.~De Loera and Christian Haase, Oberwolfach Reports.
+Vol.
+1, no.
+3.
+MR MR2144157
+
+\bibitem{mfo2007}
+Mini-workshop: Projective normality of smooth toric varieties, Oberwolfach Rep.
+4 (2007), no.
+39/2007, Abstracts from the mini-workshop held August 12-18, 2007.
+Organized by Christian Haase, Takayuki Hibi and Diane MacLagan.
+
+\bibitem{Oda1997}
+Tadao Oda. Problems on Minkowski sums of convex lattice polytopes. 
+Abstract submitted at the Oberwolfach Conference ''Combinatorial Convexity and Algebraic Geometry'' 26.10--01.11, 1997.
+Available as arXiv preprint \href{https://arxiv.org/abs/0812.1418}{arXiv:0812.1418}, 2008.
+
+\bibitem{Ogata}
+Shoetsu Ogata.
+\newblock {Multiplication maps of complete linear systems on projective toric
+  surfaces.}
+\newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006.
+
+\bibitem{White1964}
+G.~K.~White.
+Lattice tetrahedra.
+\emph{Canadian J.~Math.} 16 (1964), 389--396.
+
+\end{thebibliography}
+
+\appendix
+
+\section{Old stuff}
+
+\begin{proof}
+%Suppose by contradiction that there is no lattice point as described in the lemma.
+Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
+The same holds for $Q$ and its corresponding strip.
+Thus $P$ has two edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $e_1$ and $e_2$.
+The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.
+
+Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
+\[
+p_1=(0,-\alpha), \quad
+p_2=(0,1-\alpha), \quad
+q_1=(-\beta,0), \quad
+q_2=(1-\beta,0).
+\] 
+The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
+
+Suppose by contradiction that there is no lattice point as described in the lemma.
+Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
+The same holds for $Q$ and its corresponding strip.
+Thus $P$ has two primitive edges which each have one vertex on each side of the strip $p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
+The same holds for $Q$ and the strip $q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle$, and we call the edges  $b, t$ (with $b$ below and $t$ above).
+To make things more explicit, we assume the following coordinates for the edge-vectors:
+\[
+\vec l = (l_1,l_2), \quad
+\vec r= (r_1,r_2), \quad
+\vec b= (b_1,b_2), \quad
+\vec t= (t_1,t_2).
+\]
+Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
+
+Since the normal fan of $Q$ refines that of $P$, $Q$ will contain edges $e_1'$ and $e_2'$ parallel to $e_1$ and $e_2$ respectively.
+These edges of $Q$ are all distinct and appear alternating in $Q$, that is, they appear in the order $f_1, e_1', f_2, e_2'$.
+Indeed, consider the quadrilateral $S$ with vertices $q_1, p_1+l, q_2, p_2+l$, where $l$ is a lattice point chosen so that the segments $q_1q_2$ and $p_1+l,p_2+l$ intersect in their interior; we can further require that $\conv( q_1, p_1+l, q_2)$ is unimodular \giulia{not sure if we use it}.
+$S$ lies inside the $Q$-strip, while its translate $S-l$ lies inside the $P$-strip.
+Thus in order to have no lattice points in the strips, we must have that the edge $f_i$ separates the vertex $p_i +l$ of $S$ from the remaining three, and  the same for the edge $e_i +l$ and vertex $q_i$ of $S$.
+This is equivalent to saying that the normals to the edges $f_1, e_1, f_2, e_2$ lie in different open cones of the normal fan of $S$, in this cyclic order.
+
+- both $f_1$ or $f_2$ must meet the line through $q_1$ and $q_2$, else they would not cut out the lattice point they need to. same goes for $e_1$ and $e_2$ and the line through $p_1$ and $p_2$.
+
+claim: the segments $e_1 + l$ and $e_2 + l$ respectively intersect both $f_1$ and $f_2$.
+
+claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.
+
+%Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensibly}.
+Since these segments do not contain lattice points, they must have lattice length stricly less than $1$ (or the width wrt $q$ is less than $k$, the index of $p,q$), which means that the smallest width in direction of $p_1 p_2$ between a "consecutive" vertices of $f_1$ and of $f_2$ must also be strictly less than $k$.
+
+%Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
+\end{proof}
+
+\begin{figure}[htb]
+  \centering
+\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+  .7]
+%coordinate axes
+   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%        \tkzLabelY[orig=false,label options={font=\tiny}]
+        \tkzGrid
+%       \tkzDrawX
+%       \tkzDrawY
+
+%hexagon
+  \node[mynode] (A) at (0,-1) {}; 
+%  \node[mynode] (B) at (1,1) {};  
+  \node[mynode] (C) at (2,3) {};  
+  \node[mynode] (D) at (0,1) {};  
+%  \node[mynode] (E) at (-1,-1) {};  
+  \node[mynode] (F) at (-2,-3) {};
+
+%other points
+%  \node[mynode] () at (0,-1) {};
+
+  \draw[black] (A) -- (C) -- (D) -- (F) -- (A);
+  \draw[black, thick] (A) -- (D);
+\end{tikzpicture}
+%
+\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+  .7]
+%coordinate axes
+   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%        \tkzLabelY[orig=false,label options={font=\tiny}]
+        \tkzGrid
+%        \tkzDrawX
+%        \tkzDrawY
+
+%hexagon
+  \node[mynode] (A) at (1,0) {}; 
+  \node[mynode] (B) at (2,1) {};  
+  \node[mynode] (C) at (3,3) {};  
+  \node[mynode] (D) at (-1,0) {};  
+  \node[mynode] (E) at (-2,-1) {};  
+  \node[mynode] (F) at (-3,-3) {};
+
+  \node[mynode] (p1) at (-1,0) {}; 
+  \node[mynode] (p2) at (1,0) {};
+
+  \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
+  \draw[black, thick] (p1) -- (p2);
+\end{tikzpicture}
+
+\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
+
+\caption{Partial counter-example to \Cref{lemma:cayley}.
+With respect to the standard lattice it is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one).
+%
+With respect to the lattice of index two it is not a counter-example either, because two vertices of $Q$ are not in that lattice.
+}
+\label{fig:width3}
+\end{figure}
+
 \end{document}

### 2019-07-05 unimodular-covers.tex
added sketches of pictures

@@
 \usepackage{tkz-euclide}
+\usetikzlibrary{intersections}
+\usetkzobj{all}
 
@@
 \begin{align*}
-\width_{f_q}(l) > i, \\
-\width_{f_q}(r) > i, \\
-\width_{f_p}(t) > i, \\
+\width_{f_q}(l) , 
+\width_{f_q}(r) ,
+\width_{f_p}(t) ,
 \width_{f_p}(b) > i.
@@
 \end{proof}
+\includegraphics[width=.7\textwidth]{Claim42.jpg}
 
@@
 \begin{proof}
-\giulia{translate this into the width language!}
+
 Since $t$ and $b$ both intersect $p$, these intersections form a vertical segment which has width w.r.t $f_q$ less than $i$, the width of $p$.
@@
 
-For (b), if $r$ had negative slope, it would be impossible to fit a translated copy of $r$ in the right side of $Q$, because of (a): it would lie \giulia{finish, with picture}
+For part (b), let $r = conv(u, v)$, with $\vec r=v-u$.
+If $r$ had negative slope, it would be impossible to fit a translated copy of $r$ in the right side of $Q$: $r$ would lie inside the region to the left of $u$ delimited by $t$, the horizontal line containing $u$ and the vertical line through $u$.
+We can observe that this triangular region has width w.r.t.
+$f_q$ less than $i$ by combining part (a), which implies that the vertical segment through $u$ between $t$ and $u$ has width w.r.t $f_q$ less than $i$, with the fact that $t$ has positive slope.
 \end{proof}
+\includegraphics[ width=.7\textwidth]{Claim43.jpg}
 
@@
 
+TRYING TO MODIFY this picture:
+
+\begin{figure}[htb]
+  \centering
+\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+  .7]
+%coordinate axes
+   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%        \tkzLabelY[orig=false,label options={font=\tiny}]
+%        \tkzGrid
+%        \tkzDrawX
+%        \tkzDrawY
+
+%edges b, t
+  \node[mynode] (b1) at (-3,-2) {}; 
+  \node[mynode] (b2) at (2, 0.2) {};  
+  \node[mynode] (t1) at (-2.7, 1) {};  
+  \node[mynode] (t2) at (1.6,0.7) {};  
+  %\node[mynode] (E) at (-2,-1) {};  
+  %\node[mynode] (F) at (-3,-3) {};
+
+%p
+  \node[mynode] (p1) at (0,-1) {}; 
+  \node[mynode] (p2) at (0,1) {};
+
+%q
+  \node[mynode] (q1) at (-1,0) {}; 
+  \node[mynode] (q2) at (1,0) {};
+
+ % \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
+  \draw[name path = b, black] (b1) -- (b2); 
+  \draw[name path = t, black] (t1) -- (t2);
+  \draw[name path = p, black, thick] (p1) -- (p2);
+  \draw[name path = q, black, thick] (q1) -- (q2);
+
+ \tkzDrawLine[add=0.2 and 0.3, color=black, dashed](b1, b2);
+  \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);
+  \path [name intersections={of = p and b}];
+  \coordinate (S)  at (intersection-1);
+
+\end{tikzpicture}
+
+%\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
+
+\end{figure}
+
 \end{document}

### 2019-07-05 unimodular-covers.tex
making tikz figures

@@
 \begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  .7]
+  1]
 %coordinate axes
@@
 %edges b, t
-  \node[mynode] (b1) at (-3,-2) {}; 
-  \node[mynode] (b2) at (2, 0.2) {};  
-  \node[mynode] (t1) at (-2.7, 1) {};  
-  \node[mynode] (t2) at (1.6,0.7) {};  
+  \node[
+%label={[shift={(0,-0.67)}]$b_1$}, 
+mynode] (b1) at (-3,-2) {}; 
+  \node[
+%label={[shift={(0,-0.67)}]$b_2$}, 
+mynode] (b2) at (2, 0.2) {};  
+  \node[
+%label={[shift={(0,0)}]$t_1$}, 
+mynode] (t1) at (-2.7, 1) {};  
+  \node[
+%label={[shift={(0,0)}]$t_2$}, 
+mynode] (t2) at (1.6,0.7) {};
+
+  \draw[fill=gray!80, nearly transparent] (2.93,0.61) -- (0,0.81)  -- (0,-0.68) -- cycle;
   %\node[mynode] (E) at (-2,-1) {};  
@@
 %p
-  \node[mynode] (p1) at (0,-1) {}; 
-  \node[mynode] (p2) at (0,1) {};
+  \node[
+%label={[shift={(0,-0.67)}]$p_1$}, 
+mynode] (p1) at (0,-1) {}; 
+  \node[
+%label={[shift={(0,0)}]$p_2$}, 
+mynode] (p2) at (0,1) {};  
+%  \node[{p}]  at (-0.3,0.95)
 
 %q
-  \node[mynode] (q1) at (-1,0) {}; 
-  \node[mynode] (q2) at (1,0) {};
+  \node[
+%label={[shift={(0,-0.67)}]$q_1$}, 
+mynode] (q1) at (-1.7,-0.4) {}; 
+  \node[
+%label={[shift={(0,-0.67)}]$q_2$}, 
+mynode] (q2) at (0.3,-0.4) {};
 
- % \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
-  \draw[name path = b, black] (b1) -- (b2); 
-  \draw[name path = t, black] (t1) -- (t2);
-  \draw[name path = p, black, thick] (p1) -- (p2);
-  \draw[name path = q, black, thick] (q1) -- (q2);
+  \draw[ name path = b, black, thick] (b1) node[below, xshift=1.7cm, yshift=0.7cm] {b} -- (b2); 
+  \draw[name path = t, black, thick] (t1) node[below, xshift=1.2cm, yshift=-0.1cm] {t} -- (t2);
+  \draw[name path = p, black, thick] (p1) node[left, xshift=0 cm, yshift=1.3cm] {p} -- (p2);
+  \draw[name path = q, black, thick] (q1) node[below, xshift=0.4cm, yshift=0cm] {q} -- (q2);
 
@@
  \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);
-  \path [name intersections={of = p and b}];
-  \coordinate (S)  at (intersection-1);
+
+  \tkzDefLine[orthogonal=through p1](p1,p2)
+  \tkzDrawLine[add = 2 and 1,color=black, dashed](p1,tkzPointResult)
+
+  \tkzDefLine[orthogonal=through p2](p1,p2)
+  \tkzDrawLine[add = 2 and 1,color=black, dashed](p2,tkzPointResult)
 

### 2019-07-05 unimodular-covers.tex
made figure for claim 4.2
merged with office file

@@
 \newcommand{\Vol}{\operatorname{Vol}}
+\newcommand{\area}{\operatorname{area}}
 \newcommand{\length}{\operatorname{length}}
+
+\renewcommand{\vec}[1]{\overrightarrow#1}
+\newcommand{\vecline}[1]{\langle \vec #1 \rangle}
 
@@
 
-Our proof of \Cref{thm:cayley} is based on the following result, the proof of which we postpone to \Cref{sec:the_lemma}:
+Our proof of \Cref{thm:cayley} is based on the following result, the proof of which we postpone to \Cref{sec:the_lemma}. 
+The lemma uses the following conventions.
+If $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$.
+For a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and $\vecline s$ as the line spanned by $\vec s$.
 
@@
 Let $P$ and $Q$ be two lattice polygons with $Q$ nef with respect to $P$.
-Let $p_1p_2 \subset P$ and $q_1q_2\subset Q$ be two primitive and non-parallel segments, and assume that the prod-simplex (parallelogram)  $p_1p_2 + q_1q_2$ is not unimodular.
-Then,
-at least one of the strips
-\paco{add picture for these strips, and perhaps for the proof}
+Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the lines spanned by them.
+If the parallelogram $p + q$ is not unimodular, then at least one of the regions
 \[
-(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle) \cap P
-\qquad \text{or} \qquad
-(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle) \cap Q
+((p_1, p_2) + \vecline q ) \cap P, 
+\qquad \text{and} \qquad
+((q_1, q_2) + \vecline p ) \cap Q
 \]
-has a lattice point in its interior.
+contains a lattice point.
+
+\paco{add picture for these strips}
+
+%\[
+%(\int(p) + \vecline q) \cap P
+%\qquad \text{or} \qquad
+%\int(q + \vecline p) \cap Q
+%\]
+%contains a lattice point.
 \end{lemma}
@@
 This leaves simplices of type $(2,2)$.
-In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can I be without being confusing?}.
+In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.
 
@@
 
-Let us fix coordinates in which $p=\conv(p_1,p_2)$ and $q= \conv(q_1,q_2)$ are a vertical and horizontal segment of length one intersecting at the origin.
-That is, there are numbers $\alpha, \beta \in (0,1)$ such that
-\[
-p_1=(0,-\alpha), \quad
-p_2=(0,1-\alpha), \quad
-q_1=(-\beta,0), \quad
-q_2=(1-\beta,0).
-\] \giulia{perhaps the explicit coordinates are superfluous now}
-The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
-We can further assume (since translating the polytopes by lattice vectors will not result in any loss of generality) that $p_1$ is on the consecutive lattice line parallel to the line $q$ lies on.
+By a change of coordinates and lattice, we can assume $p=[p_1,p_2]$ and $q= [q_1,q_2]$ to be a vertical and horizontal segment .
+Since translating the polytopes by lattice vectors will not result in any loss of generality, we can further assume that $p_1$ is on the next lattice line parallel to the line spanned by $q$.
 That is, if $f_q$ is the primitive lattice functional which is zero on $q$, then $f_q(p_1)=-1$.
 Equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
-Let $f_p$ be the functional constant on $p$.
-We observe then that
+
+Let $f_p$ be the primitive lattice functional constant on $p$. Let $w=\area_\Lambda(p+q) \geq 2$, where $\area_\Lambda$ denotes the area normalized to a fundamental domain of $\Lambda$. Then 
 \[
-\width_{f_q}(p + \overrightarrow{q})=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\overrightarrow{p}) = \ind(p, q) =:i.
+w=\width_{f_q}(p + \vecline q )=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\vecline p ).
 \]
 
-Observe that $p_1$ is the unique lattice point in the strip $q +\overrightarrow{p}$ on the line $f_q(x)=-1$.
-Indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p$ and $q$ have index $1$, which is a contradiction.
-The unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.
+Observe that $p_1$ is the unique lattice point in the strip $q +\vecline p$ on the line $f_q(x)=-1$.
+Indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, which is a contradiction.
+Similarly, the unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.
 
 Suppose by contradiction that there is no lattice point as described in the lemma.
-Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
+In particular, no vertex of $P$ can be in the strip $p + \vecline q$.
 The same holds for $Q$ and its corresponding strip.
-Thus $P$ has two primitive edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
-The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges  $b, t$ (with $b$ below and $t$ above).
-Observe that one of $l$ and $r$ can coincide with $p$, if this is an edge of $P$, and similarly one of $t,b$ might be $q$, if this is an edge of $Q$.
-To make things more explicit, we assume the following coordinates for the edge-vectors:
-\[
-\vec l = (l_1,l_2), \quad
-\vec r= (r_1,r_2), \quad
-\vec b= (b_1,b_2), \quad
-\vec t= (t_1,t_2).
-\]
-Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
+Thus $P$ has two primitive segments in the boundary which each have one vertex on each side of the strip $p + \vecline q$; we will call these  $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
+The same holds for $Q$ and the strip $q + \vecline p$, and we call the segments  $b, t$ (with $b$ below and $t$ above).
+Observe that one of $l$ and $r$ can coincide with $p$, if this is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if this is on the boundary of $Q$.
+\giulia{p=P ???}
+
+%To make things more explicit, we assume the following coordinates for the edge-vectors:
+%\[
+%\vec l = (l_1,l_2), \quad
+%\vec r= (r_1,r_2), \quad
+%\vec b= (b_1,b_2), \quad
+%\vec t= (t_1,t_2).
+%\]
+%Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
 
@@
 \begin{claim}\giulia{I think we need to treat separately the case where we deal with edges (these are easier, in any case, but I couldn't find a unification)}
-The following inequalities hold, provided the segment in question is not an edge of the corresponding polytope:
+The following inequalities hold, 
 \begin{align*}
@@
 \width_{f_p}(t) ,
-\width_{f_p}(b) > i.
+\width_{f_p}(b) \geq w.
 \end{align*}
+Each inequality is strict, unless the segment in question coincides with $p$ or $q$.
 \end{claim}
@@
 \begin{proof}
-The inequality $\geq i$ follows in each case from the fact that each segment crosses the corresponding strip, the width of which (w.r.t the appropriate functional) is $i$, as observed above. 
-If one of the inequalities were not strict, then the corresponding segment, say $l$, must have vertices on the corresponding strip.
-Then the strip $(p + \langle \overrightarrow{q}\rangle) \cap P$ contains a triangle with one edge a translation of $p$ and another edge a translation of $q$.
-Such a triangle must contain a lattice point other than its vertices, and since $p_1p_2$ and $q_1q_2$ are primitive, the lattice point must be in the interior of the strip.
+The inequality $\geq w$ follows in each case from the fact that each segment crosses the corresponding strip, the width of which (w.r.t the appropriate functional) is $w$, as observed above. 
+If one of the inequalities, say the one for $\ell$, is not strict, then $\ell$ has one endpoint on each of the boundary lines of $(p + \vecline q)$.
+Unless $\ell = p$, one of the endpoints of $\ell = [l_1, l_2]$ is not an endpoint of $p$, say $l_1 \neq p_1$.
+Thus the triangle $T=\conv(p_2, p_1, l_1)$ is contained in $P$ and its edge $[p_1, l_1]$ is an integer dilation of $q$.
+Since $\width_{f_q}(T) =w \geq 2$, $T$ must contain a lattice point in the interior of the strip.
 \end{proof}
@@
 %\end{proof}
+Let $H^-$ and $H^+$ be the two halfplanes defined by the affine span of $p$, with $H^+$ to the right.
+\begin{claim}
+$f_q$ achieves its maximum over $b$ and over $t$ on the same side ($H^+$ or $H^-$). 
+\end{claim}\giulia{do we need this claim?}
 
-\begin{claim}
-The slopes of $b$ and $t$ must either be both strictly larger than the slope of $q$ or both be strictly smaller. W.l.o.g.~we assume they are larger.
-\end{claim}
+%\begin{claim}
+%The slopes of $b$ and $t$ must either be both strictly larger than the slope of $q$ or both be strictly smaller. W.l.o.g.~we assume they are larger.
+%\end{claim}
+%
+%\begin{proof}
+%If, say, $b$ has positive slope and $t$ negative. Then since $t$ intersects the vertical line $x=0$ spanned by $p$ below $p_2$, and $b$ intersects it above $p_1$, the slopes of $b$ and $t$ imply that
 
 \begin{proof}
-If, say, $b$ has positive slope and $t$ negative.
-Then since $t$ intersects the vertical line $x=0$ spanned by $p$ below $p_2$, and $b$ intersects it above $p_1$, the slopes of $b$ and $t$ imply that  $Q \cap \{x\ge 0\}$ is contained in the strip $-1<f_q(x)<i-1$, of width $i$.
-This cannot contain a translated copy of $r$, since $\width_{f_q}(r) >i$ .
+ Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $H^-$ and that on $b$ lies in $H^+$.
+
+Then $Q \cap H^ +$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$ .
 \end{proof}
-\includegraphics[width=.7\textwidth]{Claim42.jpg}
+
+\begin{figure}[htb]
+  \centering
+\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+  1]
+%coordinate axes
+   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%        \tkzLabelY[orig=false,label options={font=\tiny}]
+%        \tkzGrid
+%        \tkzDrawX
+%        \tkzDrawY
+
+%edges b, t
+  \node[
+%label={[shift={(0,-0.67)}]$b_1$}, 
+mynode] (b1) at (-3,-2) {}; 
+  \node[
+%label={[shift={(0,-0.67)}]$b_2$}, 
+mynode] (b2) at (2, 0.2) {};  
+  \node[
+%label={[shift={(0,0)}]$t_1$}, 
+mynode] (t1) at (-2.7, 1) {};  
+  \node[
+%label={[shift={(0,0)}]$t_2$}, 
+mynode] (t2) at (1.6,0.7) {};
+
+%shaded region where r can be  
+  \draw[fill=gray!80, nearly transparent] (2.93,0.61) -- (0,0.81)  -- (0,-0.68) -- cycle;
+
+%p
+  \node[
+%label={[shift={(0,-0.67)}]$p_1$}, 
+mynode] (p1) at (0,-1) {}; 
+  \node[
+%label={[shift={(0,0)}]$p_2$}, 
+mynode] (p2) at (0,1) {};  
+%  \node[{p}]  at (-0.3,0.95)
+
+%q
+  \node[
+%label={[shift={(0,-0.67)}]$q_1$}, 
+mynode] (q1) at (-1.7,-0.4) {}; 
+  \node[
+%label={[shift={(0,-0.67)}]$q_2$}, 
+mynode] (q2) at (0.3,-0.4) {};
+
+%segments b, t, p, q
+  \draw[ name path = b, black, thick] (b1) node[below, xshift=1.7cm, yshift=0.7cm] {b} -- (b2); 
+  \draw[name path = t, black, thick] (t1) node[below, xshift=1.2cm, yshift=-0.1cm] {t} -- (t2);
+  \draw[name path = p, black, thick] (p1) node[left, xshift=0 cm, yshift=1.3cm] {p} -- (p2);
+  \draw[name path = q, black, thick] (q1) node[below, xshift=0.4cm, yshift=0cm] {q} -- (q2);
+
+%prolonged lines through b and t
+ \tkzDrawLine[add=0.2 and 0.3, color=black, dashed](b1, b2);
+ \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);
+
+%functional f_q
+  \tkzDefLine[orthogonal=through p1](p1,p2)
+  \tkzDrawLine[add = 2 and 1,color=black, dashed](p1,tkzPointResult)
+  \node[label={\small $\{f_q(x)=-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab1) at (3.3, -1) {};
+
+  \tkzDefLine[orthogonal=through p2](p1,p2) 
+  \tkzDrawLine[add = 2 and 1,color=black, dashed](p2,tkzPointResult) 
+    \node[label={\small $\{f_q(x)=w-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab2) at (3.3, 1) {};
+
+\end{tikzpicture}
+\caption{Picture for Claim 4.2}
+\end{figure}
+
+We assume w.l.o.g. that the maximum on both $t$ and $b$ is achieved in $H^+$ (do we need it??).
 
@@
 \label{positive slopes}
-Assume w.l.o.g.~that $b$ has greater slope than $t$. Then, 
+Assume w.l.o.g.~that $b$ and $t$ either are parallel or their affine spans cross in $H^+$. Then, 
 \begin{enumerate}
-\item The intersection of $Q$ with any vertical line to the right of $p$ has width w.r.t. $f_q$ strictly smaller than $i$.
-\item $r$ also has positive slope.
+\item The intersection of $Q$ with any line parallel to $p$ in $H^+$ has width w.r.t. $f_q$ strictly smaller than $w$.
+\item $f_p$ and $f_q$ increase in the same direction along $r$.
 \end{enumerate}
@@
 
+Let $g$ be the functional constant on $[p_1, q_2]$. Then
 \begin{claim}
-The slopes of $t$ and $b$ are strictly smaller than the slope of $\conv(p_1, q_2)$ and that of $r$ (and $l$) is strictly greater than it.
-\end{claim}
+The slopes of $t=[t_1, t_2]$ and $b=[b_1, b_2]$ are strictly smaller than the slope of  and that of $r=[r_1, r_2]$ (and $l=[l_1, l_2]$) is strictly greater than it.
+\end{claim}\giulia{let's put some more notation in the beginning: add H+, H-, V+, V-, and say which endpoints of l, r, etc are in which halfspace.}
 
@@
 
-TRYING TO MODIFY this picture:
-
-\begin{figure}[htb]
-  \centering
-\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  1]
-%coordinate axes
-   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%        \tkzLabelY[orig=false,label options={font=\tiny}]
-%        \tkzGrid
-%        \tkzDrawX
-%        \tkzDrawY
-
-%edges b, t
-  \node[
-%label={[shift={(0,-0.67)}]$b_1$}, 
-mynode] (b1) at (-3,-2) {}; 
-  \node[
-%label={[shift={(0,-0.67)}]$b_2$}, 
-mynode] (b2) at (2, 0.2) {};  
-  \node[
-%label={[shift={(0,0)}]$t_1$}, 
-mynode] (t1) at (-2.7, 1) {};  
-  \node[
-%label={[shift={(0,0)}]$t_2$}, 
-mynode] (t2) at (1.6,0.7) {};
-
-  \draw[fill=gray!80, nearly transparent] (2.93,0.61) -- (0,0.81)  -- (0,-0.68) -- cycle;
-  %\node[mynode] (E) at (-2,-1) {};  
-  %\node[mynode] (F) at (-3,-3) {};
-
-%p
-  \node[
-%label={[shift={(0,-0.67)}]$p_1$}, 
-mynode] (p1) at (0,-1) {}; 
-  \node[
-%label={[shift={(0,0)}]$p_2$}, 
-mynode] (p2) at (0,1) {};  
-%  \node[{p}]  at (-0.3,0.95)
-
-%q
-  \node[
-%label={[shift={(0,-0.67)}]$q_1$}, 
-mynode] (q1) at (-1.7,-0.4) {}; 
-  \node[
-%label={[shift={(0,-0.67)}]$q_2$}, 
-mynode] (q2) at (0.3,-0.4) {};
-
-  \draw[ name path = b, black, thick] (b1) node[below, xshift=1.7cm, yshift=0.7cm] {b} -- (b2); 
-  \draw[name path = t, black, thick] (t1) node[below, xshift=1.2cm, yshift=-0.1cm] {t} -- (t2);
-  \draw[name path = p, black, thick] (p1) node[left, xshift=0 cm, yshift=1.3cm] {p} -- (p2);
-  \draw[name path = q, black, thick] (q1) node[below, xshift=0.4cm, yshift=0cm] {q} -- (q2);
-
- \tkzDrawLine[add=0.2 and 0.3, color=black, dashed](b1, b2);
- \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);
-
-  \tkzDefLine[orthogonal=through p1](p1,p2)
-  \tkzDrawLine[add = 2 and 1,color=black, dashed](p1,tkzPointResult)
-
-  \tkzDefLine[orthogonal=through p2](p1,p2)
-  \tkzDrawLine[add = 2 and 1,color=black, dashed](p2,tkzPointResult)
-
-\end{tikzpicture}
-
-%\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
-
-\end{figure}
-
 \end{document}

### 2019-07-06 unimodular-covers.tex
reworking section 4

@@
 \label{sec:the_lemma}
+Let $f_q$ be the primitive lattice functional constant on $q$ and $f_p$ the one constant on $p$. We assume that $f_q(p_1) < f_q(p_2)$ and $f_p(q_1) < f_p(q_2)$.
 
-%At some point I'll assume volume two for the parallelogram $p_1p_2 + q_1q_2$, but for the time being we don't.
+Observe that in the strip strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=-1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
+Since translating the polytopes by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
+That is, $f_q(p_1)=-1$, or equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
+Similarly, the unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.
 
-By a change of coordinates and lattice, we can assume $p=[p_1,p_2]$ and $q= [q_1,q_2]$ to be a vertical and horizontal segment .
-Since translating the polytopes by lattice vectors will not result in any loss of generality, we can further assume that $p_1$ is on the next lattice line parallel to the line spanned by $q$.
-That is, if $f_q$ is the primitive lattice functional which is zero on $q$, then $f_q(p_1)=-1$.
-Equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
+We let $H_1=\{f_q(x) < 0\}$ and $H_2=\{f_q(x) > 0\}$; similarly let $V_1=\{f_p(x) < 0\}$ and $V_2=\{f_p(x) > 0\}$. 
+In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants.
 
-Let $f_p$ be the primitive lattice functional constant on $p$. Let $w=\area_\Lambda(p+q) \geq 2$, where $\area_\Lambda$ denotes the area normalized to a fundamental domain of $\Lambda$. Then 
+Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain. Then 
 \[
@@
 
-Observe that $p_1$ is the unique lattice point in the strip $q +\vecline p$ on the line $f_q(x)=-1$.
-Indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, which is a contradiction.
-Similarly, the unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.
-
 Suppose by contradiction that there is no lattice point as described in the lemma.
 In particular, no vertex of $P$ can be in the strip $p + \vecline q$.
-The same holds for $Q$ and its corresponding strip.
-Thus $P$ has two primitive segments in the boundary which each have one vertex on each side of the strip $p + \vecline q$; we will call these  $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
-The same holds for $Q$ and the strip $q + \vecline p$, and we call the segments  $b, t$ (with $b$ below and $t$ above).
+Thus $P$ has two primitive segments in the boundary which each have one vertex on each side of the strip $p + \vecline q$; we will call these $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively and the convention that $f_q(l_2) >f_q(l_1)$ and $f_q(r_2) >f_q(r_1)$.
+This readily implies
+\begin{gather}
+\label{eq:widthp}
+\begin{array}{cc}
+f_q(l_2) \geq f_q(p_2), &
+f_q(l_1) \leq f_q(p_1), \\
+f_q(r_2) \geq f_q(p_2), &
+f_q(r_1) \leq f_q(p_1).
+\end{array}
+\end{gather}
+
+The same holds for $Q$ and the strip $q+\vecline p$, and we call the segments  $b=[b_1, b_2], t=[t_1, t_2]$ with $b$ and $t$ crossing the strip in $H_1$ and $H_2$ respectively. Again we have
+\begin{gather}
+\label{eq:widthq}
+\begin{array}{cc}
+f_p(t_2) \geq f_p(q_2), &
+f_p(t_1) \leq f_p(q_1) \\
+f_p(b_2) \geq f_p(q_2), &
+f_p(b_1) \leq f_p(q_1).
+\end{array}
+\end{gather}
+
 Observe that one of $l$ and $r$ can coincide with $p$, if this is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if this is on the boundary of $Q$. 
@@
 \begin{align*}
-\width_{f_q}(l) , 
+\width_{f_q}(\ell) , 
 \width_{f_q}(r) ,
@@
 \begin{proof}
-The inequality $\geq w$ follows in each case from the fact that each segment crosses the corresponding strip, the width of which (w.r.t the appropriate functional) is $w$, as observed above. 
+The inequality $\geq w$ follows in each case from \eqref{eq:widthp} and \eqref{eq:widthq}.
+
 If one of the inequalities, say the one for $\ell$, is not strict, then $\ell$ has one endpoint on each of the boundary lines of $(p + \vecline q)$.
-Unless $\ell = p$, one of the endpoints of $\ell = [l_1, l_2]$ is not an endpoint of $p$, say $l_1 \neq p_1$.
+Unless $\ell = p$, one of the endpoints of $\ell$ is not an endpoint of $p$, say $l_1 \neq p_1$.
 Thus the triangle $T=\conv(p_2, p_1, l_1)$ is contained in $P$ and its edge $[p_1, l_1]$ is an integer dilation of $q$.
@@
 %\end{proof}
-Let $H^-$ and $H^+$ be the two halfplanes defined by the affine span of $p$, with $H^+$ to the right.
+
 \begin{claim}
-$f_q$ achieves its maximum over $b$ and over $t$ on the same side ($H^+$ or $H^-$). 
+$f_q(b_2-b_1)$ and $f_q(p_2 - p_1)$ are non-zero and have the same sign. That is, $f_q$ achieves its maximum over $b$ and over $t$ on the same halfplane $V_1$ or $V_2$.
 \end{claim}\giulia{do we need this claim?}
@@
 \begin{proof}
- Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $H^-$ and that on $b$ lies in $H^+$.
+Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $V_1$ and that on $b$ lies in $V_2$.
 
-Then $Q \cap H^ +$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$ .
+Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2} .
 \end{proof}
@@
 \caption{Picture for Claim 4.2}
+\label{fig:claim2}
 \end{figure}

### 2019-07-06 unimodular-covers.tex
finished section 4

@@
 
+\Giulia{fix it to look more unimodular at the bottom}
 \begin{figure}[htb]
@@
 
-We assume w.l.o.g. that the maximum on both $t$ and $b$ is achieved in $H^+$ (do we need it??).
+We assume w.l.o.g.
+that the maximum on $t$ (and hence on $b$) is achieved in $V_2$ \giulia{do we need it??}, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).
 
@@
 \label{positive slopes}
-Assume w.l.o.g.~that $b$ and $t$ either are parallel or their affine spans cross in $H^+$. Then, 
+Assume w.l.o.g.~that $b$ and $t$ either are parallel or their affine spans cross in $V_2$. Then, 
 \begin{enumerate}
-\item The intersection of $Q$ with any line parallel to $p$ in $H^+$ has width w.r.t. $f_q$ strictly smaller than $w$.
-\item $f_p$ and $f_q$ increase in the same direction along $r$.
+\item The intersection of $Q$ with any line parallel to $p$ in $V_2$ has width w.r.t. $f_q$ strictly smaller than $w$.
+\item $f_p(r_2) > f_p(r_1)$, that is, $f_p$ achieves its maximum over $r$ in $H_2$.
 \end{enumerate}
@@
 
-Since $t$ and $b$ both intersect $p$, these intersections form a vertical segment which has width w.r.t $f_q$ less than $i$, the width of $p$.
-The assumptions on the slopes implies that this is true for any vertical segment with endpoints in $b$ and $t$ to the right of this one, which is part (a).
+Both $t$ and $b$ must intersect $p$, otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.
+Their intersections with $p$ are thus endpoints of a segment of width w.r.t $f_q$ less than $w$, the width of $p$.
+Since $t$ and $b$ cross in $V_2$, the same is true for any segment parallel to $p$ contained in $Q \cap V_2$.
 
-For part (b), let $r = conv(u, v)$, with $\vec r=v-u$.
-If $r$ had negative slope, it would be impossible to fit a translated copy of $r$ in the right side of $Q$: $r$ would lie inside the region to the left of $u$ delimited by $t$, the horizontal line containing $u$ and the vertical line through $u$.
-We can observe that this triangular region has width w.r.t.
-$f_q$ less than $i$ by combining part (a), which implies that the vertical segment through $u$ between $t$ and $u$ has width w.r.t $f_q$ less than $i$, with the fact that $t$ has positive slope.
+For part (b), If $f_p(r_2) \leq f_p(r_1)$, it would be impossible to fit a translated copy of $r$ in the correct side of $Q$: $r$ would need to lie inside the triangle delimited by the affine line $\langle t \rangle$ and the inequalities $f_q(x) \geq f_q(r_1)$, $f_p(x) \leq f_p(r_1)$.
+However, this triangular region has width less than $w$  w.r.t.
+$f_q$, by combining part (a) with the fact that $f_p$ and $f_q$ increase in the same direction along $t$, see \Cref{fig:claim3}.
 \end{proof}
+
+\begin{figure}
 \includegraphics[ width=.7\textwidth]{Claim43.jpg}
+\caption{Figure for Claim 4.3}
+\label{fig:claim3}
+\end{figure}
 
-As a conclusion of the last two claims we have that $b$, $t$ and $r$ have positive slope.
+The last two claims can be summarized as saying that in the pictures $b$, $t$ and $r$ have positive slope.
 
@@
 
-Let $g$ be the functional constant on $[p_1, q_2]$. Then
+Let $g$ be the primitive lattice functional constant on $[p_1, q_2]$ (and therefore constant also on $[q_1, q_1+q_2-p_1]$).
+By the assumption on $p_1$, the values of $g$ on these segments differer by $1$.
+We choose the sign of $g$ so that
+\[
+g([p_1, q_2])= g( [q_1, q_1+q_2-p_1]) +1r. 
+\]
+
 \begin{claim}
-The slopes of $t=[t_1, t_2]$ and $b=[b_1, b_2]$ are strictly smaller than the slope of  and that of $r=[r_1, r_2]$ (and $l=[l_1, l_2]$) is strictly greater than it.
-\end{claim}\giulia{let's put some more notation in the beginning: add H+, H-, V+, V-, and say which endpoints of l, r, etc are in which halfspace.}
+\label{claim:g}
+$g(t_1) > g(t_2)$ and $g(b_1) > g(b_2)$ while $g(r_1) < g(r_2)$ (and $g(l_1) < g(l_2)$).
+%The slopes of $t$ and $b$ are strictly smaller than the slope of  $[p_1, q_2]$ and that of $r=[r_1, r_2]$ (and $l=[l_1, l_2]$) is strictly greater than it.
+\end{claim}
 
 \begin{proof}
-Since $b$ and $t$ must respectively separate $p_1$ and $q_1+q_2-p_1$ from the remaining vertices of the parallelogram $\conv(q_1, p_1, q_2, q_1+q_2-p_1)$, they must respectively intersect its (parallel) edges $\conv(p_1, q_2)$ and $\conv(q_1, q_1+q_2-p_1)$.
-This means their slopes must be strictly less than that of these segments.
-The same argument can be applied to $l$ and $r$ if one considers the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-p_2)$, yielding that the slopes of $l$ and $r$ are strictly larger than that of $\conv(p_1, q_2)$.
+Since $b$ and $t$ must respectively separate $p_1$ and $q_1+q_2-p_1$ from the other two vertices of the parallelogram $\conv(q_1, p_1, q_2, q_1+q_2-p_1)$, they must respectively intersect its (parallel) edges $[p_1, q_2]$ and $[q_1, q_1+q_2-p_1]$, which implies the stated inequalities for $b$ and $t$.
+The same argument  applied to the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-p_2)$, yields the inequalities for $\ell$ and $r$.
 \end{proof}
 
+\begin{proof}[Proof of \Cref{lemma:cayley}]
 We are now ready to show a contradiction.
-Since $Q$ has normal fan refining that of $P$, it must have an edge which is a translated copy of $r$.
-Let $u$ and $v$ be its endpoints.
-Now consider the lattice line $d$ through $u$ parallel to  $\conv(p_1, q_2)$.
-Let $d'$ be the next parallel lattice line above $d$.
-The width w.r.t $f_q$ of the vertical segment starting at $u$ with other endpoint $v'$ on $d'$ is $i$, since this vertical segment is a translation of $p$.
-Thus $v'$ must lie outside $Q$, by claim \ref{positive slopes}.
-That is, $v'$ lies above the line spanned by $t$.
-Since the slope of $t$ is less than the slope of $d$, the same happens for every point of $d'$ to the right of $u$.
-In particular, $v$ must be below $d'$ and hence it must be on or below $d$.
-This contradicts the fact that the slope of $r$ is strictly greater than the slope of $d$.
+Since the normal fan of $Q$ refines that of $P$, $Q$ must have an edge $r'$ which is a translated copy of $r$.
+Let $r_1'$ and $r_2'$ be its endpoints.
+Now consider the lattice line $d$ through $r_1'$ parallel to  $[p_1, q_2]$, that is, $g$ is constant on $d$.
+Let $d'$ be the parallel line defined by $g(d')=g(d)+1$.
+
+Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.
+
+%By claim \ref{positive slopes}, $s_2$ lies above the affine line $\langle t \rangle$.
+
+Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,
+and in particular for $r_2'$. Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
+\end{proof}
+\giulia{add a picture!}
 
@@
 
-\appendix
-
-\section{Old stuff}
-
-\begin{proof}
-%Suppose by contradiction that there is no lattice point as described in the lemma.
+%
+%\appendix
+%
+%\section{Old stuff}
+%
+%
+%\begin{proof}
+%%Suppose by contradiction that there is no lattice point as described in the lemma.
 Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
@@
 The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.
-
-Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
-\[
-p_1=(0,-\alpha), \quad
-p_2=(0,1-\alpha), \quad
-q_1=(-\beta,0), \quad
-q_2=(1-\beta,0).
-\] 
-The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
-
-Suppose by contradiction that there is no lattice point as described in the lemma.
+%
+%Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
+%\[
+%p_1=(0,-\alpha), \quad
+%p_2=(0,1-\alpha), \quad
+%q_1=(-\beta,0), \quad
+%q_2=(1-\beta,0).
+%\] 
+%The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
+%
+%Suppose by contradiction that there is no lattice point as described in the lemma.
 Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
@@
 To make things more explicit, we assume the following coordinates for the edge-vectors:
-\[
-\vec l = (l_1,l_2), \quad
-\vec r= (r_1,r_2), \quad
-\vec b= (b_1,b_2), \quad
-\vec t= (t_1,t_2).
-\]
-Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
-
-Since the normal fan of $Q$ refines that of $P$, $Q$ will contain edges $e_1'$ and $e_2'$ parallel to $e_1$ and $e_2$ respectively.
+%\[
+%\vec l = (l_1,l_2), \quad
+%\vec r= (r_1,r_2), \quad
+%\vec b= (b_1,b_2), \quad
+%\vec t= (t_1,t_2).
+%\]
+%Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
+%
+%
+%
+%
+%
+%
+%Since the normal fan of $Q$ refines that of $P$, $Q$ will contain edges $e_1'$ and $e_2'$ parallel to $e_1$ and $e_2$ respectively.
 These edges of $Q$ are all distinct and appear alternating in $Q$, that is, they appear in the order $f_1, e_1', f_2, e_2'$.
@@
 This is equivalent to saying that the normals to the edges $f_1, e_1, f_2, e_2$ lie in different open cones of the normal fan of $S$, in this cyclic order.
-
-- both $f_1$ or $f_2$ must meet the line through $q_1$ and $q_2$, else they would not cut out the lattice point they need to. same goes for $e_1$ and $e_2$ and the line through $p_1$ and $p_2$.
-
-claim: the segments $e_1 + l$ and $e_2 + l$ respectively intersect both $f_1$ and $f_2$.
-
-claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.
-
-%Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensibly}.
+%
+%- both $f_1$ or $f_2$ must meet the line through $q_1$ and $q_2$, else they would not cut out the lattice point they need to. same goes for $e_1$ and $e_2$ and the line through $p_1$ and $p_2$.
+%
+%claim: the segments $e_1 + l$ and $e_2 + l$ respectively intersect both $f_1$ and $f_2$. 
+%
+%claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both. 
+%
+%
+%
+%
+%
+%%Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensibly}.
 Since these segments do not contain lattice points, they must have lattice length stricly less than $1$ (or the width wrt $q$ is less than $k$, the index of $p,q$), which means that the smallest width in direction of $p_1 p_2$ between a "consecutive" vertices of $f_1$ and of $f_2$ must also be strictly less than $k$.
-
-%Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
-\end{proof}
-
-\begin{figure}[htb]
-  \centering
-\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  .7]
-%coordinate axes
-   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%        \tkzLabelY[orig=false,label options={font=\tiny}]
-        \tkzGrid
-%       \tkzDrawX
-%       \tkzDrawY
-
-%hexagon
-  \node[mynode] (A) at (0,-1) {}; 
-%  \node[mynode] (B) at (1,1) {};  
-  \node[mynode] (C) at (2,3) {};  
-  \node[mynode] (D) at (0,1) {};  
-%  \node[mynode] (E) at (-1,-1) {};  
-  \node[mynode] (F) at (-2,-3) {};
-
-%other points
-%  \node[mynode] () at (0,-1) {};
-
-  \draw[black] (A) -- (C) -- (D) -- (F) -- (A);
-  \draw[black, thick] (A) -- (D);
-\end{tikzpicture}
 %
-\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  .7]
-%coordinate axes
-   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%        \tkzLabelY[orig=false,label options={font=\tiny}]
-        \tkzGrid
-%        \tkzDrawX
-%        \tkzDrawY
-
-%hexagon
-  \node[mynode] (A) at (1,0) {}; 
-  \node[mynode] (B) at (2,1) {};  
-  \node[mynode] (C) at (3,3) {};  
-  \node[mynode] (D) at (-1,0) {};  
-  \node[mynode] (E) at (-2,-1) {};  
-  \node[mynode] (F) at (-3,-3) {};
-
-  \node[mynode] (p1) at (-1,0) {}; 
-  \node[mynode] (p2) at (1,0) {};
-
-  \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
-  \draw[black, thick] (p1) -- (p2);
-\end{tikzpicture}
-
-\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
-
-\caption{Partial counter-example to \Cref{lemma:cayley}.
+%%Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
+%\end{proof}
+%
+%
+%
+%\begin{figure}[htb]
+%  \centering
+%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+%  .7]
+%%coordinate axes
+%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%%        \tkzLabelY[orig=false,label options={font=\tiny}]
+%        \tkzGrid
+%%       \tkzDrawX
+%%       \tkzDrawY
+%
+%%hexagon
+%  \node[mynode] (A) at (0,-1) {}; 
+%%  \node[mynode] (B) at (1,1) {};  
+%  \node[mynode] (C) at (2,3) {};  
+%  \node[mynode] (D) at (0,1) {};  
+%%  \node[mynode] (E) at (-1,-1) {};  
+%  \node[mynode] (F) at (-2,-3) {};  
+%
+%%other points
+%%  \node[mynode] () at (0,-1) {}; 
+%
+%
+%  \draw[black] (A) -- (C) -- (D) -- (F) -- (A);
+%  \draw[black, thick] (A) -- (D);
+%\end{tikzpicture}
+%%
+%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+%  .7]
+%%coordinate axes
+%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%%        \tkzLabelY[orig=false,label options={font=\tiny}]
+%        \tkzGrid
+%%        \tkzDrawX
+%%        \tkzDrawY
+%
+%%hexagon
+%  \node[mynode] (A) at (1,0) {}; 
+%  \node[mynode] (B) at (2,1) {};  
+%  \node[mynode] (C) at (3,3) {};  
+%  \node[mynode] (D) at (-1,0) {};  
+%  \node[mynode] (E) at (-2,-1) {};  
+%  \node[mynode] (F) at (-3,-3) {};  
+%
+%  \node[mynode] (p1) at (-1,0) {}; 
+%  \node[mynode] (p2) at (1,0) {};  
+%
+%  \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
+%  \draw[black, thick] (p1) -- (p2);
+%\end{tikzpicture}
+%
+%\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
+%
+%\caption{Partial counter-example to \Cref{lemma:cayley}.
 With respect to the standard lattice it is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one).
-%
-With respect to the lattice of index two it is not a counter-example either, because two vertices of $Q$ are not in that lattice.
-}
-\label{fig:width3}
-\end{figure}
+%%
+%With respect to the lattice of index two it is not a counter-example either, because two vertices of $Q$ are not in that lattice.
+%}
+%\label{fig:width3}
+%\end{figure}
 

### 2019-07-06 unimodular-covers.tex
reviewed introduction

@@
 %
-Oda~\cite{Oda1997} posed several questions regarding IDP polytopes which are now considered conjectures~\cite{HNPS2008,mfo2007}. 
-Among them:
+Oda~\cite{Oda1997} posed several questions regarding IDP polytopes. The following variations of them are now considered conjectures~\cite{HNPS2008,mfo2007}:
 \begin{conjecture}
@@
 
-This not only implies \Cref{conj:Oda}(1) in dimension three, but also \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
+This not only implies \Cref{conj:Oda}(1) in dimension three, but also, as we show in \Cref{sec:cayley}, \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
 
@@
 %\section{Preliminaries}
-%\label{sec:prelim}
+%\label{Fprelim}
 
@@
 w=\width_{f_q}(p + \vecline q )=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\vecline p ). 
-\]
+\] \giulia{we want a picture for the setup}
 
+\begin{proof}[Proof of \Cref{lemma:cayley}]
 Suppose by contradiction that there is no lattice point as described in the lemma.
@@
 \begin{claim}
-$f_q(b_2-b_1)$ and $f_q(p_2 - p_1)$ are non-zero and have the same sign. That is, $f_q$ achieves its maximum over $b$ and over $t$ on the same halfplane $V_1$ or $V_2$.
-\end{claim}\giulia{do we need this claim?}
+$f_q(b_2-b_1)$ and $f_q(t_2 - t_1)$ are non-zero and have the same sign. That is, $f_q$ achieves its maximum over $b$ and over $t$ on the same halfplane $V_1$ or $V_2$.
+\end{claim}
 
@@
 The last two claims can be summarized as saying that in the pictures $b$, $t$ and $r$ have positive slope.
+Observe that this implies that $q$ is not in the boundary of $Q$ and $p \neq r$, so both $P$ and $Q$ are full dimensional.
 
@@
 
-\begin{proof}[Proof of \Cref{lemma:cayley}]
 We are now ready to show a contradiction.

### 2019-07-07 unimodular-covers.tex
reviewed introduction + final changes to sec4

@@
 
-This not only implies \Cref{conj:Oda}(1) in dimension three, but also, as we show in \Cref{sec:cayley}, \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
+This not only implies \Cref{conj:Oda}(1) in dimension three, but also the following two statements, that we prove in \Cref{sec:cayley}.
+A $3$-dimensional \emph{prismatoids} is a polytopes whose vertices all lie in two parallel facets.
+A special case is a polytope of width $1$, whose vertices lie in two consecutive parallel lattice hyperplanes.
 
@@
 
-\Paco{another corollary, suggested by Christian Haase: the $k$-th dilation of an emptyy simplex is very easy to decompose into slices of width one all of which fit into our description ($P$ and $Q$ are either pararllel parallelograms, one of which perhaps degenerates to a segment).
-In particular, we recover as a corollary that the $k$-th dilation of every empty tetrahedron (and, a fortiori, the $k$-th dilation of every lattice $3$-polytope) has a unimodular cover for every $k\ge 2$
+\begin{corollary}
+Every integer dilation of a lattice $3$-polytope of width $1$ has a unimodular cover.
+\end{corollary}
 
-EXCEPT FOR THIS TO BE TRUE WE NEED TO ALLOW $Q$ TO DEGENERATE TO AN EDGE IN \Cref{hm:cayley}. CAN WE?
-}
+We believe that the $3$-polytopes in all these statements all have unimodular triangulations, but this remains an open question.
 
-We believe that the $3$-polytopes in these four statements all have unimodular triangulations, but we do not have a proof.
-
-\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, Gaku Liu, Johannes Hofscheier, Christian Haase, .... et al.
-
-%\section{Preliminaries}
-%\label{Fprelim}
+\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and 
+Christian Haase for helpful discussions.
 
@@
 Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.  
-
-%By claim \ref{positive slopes}, $s_2$ lies above the affine line $\langle t \rangle$.
-
 Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,

### 2019-07-07 unimodular-covers.tex
edited section 2

--- 2019-07-06 unimodular-covers.tex
@@
 \usepackage[colorlinks=true]{hyperref}
+%\usepackage{pictex}
 
@@
 
-\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions?}
+\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions? Probably not}
 A lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice. 
 %
-Oda~\cite{Oda1997} posed several questions regarding IDP polytopes. The following variations of them are now considered conjectures~\cite{HNPS2008,mfo2007}:
+Oda~\cite{Oda1997} posed several questions regarding smoothness and IDP for lattice polytopes. The following versions of them are now considered conjectures~\cite{HNPS2008,mfo2007}:
 \begin{conjecture}
@@
 \begin{enumerate}
-\item (Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
+\item 
+\label{itm:mixedIDP}
+(Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
 \begin{align*}
@@
 
-\item (Related to problems 2 and 5 in \cite{Oda1997})
+\item 
+\label{itm:smoothIDP}
+(Related to problems 2 and 5 in \cite{Oda1997})
 Every smooth lattice polytope is IDP.
@@
 
-Another case of Oda's questions that has been solved is the two-dimensional case of \Cref{conj:Oda}(1), with three different proofs by
+Another case of Oda's questions that has been solved is the two-dimensional case of \Cref{conj:Oda}\eqref{itm:mixedIDP}, with three different proofs by
 Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth. 
@@
 
-An alternative approach to \Cref{conj:Oda}(1) is via Cayley sums, that we discuss in  \Cref{sec:cayley}. 
+An alternative approach to \Cref{conj:Oda}\eqref{itm:mixedIDP} is via Cayley sums, that we discuss in  \Cref{sec:cayley}. 
 As we note there (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
@@
 
-This not only implies \Cref{conj:Oda}(1) in dimension three, but also, as we show in \Cref{sec:cayley}, \Cref{conj:Oda}(2) for $3$-dimensional \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
+This not only implies \Cref{conj:Oda}\eqref{itm:mixedIDP} in dimension two but also, as we show in \Cref{sec:cayley}, \Cref{conj:Oda}\eqref{itm:smoothIDP} for $3$-dimensional smooth \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
 
@@
 
-The main tool is what we call the parallelepiped circumscribed to a given tetrahedron, defined as follows:
+The main tool for the proof of \Cref{thm:parallelepipeds} is what we call the parallelepiped circumscribed to a given tetrahedron, defined as follows:
 
 \begin{definition}
-Let $T$ be a tetrahedron with vertex set $p_1,p_2,p_3,p_4$. Consider points $q_1,q_2,q_3,q_4$ defined as $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ and define
-
+\label{def:circunpara}
+Let $T$ be a tetrahedron with vertices $p_1$, $p_2$, $p_3$, and $p_4$. Consider the points $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$, $i\ n [4]$, and let
 \[
@@
 \] 
-
-$C(T)$ is a parallelepiped whose faces are of the form $\conv(p_i, p_j, q_k, q_l)$ for all $\{i,j,k,l\}=[4]$, which we call the \emph{parallelepiped circumscribed} to $T$.
-
-For each $i \in [4]$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=[4]$; we call these $T_i$ the \emph{corner tetrahedra} of $C(T)$. These along with $T$ form a triangulation of $C(T)$.
+$C(T)$ is a parallelepiped with facets $\conv(p_i, p_j, q_k, q_l)$ for all choices of $\{i,j,k,l\}=[4]$. We call it the \emph{parallelepiped circumscribed} to $T$.
+
+For each $i \in [4]$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=[4]$; we call these $T_i$ the \emph{corner tetrahedra} of $C(T)$. Together with $T$ they triangulate of $C(T)$.
 \end{definition}
-\paco{define the ``corner tetrahedra''}
-
-The idea to have in mind is that of the regular tetrahedron inscribed in a cube. Modulo an affine transformation this is exactly the situation of $T$ and $S$.
+
+The idea to have in mind is that of the regular tetrahedron inscribed in a cube. Modulo an affine transformation this is exactly the situation of $T$ and $C(T)$.
 \giulia{we should add a picture of the standard cube}
-
-\Giulia{not sure if we need this remark:
-$S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
+\paco{perhaps, but not necessarily}
+
+%\Giulia{not sure if we need this remark:
+%$S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
 Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it.
 (Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).
-}
+%}
 
@@
 \label{lemma:corner}
-Let $T=\conv\{p_1,p_2,p_3,p_4\}$ be an empty lattice tetrahedron that is not unimodular and $C(T)$ be the parallelepiped circumscribed to $T$.
-Let $T_1, T_2,T_3$ and $T_4$ be the corresponding corner tetrahedra in $C(T)$.
+Let $T=\conv\{p_1,p_2,p_3,p_4\}$ be an empty lattice tetrahedron that is not unimodular.
+Let $C(T)$ be the parallelepiped circumscribed to $T$ and let $T_1, T_2,T_3$ and $T_4$ be the corresponding corner tetrahedra in $C(T)$.
 Then, every $T_i$ contains at least one lattice point different from $\{p_1,\dots,p_4\}$.
@@
 \paco{picture of the quadrilaterals $p_1q_4p_2q_3$ and $q_2p_3q_1p_4$ would be good.}
+\paco{perhaps, but not necessarily}
 \begin{align*}
@@
 
-Then, the inequalities $b\ge 1+a$ and $b\ge 2$ imply:
-\[
-(1,1,0)\in \conv(p_1p_2q_3) \subset T_4, \quad
-(0,-1,0)\in \conv(p_1p_2q_4) \subset T_3.
-\]
-\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
-Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.
+Then, the inequalities $b\ge 1+a \ge 2$ imply:
+\[
+u:=(1,1,0)\in \conv(p_1p_2q_3) \subset T_4, \quad
+v:=(0,-1,0)\in \conv(p_1p_2q_4) \subset T_3.
+\]
+
+Observe that $u+v=p_1+p_2=q_3+q_4$.
+Now, the translation of vector $\left(\frac{a-1}2,\frac{b}2,1\right)$ sends $\conv(p_1p_2q_3q_4)$ to $\conv(q_2q_1p_4p_3)$, so that the translated points $u + \left(\frac{a-1}2,\frac{b}2,1\right)$ and $u + \left(\frac{a-1}2,\frac{b}2,1\right)$ lie in $\conv(q_2q_1p_4p_3)$ and are centrally symmetrically placed in it.
+thus, one of them belongs to $\conv(q_1p_4p_3)$ and the other to $\conv(q_2p_4p_3)$.
+%\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
+%Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.
 Hence, its translate $\conv(q_2p_3q_1p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$.
@@
 \begin{proof}
-We denote the vertices of $T$ by $p_1, p_2, p_3, p_4$ and by $q_1, q_2, q_3, q_4$ the vertices of $C(T)$ not in $T$, as before.
-Let $f\in (\R^3)^*$ be any functional.
-Fix $i \in [4]$; observe that if $f(q_i) > f( p_j)$ for all $j\neq i$, then
-\begin{gather*}
-f(q_i) \overset{(i)}{>} f(q_j) \overset{(ii)}{>} f(p_i) \\
-f(q_i) \overset{(iii)}{>} f(p_j) \overset{(iv)}{>} f(p_i) \\
-f(p_j)\overset{(v)}{>}f(q_k)
-\end{gather*}
-for all $j \neq i$, $k \notin \{j, i\}$.
-
-Indeed, for any $\{i,j,k,l\}=[4]$ we have
-
-\begin{align*}
-f(q_i) > f( p_j) \Leftrightarrow \frac12 f(p_l+&p_k -p_j-p_i) > 0  \Leftrightarrow f(q_j) > f(p_i)\\
-& \Updownarrow \\
- f(q_l) < f(p_k) \Leftrightarrow \frac12 f(-p_l-&p_k +p_j+p_i) < 0 \Leftrightarrow f(q_k) < f(p_l),
-\end{align*}
-which shows the equivalence of $(i)$, $(iii)$ and $(v)$; note that $(iii)$ is our assumption.
+Let us denote the vertices of $T$ by $p_1, p_2, p_3, p_4$ and the vertices of $C(T)$ not in $T$ by $q_1, q_2, q_3, q_4$, with the conventions of \Cref{def:circunpara}.
+
+\paco{changed this}
+The parallelepiped $P$ is the intersection of three bands\paco{bands?}, where each band is a region defined as $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*$ and closed interval $[\alpha,\beta]\subset \R$.
+We claim that a band containing $T$ must contain at least three of the $q_i$s, which implies that the intersection $P$ of three of them contains at least one of the $q_i$s and hence $P$ contains the corresponding $T_i$.
+
+To prove the claim suppose that $q_1\not\in B:= f^{-1}([\alpha,\beta])$ for a certain band $B$. 
+Without loss of generality, say $f(q_1)<\alpha$. Then the equalities $q_1+q_2=p_3+p_4$ and $q_1+p_1=q_2+p_2$ respectively give:
+\begin{gather}
+f(q_2) = f(p_3+p_4-q_1) = f(p_3)+f(p_4)-f(q_1) > 2\alpha-\alpha=\alpha,\\
+f(q_2) = f(q_2+p_2-p_1) = f(q_2)+(f(p_2)-f(p_1)) < \alpha + (\beta-\alpha) = \beta,
+\end{gather}
+so that $q_2 \in B$.
+
+The first equation also implies
+\[
+f(q_1) <  f(p_i) < f(q_2), \quad\text{ for $i=3,4$}.
+\]
+By the translation sending $q_1,q_2,p_3,p_4$ to $p_2,p_1,q_4,q_3$ (in this order) we then have
+\[
+\alpha \le f(p_2) < f(q_i) < f(p_1) \le \beta, \quad\text{ for $i=3,4$},
+\]
+so that $q_3,q_4 \in B$.
+%
+This finishes the proof of the claim, and of the lemma.
+%
+%Let $f\in (\R^3)^*$ be any functional. Fix $i \in [4]$; observe that if $f(q_i) > f( p_j)$ for all $j\neq i$, then 
+%\begin{gather*}
+%f(q_i) \overset{(i)}{>} f(q_j) \overset{(ii)}{>} f(p_i) \\
+%f(q_i) \overset{(iii)}{>} f(p_j) \overset{(iv)}{>} f(p_i) \\
+%f(p_j)\overset{(v)}{>}f(q_k)
+%\end{gather*}
+%for all $j \neq i$, $k \notin \{j, i\}$. 
+%
+%Indeed, for any $\{i,j,k,l\}=[4]$ we have 
+%
+%\begin{align*}
+%f(q_i) > f( p_j) \Leftrightarrow \frac12 f(p_l+&p_k -p_j-p_i) > 0  \Leftrightarrow f(q_j) > f(p_i)\\
+%& \Updownarrow \\
+% f(q_l) < f(p_k) \Leftrightarrow \frac12 f(-p_l-&p_k +p_j+p_i) < 0 \Leftrightarrow f(q_k) < f(p_l),
+%\end{align*}
+%which shows the equivalence of $(i)$, $(iii)$ and $(v)$; note that $(iii)$ is our assumption.
 Inequalities $(ii)$ and $(iv)$ follow by taking the sum of two inequalities of the form $\frac12 f(p_l+p_k -p_j-p_i) > 0$.
-
-Using this observation, we can show that each of pairs of parallel facets of $P$ cuts out at most one of the $q_i$s:
-if $q_i$ lies outside of $P$, it must violate a facet-defining inequality $f(x) \leq a$.Thus $f(q_i) > a \geq f(p_j)$, and by our observation, it follows that all remaining $q_j$ satisfy the inequality $f(x) \leq a$.
+%
+%Using this observation, we can show that each of pairs of parallel facets of $P$ cuts out at most one of the $q_i$s:
+%if $q_i$ lies outside of $P$, it must violate a facet-defining inequality $f(x) \leq a$.Thus $f(q_i) > a \geq f(p_j)$, and by our observation, it follows that all remaining $q_j$ satisfy the inequality $f(x) \leq a$.
 Since $P$ is a parallelogram, there is a facet with inequality $f(x) \geq b$; by what is said above, the minimum value of $f(x)$ is attained at $p_i$, which is in $P$.
 Thus $f(q_j) > f(p_i) \geq b$, that is, the inequality is satisfied by all the $q_j$.
-
-Since at most one of the $q_i$ does not satisfy the inequalities corresponding to a pairs of parallel facets of $P$, and there are only three pairs of facets, one of $q_i$s must satisfy all the inequalities and therefore be inside of $P$.
+%
+%Since at most one of the $q_i$ does not satisfy the inequalities corresponding to a pairs of parallel facets of $P$, and there are only three pairs of facets, one of $q_i$s must satisfy all the inequalities and therefore be inside of $P$.
 \end{proof}
-\giulia{here it may also be nice to add a picture}
-
-\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
-Let $P \subseteq \R^3$ be a parallelopiped.
-It can be triangulated into empty lattice tetrahedra, simply by taking a triangulation which uses all lattice points as vertices.
-If all the tetrahedra in the triangulation are unimodular we are done.
-Otherwise, suppose $T$ is be empty tetrahedron in the triangulation which is not unimodular.
-\Cref{lemma:3<4} guarantees that one of the corner tetrahedra $T_i$ of the parallelepiped $C(T)$ circumscribed to $T$ is contained in $P$.
+%\giulia{here it may also be nice to add a picture}
+
+\begin{corollary}
+\label{coro:coverpara}
+\paco{added this intermediate result}
+Let $T$ be a lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
+\end{corollary}
+
+\begin{proof}
+We proceed by induction on the (normalized) volume of $T$, which is a positive integer.
+If this volume equals $1$ then $T$ is unimodular and there is nothing to prove, so we assume $T$ is not unimodular.
+
+\Cref{lemma:3<4} guarantees that one of the corner tetrahedra $T_i$ of the parallelepiped $C(T)$ is contained in $P$.
 Without loss of generality, suppose $T_4 = \conv(p_1, p_2, p_3,q_4)$ is in $P$.
-By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point, which we denote by $l$.
-Observe that from the proof of \Cref{lemma:corner} we know more: $l$ lies in the face $\conv(p_1, p_2, q_4)$.\giulia{we need this for the volume computation below.
+By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point other than the $p_i$s, which we denote by $u$.
+%Observe that from the proof of \Cref{lemma:corner} we know more: $u$ lies in the face $\conv(p_1, p_2, q_4)$.\giulia{we need this for the volume computation below.
 perhaps it should be made explicit in the statemente of the lemma...}
+\paco{removed assumption the $u$ lies in a face. I think we do not need it}
 
@@
 \]
-
-Each of the tetrahedra $S_i$ has lattice volume strictly smaller than $T$: indeed, $\Vol(S_3)=\Vol(T'_4)$, since these tetrahedra have a common face $\conv(p_1, p_2, l)$ and the remaining vertex is at distance one from this face in both.
+Each of the tetrahedra $S_i$ has lattice volume strictly smaller than $T$, because, for each $i$, $p_i$ is the unique point of $C(T)$ maximizing the distance to the opposite facet $\conv(p_j,p_k,p_l)$ of $T$.
+Thus, $S_1$, $S_2$ and $S_3$ cover $T$ and have volume strictly smaller than $T$.
+By inductive hypothesis they can be covered unimodularly, and we are done,
+
+Observe that  $T'_4$ and/or one of the $S_i$ can degenerate to a triangle, if $u$ lies in $\conv(p_1,p_2,p_3)$, 
+but this does not affect the argument.
+%
+%: indeed, $\Vol(S_3)=\Vol(T'_4)$, since these tetrahedra have a common face $\conv(p_1, p_2, l)$ and the remaining vertex is at distance one from this face in both.
 Now $T'_4$ is contained in $T_4$, and we know that each corner tetrahedron $T_i$ has volume half that of $T$.
 The remaining two tetrahedra $S_1$ and $S_2$ must therefore have volume summing to that of $T$, and therefore each has volume strictly less than $T$.
-
-This allows us to cover $T$ with lattice simplices of smaller volume recursively, until it is covered by unimodular ones.
 \end{proof}
 
-\begin{remark}
-\label{rem:octahedra_and_tri-prisms}
+\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
+Arbitrarily triangulate the parallelepiped into lattice tetrahedra and apply \Cref{coro:coverpara} to these tetrahedra.
+%Let $P \subseteq \R^3$ be a lattice parallelepiped.
+It can be triangulated into empty lattice tetrahedra, simply by taking a triangulation which uses all lattice points as vertices.
+If all the tetrahedra in the triangulation are unimodular we are done.
+Otherwise, suppose $T$ is an empty tetrahedron in the triangulation which is not unimodular.
+\Cref{lemma:3<4} guarantees that one of the corner tetrahedra $T_i$ of the parallelepiped $C(T)$ circumscribed to $T$ is contained in $P$.
+Without loss of generality, suppose $T_4 = \conv(p_1, p_2, p_3,q_4)$ is in $P$.
+By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point, which we denote by $u$.
+Observe that from the proof of \Cref{lemma:corner} we know more: $u$ lies in the face $\conv(p_1, p_2, q_4)$.\giulia{we need this for the volume computation below.
+perhaps it should be made explicit in the statemente of the lemma...}
+%
+%
+%This allows us to cover $T$ with lattice simplices of smaller volume recursively, until it is covered by unimodular ones.
+\end{proof}
+
+%\begin{remark}
+%\label{rem:octahedra_and_tri-prisms}
+
+We finish with a question which, if answered in the positive, would imply \Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three.
 Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is
-contained in $P$''. Then $P$ has a unimodular cover, since the proof of \Cref{thm:parallelepipeds} works for $P$.
-
-In turn, our proof of \Cref{lemma:3<4} relies on the fact that parallelepipeds have only three (pairs of) normal vectors.
-The proof fails if $P$ has four of them, as the following non-IDP octahedron and triangular prism show:
-
-\begin{example}
-\label{ex:non-IDP}
-The following lattice octahedron is not IDP:
-\[
-Q= \conv((0,1,1),(1,0,1),(1,1,0),(0,-1,-1),(-1,0,-1),(-1,-1,0))
-\]
-
-The following lattice triangular prism is not IDP:
-\[
-P=\conv((0,1,1),(1,0,1),(1,1,0),(-1,0,0),(0,-1,0),(0,0,-1)).
-\]
-Indeed, in both polytopes the only lattice points are the six vertices and the origin. The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
-\giulia{is it necessary to be more explicit, for example show that these are the only lattice points?}
-\end{example}
-
-\end{remark}
+contained in $P$''. If this holds then $P$ has a unimodular cover, since then the proofs of \Cref{coro:coverpara} and \Cref{thm:parallelepipeds} work for $P$.
 
@@
 
+Our proof that parallelepipeds have the property (\Cref{lemma:3<4}) is based on the fact that they have only three (pairs of) normal vectors.
+The proof, and the property of being IDP, fail if there are four of them:
+
+\begin{example}[Non-IDP octahedron and triangular prism]
+\label{ex:non-IDP}
+The following lattice octahedron $Q$ and triangular prism $P$ are not IDP:
+\begin{gather}
+Q= \conv((0,1,1),(1,0,1),(1,1,0),(0,-1,-1),(-1,0,-1),(-1,-1,0)),\\
+P=\conv((0,1,1),(1,0,1),(1,1,0),(-1,0,0),(0,-1,0),(0,0,-1)).
+\end{gather}
+
+Indeed, in both polytopes the only lattice points are the six vertices and the origin.
+The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
+Hence, they are not IDP, whcih implies they do not admit unimodular covers.
+\giulia{is it necessary to be more explicit, for example show that these are the only lattice points?}
+\paco{I think it is ok as is. It is only an example}
+\end{example}
+
 \section{nef Cayley sums}
@@
 
-In this language, \Cref{conj:Oda}(1) translates to:
+In this language, \Cref{conj:Oda}\eqref{itm:mixedIDP} translates to:
 \emph{If $Q$ is nef for $P$ and $P$ is smooth then the pair $(P,Q)$ mixed IDP.}
@@
 \]
-contains a lattice point.
-
-\paco{add picture for these strips}
-
+contains a lattice point. See \Cref{fig:strips}.
 %\[
@@
 
+\begin{figure}[htb]
+\scalebox{.75}{\input{strips.pdf_t}}
+\caption{???}
+\label{fig:strips}
+\end{figure}
+
 \begin{proof}[Proof of \Cref{thm:cayley}]
@@
 
-Observe that in the strip strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=-1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
+Observe that in the strip strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
+%We call that point $q^+$.
+Similarly, $q +\vecline p$ has a unique lattice point at $f_q(x)=-1$, which we call $q_-$.
+Observe that $q_++q_-=q_1+q_2$ and that the triangles $\conv(q_1, q_2, q_+)$ and $\conv(q_1, q_2, q_1)$ are unimodular.
 Since translating the polytopes by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
@@
 \begin{array}{cc}
-f_q(l_2) \geq f_q(p_2), &
-f_q(l_1) \leq f_q(p_1), \\
-f_q(r_2) \geq f_q(p_2), &
-f_q(r_1) \leq f_q(p_1).
+f_q(l_1) \leq f_q(p_1), &
+f_q(l_2) \geq f_q(p_2), \\
+f_q(r_1) \leq f_q(p_1),&
+f_q(r_2) \geq f_q(p_2).
 \end{array}
@@
 \begin{array}{cc}
-f_p(t_2) \geq f_p(q_2), &
-f_p(t_1) \leq f_p(q_1) \\
-f_p(b_2) \geq f_p(q_2), &
-f_p(b_1) \leq f_p(q_1).
+f_p(t_1) \leq f_p(q_1), &
+f_p(t_2) \geq f_p(q_2), \\
+f_p(b_1) \leq f_p(q_1), &
+f_p(b_2) \geq f_p(q_2).
 \end{array}
@@
 
-Observe that one of $l$ and $r$ can coincide with $p$, if this is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if this is on the boundary of $Q$. 
+Observe that a priori one of $l$ and $r$ can coincide with $p$, if this is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if this is on the boundary of $Q$. 
 \giulia{p=P ???}
@@
 %Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
-
-Then:
 
@@
 \begin{claim}
+\label{claim:b_and_t}
 $f_q(b_2-b_1)$ and $f_q(t_2 - t_1)$ are non-zero and have the same sign. That is, $f_q$ achieves its maximum over $b$ and over $t$ on the same halfplane $V_1$ or $V_2$.
@@
 \begin{proof}
-Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $V_1$ and that on $b$ lies in $V_2$.
+Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $V_1$ and that the maximum on $b$ lies in $V_2$.
 
@@
 
+\begin{figure}[htb]
+\scalebox{.75}{\input{claim2.pdf_t}}
+\caption{Illustration of the proof of \Cref{claim:b_and_t}}
+\label{fig:claim2}
+\end{figure}
 \Giulia{fix it to look more unimodular at the bottom}
@@
 \begin{claim}
-\label{positive slopes}
+\label{claim:r}
 Assume w.l.o.g.~that $b$ and $t$ either are parallel or their affine spans cross in $V_2$. Then, 
@@
 \begin{proof}
-
 Both $t$ and $b$ must intersect $p$, otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.
@@
 
-For part (b), If $f_p(r_2) \leq f_p(r_1)$, it would be impossible to fit a translated copy of $r$ in the correct side of $Q$: $r$ would need to lie inside the triangle delimited by the affine line $\langle t \rangle$ and the inequalities $f_q(x) \geq f_q(r_1)$, $f_p(x) \leq f_p(r_1)$.
+%\begin{figure}
+%\includegraphics[ width=.7\textwidth]{Claim43.jpg}
+%\caption{Figure for Claim 4.3}
+%\label{fig:claim3}
+%\end{figure}
+\begin{figure}[htb]
+\scalebox{.75}{\input{claim3.pdf_t}}
+\caption{Illustration of the proof of \Cref{claim:r}}
+\label{fig:claim3}
+\end{figure}
+
+For part (b), If $f_p(r_2) \leq f_p(r_1)$, it would be impossible to fit a translated copy $r'$ of $r$ in the correct side of $Q$: $r$ would need to lie inside the triangle delimited by the affine line $\langle t \rangle$ and the inequalities $f_q(x) \geq f_q(r_1)$, $f_p(x) \leq f_p(r_1)$.
 However, this triangular region has width less than $w$  w.r.t.
@@
 \end{proof}
-
-\begin{figure}
-\includegraphics[ width=.7\textwidth]{Claim43.jpg}
-\caption{Figure for Claim 4.3}
-\label{fig:claim3}
-\end{figure}
 
@@
 \[
-g([p_1, q_2])= g( [q_1, q_1+q_2-p_1]) +1r. 
+g([p_1, q_2])= g( [q_1, q_1+q_2-p_1]) +1. 
 \]
@@
 \label{claim:g}
-$g(t_1) > g(t_2)$ and $g(b_1) > g(b_2)$ while $g(r_1) < g(r_2)$ (and $g(l_1) < g(l_2)$).
+$g(t_1) > g(t_2)$, $g(b_1) > g(b_2)$, and $g(r_1) < g(r_2)$.
 %The slopes of $t$ and $b$ are strictly smaller than the slope of  $[p_1, q_2]$ and that of $r=[r_1, r_2]$ (and $l=[l_1, l_2]$) is strictly greater than it.
@@
 \end{proof}
+
+\begin{figure}[htb]
+\scalebox{.75}{\input{claim4.pdf_t}}
+\caption{Illustration of the proof of \Cref{claim:g}}
+\label{fig:claim4}
+\end{figure}
 
@@
 
-%By claim \ref{positive slopes}, $s_2$ lies above the affine line $\langle t \rangle$.
+%By claim \ref{claim:r}, $s_2$ lies above the affine line $\langle t \rangle$.
 
@@
 \newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006.
+
+\bibitem{Santos-hirsch}
+\paco{add reference}
 

### 2019-07-07 unimodular-covers.tex

@@
 
-This not only implies \Cref{conj:Oda}\eqref{itm:mixedIDP} in dimension two but also, as we show in \Cref{sec:cayley}, \Cref{conj:Oda}\eqref{itm:smoothIDP} for $3$-dimensional smooth \emph{prismatoids}, polytopes whose vertices all lie in two parallel facets:
+This not only implies \Cref{conj:Oda}(1) in dimension three, but also the following two statements, that we prove in \Cref{sec:cayley}.
+A $3$-dimensional \emph{prismatoids} is a polytopes whose vertices all lie in two parallel facets.
+A special case is a polytope of width $1$, whose vertices lie in two consecutive parallel lattice hyperplanes.
 
@@
 
-\Paco{another corollary, suggested by Christian Haase: the $k$-th dilation of an emptyy simplex is very easy to decompose into slices of width one all of which fit into our description ($P$ and $Q$ are either pararllel parallelograms, one of which perhaps degenerates to a segment).
-In particular, we recover as a corollary that the $k$-th dilation of every empty tetrahedron (and, a fortiori, the $k$-th dilation of every lattice $3$-polytope) has a unimodular cover for every $k\ge 2$
-
-EXCEPT FOR THIS TO BE TRUE WE NEED TO ALLOW $Q$ TO DEGENERATE TO AN EDGE IN \Cref{hm:cayley}. CAN WE?
-}
-
-We believe that the $3$-polytopes in these four statements all have unimodular triangulations, but we do not have a proof.
-
-\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, Gaku Liu, Johannes Hofscheier, Christian Haase, .... et al.
-
-%\section{Preliminaries}
-%\label{Fprelim}
+\begin{corollary}
+Every integer dilation of a lattice $3$-polytope of width $1$ has a unimodular cover.
+\end{corollary}
+
+We believe that the $3$-polytopes in all these statements all have unimodular triangulations, but this remains an open question.
+
+\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and 
+Christian Haase for helpful discussions.
 
@@
 Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.  
-
-%By claim \ref{claim:r}, $s_2$ lies above the affine line $\langle t \rangle$.
-
 Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,

### 2019-07-07 unimodular-covers.tex
made changes mostly in Section 3

@@
 
-\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions? Probably not}
+\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions in the intro? Probably not}
 A lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice. 
@@
 \item 
+\label{itm:smoothIDP}
+(Related to problems 2 and 5 in \cite{Oda1997})
+Every smooth lattice polytope is IDP.
+\item 
 \label{itm:mixedIDP}
-(Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $Q$ refining that of $P$, and $P$ smooth. Then
+(Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $P$ refining that of $Q$, and $Q$ smooth. Then
 \begin{align*}
@@
 \paco{perhaps we can call this formula ``being mixed IDP''. The property of ``$N(Q)$ refines $N(P)$'' is sometimes called ``$P$ is a Minkowski summand of $Q$''.}
-
-\item 
-\label{itm:smoothIDP}
-(Related to problems 2 and 5 in \cite{Oda1997})
-Every smooth lattice polytope is IDP.
 \end{enumerate}
@@
 parallelepipeds and unimodular simplices, both of which are trivially IDP.
-
-In \Cref{sec:parallelepipeds} we show the following stronger statement:
+%
+In \Cref{sec:parallelepipeds} we show:
 
@@
 
-As a consequence:
+This, together with the mentioned result from~\cite{BHHHJKM2019}, gives:
 
@@
 An alternative approach to \Cref{conj:Oda}\eqref{itm:mixedIDP} is via Cayley sums, that we discuss in  \Cref{sec:cayley}. 
+\paco{added def}
+Recall that the \emph{Cayley sum} of two lattice polytopes $Q_1,Q_2\subset \R^d$ is the lattice polytope
+\[
+\cayley(Q_1,Q_2) := \conv(Q_1\times\{0\} \cup Q_2 \times \{1\}) \subset \R^3.
+\]
+\paco{new sentence. In fact, this implies that Cayley sums are the same as polytopes of width 1}
+We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can remove coordinates) but this does not need $Q_1$ and $Q_2$ to be full-dimensional.
+It only requires the linear subspaces parallel to them to span $\R^d$.
+
 As we note there (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
@@
 
-This not only implies \Cref{conj:Oda}(1) in dimension three, but also the following two statements, that we prove in \Cref{sec:cayley}.
-A $3$-dimensional \emph{prismatoids} is a polytopes whose vertices all lie in two parallel facets.
-A special case is a polytope of width $1$, whose vertices lie in two consecutive parallel lattice hyperplanes.
+This also implies the following two statements, that we prove in \Cref{sec:cayley}.
+A  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
+\paco{Wait; ``width one'' is only a special case of ``prismatoid'' if we allow prismatoids to have vertices in two parallel \emph{faces}, not \emph{facets}.
+But \Cref{coro:prismatoid} is not so obvious in that more general case.
+Either prove it or remove the phrase ``special case''}
+A special case is a polytope of width $1$, whose vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is equivalent to being $SL(\Z,d)$-equivalent to a Cayley sum.
 
@@
 \begin{corollary}
-Every integer dilation of a lattice $3$-polytope of width $1$ has a unimodular cover.
+\label{coro:width1}
+Every integer dilation $kP$, $k\ge 2$, of a lattice $3$-polytope $P$ of width $1$ has a unimodular cover.
 \end{corollary}
 
-We believe that the $3$-polytopes in all these statements all have unimodular triangulations, but this remains an open question.
+A special case of the latter are integer dilations of empty tetrahedra. That their dilations have
+unimodular covers is \cite[Cor.~4.2]{SantosZiegler} (and is also implicit in \cite{KantorSarkaria}).
+
+We believe that the $3$-polytopes in all these statements have unimodular triangulations, but this remains an open question.
 
@@
 We now turn our attention to \emph{prismatoids}, defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets.
-Since we are only interested in lattice polytopes, we here use \emph{prismatoid} meaning any $d$-polytope of the form
+In particular, a \emph{lattice prismatoid} is any $d$-polytope $SL(\Z,d)$-equivalent to one of the form
 \[
@@
 \]
-with $k\in \Z_{>0}$ and $Q_1,Q_2$ being lattice $(d-1)$-polytopes. This generalizes Cayley sums, which are the case $k=1$.
+where $Q_1,Q_2$ are lattice $(d-1)$-polytopes and $k\in \Z_{>0}$. This generalizes Cayley sums, which are the case $k=1$.
 
@@
 \label{prop:prismatoid}
-Let $Q_1$ and $Q_2$ be two lattice polygons and consider the prismatoid 
+Let $Q_1$, $Q_2$ be two lattice polygons and consider the prismatoid 
 \[
@@
 
-\Cref{coro:prismatoid} follows from \Cref{prop:prismatoid} since smoothness of a prismatoid implies $P\cap(\R^2\times\{i\})$ to be a lattice polytope for every $i$.
+This statement implies \Cref{coro:prismatoid} and \Cref{coro:width1}. 
+For the former, smoothness of a prismatoid implies $P\cap(\R^2\times\{1\})$ to be a lattice polytope. \paco{details?}
+For the latter, if $P$ has width one then $P\cong \cayley(Q_1, Q_2)$ and, for every $k\in \Z_{>0}$, we have
+\[
+kP \cap(\R^2\times\{1\}) = (k-1)Q_1 + Q_2.
+\]
 
@@
 \end{figure}
-\Giulia{fix it to look more unimodular at the bottom}
-\begin{figure}[htb]
-  \centering
-\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-  1]
-%coordinate axes
-   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%        \tkzLabelY[orig=false,label options={font=\tiny}]
-%        \tkzGrid
-%        \tkzDrawX
-%        \tkzDrawY
-
-%edges b, t
-  \node[
-%label={[shift={(0,-0.67)}]$b_1$}, 
-mynode] (b1) at (-3,-2) {}; 
-  \node[
-%label={[shift={(0,-0.67)}]$b_2$}, 
-mynode] (b2) at (2, 0.2) {};  
-  \node[
-%label={[shift={(0,0)}]$t_1$}, 
-mynode] (t1) at (-2.7, 1) {};  
-  \node[
-%label={[shift={(0,0)}]$t_2$}, 
-mynode] (t2) at (1.6,0.7) {};
-
-%shaded region where r can be  
-  \draw[fill=gray!80, nearly transparent] (2.93,0.61) -- (0,0.81)  -- (0,-0.68) -- cycle;
-
-%p
-  \node[
-%label={[shift={(0,-0.67)}]$p_1$}, 
-mynode] (p1) at (0,-1) {}; 
-  \node[
-%label={[shift={(0,0)}]$p_2$}, 
-mynode] (p2) at (0,1) {};  
-%  \node[{p}]  at (-0.3,0.95)
-
-%q
-  \node[
-%label={[shift={(0,-0.67)}]$q_1$}, 
-mynode] (q1) at (-1.7,-0.4) {}; 
-  \node[
-%label={[shift={(0,-0.67)}]$q_2$}, 
-mynode] (q2) at (0.3,-0.4) {};
-
-%segments b, t, p, q
-  \draw[ name path = b, black, thick] (b1) node[below, xshift=1.7cm, yshift=0.7cm] {b} -- (b2); 
-  \draw[name path = t, black, thick] (t1) node[below, xshift=1.2cm, yshift=-0.1cm] {t} -- (t2);
-  \draw[name path = p, black, thick] (p1) node[left, xshift=0 cm, yshift=1.3cm] {p} -- (p2);
-  \draw[name path = q, black, thick] (q1) node[below, xshift=0.4cm, yshift=0cm] {q} -- (q2);
-
-%prolonged lines through b and t
- \tkzDrawLine[add=0.2 and 0.3, color=black, dashed](b1, b2);
- \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);
-
-%functional f_q
-  \tkzDefLine[orthogonal=through p1](p1,p2)
-  \tkzDrawLine[add = 2 and 1,color=black, dashed](p1,tkzPointResult)
-  \node[label={\small $\{f_q(x)=-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab1) at (3.3, -1) {};
-
-  \tkzDefLine[orthogonal=through p2](p1,p2) 
-  \tkzDrawLine[add = 2 and 1,color=black, dashed](p2,tkzPointResult) 
-    \node[label={\small $\{f_q(x)=w-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab2) at (3.3, 1) {};
-
-\end{tikzpicture}
-\caption{Picture for Claim 4.2}
-\label{fig:claim2}
-\end{figure}
+
+%\Giulia{fix it to look more unimodular at the bottom}
+%\begin{figure}[htb]
+%  \centering
+%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
+%  1]
+%%coordinate axes
+%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
+%%        \tkzLabelX[orig=false,label options={font=\tiny}]
+%%        \tkzLabelY[orig=false,label options={font=\tiny}]
+%%        \tkzGrid
+%%        \tkzDrawX
+%%        \tkzDrawY
+%
+%%edges b, t
+%  \node[
+%%label={[shift={(0,-0.67)}]$b_1$}, 
+%mynode] (b1) at (-3,-2) {}; 
+%  \node[
+%%label={[shift={(0,-0.67)}]$b_2$}, 
+%mynode] (b2) at (2, 0.2) {};  
+%  \node[
+%%label={[shift={(0,0)}]$t_1$}, 
+%mynode] (t1) at (-2.7, 1) {};  
+%  \node[
+%%label={[shift={(0,0)}]$t_2$}, 
+%mynode] (t2) at (1.6,0.7) {};  
+%
+%%shaded region where r can be  
+%  \draw[fill=gray!80, nearly transparent] (2.93,0.61) -- (0,0.81)  -- (0,-0.68) -- cycle;
+%
+%
+%%p
+%  \node[
+%%label={[shift={(0,-0.67)}]$p_1$}, 
+%mynode] (p1) at (0,-1) {}; 
+%  \node[
+%%label={[shift={(0,0)}]$p_2$}, 
+%mynode] (p2) at (0,1) {};  
+%%  \node[{p}]  at (-0.3,0.95)
+%
+%
+%%q
+%  \node[
+%%label={[shift={(0,-0.67)}]$q_1$}, 
+%mynode] (q1) at (-1.7,-0.4) {}; 
+%  \node[
+%%label={[shift={(0,-0.67)}]$q_2$}, 
+%mynode] (q2) at (0.3,-0.4) {};  
+%
+%%segments b, t, p, q
+%  \draw[ name path = b, black, thick] (b1) node[below, xshift=1.7cm, yshift=0.7cm] {b} -- (b2); 
+%  \draw[name path = t, black, thick] (t1) node[below, xshift=1.2cm, yshift=-0.1cm] {t} -- (t2);
+%  \draw[name path = p, black, thick] (p1) node[left, xshift=0 cm, yshift=1.3cm] {p} -- (p2);
+%  \draw[name path = q, black, thick] (q1) node[below, xshift=0.4cm, yshift=0cm] {q} -- (q2);
+%
+%
+%%prolonged lines through b and t
+% \tkzDrawLine[add=0.2 and 0.3, color=black, dashed](b1, b2);
+% \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);
+% 
+%%functional f_q
+%  \tkzDefLine[orthogonal=through p1](p1,p2)
+%  \tkzDrawLine[add = 2 and 1,color=black, dashed](p1,tkzPointResult)
+%  \node[label={\small $\{f_q(x)=-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab1) at (3.3, -1) {};
+%
+%
+%  \tkzDefLine[orthogonal=through p2](p1,p2) 
+%  \tkzDrawLine[add = 2 and 1,color=black, dashed](p2,tkzPointResult) 
+%    \node[label={\small $\{f_q(x)=w-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab2) at (3.3, 1) {};
+%
+%\end{tikzpicture}
+%\caption{Picture for Claim 4.2}
+%\label{fig:claim2}
+%\end{figure}
 
@@
 J. A. De Loera, J. Rambau, F. Santos
-Triangulations: Structures for Algorithms and Applications, 539 pp.
+\emph{Triangulations: Structures for Algorithms and Applications}, 539 pp.
 Algorithms and Computation in Mathematics, Vol. 25, Springer-Verlag. 
@@
 Najmuddin Fakhruddin.
-\newblock Multiplication maps of linear systems on smooth projective toric
-  surfaces, arXiv:0208.5178, August 2002.
+\newblock Multiplication maps of linear systems on smooth projective toric surfaces.
+  Preprint, arXiv:0208.5178, August 2002.
 
@@
 Christian Haase, Benjamin Nill, Andreas Paffenholz, and Francisco Santos, Lattice points in Minkowski sums, 
-Electron J. Combin., 15 (2008), no. 1, Note 11, 5 pp.
+\emph{Electron.~J.~Combin.}, 15 (2008), no. 1, Note 11, 5 pp.
+
+\bibitem{KantorSarkaria}
+Jean-Michel Kantor and Karanbir~S.\ Sarkaria.
+On primitive subdivisions of an elementary tetrahedron, 
+\emph{Pacific J.~Math.} \textbf{211} (2003), no.~1,  123--155.
 
@@
 Shoetsu Ogata.
-\newblock {Multiplication maps of complete linear systems on projective toric
-  surfaces.}
+\newblock {Multiplication maps of complete linear systems on projective toric surfaces.}
 \newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006.
@@
 \bibitem{Santos-hirsch}
-\paco{add reference}
+Francisco Santos.
+A counter-example to the Hirsch Conjecture.
+\emph{Ann.~Math.~(2)}, 176 (July 2012), 383--412. 
+DOI: 10.4007/annals.2012.176.1.7
+
+\bibitem{SantosZiegler}
+Francisco Santos and G{\"u}nter~M. Ziegler.
+\newblock {Unimodular triangulations of dilated 3-polytopes}, 
+\newblock {\em Trans. Moscow Math. Soc.} (2013), 293--311. % \MR{3235802}
 

### 2019-07-08 unimodular-covers.tex
minor changes
minor change to figure
added picture for the setup
corrections and comments for sect 1 and 2

@@
 
-Another case of Oda's questions that has been solved is the two-dimensional case of \Cref{conj:Oda}\eqref{itm:mixedIDP}, with three different proofs by
-Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth. 
+Another case of Oda's questions that has been solved is the two-dimensional case of \Cref{conj:Oda}\eqref{itm:mixedIDP}, with three different proofs by Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth.
 In dimension three however the statement fails without the smoothness assumption, as shown by letting $P=Q$ be any non-unimodular empty tetrahedron.
 
-An alternative approach to \Cref{conj:Oda}\eqref{itm:mixedIDP} is via Cayley sums, that we discuss in  \Cref{sec:cayley}. 
+An alternative approach to \Cref{conj:Oda}\eqref{itm:mixedIDP} is via Cayley sums, which we discuss in  \Cref{sec:cayley}. 
 \paco{added def}
@@
 \]
-In particular, the following statement that we prove in \Cref{sec:cayley} is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:
+In particular, the following statement, proved in \Cref{sec:cayley}, is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:
 
@@
 
-This also implies the following two statements, that we prove in \Cref{sec:cayley}.
+This also implies the following two statements, also proved in \Cref{sec:cayley}.
 A  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
@@
 But \Cref{coro:prismatoid} is not so obvious in that more general case.
-Either prove it or remove the phrase ``special case''}
-A special case is a polytope of width $1$, whose vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is equivalent to being $SL(\Z,d)$-equivalent to a Cayley sum.
+Either prove it or remove the phrase ``special case''} \giulia{true...}
+A polytope has width $1$ if its vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is the same as being $SL(\Z,d)$-equivalent to a Cayley sum.
 
@@
 Observe that $u+v=p_1+p_2=q_3+q_4$.
-Now, the translation of vector $\left(\frac{a-1}2,\frac{b}2,1\right)$ sends $\conv(p_1p_2q_3q_4)$ to $\conv(q_2q_1p_4p_3)$, so that the translated points $u + \left(\frac{a-1}2,\frac{b}2,1\right)$ and $u + \left(\frac{a-1}2,\frac{b}2,1\right)$ lie in $\conv(q_2q_1p_4p_3)$ and are centrally symmetrically placed in it.
-thus, one of them belongs to $\conv(q_1p_4p_3)$ and the other to $\conv(q_2p_4p_3)$.
+Now, the translation of vector $\left(\frac{a-1}2,\frac{b}2,1\right)$ sends the facet $\conv(p_1p_2q_3q_4)$ of $C(T)$ to the opposite facet $\conv(q_2q_1p_4p_3)$, so that the translated points $u + \left(\frac{a-1}2,\frac{b}2,1\right)$ and $v + \left(\frac{a-1}2,\frac{b}2,1\right)$ lie in $\conv(q_2q_1p_4p_3)$ and are centrally symmetrically placed in it.
+Thus, one of them belongs to $\conv(q_1p_4p_3) \subset T_2$ and the other to $\conv(q_2p_4p_3) \subset T_1$.
 %\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
@@
 \paco{changed this}
-The parallelepiped $P$ is the intersection of three bands\paco{bands?}, where each band is a region defined as $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*$ and closed interval $[\alpha,\beta]\subset \R$.
-We claim that a band containing $T$ must contain at least three of the $q_i$s, which implies that the intersection $P$ of three of them contains at least one of the $q_i$s and hence $P$ contains the corresponding $T_i$.
-
-To prove the claim suppose that $q_1\not\in B:= f^{-1}([\alpha,\beta])$ for a certain band $B$. 
+The parallelepiped $P$ is the intersection of three regions of the form $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*$ and closed interval $[\alpha,\beta]\subset \R$.
+We will call such regions \emph{bands}.
+We claim that any band containing $T$ must contain at least three of the $q_i$s.
+This claim implies that the intersection $P$ of three bands contains at least one of the $q_i$s and hence $P$ contains the corresponding $T_i$.
+
+To prove the claim, suppose that $q_1\not\in B:= f^{-1}([\alpha,\beta])$ for a certain band $B \supset T$. 
 Without loss of generality, say $f(q_1)<\alpha$. Then the equalities $q_1+q_2=p_3+p_4$ and $q_1+p_1=q_2+p_2$ respectively give:
 \begin{gather}
+\label{eq:first}
 f(q_2) = f(p_3+p_4-q_1) = f(p_3)+f(p_4)-f(q_1) > 2\alpha-\alpha=\alpha,\\
+\label{eq:second}
 f(q_2) = f(q_2+p_2-p_1) = f(q_2)+(f(p_2)-f(p_1)) < \alpha + (\beta-\alpha) = \beta,
@@
 
-The first equation also implies
-\[
+The inequality \eqref{eq:first} also implies
+\begin{equation}
+\label{eq:third}
 f(q_1) <  f(p_i) < f(q_2), \quad\text{ for $i=3,4$}.
-\]
-By the translation sending $q_1,q_2,p_3,p_4$ to $p_2,p_1,q_4,q_3$ (in this order) we then have
+\end{equation}
+
+The translation of vector $\frac12 (p_1+p_2-p_3-p_4)$ sends $q_1,q_2,p_3,p_4$ to $p_2,p_1,q_4,q_3$ (in this order). By applying this to inequality \eqref{eq:third}, we obtain
 \[
@@
 Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is
-contained in $P$''. If this holds then $P$ has a unimodular cover, since then the proofs of \Cref{coro:coverpara} and \Cref{thm:parallelepipeds} work for $P$.
+contained in $P$''. 
+If this holds then $P$ has a unimodular cover, since then the proofs of \Cref{coro:coverpara} and \Cref{thm:parallelepipeds} work for $P$.
 
@@
 \scalebox{.75}{\input{strips.pdf_t}}
-\caption{???}
+\caption{The strips  of Lemma \ref{lemma:cayley}}
 \label{fig:strips}
@@
 \] \giulia{we want a picture for the setup}
+
+\begin{figure}
+\includegraphics[scale=.3]{setup.png}
+\caption{Setup for \Cref{lemma:cayley}}
+\end{figure}
 

### 2019-07-09 unimodular-covers.tex
small corrections

@@
 
-The inequality \eqref{eq:first} also implies
+Inequality \eqref{eq:first} also implies
 \begin{equation}
@@
 \paco{added this intermediate result}
-Let $T$ be a lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
+\giulia{We are using lemmas 2 and 3 which are for empty tetrahedra; either we add a step in this corollary saying that we can reduce to empty tetrahedra, or we keep this for empty and triangulate into empty tetrah in the proof of the theorem (which makes more sense to me)}
+Let $T$ be a empty lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
 \end{corollary}
@@
 If this volume equals $1$ then $T$ is unimodular and there is nothing to prove, so we assume $T$ is not unimodular.
+As before, assume $T$ has vertices $p_1, p_2, p_3, p_4$.
 
@@
 perhaps it should be made explicit in the statemente of the lemma...}
-\paco{removed assumption the $u$ lies in a face. I think we do not need it}
-
-Then $S=\conv(T\cup \{l\})$ can be triangulated in two different ways: $S=T \cup T'_4 = S_1 \cup S_2 \cup S_3$, where $T'_4 = \conv(p_1, p_2, p_3, l) \subseteq T_4$ and
-\[
-S_1= \conv(p_2,p_3,p_4, l),
-S_2=\conv(p_1,p_3,p_4, l),
-S_3=\conv(p_1,p_2,p_4, l).
+
+Then $S=\conv(T\cup \{u\})$ can be triangulated in two different ways: $S=T \cup T'_4$, where $T'_4 = \conv(p_1, p_2, p_3, u) \subseteq T_4$ and $S= S_1 \cup S_2 \cup S_3$, with
+\[
+S_1= \conv(p_2,p_3,p_4, u),
+S_2=\conv(p_1,p_3,p_4, u),
+S_3=\conv(p_1,p_2,p_4, u).
 \]
@@
 
-Observe that  $T'_4$ and/or one of the $S_i$ can degenerate to a triangle, if $u$ lies in $\conv(p_1,p_2,p_3)$, 
-but this does not affect the argument.
+%Observe that  $T'_4$ and/or one of the $S_i$ can degenerate to a triangle, if $u$ lies in $\conv(p_1,p_2,p_3)$, but this does not affect the argument.
 %
@@
 
-\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
+\begin{proof}[Proof of \Cref{thm:parallelepipeds}]\giulia{see comment above}
 Arbitrarily triangulate the parallelepiped into lattice tetrahedra and apply \Cref{coro:coverpara} to these tetrahedra.
@@
 By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point, which we denote by $u$.
-Observe that from the proof of \Cref{lemma:corner} we know more: $u$ lies in the face $\conv(p_1, p_2, q_4)$.\giulia{we need this for the volume computation below.
-perhaps it should be made explicit in the statemente of the lemma...}
 %
@@
 The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
-Hence, they are not IDP, whcih implies they do not admit unimodular covers.
-\giulia{is it necessary to be more explicit, for example show that these are the only lattice points?}
-\paco{I think it is ok as is. It is only an example}
+Hence, they are not IDP, which implies they do not admit unimodular covers.
 \end{example}
@@
 
-Throughout this section let $P$ and $Q$ be two lattice polytopes in $\R^d$.
+Throughout this section \giulia{I think this may be misleading, in the sense that later we specify to $d=2$ and we should maybe highlight this} let $P$ and $Q$ be two lattice polytopes in $\R^d$.
 We do not require them to be full-dimensional, but we assume their Minkowski sum is.
@@
 
-We introduce the following two concepts.
-
 \begin{definition}
@@
 We say that the pair $(P,Q)$ is \emph{mixed IDP} if 
-\paco{``mixed IDP'' is not the right property, since this is weaker than IDP}
+\paco{``mixed IDP'' is not the right property, since this is weaker than IDP} \giulia{do we give it a name at all then?}
 \[

### 2019-07-12 unimodular-covers.tex
adding  some questions

@@
 \begin{definition}
-In the above conditions, we say that $Q$ \emph{is nef for $P$} if the normal fan of $Q$ refines that of $P$. If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.
+In the above conditions, we say that $Q$ \emph{is a weak Minkowski summand of $P$} if the normal fan of $Q$ refines that of $P$. 
+If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.\giulia{I guess there's no nice term for this now...}
 
@@
 which induces the following canonical bijections:
+
+\begin{center}
+  \begin{tabular}{ c | c }
+    $\cayley(P,Q)$ & $P+Q$\\
+    \hline
+   polyhedral subdivisions & mixed subdivisions \\
+    triangulations & fine mixed subdivisions \\
+   unimodular simplices & unimodular prod-simplices \\
+  \end{tabular}
+\end{center}\giulia{not sure if the table makes it clear that these are bijections}
+
 \begin{align*}
@@
 \begin{figure}
-\includegraphics[scale=.3]{setup.png}
+\includegraphics[scale=.3]{setup'.png}
 \caption{Setup for \Cref{lemma:cayley}}

### 2019-07-13 unimodular-covers.tex
added grunbaum and cls refs

@@
 \end{align*}
-\paco{perhaps we can call this formula ``being mixed IDP''. The property of ``$N(Q)$ refines $N(P)$'' is sometimes called ``$P$ is a Minkowski summand of $Q$''.}
+
+The condition that the normal fan of $Q$ is finer than that of $P$ is sometimes dubbed \emph{$P$ is a weak Minkowski summand of $Q$}, since it is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$~\cite[??]{Grunbaum}.
+It has the following algebraic implication for the projective toric variety $X_Q$ corresponding to (the normal fan of) $Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the corresponding Cartier divisor $D_P$ on  $X_Q$ is \emph{numerically effective}, or ``nef''~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}; (observe that~\cite{CLS} use ``Minkowski summand'' meaning what we call ``weak Minkowski summand'').
+\paco{added this explanation}
 \end{enumerate}
@@
 
-\section{nef Cayley sums}
+\section{ Cayley sums}
 \label{sec:cayley}
@@
 \]
-The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ is a $(d+1)$-dimensional.
+The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ being $(d+1)$-dimensional.
 
@@
 If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.\giulia{I guess there's no nice term for this now...}
+\paco{not sure; maybe we can keep nef after all; another (unclear) option could be ``weak prism over $Q$'' or ``degenerate prism over $Q$''.}
 
@@
 \paco{``mixed IDP'' is not the right property, since this is weaker than IDP} \giulia{do we give it a name at all then?}
+\paco{perhaps mixed-IDP}
 \[
@@
 \begin{figure}
-\includegraphics[scale=.3]{setup'.png}
+\includegraphics[scale=.3]{setup.png}
 \caption{Setup for \Cref{lemma:cayley}}
@@
 
+\bibitem{CLS}
+David A.~Cox, John B.~Little, and Hal K.~Schenck.
+{\em {T}oric {V}arieties}.
+AMS, Providence, 2011.
+
  \bibitem{DLRS2010}
@@
   Preprint, arXiv:0208.5178, August 2002.
+
+\bibitem{Grunbaum}
+{Branko Gr\"unbaum}.
+{\em {C}onvex {P}olytopes}.
+Wiley, London, 1967.
 

### 2019-07-13 unimodular-covers.tex
checking comments

@@
 
-\paco{define mixed IDP, Cayley sums, Minkowski sums, mixed subdivisions in the intro? Probably not}
 A lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice. 
@@
 \end{align*}
-
-The condition that the normal fan of $Q$ is finer than that of $P$ is sometimes dubbed \emph{$P$ is a weak Minkowski summand of $Q$}, since it is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$~\cite[??]{Grunbaum}.
+\end{enumerate}
+\end{conjecture}
+
+Motivated by  \Cref{conj:Oda}\eqref{itm:mixedIDP}, we recall the following definitions:
+
+\begin{definition}
+We say that $P$ is a weak Minkowski summand of $Q$ if the normal fan of $Q$ refines that of $P$.
+\end{definition}
+The name is due to the fact that the condition on the refinement of the normal fans is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$~\cite[??]{Grunbaum}.
 It has the following algebraic implication for the projective toric variety $X_Q$ corresponding to (the normal fan of) $Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the corresponding Cartier divisor $D_P$ on  $X_Q$ is \emph{numerically effective}, or ``nef''~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}; (observe that~\cite{CLS} use ``Minkowski summand'' meaning what we call ``weak Minkowski summand'').
 \paco{added this explanation}
-\end{enumerate}
-\end{conjecture}
+
+\begin{definition}
+The pair $(P, Q)$ has the integer decomposition property, or is IDP, if 
+\begin{align*}
+(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
+\end{align*}
+\end{definition}
+
+In this language,  \Cref{conj:Oda}\eqref{itm:mixedIDP} can be rephrased as if $P$ is a weak Minkowski summand of $Q$ and $Q$ is smooth, then the pair $(P,Q)$ is IDP.
 

### 2019-07-13 unimodular-covers.tex
checking comments

@@
 \begin{definition}
-We say that $P$ is a weak Minkowski summand of $Q$ if the normal fan of $Q$ refines that of $P$.
+We say that $P$ is a \emph{weak Minkowski summand} of $Q$ if the normal fan of $Q$ refines that of $P$.
 \end{definition}
@@
 \begin{definition}
-The pair $(P, Q)$ has the integer decomposition property, or is IDP, if 
+The pair $(P, Q)$ has the integer decomposition property, or \emph{the pair $(P,Q)$ is IDP}, if 
 \begin{align*}
@@
 
-In this language,  \Cref{conj:Oda}\eqref{itm:mixedIDP} can be rephrased as if $P$ is a weak Minkowski summand of $Q$ and $Q$ is smooth, then the pair $(P,Q)$ is IDP.
+In this language,  \Cref{conj:Oda}\eqref{itm:mixedIDP} can be rephrased as: if $P$ is a weak Minkowski summand of $Q$ and $Q$ is smooth, then the pair $(P,Q)$ is IDP.
 
@@
 A  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
-\paco{Wait; ``width one'' is only a special case of ``prismatoid'' if we allow prismatoids to have vertices in two parallel \emph{faces}, not \emph{facets}.
-But \Cref{coro:prismatoid} is not so obvious in that more general case.
-Either prove it or remove the phrase ``special case''} \giulia{true...}
 A polytope has width $1$ if its vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is the same as being $SL(\Z,d)$-equivalent to a Cayley sum.
@@
 \label{lemma:3<4}
-Let $P$ be a lattice parallelepiped and let $T\subset P$ be an empty tetrahedron.
-Then, at least one of the four corner tetrahedra $T_i$ of the circumscribed parallelogram $C(T)$ is fully contained in $P$.
+Let $P$ be a lattice parallelepiped and let $T\subset P$ be a tetrahedron. Then, at least one of the four corner tetrahedra $T_i$ of the circumscribed parallelogram $C(T)$ is fully contained in $P$.
 \end{lemma}
@@
 Thus, $S_1$, $S_2$ and $S_3$ cover $T$ and have volume strictly smaller than $T$.
-By inductive hypothesis they can be covered unimodularly, and we are done,
+The $S_i$ may not be empty, but we can triangulate them into empty tetrahedra, which by inductive hypothesis they can be covered unimodularly, and we are done.
 
@@
 \begin{proof}[Proof of \Cref{thm:parallelepipeds}]\giulia{see comment above}
-Arbitrarily triangulate the parallelepiped into lattice tetrahedra and apply \Cref{coro:coverpara} to these tetrahedra.
+Arbitrarily triangulate the parallelepiped into empty lattice tetrahedra and apply \Cref{coro:coverpara} to these tetrahedra.
 %Let $P \subseteq \R^3$ be a lattice parallelepiped.

### 2019-07-13 unimodular-covers.tex
checking comments

@@
 \label{coro:coverpara}
-\paco{added this intermediate result}
-\giulia{We are using lemmas 2 and 3 which are for empty tetrahedra; either we add a step in this corollary saying that we can reduce to empty tetrahedra, or we keep this for empty and triangulate into empty tetrah in the proof of the theorem (which makes more sense to me)}
 Let $T$ be a empty lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
@@
 
-\section{ Cayley sums}
+\section{ Cayley sums} \giulia{Change notations: remove nef and make sure everything works}
 \label{sec:cayley}
-
-Throughout this section \giulia{I think this may be misleading, in the sense that later we specify to $d=2$ and we should maybe highlight this} let $P$ and $Q$ be two lattice polytopes in $\R^d$.
+Let $P$ and $Q$ be two lattice polytopes in $\R^d$.
 We do not require them to be full-dimensional, but we assume their Minkowski sum is.
@@
 
+\medskip
+\giulia{move first paragraph of proof of theorem here, and explain simplices, prodsimplices...}
+
+We now turn our attention to dimension $d=2$.
+The unimodular prod-simplices are unimodular parallelograms $p+q$ with $p$ a segment of $P$ and $q$ a segment of $Q$, and unimodular triangles $T+u$, where $T \subset P$ and $u \in Q$ or viceversa.
+
+Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$.
+Observe that any simplex in $T$ has vertices in $P \times \{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$.
+Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$).
+
+\giulia{fix this}
+
 %As a consequence, we also have that:
@@
 \begin{proof}[Proof of \Cref{thm:cayley}]
-Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$.
-Observe that any simplex in $T$ has vertices in $Px\{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$.
-Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$).
-\giulia{is it usually denoted by dimension or cardinality?
-that is, (2,0) or (3,1)?} A triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained by simply unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$).
+
+A unimodular triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained simply by unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$).
 This leaves simplices of type $(2,2)$.
@@
 
-\begin{claim}\giulia{I think we need to treat separately the case where we deal with edges (these are easier, in any case, but I couldn't find a unification)}
+\begin{claim}
 The following inequalities hold, 
@@
 \end{proof}
-\giulia{add a picture!}
+\giulia{add a picture?}
+
+\Giulia{add references to Haase&Hofmann and Tsuchiya}
 

### 2019-07-13 unimodular-covers.tex
checking comments
several edits, mostly intro and Section 3

@@
 
-A lattice polytope is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice. 
-%
-Oda~\cite{Oda1997} posed several questions regarding smoothness and IDP for lattice polytopes. The following versions of them are now considered conjectures~\cite{HNPS2008,mfo2007}:
+Oda~\cite{Oda1997} posed several questions regarding smoothness and the IDP property for lattice polytopes.
+%
+\paco{I think this definition goes better here, and decided to "undeclare" it and the definition of weak Mink sum}
+Following \cite{HaaseHof, Tsuchiya} we say that a pair $(P, Q)$ of lattice polytopes has the integer decomposition property, or that \emph{the pair $(P,Q)$ is IDP}, if 
+\begin{align*}
+\label{eq:mixedIDP}
+(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
+\end{align*}
+%
+A lattice polytope $Q$ is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth.
+The following versions of Oda's questions are now considered conjectures~\cite{HNPS2008,mfo2007}, and they are open even in dimension three:
 \begin{conjecture}
@@
 \label{itm:mixedIDP}
-(Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Let $P, Q\subset \R^d$ be lattice polytopes with the normal fan of $P$ refining that of $Q$, and $Q$ smooth. Then
-\begin{align*}
-\label{eq:mixedIDP}
-(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
-\end{align*}
+(Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Every pair $(P,Q)$ of lattice polytopes with $Q$ smooth and the normal fan of $Q$ refining that of $P$ is IDP.
+%\begin{align}
+%\label{eq:mixedIDP}
+%(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
+%\end{align}
 \end{enumerate}
@@
 
-Motivated by  \Cref{conj:Oda}\eqref{itm:mixedIDP}, we recall the following definitions:
-
-\begin{definition}
-We say that $P$ is a \emph{weak Minkowski summand} of $Q$ if the normal fan of $Q$ refines that of $P$.
-\end{definition}
-The name is due to the fact that the condition on the refinement of the normal fans is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$~\cite[??]{Grunbaum}.
-It has the following algebraic implication for the projective toric variety $X_Q$ corresponding to (the normal fan of) $Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the corresponding Cartier divisor $D_P$ on  $X_Q$ is \emph{numerically effective}, or ``nef''~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}; (observe that~\cite{CLS} use ``Minkowski summand'' meaning what we call ``weak Minkowski summand'').
-\paco{added this explanation}
-
-\begin{definition}
-The pair $(P, Q)$ has the integer decomposition property, or \emph{the pair $(P,Q)$ is IDP}, if 
-\begin{align*}
-(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
-\end{align*}
-\end{definition}
-
-In this language,  \Cref{conj:Oda}\eqref{itm:mixedIDP} can be rephrased as: if $P$ is a weak Minkowski summand of $Q$ and $Q$ is smooth, then the pair $(P,Q)$ is IDP.
-
+When the normal fan of a $Q$ refines that of $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$\paco{did not find this in the google books version of \cite{Grunbaum}; found it in Postnikov-Reiner-Williams, Tsuchiya, a survey by Rekha Thomas, and other places.
+Perhaps we can leave it wothout reference}.
+This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor on  $X_Q$ defined by $P$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that there ``Minkowski summand'' is used meaning what we call ``weak Minkowski summand'').
+%\paco{added this explanation}
+%
+%\begin{definition}
+%Following \cite{HaaseHof, Tsuchiya} we say that the pair $(P, Q)$ has the integer decomposition property, or \emph{the pair $(P,Q)$ is IDP}, if 
+%\begin{align*}
+%(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
+%\end{align*}
+%\end{definition}
+% 
+%In this language,  \Cref{conj:Oda}\eqref{itm:mixedIDP} can be rephrased as: if $P$ is a weak Minkowski summand of a smooth polytope $Q$ then the pair $(P,Q)$ is IDP.
+
+\medskip 
 Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes. 
@@
 
-Another case of Oda's questions that has been solved is the two-dimensional case of \Cref{conj:Oda}\eqref{itm:mixedIDP}, with three different proofs by Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}, the latter without assuming $P$ smooth.
-In dimension three however the statement fails without the smoothness assumption, as shown by letting $P=Q$ be any non-unimodular empty tetrahedron.
+The two-dimensional case of \Cref{conj:Oda}\eqref{itm:mixedIDP} is known to hold, with three different proofs by Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}.
+This last one actually shows that smoothness of $Q$ is not needed.
+In dimension three, however, the conjecture fails without the smoothness assumption, as shown by letting $P=Q$ be any non-unimodular \emph{empty tetrahedron}, by which we mean a lattice tetrahedron with no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for details on them).
+\paco{added def of empty tetrahedron}
 
 An alternative approach to \Cref{conj:Oda}\eqref{itm:mixedIDP} is via Cayley sums, which we discuss in  \Cref{sec:cayley}. 
-\paco{added def}
 Recall that the \emph{Cayley sum} of two lattice polytopes $Q_1,Q_2\subset \R^d$ is the lattice polytope
@@
 \]
-\paco{new sentence. In fact, this implies that Cayley sums are the same as polytopes of width 1}
-We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can remove coordinates) but this does not need $Q_1$ and $Q_2$ to be full-dimensional.
+We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but for this does not need $Q_1$ and $Q_2$ to be full-dimensional.
 It only requires the linear subspaces parallel to them to span $\R^d$.
 
-As we note there (\Cref{prop:mixedIDP}), if $P,Q\subset \R^d$ are two lattice polytopes one has
-\[
-\cayley(P,Q) \text{ is IDP} 
-\Rightarrow
-(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
-\]
-In particular, the following statement, proved in \Cref{sec:cayley}, is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:
+As we note in \Cref{prop:mixedIDP}, if the Cayley sum of $P$ and $Q$ is IDP then the pair $(P,Q)$ is IDP.
+In particular, the following statement from \Cref{sec:cayley} is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:
 
@@
 
-This also implies the following two statements, also proved in \Cref{sec:cayley}.
+This has the following two consequences, also proved in \Cref{sec:cayley}.
 A  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
-A polytope has width $1$ if its vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is the same as being $SL(\Z,d)$-equivalent to a Cayley sum.
+A polytope has width $1$ if its vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is the same as being ($SL(\Z,d)$-equivalent to) a Cayley sum.
 
@@
 
+\medskip
 We believe that the $3$-polytopes in all these statements have unimodular triangulations, but this remains an open question.
@@
 \label{def:circunpara}
-Let $T$ be a tetrahedron with vertices $p_1$, $p_2$, $p_3$, and $p_4$. Consider the points $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$, $i\ n [4]$, and let
+Let $T$ be a tetrahedron with vertices $p_1$, $p_2$, $p_3$, and $p_4$. Consider the points $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$, $i\in [4]$, and let
 \[
@@
 
-The idea to have in mind is that of the regular tetrahedron inscribed in a cube. Modulo an affine transformation this is exactly the situation of $T$ and $C(T)$.
+The idea to have in mind is that of the regular tetrahedron inscribed in a cube. 
+Modulo an affine transformation this is exactly the situation of $T$ and $C(T)$.
 \giulia{we should add a picture of the standard cube}
@@
 \paco{changed this}
-The parallelepiped $P$ is the intersection of three regions of the form $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*$ and closed interval $[\alpha,\beta]\subset \R$.
-We will call such regions \emph{bands}.
+We call \emph{band} any region of the form $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*$ and closed interval $[\alpha,\beta]\subset \R$.
 We claim that any band containing $T$ must contain at least three of the $q_i$s. 
-This claim implies that the intersection $P$ of three bands contains at least one of the $q_i$s and hence $P$ contains the corresponding $T_i$.
+This claim implies that the parallelepiped $P$, which is the intersection of three bands, contains at least one of the $q_i$s and hence it fully contains the corresponding $T_i$.
 
@@
 \label{coro:coverpara}
-Let $T$ be a empty lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
+Let $T$ be an empty lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
 \end{corollary}
@@
 If this volume equals $1$ then $T$ is unimodular and there is nothing to prove, so we assume $T$ is not unimodular.
-As before, assume $T$ has vertices $p_1, p_2, p_3, p_4$.
+Let $p_1, p_2, p_3, p_4$ denote the vertices of $T$.
 
@@
 perhaps it should be made explicit in the statemente of the lemma...}
-
+%
 Then $S=\conv(T\cup \{u\})$ can be triangulated in two different ways: $S=T \cup T'_4$, where $T'_4 = \conv(p_1, p_2, p_3, u) \subseteq T_4$ and $S= S_1 \cup S_2 \cup S_3$, with
@@
 \]
-Each of the tetrahedra $S_i$ has lattice volume strictly smaller than $T$, because, for each $i$, $p_i$ is the unique point of $C(T)$ maximizing the distance to the opposite facet $\conv(p_j,p_k,p_l)$ of $T$.
+
+Each of the tetrahedra $S_i$ has lattice volume strictly smaller than $T$ because, for each $i$, $p_i$ is the unique point of $C(T)$ maximizing the distance to the opposite facet $\conv(p_j,p_k,p_l)$ of $T$.
 Thus, $S_1$, $S_2$ and $S_3$ cover $T$ and have volume strictly smaller than $T$.
-The $S_i$ may not be empty, but we can triangulate them into empty tetrahedra, which by inductive hypothesis they can be covered unimodularly, and we are done.
-
+The $S_i$ may not be empty, but we can triangulate them into empty tetrahedra, which by inductive hypothesis they can be covered unimodularly.
+%
 %Observe that  $T'_4$ and/or one of the $S_i$ can degenerate to a triangle, if $u$ lies in $\conv(p_1,p_2,p_3)$, but this does not affect the argument.
@@
 
-We finish with a question which, if answered in the positive, would imply \Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three.
+%We finish with a question which, if answered in the positive, would imply \Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three.
 Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is
@@
 If this holds then $P$ has a unimodular cover, since then the proofs of \Cref{coro:coverpara} and \Cref{thm:parallelepipeds} work for $P$.
+Hence, a positive answer to the following question would imply that every smooth $3$-polytope has a unimodular cover, which in turn implies \Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three.
 
 \begin{question}
-Does every smooth 3-polytope have the circumscribed parallelepiped property? By the above remark, if the answer is yes then every smooth $3$-polytope has a unimodular cover.
+Does every smooth 3-polytope have the circumscribed parallelepiped property? 
 \end{question}
@@
 We do not require them to be full-dimensional, but we assume their Minkowski sum is.
-Remember that the Minkowski sum $P+Q$ of $P$ and $Q$ is defined as:
-\[
-P + Q := \{ p+q \in \R^d: p\in P, q \in Q\}.
-\]
-The \emph{Cayley sum} of $P$ and $Q$ is defined as
-\[
+Remember that the \emph{Minkowski sum} $P+Q$ and the \emph{Cayley sum} of $P$ and $Q$ are defined as:
+\begin{gather*}
+P + Q := \{ p+q \in \R^d: p\in P, q \in Q\}.\\
 \cayley (P,Q) = \conv( P\times\{0\} \cup Q\times \{1\}) \in \R^{d+1}.
-\]
-The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ being $(d+1)$-dimensional.
-
-\begin{definition}
-In the above conditions, we say that $Q$ \emph{is a weak Minkowski summand of $P$} if the normal fan of $Q$ refines that of $P$. 
-If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.\giulia{I guess there's no nice term for this now...}
-\paco{not sure; maybe we can keep nef after all; another (unclear) option could be ``weak prism over $Q$'' or ``degenerate prism over $Q$''.}
-
-We say that the pair $(P,Q)$ is \emph{mixed IDP} if 
-\paco{``mixed IDP'' is not the right property, since this is weaker than IDP} \giulia{do we give it a name at all then?}
-\paco{perhaps mixed-IDP}
-\[
-(P\cap \Z^d) + (Q\cap \Z^d) = (P+Q)\cap \Z^d.
-\]
-\end{definition}
-
-In this language, \Cref{conj:Oda}\eqref{itm:mixedIDP} translates to:
-\emph{If $Q$ is nef for $P$ and $P$ is smooth then the pair $(P,Q)$ mixed IDP.}
-
-\begin{proposition}
+\end{gather*}
+
+\paco{removed definitions of "nef cayley", "mixed IDP" and "weak Minkowski sum" from here}
+%\begin{definition}
+%In the above conditions, we say that $Q$ \emph{is a weak Minkowski summand of $P$} if the normal fan of $Q$ refines that of $P$. 
+%If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.\giulia{I guess there's no nice term for this now...}
+%\paco{not sure; maybe we can keep nef after all; another (unclear) option could be ``weak prism over $Q$'' or ``degenerate prism over $Q$''.}
+%
+%We say that the pair $(P,Q)$ is \emph{mixed IDP} if 
+%\paco{``mixed IDP'' is not the right property, since this is weaker than IDP} \giulia{do we give it a name at all then?}
+%\paco{perhaps mixed-IDP}
+%\[
+%(P\cap \Z^d) + (Q\cap \Z^d) = (P+Q)\cap \Z^d.
+%\]
+%\end{definition}
+%
+%In this language, \Cref{conj:Oda}\eqref{itm:mixedIDP} translates to:
+%\emph{If $Q$ is nef for $P$ and $P$ is smooth then the pair $(P,Q)$ mixed IDP.}
+%
+The so-called \emph{Cayley Trick} is the isomorphism
+\[
+2\cayley(P, Q) \cap (\R^d\times \{1\}) \cong P+Q,
+\]
+which easily implies:
+
+\begin{proposition}[see, e.g.~\protect{\cite[Thm.~0.4]{Tsuchiya}}]
 \label{prop:mixedIDP}
@@
 
-The so-called \emph{Cayley Trick} is the isomorphism
-\[
-2\cayley(P, Q) \cap (\R^d\times \{1\}) \cong P+Q,
-\]
-which induces the following canonical bijections:
+%The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ being $(d+1)$-dimensional.
+
+The Cayley Trick also provides the following canonical bijections:
 
 \begin{center}
-  \begin{tabular}{ c | c }
-    $\cayley(P,Q)$ & $P+Q$\\
+  \begin{tabular}{ c c c }
+    $\cayley(P,Q)$ && $P+Q$\\
     \hline
-   polyhedral subdivisions & mixed subdivisions \\
-    triangulations & fine mixed subdivisions \\
-   unimodular simplices & unimodular prod-simplices \\
+   polyhedral subdivisions &$\leftrightarrow$& mixed subdivisions \\
+    triangulations &$\leftrightarrow$& fine mixed subdivisions \\
+   unimodular simplices &$\leftrightarrow$& unimodular prod-simplices \\
   \end{tabular}
@@
 
-\begin{align*}
-\text{polyhedral subdivisions of $\cayley(P,Q)$} &\leftrightarrow \text{mixed subdivisions of $P + Q$}\\
-\text{triangulations of $\cayley(P,Q)$} &\leftrightarrow \text{fine mixed subdivisions of $P + Q$}\\
-\text{unimodular simplices in $\cayley(P,Q)$} &\leftrightarrow \text{unimodular prod-simplices in $P + Q$}.
-\end{align*}
+\[
+\begin{array}{ccc}
+\text{polyhedral subdivisions of $\cayley(P,Q)$} &\leftrightarrow& \text{mixed subdivisions of $P + Q$}\\
+\text{triangulations of $\cayley(P,Q)$} &\leftrightarrow& \text{fine mixed subdivisions of $P + Q$}\\
+\text{unimodular simplices in $\cayley(P,Q)$} &\leftrightarrow& \text{unimodular prod-simplices in $P + Q$}.
+\end{array}
+\]
 See \cite{DLRS2010} for more details on the Cayley Trick and on triangulations and polyhedral subdivisions of polytopes.
@@
 
-We now turn our attention to dimension $d=2$.
-The unimodular prod-simplices are unimodular parallelograms $p+q$ with $p$ a segment of $P$ and $q$ a segment of $Q$, and unimodular triangles $T+u$, where $T \subset P$ and $u \in Q$ or viceversa.
-
-Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$.
+We now turn our attention to $d=2$, in order to prove \Cref{thm:cayley}.
+In a triangulation of $\cayley(P,Q)\subset \R^3$ there are tetrahedra of types $(1,3)$, $(2,2)$ and $(3,1)$, where the type denotes how many vertices they have in $P$ and in $Q$.
+Empty tetrahedra of types $(1,3)$ or $(3,1)$ are automatically unimodular, so we only need to care with those of type $(2,2)$.
+These are the Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
+%
+%
+%The unimodular prod-simplices are unimodular parallelograms $p+q$ with $p$ a segment of $P$ and $q$ a segment of $Q$, and unimodular triangles $T+u$, where $T \subset P$ and $u \in Q$ or viceversa. 
+%
+%Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$.
 Observe that any simplex in $T$ has vertices in $P \times \{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$.
 Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$).
-
+%
 \giulia{fix this}
-
+\paco{done, but check it}
+%
 %As a consequence, we also have that:
@@
 %\end{align*}
-
-Our proof of \Cref{thm:cayley} is based on the following result, the proof of which we postpone to \Cref{sec:the_lemma}. 
+For this the following result, the proof of which we postpone to \Cref{sec:the_lemma}, is crucial. 
 The lemma uses the following conventions.
@@
 \label{lemma:cayley}
-Let $P$ and $Q$ be two lattice polygons with $Q$ nef with respect to $P$.
+\paco{$Q$ is 2-dim, but $P$ may not}
+Let $Q$ be a two-dimensional lattice polytope and $P$ a weak Minkowski summand of it. 
 Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the lines spanned by them.
@@
 \end{lemma}
-\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in the interior of $\{p_1, p_2\} + \{q_1, q_2\}$ (or symmetric for $Q$)}
+%\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in the interior of $\{p_1, p_2\} + \{q_1, q_2\}$ (or symmetric for $Q$)}
 
@@
 
-\begin{proof}[Proof of \Cref{thm:cayley}]
-
-A unimodular triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained simply by unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$).
+\begin{corollary}
+\label{coro:covercayley}
+\paco{inserted this corollary to make proof clearer, and to parallel section 2}
+Let $T$ be an empty lattice tetrahedron contained in the Cayley sum $\cayley(P,Q)$, where $P$ is a weak Minkowski summand of $Q$.
+Then, $T$ can be covered by unimodular tetrahedra contained in $\cayley(P,Q)$.
+\end{corollary}
+
+\begin{proof}
+The proof is by induction on the normalized volume of $T$, which we assume to be at least $2$. This implies that $T$ is of type $(2,2)$,
+since empty tetrahedra of types $(1,3)$ and $(3,1)$ are unimodular. Thus, $T$ is the Cayley sum of primitive segments $p=[p_1,p_2]\subset P$ and $q=[q_1,q_2]\subset Q$.  
+Let $u$ be the lattice point whose existence is guaranteed by \Cref{lemma:cayley}. Assume  (the other case is similar) that 
+\[
+u \in ((p_1, p_2) + \vecline q ) \cap P,
+\]
+and call $t$ the triangle $t=\conv( u,  p_1,  p_2)\subset P$.
+
+Let us denote $\tilde u$, $\tilde p_1$, $\tilde p_2$, $\tilde q_1$, $\tilde q_2$ the points corresponding to $u, p_1, p_2, q_1, q_2$ in $\cayley(P,Q)$.
+That is, $\tilde p_i = p\times\{0\}$, $\tilde q_i = p\times\{1\}$, and $\tilde u = u\times\{0\}$.
+Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that of the segments $[u,q_i]$ crosses the triangle $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$.
+In turn, this means that the polytope $\conv(\tilde u, \tilde p_1, \tilde p_2, \tilde q_1, \tilde q_2) = \cayley(t,q)$
+ has the following two triangulations:
+\begin{gather*}
+\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_i\}) \right\},
+\\
+\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_j\}) \}.
+\end{gather*}
+The tetrahedra $\cayley(t, \{q_j\})$ and $\cayley(t, \{q_i\})$ are unimodular, which implies that $T=\cayley(p,q)$ has volume equal to the sum of the volumes of $\cayley([p_1,u],q)$ and $\cayley([p_2,u],q)$.
+In particular, we have covered $T$ by the three tetrahedra in $\mathcal T^-$, which are of smaller volume and hence have unimodular covers by inductive assumption.
+%
+%A unimodular triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained simply by unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$).
 This leaves simplices of type $(2,2)$.
 In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.
-
-Let $S=\{p_1, p_2\} + \{q_1, q_2\}$ be such a parallelogram.
+%
+%Let $S=\{p_1, p_2\} + \{q_1, q_2\}$ be such a parallelogram.
 By \Cref{lemma:cayley} we know that there is a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $S$, w.l.o.g.
@@
 
-We now turn our attention to \emph{prismatoids}, defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets.
+\begin{proof}[Proof of \Cref{thm:cayley}]
+Arbitrarily triangulate $\cayley(P,Q)$ into empty lattice tetrahedra and apply \Cref{coro:covercayley} to these tetrahedra.
+\end{proof}
+
+Let us show how to derive \Cref{coro:prismatoid,coro:width1} from this.
+Recall that \emph{prismatoids} were defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets.
 In particular, a \emph{lattice prismatoid} is any $d$-polytope $SL(\Z,d)$-equivalent to one of the form
@@
 \]
-where $Q_1,Q_2$ are lattice $(d-1)$-polytopes and $k\in \Z_{>0}$. This generalizes Cayley sums, which are the case $k=1$.
+where $Q_1,Q_2$ are lattice $(d-1)$-polytopes and $k\in \Z_{>0}$.
+This generalizes Cayley sums, which are the case $k=1$, except the definition of prismatoid requires $Q_1$ and $Q_2$ to be full-dimensional, while the Cayley sum only requires this for $Q_1+Q_2$.
 
@@
 \begin{proof}
-\paco{perhaps add details}
+\paco{ added details}
 The condition that $P\cap(\R^2\times\{1\})$ is a lattice polygon implies the same for $P\cap(\R^2\times\{i\})$, for every $i$. 
-Then, each slice
-\[
-P \cap (\R^2\times[i,i+1])
-\]
-is a nef Cayley polytope, hence it has a unimodular cover by \Cref{thm:cayley}.
-\end{proof}
-
-This statement implies \Cref{coro:prismatoid} and \Cref{coro:width1}. 
-For the former, smoothness of a prismatoid implies $P\cap(\R^2\times\{1\})$ to be a lattice polytope. \paco{details?}
-For the latter, if $P$ has width one then $P\cong \cayley(Q_1, Q_2)$ and, for every $k\in \Z_{>0}$, we have
+Indeed, the condition implies that every edge of $\cayley(P,Q)$ of the form $[u\times \{0\}, v\times \{k\}]$ has a lattice point in $\R^2\times\{i\}$, and hence it has a lattice point in $P\cap(\R^2\times\{i\})$, for every $i$.
+
+Observe that for every $i\in \{1,\dots,k-1\}$ the intersection $P\cap(\R^2\times\{i\})$ has the same normal fan as $Q_1+Q_2$. Thus, each slice
+\[
+P \cap (\R^2\times[i-1,i])
+\]
+is a Cayley polytope in which both bases have the same normal fan (if $i\in\{2,\dots,k-1\}$ or one is a weak Minkowski summand of the other (if $i\in \{1,k\}$).
+\end{proof}
+
+\begin{proof}[Proof of \Cref{coro:prismatoid,coro:width1}]
+In both cases the polytope under study satisfies the hypotheses of \Cref{prop:prismatoid}.
+For the first, smoothness of a prismatoid implies that every edge of the form $[u\times \{0\}, v\times \{k\}]$ has lattice points in all slices.
+For the second, if $P$ has width one then $P\cong \cayley(Q_1, Q_2)$ for some $Q_1$ and $Q_2$.
+Hence,
 \[
 kP \cap(\R^2\times\{1\}) = (k-1)Q_1 + Q_2.
-\]
+\qedhere
+\] 
+\end{proof}
 
@@
 
-Observe that in the strip strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
+Observe that in the strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
 %We call that point $q^+$.
@@
 w=\width_{f_q}(p + \vecline q )=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\vecline p ). 
-\] \giulia{we want a picture for the setup}
+\] 
+\paco{picture for setup needs change}
 
@@
 
-\Giulia{add references to Haase&Hofmann and Tsuchiya}
-
 \begin{thebibliography}{99}
@@
 Wiley, London, 1967.
+
+\bibitem{HaaseHof}
+Christian Haase and Jan Hofmann, 
+Convex-normal (Pairs of) Polytopes, 
+\emph{Canad. Math. Bull.} 60 (2017), 510--521.
 
@@
 \newblock {\em Trans. Moscow Math. Soc.} (2013), 293--311. % \MR{3235802}
+
+\bibitem{Tsuchiya}
+Akiyoshi Tsuchiya.
+Cayley sums and Minkowski sums of 2-convex-normal lattice polytopes.
+Preprint, arXiv:1804.10538, April 2018.
 

### 2019-07-22 unimodular-covers.tex
comprehensive revision of sect 1,2,3

@@
 
-Oda~\cite{Oda1997} posed several questions regarding smoothness and the IDP property for lattice polytopes.
+Oda (\cite{Oda1997}) posed several questions regarding smoothness and the IDP property for lattice polytopes.
 %
 \paco{I think this definition goes better here, and decided to "undeclare" it and the definition of weak Mink sum}
-Following \cite{HaaseHof, Tsuchiya} we say that a pair $(P, Q)$ of lattice polytopes has the integer decomposition property, or that \emph{the pair $(P,Q)$ is IDP}, if 
+Following \cite{HaaseHof, Tsuchiya}, we say that a pair $(P, Q)$ of lattice polytopes has the integer decomposition property, or that \emph{the pair $(P,Q)$ is IDP}, if 
 \begin{align*}
@@
 %
-A lattice polytope $Q$ is called \emph{smooth} if it is simple and the edge generators at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth.
+A lattice polytope $Q$ is called \emph{smooth} if it is simple and the primitive edge directions \giulia{changed this from "edge generators", which to me was unclear} at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth.
 The following versions of Oda's questions are now considered conjectures~\cite{HNPS2008,mfo2007}, and they are open even in dimension three:
@@
 Perhaps we can leave it wothout reference}.
-This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor on  $X_Q$ defined by $P$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that there ``Minkowski summand'' is used meaning what we call ``weak Minkowski summand'').
+This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor on  $X_Q$ defined by $P$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that there, what we call  ``weak Minkowski summand'' is called simply ``Minkowski summand'').
 %\paco{added this explanation}
@@
 This last one actually shows that smoothness of $Q$ is not needed.
-In dimension three, however, the conjecture fails without the smoothness assumption, as shown by letting $P=Q$ be any non-unimodular \emph{empty tetrahedron}, by which we mean a lattice tetrahedron with no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for details on them).
-\paco{added def of empty tetrahedron}
+In dimension three, however, the conjecture fails without the smoothness assumption.
+Indeed, it fails if $P=Q$ is any non-unimodular \emph{empty tetrahedron}, by which we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of empty tetrahedra).
+\paco{added def of empty tetrahedron}\giulia{I didn't like "details on them" and changed it to  "classification of empty tetrahedra". Is this ok?}
 
@@
 \label{thm:cayley}
-Let $P$ and $Q$ be lattice polygons with the normal fan of $Q$ refining that of $P$. Then the Cayley sum $\cayley(P,Q)$ has a unimodular cover.
+Let $P$ and $Q$ be lattice polygons, with $P$ a weak Minkowski summand of $Q$. Then the Cayley sum $\cayley(P,Q)$ has a unimodular cover.
 \end{theorem}
@@
 
-\paco{changed this}
 We call \emph{band} any region of the form $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*$ and closed interval $[\alpha,\beta]\subset \R$.
@@
 
-\begin{proof}[Proof of \Cref{thm:parallelepipeds}]\giulia{see comment above}
+\begin{proof}[Proof of \Cref{thm:parallelepipeds}]
 Arbitrarily triangulate the parallelepiped into empty lattice tetrahedra and apply \Cref{coro:coverpara} to these tetrahedra.
@@
 
-\section{ Cayley sums} \giulia{Change notations: remove nef and make sure everything works}
+\section{ Cayley sums}
 \label{sec:cayley}
@@
 
-\begin{center}
-  \begin{tabular}{ c c c }
-    $\cayley(P,Q)$ && $P+Q$\\
-    \hline
-   polyhedral subdivisions &$\leftrightarrow$& mixed subdivisions \\
-    triangulations &$\leftrightarrow$& fine mixed subdivisions \\
-   unimodular simplices &$\leftrightarrow$& unimodular prod-simplices \\
-  \end{tabular}
-\end{center}\giulia{not sure if the table makes it clear that these are bijections}
-
 \[
@@
 \medskip
-\giulia{move first paragraph of proof of theorem here, and explain simplices, prodsimplices...}
 
 We now turn our attention to $d=2$, in order to prove \Cref{thm:cayley}.
-In a triangulation of $\cayley(P,Q)\subset \R^3$ there are tetrahedra of types $(1,3)$, $(2,2)$ and $(3,1)$, where the type denotes how many vertices they have in $P$ and in $Q$.
-Empty tetrahedra of types $(1,3)$ or $(3,1)$ are automatically unimodular, so we only need to care with those of type $(2,2)$.
-These are the Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
+A triangulation of $\cayley(P,Q)\subset \R^3$ consists of tetrahedra of types $(1,3)$, $(2,2)$ and $(3,1)$, where the type denotes how many vertices they have in $P$ and in $Q$.
+Empty tetrahedra of types $(1,3)$ or $(3,1)$, which are Cayley sums of a triangle in $P$ and a point in $Q$, or viceversa, are automatically unimodular.
+The case that we need to study are therefore tetrahedra of type $(2,2)$, which are Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
 %
@@
 %\end{align*}
-For this the following result, the proof of which we postpone to \Cref{sec:the_lemma}, is crucial. 
-The lemma uses the following conventions.
-If $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$.
-For a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and $\vecline s$ as the line spanned by $\vec s$.
+The following lemma, whose proof we postpone to \Cref{sec:the_lemma}, is crucial to understand how to unimodularly cover these tetrahedra. 
+%For this the following result, the proof of which we postpone to \Cref{sec:the_lemma}, is crucial. 
+We use the following conventions: if $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$.
+Given a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and  the line spanned by $\vec s$ by $\vecline s$.
 
@@
 \label{lemma:cayley}
-\paco{$Q$ is 2-dim, but $P$ may not}
+%\paco{$Q$ is 2-dim, but $P$ may not}
 Let $Q$ be a two-dimensional lattice polytope and $P$ a weak Minkowski summand of it. 
@@
 \label{coro:covercayley}
-\paco{inserted this corollary to make proof clearer, and to parallel section 2}
+\paco{inserted this corollary to make proof clearer, and to parallel section 2} \giulia{picture?}
 Let $T$ be an empty lattice tetrahedron contained in the Cayley sum $\cayley(P,Q)$, where $P$ is a weak Minkowski summand of $Q$.
@@
 
-Let us show how to derive \Cref{coro:prismatoid,coro:width1} from this.
-Recall that \emph{prismatoids} were defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets.
+Let us now show how to derive \Cref{coro:prismatoid,coro:width1} from this theorem.
+\emph{Prismatoids} were defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets.
 In particular, a \emph{lattice prismatoid} is any $d$-polytope $SL(\Z,d)$-equivalent to one of the form
@@
 where $Q_1,Q_2$ are lattice $(d-1)$-polytopes and $k\in \Z_{>0}$.
-This generalizes Cayley sums, which are the case $k=1$, except the definition of prismatoid requires $Q_1$ and $Q_2$ to be full-dimensional, while the Cayley sum only requires this for $Q_1+Q_2$.
+This is almost a generalization of Cayley sums, which would be the case $k=1$, except the definition of prismatoid requires $Q_1$ and $Q_2$ to be full-dimensional, while the Cayley sum only requires this for $Q_1+Q_2$.
+\giulia{I didn't like "this is a generalization except...}
 
@@
 \begin{proof}
-\paco{ added details}
 The condition that $P\cap(\R^2\times\{1\})$ is a lattice polygon implies the same for $P\cap(\R^2\times\{i\})$, for every $i$. 
@@
 \]
-is a Cayley polytope in which both bases have the same normal fan (if $i\in\{2,\dots,k-1\}$ or one is a weak Minkowski summand of the other (if $i\in \{1,k\}$).
+is a Cayley polytope.
+If $i\in\{2,\dots,k-1\}$, both bases have the same normal fan (and therefore each is a weak Minkowski summand of the other), while for $i\in \{1,k\}$ one base is a weak Minkowski summand of the other.
+We can therefore apply \Cref{thm:cayley} to each slice and combine the covers thus obtained to get a unimodular cover of $P$.
 \end{proof}
@@
 \begin{proof}[Proof of \Cref{coro:prismatoid,coro:width1}]
-In both cases the polytope under study satisfies the hypotheses of \Cref{prop:prismatoid}.
-For the first, smoothness of a prismatoid implies that every edge of the form $[u\times \{0\}, v\times \{k\}]$ has lattice points in all slices.
-For the second, if $P$ has width one then $P\cong \cayley(Q_1, Q_2)$ for some $Q_1$ and $Q_2$.
+In both cases the polytope under study satisfies the hypotheses of \Cref{prop:prismatoid}: in \Cref{coro:prismatoid}, the smoothness of the prismatoid implies that every edge of the form $[u\times \{0\}, v\times \{k\}]$ has lattice points in all slices.
+In \Cref{coro:width1}, since $P$ has width one, $P\cong \cayley(Q_1, Q_2)$ for some $Q_1$ and $Q_2$.
 Hence,

### 2019-07-23 unimodular-covers.tex
revision of sect 4

@@
 
-Observe that in the strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
-%We call that point $q^+$.
-Similarly, $q +\vecline p$ has a unique lattice point at $f_q(x)=-1$, which we call $q_-$.
-Observe that $q_++q_-=q_1+q_2$ and that the triangles $\conv(q_1, q_2, q_+)$ and $\conv(q_1, q_2, q_1)$ are unimodular.
+Observe that in the strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=-1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
 Since translating the polytopes by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
@@
 
-We let $H_1=\{f_q(x) < 0\}$ and $H_2=\{f_q(x) > 0\}$; similarly let $V_1=\{f_p(x) < 0\}$ and $V_2=\{f_p(x) > 0\}$. 
+We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq 0\}$ and $V_2=\{f_p(x) \geq 0\}$.
+\giulia{we need to make these closed, so that the case $P$ one dimensional fits into this same proof}
 In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants.
@@
 Suppose by contradiction that there is no lattice point as described in the lemma.
-In particular, no vertex of $P$ can be in the strip $p + \vecline q$.
-Thus $P$ has two primitive segments in the boundary which each have one vertex on each side of the strip $p + \vecline q$; we will call these $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively and the convention that $f_q(l_2) >f_q(l_1)$ and $f_q(r_2) >f_q(r_1)$.
+In particular, no lattice point on the boundary of $P$ can be in the interior of the strip $p + \vecline q$.
+Thus the boundary of $P$ contains two primitive segments which each have one vertex on each side of the strip $p + \vecline q$; we will call these $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively and the convention that $f_q(l_2) >f_q(l_1)$ and $f_q(r_2) >f_q(r_1)$.
 This readily implies
@@
 Observe that a priori one of $l$ and $r$ can coincide with $p$, if this is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if this is on the boundary of $Q$.
-\giulia{p=P ???}
-
-%To make things more explicit, we assume the following coordinates for the edge-vectors:
-%\[
-%\vec l = (l_1,l_2), \quad
-%\vec r= (r_1,r_2), \quad
-%\vec b= (b_1,b_2), \quad
-%\vec t= (t_1,t_2).
-%\]
-%Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
-
-%\begin{claim}
-%$l_2, r_2, b_1, t_1 >1$.
-%\end{claim}
 

### 2019-07-24 unimodular-covers.tex
last changes plus figures

@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 
-\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and 
+\subsection*{Acknowledgements:} We would like to thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and 
 Christian Haase for helpful discussions.
@@
 
-The idea to have in mind is that of the regular tetrahedron inscribed in a cube. 
+The idea to have in mind is that of the regular tetrahedron inscribed in a cube, see \Cref{fig:circumscribed_parall}. 
 Modulo an affine transformation this is exactly the situation of $T$ and $C(T)$.
-\giulia{we should add a picture of the standard cube}
-\paco{perhaps, but not necessarily}
-
-%\Giulia{not sure if we need this remark:
-%$S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique).
-Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it.
-(Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).
-%}
+
+\begin{figure}[htb]
+\includegraphics[width=.7\textwidth]{circumscribed_parall}
+\caption{In red we have the tetrahedron $T$, in black the circumscribed parallelepiped $C(T)$, and in blue the corner simplex $T_4$.}
+\label{fig:circumscribed_parall}
+\end{figure}
+\giulia{I added a picture, now we can decide if we want it in the paper or not.}
 
@@
 where $b\ge 2$ is the (normalized) volume of $T$, and $a\in \{1,\dots,b-1\}$ satisfies $\gcd(a,b)=1$. This gives 
-\paco{picture of the quadrilaterals $p_1q_4p_2q_3$ and $q_2p_3q_1p_4$ would be good.}
-\paco{perhaps, but not necessarily}
 \begin{align*}
@@
 
-Observe that $u+v=p_1+p_2=q_3+q_4$.
-Now, the translation of vector $\left(\frac{a-1}2,\frac{b}2,1\right)$ sends the facet $\conv(p_1p_2q_3q_4)$ of $C(T)$ to the opposite facet $\conv(q_2q_1p_4p_3)$, so that the translated points $u + \left(\frac{a-1}2,\frac{b}2,1\right)$ and $v + \left(\frac{a-1}2,\frac{b}2,1\right)$ lie in $\conv(q_2q_1p_4p_3)$ and are centrally symmetrically placed in it.
-Thus, one of them belongs to $\conv(q_1p_4p_3) \subset T_2$ and the other to $\conv(q_2p_4p_3) \subset T_1$.
-%\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
-%Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.
+See \Cref{fig:corner}. Observe that $u+v=p_1+p_2=q_3+q_4$.
+Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.
 Hence, its translate $\conv(q_2p_3q_1p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$.
 By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
+
+\begin{figure}[htb]
+\includegraphics[width=.9\textwidth]{Whites}
+\caption{The parallelograms $\conv(p_1p_2q_3q_4)$ and $\conv(q_2q_1p_4p_3)$. The red points are lattice points.}
+\label{fig:corner}
+\end{figure}
+\giulia{ditto}
+
+%\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
+
 \end{proof}
@@
 
-\paco{removed definitions of "nef cayley", "mixed IDP" and "weak Minkowski sum" from here}
-%\begin{definition}
-%In the above conditions, we say that $Q$ \emph{is a weak Minkowski summand of $P$} if the normal fan of $Q$ refines that of $P$. 
-%If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.\giulia{I guess there's no nice term for this now...}
-%\paco{not sure; maybe we can keep nef after all; another (unclear) option could be ``weak prism over $Q$'' or ``degenerate prism over $Q$''.}
-%
-%We say that the pair $(P,Q)$ is \emph{mixed IDP} if 
-%\paco{``mixed IDP'' is not the right property, since this is weaker than IDP} \giulia{do we give it a name at all then?}
-%\paco{perhaps mixed-IDP}
-%\[
-%(P\cap \Z^d) + (Q\cap \Z^d) = (P+Q)\cap \Z^d.
-%\]
-%\end{definition}
-%
-%In this language, \Cref{conj:Oda}\eqref{itm:mixedIDP} translates to:
-%\emph{If $Q$ is nef for $P$ and $P$ is smooth then the pair $(P,Q)$ mixed IDP.}
-%
 The so-called \emph{Cayley Trick} is the isomorphism
@@
 The case that we need to study are therefore tetrahedra of type $(2,2)$, which are Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
-%
-%
-%The unimodular prod-simplices are unimodular parallelograms $p+q$ with $p$ a segment of $P$ and $q$ a segment of $Q$, and unimodular triangles $T+u$, where $T \subset P$ and $u \in Q$ or viceversa. 
-%
-%Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$.
-Observe that any simplex in $T$ has vertices in $P \times \{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$.
-Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$).
-%
 \giulia{fix this}
 \paco{done, but check it}
-%
-%As a consequence, we also have that:
-%\begin{align*}
-%\text{$\cayley(P,Q)$ has a unimodular cover} &\leftrightarrow \text{$P+Q$ can be covered by unimodular prod-simplices}.
-%\end{align*}
+\giulia{just changed small things}
 The following lemma, whose proof we postpone to \Cref{sec:the_lemma}, is crucial to understand how to unimodularly cover these tetrahedra. 
-%For this the following result, the proof of which we postpone to \Cref{sec:the_lemma}, is crucial. 
 We use the following conventions: if $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$.
@@
 \end{lemma}
-%\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in the interior of $\{p_1, p_2\} + \{q_1, q_2\}$ (or symmetric for $Q$)}
 
@@
 \end{figure}
+\giulia{added picture, we may not want it}
 
@@
 \label{coro:covercayley}
-\paco{inserted this corollary to make proof clearer, and to parallel section 2} \giulia{picture?}
 Let $T$ be an empty lattice tetrahedron contained in the Cayley sum $\cayley(P,Q)$, where $P$ is a weak Minkowski summand of $Q$.
@@
 That is, $\tilde p_i = p\times\{0\}$, $\tilde q_i = p\times\{1\}$, and $\tilde u = u\times\{0\}$.
-Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that of the segments $[u,q_i]$ crosses the triangle $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$.
+Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that of the segments $[u,q_i]$ crosses the triangle $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$, see \Cref{fig:flip}.
+
+\begin{figure}[htb]
+\includegraphics[width=.9\textwidth]{flip}
+\caption{$[u,q_2]$ intersects $\conv(p_1,p_2,q_1)$}
+\label{fig:flip}
+\end{figure}
+
 In turn, this means that the polytope $\conv(\tilde u, \tilde p_1, \tilde p_2, \tilde q_1, \tilde q_2) = \cayley(t,q)$
@@
 
-Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain. Then 
+Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain. Then:
 \[
@@
 \]
-\paco{picture for setup needs change}
 
@@
 Suppose by contradiction that there is no lattice point as described in the lemma.
-In particular, no lattice point on the boundary of $P$ can be in the interior of the strip $p + \vecline q$.
-Thus the boundary of $P$ contains two primitive segments which each have one vertex on each side of the strip $p + \vecline q$; we will call these $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively and the convention that $f_q(l_2) >f_q(l_1)$ and $f_q(r_2) >f_q(r_1)$.
+In particular, no lattice point on the boundary of $Q$ can be in the interior of the strip $q + \vecline p$.
+Thus the boundary of $Q$ contains two primitive segments which each have one vertex on each side of the strip $q + \vecline p$; we will call these  $b=[b_1, b_2], t=[t_1, t_2]$, with $b$ and $t$ crossing the strip in $H_1$ and $H_2$ respectively and the convention that $f_p(b_2) >f_p(b_1)$ and $f_p(t_2) >f_p(t_1)$.
 This readily implies
+
+\begin{gather}
+\label{eq:widthq}
+\begin{array}{cc}
+f_p(t_1) \leq f_p(q_1), &
+f_p(t_2) \geq f_p(q_2), \\
+f_p(b_1) \leq f_p(q_1), &
+f_p(b_2) \geq f_p(q_2).
+\end{array}
+\end{gather}
+
+The same holds for $P$ and the strip $p+\vecline q$, and we call the segments $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively.
+The only difference is that in the case that $P$ is one dimensional we have $\ell=r=p$.
+Again we have
 \begin{gather}
@@
 f_q(r_2) \geq f_q(p_2).
-\end{array}
-\end{gather}
-
-The same holds for $Q$ and the strip $q+\vecline p$, and we call the segments  $b=[b_1, b_2], t=[t_1, t_2]$ with $b$ and $t$ crossing the strip in $H_1$ and $H_2$ respectively. Again we have
-\begin{gather}
-\label{eq:widthq}
-\begin{array}{cc}
-f_p(t_1) \leq f_p(q_1), &
-f_p(t_2) \geq f_p(q_2), \\
-f_p(b_1) \leq f_p(q_1), &
-f_p(b_2) \geq f_p(q_2).
 \end{array}
@@
 
-Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2} .
+Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$.
+This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2}.
+This is a contradiction, since $P$ is a Minkowski summand of $Q$ and therefore $Q$ must have an edge parallel to $t$.
 \end{proof}
@@
 
-We assume w.l.o.g.
-that the maximum on $t$ (and hence on $b$) is achieved in $V_2$ \giulia{do we need it??}, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).
+We assume w.l.o.g. that the maximum on $t$ (and hence on $b$) is achieved in $V_2$, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).
 
@@
 \end{proof}
-\giulia{add a picture?}
 

### 2019-07-25 unimodular-covers.tex
wrong file added
went through Giulia's changes

@@
 \begin{abstract}
-We show that the following classes of three-dimensional lattice polytopes have unimodular covers: the class of all parallelepipeds, the class of all centrally symmetric polytopes, and the class of all three-dimensional Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$.
+We show that the following classes of lattice polytopes have unimodular covers, in dimension three: the class of parallelepipeds, the class of centrally symmetric polytopes, and the class of Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$.
 This improves results of Beck et al.~(2018) and Haase et al.~(2008) where the last two classes were shown to be IDP.
@@
 
-Oda (\cite{Oda1997}) posed several questions regarding smoothness and the IDP property for lattice polytopes.
-%
-\paco{I think this definition goes better here, and decided to "undeclare" it and the definition of weak Mink sum}
+Oda \cite{Oda1997} posed several questions regarding smoothness and the IDP property for lattice polytopes.
+%
 Following \cite{HaaseHof, Tsuchiya}, we say that a pair $(P, Q)$ of lattice polytopes has the integer decomposition property, or that \emph{the pair $(P,Q)$ is IDP}, if 
@@
 %
-A lattice polytope $Q$ is called \emph{smooth} if it is simple and the primitive edge directions \giulia{changed this from "edge generators", which to me was unclear} at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth.
+A lattice polytope $Q$ is called \emph{smooth} if it is simple and the primitive edge directions at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth.
 The following versions of Oda's questions are now considered conjectures~\cite{HNPS2008,mfo2007}, and they are open even in dimension three:
@@
 
-When the normal fan of a $Q$ refines that of $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$\paco{did not find this in the google books version of \cite{Grunbaum}; found it in Postnikov-Reiner-Williams, Tsuchiya, a survey by Rekha Thomas, and other places.
+When the normal fan of a polytope $Q$ refines that of another polytope $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$\paco{did not find this in the google books version of \cite{Grunbaum}; found it in Postnikov-Reiner-Williams, Tsuchiya, a survey by Rekha Thomas, and other places.
 Perhaps we can leave it wothout reference}.
-This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor on  $X_Q$ defined by $P$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that there, what we call  ``weak Minkowski summand'' is called simply ``Minkowski summand'').
+This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor defined by $P$ on  $X_Q$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that what we here call  ``weak Minkowski summand'' is simply called ``Minkowski summand'' there).
 %\paco{added this explanation}
@@
 In dimension three, however, the conjecture fails without the smoothness assumption.
-Indeed, it fails if $P=Q$ is any non-unimodular \emph{empty tetrahedron}, by which we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of empty tetrahedra).
-\paco{added def of empty tetrahedron}\giulia{I didn't like "details on them" and changed it to  "classification of empty tetrahedra". Is this ok?}
+Indeed, if we let $P=Q$ be any non-unimodular \emph{empty tetrahedron}, then $P$ is obviously a weak Minkowski summand of $Q$ but the pair $(P.Q)$ is not IDP.
+By an empty tetrahedron we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of them).
+\giulia{I didn't like "details on them" and changed it to  "classification of empty tetrahedra". Is this ok?}
+\paco{ok. But left it as "classification of them"}
 
@@
 \]
-We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but for this does not need $Q_1$ and $Q_2$ to be full-dimensional.
+We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but this does not need $Q_1$ and $Q_2$ to be full-dimensional.
 It only requires the linear subspaces parallel to them to span $\R^d$.
@@
 
-This has the following two consequences, also proved in \Cref{sec:cayley}.
+This has the following two corollaries, also proved in \Cref{sec:cayley}.
 A  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
@@
 
-For each $i \in [4]$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=[4]$; we call these $T_i$ the \emph{corner tetrahedra} of $C(T)$. Together with $T$ they triangulate of $C(T)$.
+For each $i \in [4]$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=[4]$; we call these $T_i$ the \emph{corner tetrahedra} of $C(T)$. Together with $T$ they triangulate $C(T)$.
 \end{definition}
 
-The idea to have in mind is that of the regular tetrahedron inscribed in a cube, see \Cref{fig:circumscribed_parall}. 
-Modulo an affine transformation this is exactly the situation of $T$ and $C(T)$.
-
+Modulo an affine transformation, the situation of $T$ and $C(T)$ is exactly that of the regular tetrahedron inscribed in a cube; see \Cref{fig:circumscribed_parall}. 
+%
 \begin{figure}[htb]
-\includegraphics[width=.7\textwidth]{circumscribed_parall}
-\caption{In red we have the tetrahedron $T$, in black the circumscribed parallelepiped $C(T)$, and in blue the corner simplex $T_4$.}
+\includegraphics[scale=.25]{circumscribed_parall}
+\caption{In red we have a tetrahedron $T$, in black its circumscribed parallelepiped $C(T)$, and in blue the corner simplex $T_4$.}
 \label{fig:circumscribed_parall}
 \end{figure}
-\giulia{I added a picture, now we can decide if we want it in the paper or not.}
 
@@
 By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
-
+%
+\giulia{ditto}
+\paco{not sure about this one. It looks like the red dots lie on edges of $C(T)$}
 \begin{figure}[htb]
-\includegraphics[width=.9\textwidth]{Whites}
-\caption{The parallelograms $\conv(p_1p_2q_3q_4)$ and $\conv(q_2q_1p_4p_3)$. The red points are lattice points.}
+\includegraphics[scale=.25]{Whites}
+\caption{The parallelograms $\conv(p_1p_2q_3q_4)$ and $\conv(q_2q_1p_4p_3)$. The large red dots are lattice points.}
 \label{fig:corner}
 \end{figure}
-\giulia{ditto}
-
 %\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
-
 \end{proof}
@@
 The case that we need to study are therefore tetrahedra of type $(2,2)$, which are Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
-\giulia{fix this}
-\paco{done, but check it}
-\giulia{just changed small things}
 The following lemma, whose proof we postpone to \Cref{sec:the_lemma}, is crucial to understand how to unimodularly cover these tetrahedra. 
@@
 
+%\giulia{added picture, we may not want it}
+%\paco{looks ok}
 \begin{figure}[htb]
@@
 \end{figure}
-\giulia{added picture, we may not want it}
 
@@
 \begin{figure}[htb]
-\includegraphics[width=.9\textwidth]{flip}
+\includegraphics[scale=.3]{flip}
 \caption{$[u,q_2]$ intersects $\conv(p_1,p_2,q_1)$}
@@
 This is almost a generalization of Cayley sums, which would be the case $k=1$, except the definition of prismatoid requires $Q_1$ and $Q_2$ to be full-dimensional, while the Cayley sum only requires this for $Q_1+Q_2$.
-\giulia{I didn't like "this is a generalization except...}
 
@@
 We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq 0\}$ and $V_2=\{f_p(x) \geq 0\}$. 
-\giulia{we need to make these closed, so that the case $P$ one dimensional fits into this same proof}
+\giulia{we need to make these closed, so that the case $P$ one dimensional fits into this same proof}\paco{ok}
 In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants. 
+See \Cref{fig:setup}.
+%
+\begin{figure}[htb]
+\includegraphics[scale=.3]{setup.png}
+\caption{Setup for the proof of \Cref{lemma:cayley}}
+\label{fig:setup}
+\end{figure}
 
@@
 \]
-
-\begin{figure}
-\includegraphics[scale=.3]{setup.png}
-\caption{Setup for \Cref{lemma:cayley}}
-\end{figure}
 
@@
 This readily implies
-
 \begin{gather}

### 2019-07-25 unimodular-covers.tex
minor edits

@@
 
-Oda \cite{Oda1997} posed several questions regarding smoothness and the IDP property for lattice polytopes.
+Oda (\cite{Oda1997}) posed several questions regarding smoothness and the IDP property for lattice polytopes.
 %
@@
 is a Cayley polytope.
-If $i\in\{2,\dots,k-1\}$, both bases have the same normal fan (and therefore each is a weak Minkowski summand of the other), while for $i\in \{1,k\}$ one base is a weak Minkowski summand of the other.
+For $i\in\{2,\dots,k-1\}$,  both bases have the same normal fan (and therefore each is a weak Minkowski summand of the other); for $i\in \{1,k\}$ one base is a weak Minkowski summand of the other.
 We can therefore apply \Cref{thm:cayley} to each slice and combine the covers thus obtained to get a unimodular cover of $P$.

### 2019-07-26 unimodular-covers.tex
changed statement of theorem 1.6

@@
 (Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Every pair $(P,Q)$ of lattice polytopes with $Q$ smooth and the normal fan of $Q$ refining that of $P$ is IDP.
-%\begin{align}
-%\label{eq:mixedIDP}
-%(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
-%\end{align}
 \end{enumerate}
@@
 
-When the normal fan of a polytope $Q$ refines that of another polytope $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$\paco{did not find this in the google books version of \cite{Grunbaum}; found it in Postnikov-Reiner-Williams, Tsuchiya, a survey by Rekha Thomas, and other places.
-Perhaps we can leave it wothout reference}.
+When the normal fan of a polytope $Q$ refines that of another polytope $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is easily seen to be equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$.
+%\paco{did not find this in the google books version of \cite{Grunbaum}; found it in Postnikov-Reiner-Williams, Tsuchiya, a survey by Rekha Thomas, and other places.
+Perhaps we can leave it wothout reference}
 This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor defined by $P$ on  $X_Q$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that what we here call  ``weak Minkowski summand'' is simply called ``Minkowski summand'' there).
-%\paco{added this explanation}
-%
-%\begin{definition}
-%Following \cite{HaaseHof, Tsuchiya} we say that the pair $(P, Q)$ has the integer decomposition property, or \emph{the pair $(P,Q)$ is IDP}, if 
-%\begin{align*}
-%(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
-%\end{align*}
-%\end{definition}
-% 
-%In this language,  \Cref{conj:Oda}\eqref{itm:mixedIDP} can be rephrased as: if $P$ is a weak Minkowski summand of a smooth polytope $Q$ then the pair $(P,Q)$ is IDP.
 
@@
 
-%\begin{proof}
-%Let $P$ be a smooth centrally symmetric lattice $3$-polytope.
-%Beck et al.~\cite{BHHHJKM2019} show that $P$ is IDP by covering it with parallelepipeds and unimodular simplices.
-Once we know parallelepipeds have unimodular covers (\Cref{thm:parallelepipeds}), we conclude $P$ has a unimodular cover too.
-%\end{proof}
-
 The two-dimensional case of \Cref{conj:Oda}\eqref{itm:mixedIDP} is known to hold, with three different proofs by Fakhruddin~\cite{Fakhruddin}, Ogata~\cite{Ogata} and Haase et al.~\cite{HNPS2008}.
@@
 Indeed, if we let $P=Q$ be any non-unimodular \emph{empty tetrahedron}, then $P$ is obviously a weak Minkowski summand of $Q$ but the pair $(P.Q)$ is not IDP.
-By an empty tetrahedron we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of them).
+By an empty tetrahedron we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of these).
 \giulia{I didn't like "details on them" and changed it to  "classification of empty tetrahedra". Is this ok?}
 \paco{ok. But left it as "classification of them"}
+\giulia{is it of them or of these!? my grammar needs some refresh}
 
 An alternative approach to \Cref{conj:Oda}\eqref{itm:mixedIDP} is via Cayley sums, which we discuss in  \Cref{sec:cayley}. 
-Recall that the \emph{Cayley sum} of two lattice polytopes $Q_1,Q_2\subset \R^d$ is the lattice polytope
-\[
-\cayley(Q_1,Q_2) := \conv(Q_1\times\{0\} \cup Q_2 \times \{1\}) \subset \R^3.
-\]
-We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but this does not need $Q_1$ and $Q_2$ to be full-dimensional.
+Recall that the \emph{Cayley sum} of two lattice polytopes $P,Q\subset \R^d$ is the lattice polytope
+\[
+\cayley(P,Q) := \conv(P\times\{0\} \cup Q \times \{1\}) \subset \R^3.
+\]
+We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but this does not need $P$ and $Q$ to be full-dimensional.
 It only requires the linear subspaces parallel to them to span $\R^d$.
@@
 \label{thm:cayley}
-Let $P$ and $Q$ be lattice polygons, with $P$ a weak Minkowski summand of $Q$. Then the Cayley sum $\cayley(P,Q)$ has a unimodular cover.
-\end{theorem}
+Let $Q$ be lattice polygon, and $P$ a weak Minkowski summand of $Q$. Then the Cayley sum $\cayley(P,Q)$ has a unimodular cover.
+\end{theorem}\giulia{changed this so that only $Q$ need be a lattice polygon}
 
@@
 %
-\giulia{ditto}
-\paco{not sure about this one. It looks like the red dots lie on edges of $C(T)$}
-\begin{figure}[htb]
-\includegraphics[scale=.25]{Whites}
-\caption{The parallelograms $\conv(p_1p_2q_3q_4)$ and $\conv(q_2q_1p_4p_3)$. The large red dots are lattice points.}
-\label{fig:corner}
-\end{figure}
-%\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
+
+%\begin{figure}[htb]
+%\includegraphics[scale=.28]{Whites}
+%\caption{The parallelograms $\conv(p_1p_2q_3q_4)$ and $\conv(q_2q_1p_4p_3)$. The large red dots are lattice points.}
+%\label{fig:corner}
+%\end{figure}
+
 \end{proof}
@@
 \begin{gather*}
-P + Q := \{ p+q \in \R^d: p\in P, q \in Q\}.\\
-\cayley (P,Q) = \conv( P\times\{0\} \cup Q\times \{1\}) \in \R^{d+1}.
+P + Q := \{ p+q \in \R^d: p\in P, q \in Q\} \subset \R^d,\\
+\cayley (P,Q) = \conv( P\times\{0\} \cup Q\times \{1\}) \subset \R^{d+1}.
 \end{gather*}
@@
 
-%\giulia{added picture, we may not want it}
-%\paco{looks ok}
 \begin{figure}[htb]
@@
 \label{coro:covercayley}
-Let $T$ be an empty lattice tetrahedron contained in the Cayley sum $\cayley(P,Q)$, where $P$ is a weak Minkowski summand of $Q$.
+Let $T$ be an empty lattice tetrahedron contained in the Cayley sum $\cayley(P,Q)$, where $Q$ is a lattice polygon and $P$ is a weak Minkowski summand of $Q$.
 Then, $T$ can be covered by unimodular tetrahedra contained in $\cayley(P,Q)$.
@@
 We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq 0\}$ and $V_2=\{f_p(x) \geq 0\}$. 
-\giulia{we need to make these closed, so that the case $P$ one dimensional fits into this same proof}\paco{ok}
+%\giulia{we need to make these closed, so that the case $P$ one dimensional fits into this same proof}\paco{ok}
 In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants. 
@@
 
-\bibitem{Grunbaum}
-{Branko Gr\"unbaum}.
-{\em {C}onvex {P}olytopes}.
-Wiley, London, 1967.
+%\bibitem{Grunbaum}
+%{Branko Gr\"unbaum}.
+%{\em {C}onvex {P}olytopes}.
+%Wiley, London, 1967.
 

### 2019-07-29 unimodular-covers.tex
cleaned up comments

@@
 Indeed, if we let $P=Q$ be any non-unimodular \emph{empty tetrahedron}, then $P$ is obviously a weak Minkowski summand of $Q$ but the pair $(P.Q)$ is not IDP.
-By an empty tetrahedron we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of these).
-\giulia{I didn't like "details on them" and changed it to  "classification of empty tetrahedra". Is this ok?}
-\paco{ok. But left it as "classification of them"}
-\giulia{is it of them or of these!? my grammar needs some refresh}
+By an empty tetrahedron we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of them).
 
@@
 Let $Q$ be lattice polygon, and $P$ a weak Minkowski summand of $Q$. Then the Cayley sum $\cayley(P,Q)$ has a unimodular cover.
-\end{theorem}\giulia{changed this so that only $Q$ need be a lattice polygon}
+\end{theorem}
 
@@
 
-See \Cref{fig:corner}. Observe that $u+v=p_1+p_2=q_3+q_4$.
+Observe that $u+v=p_1+p_2=q_3+q_4$.
 Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.

### 2021-02-25 unimodular-covers.tex
updated Paco's grant info
comment on corollary

@@
 \thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
-Work of F. Santos is also supported by project MTM2017-83750-P of the Spanish Ministry of Science (AEI/FEDER, UE)}
+Work of F. Santos is also supported by grants  MTM2017-83750-P/AEI/10.13039/501100011033 and PID2019-106188GB-I00/AEI/10.13039/501100011033 of the Spanish State Research Agency}
 
@@
 \begin{abstract}
-We show that the following classes of lattice polytopes have unimodular covers, in dimension three: the class of parallelepipeds, the class of centrally symmetric polytopes, and the class of Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$.
+We show that the following classes of lattice polytopes have unimodular covers, in dimension three:  parallelepipeds,  smooth centrally symmetric polytopes, and  Cayley sums $\cayley(P,Q)$ where the normal fan of $Q$ refines that of $P$.
 This improves results of Beck et al.~(2018) and Haase et al.~(2008) where the last two classes were shown to be IDP.
 \end{abstract}
+\paco{added ``smooth'' to c.s. polytopes in abstract}
 
@@
 
+\Paco{
+I am a bit puzzled.
+After the next corollary we say that \cite[Cor.~4.2]{SantosZiegler} proves a case of this, but in fact it proves the whole of it, in a stronger form that does not need width one, and that gives a "dissection" and not a mere cover (dissection= cover by simplices with disjoint interiors).
+
+I propose to change it to the following, related to a recent preprint of Gubeladze:
+
+\noindent{\bf Corollary}
+%\label{coro:gorenstein}
+Let $P$ be a lattice $3$-polytope with the following property: for every facet $F$ of $P$ and every edge $E$ of $P$ with (exactly) an end-point in $F$, there is a lattice point $v\in E$ at distance $1$ from $P$.
+Then, there is a unimodular cover of its ``lattice boundary''.
+%\end{corollary}
+
+Here ee call ``lattice boundary'' of a polytope the union of all the prismatoids obtained translating each facet hyperplane one lattice unit towards the interior.
+Put differently, if $\{a_i x \le b_i, i=1,\dots,m\}$ integer is an integral facet description of $P$, the lattice boundary is $P$ minus the interior of $P'=\{a_i x \le b_i-1, i=1,\dots,m\}$.
+Remark: $P'$ is related to (but not exactly the same as) the ``Fine interior'' of $P$.
+See eg, some papers by Batyrev on this.
+
+This corollary is related (but neither implies nor is implied by) Theorem (a) in Gubeladze's recent arXiv:2012.11846
+}
+
 \begin{corollary}
@@
 \end{question}
+\Paco{
+The following is a counter-example: Let $P$ be the Cayley embedding of a long horizontal rectangle and a long vertical rectangle. That is,
+$P=conv([0,a]\times [0,1] \times\{0\} \cup [0,1]\times [0,b] \times\{1\})$, for big $a$ and $b$.
+This is smooth and contains the big empty tetrahedron $T$ with vertices (0,0,0), (a,1,0), (0,0,1), (1,b,1) which occupies most of its volume.
+In particular, none of the corner tetrahedra of $T$ is contained in $P$.
+
+A weaker version of the question is:
+
+Question 2.5Õ.
+If $T$ is an empty tetrahedron contained in a smooth 3-polytope $P$, can one guarantee that there is a lattice point of $P$ in the circumscribed parallelepiped of $T$ (apart of the vertices of $T$)?
+
+A positive answer would still imply that $P$ can be unimodularly covered.
+}
 

### 2021-06-23 unimodular-covers.tex
added Gubeladze, removed Coro. 1.8

@@
 
-Oda (\cite{Oda1997}) posed several questions regarding smoothness and the IDP property for lattice polytopes.
+Oda \cite{Oda1997} posed several questions regarding smoothness and the IDP property for lattice polytopes.
 %
@@
 \medskip 
-Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes. 
+Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes,
+with special attention to dimension $3$ (in dimension 2 is it straightforward that ever lattice polygon has unimodular triangulations). 
 %
@@
 
-This has the following two corollaries, also proved in \Cref{sec:cayley}.
+This has the following consequence, also proved in \Cref{sec:cayley}.
 A  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
@@
 I am a bit puzzled.
-After the next corollary we say that \cite[Cor.~4.2]{SantosZiegler} proves a case of this, but in fact it proves the whole of it, in a stronger form that does not need width one, and that gives a "dissection" and not a mere cover (dissection= cover by simplices with disjoint interiors).
-
-I propose to change it to the following, related to a recent preprint of Gubeladze:
-
-\noindent{\bf Corollary}
-%\label{coro:gorenstein}
-Let $P$ be a lattice $3$-polytope with the following property: for every facet $F$ of $P$ and every edge $E$ of $P$ with (exactly) an end-point in $F$, there is a lattice point $v\in E$ at distance $1$ from $P$.
+After the next corollary we say that \cite[Cor.~4.2]{SantosZiegler} proves a case of this, but in fact it proves the whole of it, in a stronger form: it does not need width one, and it gives a "dissection" and not a mere cover (dissection= cover by simplices with disjoint interiors).
+I propose to delete it:
+%
+%\noindent{\bf Corollary}
+%%\label{coro:gorenstein}
+%Let $P$ be a lattice $3$-polytope with the following property: for every facet $F$ of $P$ and every edge $E$ of $P$ with (exactly) an end-point in $F$, there is a lattice point $v\in E$ at distance $1$ from $P$.
 Then, there is a unimodular cover of its ``lattice boundary''.
-%\end{corollary}
-
-Here ee call ``lattice boundary'' of a polytope the union of all the prismatoids obtained translating each facet hyperplane one lattice unit towards the interior.
+%%\end{corollary}
+%
+%Here we call ``lattice boundary'' of a polytope the union of all the prismatoids obtained translating each facet hyperplane one lattice unit towards the interior.
 Put differently, if $\{a_i x \le b_i, i=1,\dots,m\}$ integer is an integral facet description of $P$, the lattice boundary is $P$ minus the interior of $P'=\{a_i x \le b_i-1, i=1,\dots,m\}$.
@@
 See eg, some papers by Batyrev on this.
-
-This corollary is related (but neither implies nor is implied by) Theorem (a) in Gubeladze's recent arXiv:2012.11846
+%
+%This corollary is related (but neither implies nor is implied by) Theorem (a) in Gubeladze's recent arXiv:2012.11846
 }
-
+{\color{red}
 \begin{corollary}
-\label{coro:width1}
+%\label{coro:width1}
 Every integer dilation $kP$, $k\ge 2$, of a lattice $3$-polytope $P$ of width $1$ has a unimodular cover.
@@
 unimodular covers is \cite[Cor.~4.2]{SantosZiegler} (and is also implicit in \cite{KantorSarkaria}).
-
-\medskip
-We believe that the $3$-polytopes in all these statements have unimodular triangulations, but this remains an open question.
-
-\subsection*{Acknowledgements:} We would like to thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and 
-Christian Haase for helpful discussions.
+}
+
+\paco{added ref to Gubeladze}
+Let us also mention the following recent results of Gubeladze~\cite{Gubeladze}.
+In the second one, a polytope $P$ is called \emph{very ample} if the vertex cone $\cone(P-v)$ is contained in the translation $P-v$, for every vertex $v$ of $P$.
+
+\begin{theorem}[Gubeladze~\cite{Gubeladze}]
+Let $P$ be a lattice $3$-polytope. 
+\begin{enumerate}
+\item If $P=E\cap \Z^3$ for some (solid) ellipsoid $E$, then $P$ has a unimodular cover.
+\item If $P$ is \emph{very ample} then the boundary of $P$ has a unimodular cover. That is, there are unimodular simplices contained in $P$ whose union contains $\partial P$.
+\end{enumerate}
+\end{theorem}
+
+\paco{removed closing sentence}
+%\medskip
+%We believe that the $3$-polytopes in all these statements have unimodular triangulations, but this remains an open question.
+
+\subsection*{Acknowledgements:} We would like to thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and Christian Haase for helpful discussions.
 
@@
 
-Let us now show how to derive \Cref{coro:prismatoid,coro:width1} from this theorem.
+\paco{removed references to Coro 1.8}
+
+Let us now show how to derive \Cref{coro:prismatoid} from this theorem.
 \emph{Prismatoids} were defined in~\cite{Santos-hirsch} as polytopes whose vertices all lie in two parallel facets.
@@
 
-\begin{proof}[Proof of \Cref{coro:prismatoid,coro:width1}]
-In both cases the polytope under study satisfies the hypotheses of \Cref{prop:prismatoid}: in \Cref{coro:prismatoid}, the smoothness of the prismatoid implies that every edge of the form $[u\times \{0\}, v\times \{k\}]$ has lattice points in all slices.
-In \Cref{coro:width1}, since $P$ has width one, $P\cong \cayley(Q_1, Q_2)$ for some $Q_1$ and $Q_2$.
+\begin{proof}[Proof of \Cref{coro:prismatoid}]
+The polytope under study satisfies the hypotheses of \Cref{prop:prismatoid}: %in \Cref{coro:prismatoid}, 
+the smoothness of the prismatoid implies that every edge of the form $[u\times \{0\}, v\times \{k\}]$ has lattice points in all slices. 
+%In \Cref{coro:width1}, since $P$ has width one, $P\cong \cayley(Q_1, Q_2)$ for some $Q_1$ and $Q_2$. 
 Hence,
@@
 \newblock Multiplication maps of linear systems on smooth projective toric surfaces.
-  Preprint, arXiv:0208.5178, August 2002.
+  Preprint, 
+  \href{https://arxiv.org/abs/math/0208178}{arXiv:math/0208178}, August 2002.
 
@@
 %Wiley, London, 1967.
+
+\bibitem{Gubeladze}
+Joseph Gubeladze,
+Normal polytopes and ellipsoids,
+preprint, \href{https://arxiv.org/abs/2012.11846}{arXiv:2012.11846}, December 2012.
 
@@
 Cayley sums and Minkowski sums of 2-convex-normal lattice polytopes.
-Preprint, arXiv:1804.10538, April 2018.
+Preprint,  \href{https://arxiv.org/abs/1804.10538}{arXiv:1804.10538}, April 2018.
 

### 2022-02-22 unimodular-covers.tex
cleaned (some) comments and updated grant info

@@
 \thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
-Work of F. Santos is also supported by grants  MTM2017-83750-P/AEI/10.13039/501100011033 and PID2019-106188GB-I00/AEI/10.13039/501100011033 of the Spanish State Research Agency}
+Work of F.
+Santos is also supported by grants  MTM2017-83750-P/AEI/10.13039/501100011033 and PID2019-106188GB-I00/AEI/10.13039/501100011033 of the Spanish State Research Agency and by project CLaPPo (21.SI03.64658) of Universisdad de Cantabria and Banco Santander}
 
@@
 \end{abstract}
-\paco{added ``smooth'' to c.s. polytopes in abstract}
 
@@
 %\paco{did not find this in the google books version of \cite{Grunbaum}; found it in Postnikov-Reiner-Williams, Tsuchiya, a survey by Rekha Thomas, and other places.
-Perhaps we can leave it wothout reference}
+Perhaps we can leave it without reference}
 This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor defined by $P$ on  $X_Q$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that what we here call  ``weak Minkowski summand'' is simply called ``Minkowski summand'' there).
@@
 
-\Paco{
-I am a bit puzzled.
+\paco{deleted former corollary}
+%\Paco{
+%I am a bit puzzled.
 After the next corollary we say that \cite[Cor.~4.2]{SantosZiegler} proves a case of this, but in fact it proves the whole of it, in a stronger form: it does not need width one, and it gives a "dissection" and not a mere cover (dissection= cover by simplices with disjoint interiors).
-I propose to delete it:
-%
-%\noindent{\bf Corollary}
-%%\label{coro:gorenstein}
-%Let $P$ be a lattice $3$-polytope with the following property: for every facet $F$ of $P$ and every edge $E$ of $P$ with (exactly) an end-point in $F$, there is a lattice point $v\in E$ at distance $1$ from $P$.
-Then, there is a unimodular cover of its ``lattice boundary''.
-%%\end{corollary}
-%
-%Here we call ``lattice boundary'' of a polytope the union of all the prismatoids obtained translating each facet hyperplane one lattice unit towards the interior.
-Put differently, if $\{a_i x \le b_i, i=1,\dots,m\}$ integer is an integral facet description of $P$, the lattice boundary is $P$ minus the interior of $P'=\{a_i x \le b_i-1, i=1,\dots,m\}$.
-Remark: $P'$ is related to (but not exactly the same as) the ``Fine interior'' of $P$.
-See eg, some papers by Batyrev on this.
-%
-%This corollary is related (but neither implies nor is implied by) Theorem (a) in Gubeladze's recent arXiv:2012.11846
-}
-{\color{red}
-\begin{corollary}
-%\label{coro:width1}
-Every integer dilation $kP$, $k\ge 2$, of a lattice $3$-polytope $P$ of width $1$ has a unimodular cover.
-\end{corollary}
-
-A special case of the latter are integer dilations of empty tetrahedra. That their dilations have
-unimodular covers is \cite[Cor.~4.2]{SantosZiegler} (and is also implicit in \cite{KantorSarkaria}).
-}
+%I propose to delete it:
+%}
+%{\color{red}
+%\begin{corollary}
+%%\label{coro:width1}
+%Every integer dilation $kP$, $k\ge 2$, of a lattice $3$-polytope $P$ of width $1$ has a unimodular cover.
+%\end{corollary}
+%
+%A special case of the latter are integer dilations of empty tetrahedra. That their dilations have
+%unimodular covers is \cite[Cor.~4.2]{SantosZiegler} (and is also implicit in \cite{KantorSarkaria}).
+%}
 
@@
 Let us also mention the following recent results of Gubeladze~\cite{Gubeladze}.
-In the second one, a polytope $P$ is called \emph{very ample} if the vertex cone $\cone(P-v)$ is contained in the translation $P-v$, for every vertex $v$ of $P$.
+In part (2), a polytope $P$ is called \emph{very ample} if the vertex cone $\cone(P-v)$ is contained in the translation $P-v$, for every vertex $v$ of $P$.
 
@@
 If this holds then $P$ has a unimodular cover, since then the proofs of \Cref{coro:coverpara} and \Cref{thm:parallelepipeds} work for $P$.
-Hence, a positive answer to the following question would imply that every smooth $3$-polytope has a unimodular cover, which in turn implies \Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three.
-
-\begin{question}
-Does every smooth 3-polytope have the circumscribed parallelepiped property? 
-\end{question}
-\Paco{
-The following is a counter-example: Let $P$ be the Cayley embedding of a long horizontal rectangle and a long vertical rectangle. That is,
-$P=conv([0,a]\times [0,1] \times\{0\} \cup [0,1]\times [0,b] \times\{1\})$, for big $a$ and $b$.
+In turn, our proof that parallelepipeds have the property (\Cref{lemma:3<4}) is based on the fact that they have only three (pairs of) normal vectors. 
+In the following two examples we show a smooth $3$-polytope and two $3$-polytope with four normal vectors that do not have the property. The latter are not IDP:
+
+\begin{example}[A smooth 3-polytope without the circumscribed parallelepiped property]
+ Let $P$ be the Cayley embedding of a long horizontal rectangle and a long vertical rectangle. That is,
+$P=\conv([0,a]\times [0,1] \times\{0\} \cup [0,1]\times [0,b] \times\{1\})$, for big $a$ and $b$.
 This is smooth and contains the big empty tetrahedron $T$ with vertices (0,0,0), (a,1,0), (0,0,1), (1,b,1) which occupies most of its volume.
 In particular, none of the corner tetrahedra of $T$ is contained in $P$.
-
-A weaker version of the question is:
-
-Question 2.5Õ.
-If $T$ is an empty tetrahedron contained in a smooth 3-polytope $P$, can one guarantee that there is a lattice point of $P$ in the circumscribed parallelepiped of $T$ (apart of the vertices of $T$)?
-
-A positive answer would still imply that $P$ can be unimodularly covered.
-}
-
-Our proof that parallelepipeds have the property (\Cref{lemma:3<4}) is based on the fact that they have only three (pairs of) normal vectors.
-The proof, and the property of being IDP, fail if there are four of them:
+\end{example}
 
@@
 \end{example}
+
+An affirmative answer to the following question (weaker than the circumscribed parallelepiped property) would still  imply that smooth $3$-polytopes can be unimodularly covered, and hence\Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three is:
+
+\begin{question}
+If $T$ is an empty tetrahedron contained in a smooth 3-polytope $P$, can one guarantee that there is a lattice point of $P$ in the circumscribed parallelepiped of $T$ (apart of the vertices of $T$)?
+\end{question}
 
@@
 
-\paco{removed references to Coro 1.8}
-
 Let us now show how to derive \Cref{coro:prismatoid} from this theorem.
@@
 We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq 0\}$ and $V_2=\{f_p(x) \geq 0\}$. 
-%\giulia{we need to make these closed, so that the case $P$ one dimensional fits into this same proof}\paco{ok}
 In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants. 

### 2022-02-25 unimodular-covers.tex
Add files via upload

@@
 \usetikzlibrary{intersections}
-\usetkzobj{all}
+%\usetkzobj{all}
 
@@
 
-These results leave open the following important questions:
+These results leave open the following important questions regarding parallelotopes:
 
@@
 \paco{added ref to Gubeladze}
-Let us also mention the following recent results of Gubeladze~\cite{Gubeladze}.
-In part (2), a polytope $P$ is called \emph{very ample} if the vertex cone $\cone(P-v)$ is contained in the translation $P-v$, for every vertex $v$ of $P$.
-
-\begin{theorem}[Gubeladze~\cite{Gubeladze}]
-Let $P$ be a lattice $3$-polytope. 
-\begin{enumerate}
-\item If $P=E\cap \Z^3$ for some (solid) ellipsoid $E$, then $P$ has a unimodular cover.
-\item If $P$ is \emph{very ample} then the boundary of $P$ has a unimodular cover. That is, there are unimodular simplices contained in $P$ whose union contains $\partial P$.
-\end{enumerate}
-\end{theorem}
+\giulia{changed it}
+
+To date there are no known example of IDP $3$-polytopes without a unimodular cover.
+Recent work of Gubeladze~\cite{Gubeladze} shows that another class of IDP 3-polytopes has unimodular covers, lattice polytopes which are convex hulls of all lattice points inside an ellipsoid.
+It remains open in general whether the classes are distinct in dimension $3$.
+%Let us also mention the following recent results of Gubeladze~\cite{Gubeladze}.
+%In part (2), a polytope $P$ is called \emph{very ample} if the vertex cone $\cone(P-v)$ is contained in the translation $P-v$, for every vertex $v$ of $P$.
+
+%\begin{theorem}[Gubeladze~\cite{Gubeladze}]
+%Let $P$ be a lattice $3$-polytope. 
+%\begin{enumerate}
+%\item If $P=E\cap \Z^3$ for some (solid) ellipsoid $E$, then $P$ has a unimodular cover.
+%\item If $P$ is \emph{very ample} then the boundary of $P$ has a unimodular cover. That is, there are unimodular simplices contained in $P$ whose union contains $\partial P$.
+%\end{enumerate}
+%\end{theorem}
 

### 2022-02-26 unimodular-covers.tex
edited ref to Gubeladze and updated bibliography

@@
 \usepackage[colorlinks=true]{hyperref}
+\usepackage{doi}
 %\usepackage{pictex}
@@
 In dimension three, however, the conjecture fails without the smoothness assumption.
-Indeed, if we let $P=Q$ be any non-unimodular \emph{empty tetrahedron}, then $P$ is obviously a weak Minkowski summand of $Q$ but the pair $(P.Q)$ is not IDP.
+Indeed, if we let $P=Q$ be any non-unimodular \emph{empty tetrahedron}, then $P$ is obviously a weak Minkowski summand of $Q$ but the pair $(P, Q)$ is not IDP.
 By an empty tetrahedron we mean a lattice tetrahedron containing no lattice points other than its vertices (see the proof of \Cref{lemma:corner} for a classification of them).
@@
 As we note in \Cref{prop:mixedIDP}, if the Cayley sum of $P$ and $Q$ is IDP then the pair $(P,Q)$ is IDP.
-In particular, the following statement from \Cref{sec:cayley} is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:
+Hence, the following statement proved in \Cref{sec:cayley} is stronger than the afore-mentioned result of \cite{Fakhruddin,HNPS2008,Ogata}:
 
@@
 This has the following consequence, also proved in \Cref{sec:cayley}.
-A  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
-A polytope has width $1$ if its vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is the same as being ($SL(\Z,d)$-equivalent to) a Cayley sum.
+Here a  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets. 
+%A polytope has width $1$ if its vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is the same as being ($SL(\Z,d)$-equivalent to) a Cayley sum.
 
@@
 
-\paco{deleted former corollary}
+%\paco{deleted former corollary}
 %\Paco{
@@
 \giulia{changed it}
-
-To date there are no known example of IDP $3$-polytopes without a unimodular cover.
-Recent work of Gubeladze~\cite{Gubeladze} shows that another class of IDP 3-polytopes has unimodular covers, lattice polytopes which are convex hulls of all lattice points inside an ellipsoid.
-It remains open in general whether the classes are distinct in dimension $3$.
+\paco{changed again, adding more references}
+
+Let us mention that recent work of Gubeladze~\cite{Gubeladze} shows another class of 3-polytopes admitting unimodular covers: the convex hulls of all lattice points inside an ellipsoid; these had previously been shown to be IDP by Bruns, Gubeladze and Michalek~\cite{BGM}.
+To date there are no known examples of IDP $3$-polytopes without a unimodular cover, although such polytopes exist in higher dimension~\cite{BG}.
 %Let us also mention the following recent results of Gubeladze~\cite{Gubeladze}.
@@
 
-\paco{removed closing sentence}
+%\paco{removed closing sentence}
 %\medskip
@@
 
-Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain. Then:
+Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain.
+In what follows, the width of a functional $f$ on a set $S$, denoted $\width_f(S)$ is defined as the difference $\sup_{x\in S} f(x)-\inf_{x\in S} f(x)$
+Then:
 \[
@@
 Smooth Centrally Symmetric Polytopes in Dimension 3 are IDP,
-\emph{Ann.~Combin.}, in press, 2019.
-\url{https://doi.org/10.1007/s00026-019-00418-x}
+\emph{Ann.~Combin.}, 23 (2019), 255--262.
+\doi{10.1007/s00026-019-00418-x}
+
+\bibitem{BG}
+Winfried Bruns and Joseph Gubeladze, Normality and covering properties of affine semigroups, \emph{J. Reine Angew. Math.} 510 (1999), 151--178. \doi{10.1515/crll.1999.044}
 
@@
   Springer-Verlag, 2009, XIV, 461 p. 52 illus.
+
+\bibitem{BGM}
+  Winfried Bruns, Joseph Gubeladze, and Mateusz Michalek. Quantum Jumps of Normal Polytopes. \emph{Discrete Comput. Geom.}, 56(1) (2016), 181--215. \doi{10.1007/s00454-016-9773-7}
 
@@
 Normal polytopes and ellipsoids,
-preprint, \href{https://arxiv.org/abs/2012.11846}{arXiv:2012.11846}, December 2012.
+\emph{Elect. J. Combin.}, 28:4 (2021), P4.7.
+\doi{10.37236/10338}
 
@@
 Convex-normal (Pairs of) Polytopes, 
-\emph{Canad. Math. Bull.} 60 (2017), 510--521.
+\emph{Canad. Math. Bull.} 60:3 (2017), 510--521. \doi{10.4153/CMB-2016-057-0}
 
@@
 Christian Haase, Andreas Paffenholz, Lindsay C. Piechnik, Francisco Santos. Existence of unimodular triangulations - positive results. 
-%Preprint May 2014, updated December 2017, 89 pages. 
- \emph{Mem. Amer. Math. Soc.}, to appear.
- Available as arXiv preprint \href{https://arxiv.org/abs/1405.1687}{arXiv:1405.16878}
+ \emph{Mem. Amer. Math. Soc.},  Vol.~270 1N.~321, American Math. Society, 2021. DOI: \doi{10.1090/memo/1321}
+%ISBNs: 978-1-4704-4716-8 (print); 978-1-4704-6530-8 (online)
 
@@
 Christian Haase, Benjamin Nill, Andreas Paffenholz, and Francisco Santos, Lattice points in Minkowski sums, 
-\emph{Electron.~J.~Combin.}, 15 (2008), no. 1, Note 11, 5 pp.
+\emph{Electron.~J.~Combin.}, 15:1 (2008), Note 11, 5 pp. \doi{10.37236/886}
 
@@
 On primitive subdivisions of an elementary tetrahedron, 
-\emph{Pacific J.~Math.} \textbf{211} (2003), no.~1,  123--155.
+\emph{Pacific J.~Math.} \textbf{211} (2003), no.~1,  123--155. \doi{10.2140/pjm.2003.211.123}
 
@@
 3.
-MR MR2144157
+\doi{10.14760/OWR-2004-39}
 
@@
 Organized by Christian Haase, Takayuki Hibi and Diane MacLagan.
+\doi{10.14760/OWR-2007-39}
 
@@
 \newblock {Multiplication maps of complete linear systems on projective toric surfaces.}
-\newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006.
+\newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006. \doi{10.4036/iis.2006.93}
 
@@
 \emph{Ann.~Math.~(2)}, 176 (July 2012), 383--412. 
-DOI: 10.4007/annals.2012.176.1.7
+\doi{10.4007/annals.2012.176.1.7}
 
@@
 \newblock {Unimodular triangulations of dilated 3-polytopes}, 
-\newblock {\em Trans. Moscow Math. Soc.} (2013), 293--311. % \MR{3235802}
+\newblock {\em Trans. Moscow Math. Soc.} (2013), 293--311. \doi{10.1007/s00493-014-2959-9}
 
@@
 Lattice tetrahedra.
-\emph{Canadian J.~Math.} 16 (1964), 389--396.
+\emph{Canadian J.~Math.} 16 (1964), 389--396. \doi{10.4153/CJM-1964-040-2}
 

### 2022-02-26 unimodular-covers.tex
updated paco's grant info

@@
 Work of F.
-Santos is also supported by grants  MTM2017-83750-P/AEI/10.13039/501100011033 and PID2019-106188GB-I00/AEI/10.13039/501100011033 of the Spanish State Research Agency and by project CLaPPo (21.SI03.64658) of Universisdad de Cantabria and Banco Santander}
+Santos is also supported by grants PID2019-106188GB-I00 funded by MCIN/AEI/10.13039/501100011033 and by FPU19/04163 of the Spanish Government and by project CLaPPo (21.SI03.64658) of Universidad de Cantabria and Banco Santander}
 

### 2022-03-01 unimodular-covers-CombTheory-submitted.tex
Add files via upload

--- 2022-02-26 unimodular-covers.tex
@@
 
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\giulia}[1]{\todo[size=\tiny,color=blue!30]{#1 \\ \hfill --- G.}}
-\newcommand{\Giulia}[1]{\todo[size=\tiny,inline,color=blue!30]{#1 \\ \hfill --- G.}}
-
-\newcommand{\paco}[1]{\todo[size=\tiny,color=green!30]{#1 \\ \hfill --- P.}}
-\newcommand{\Paco}[1]{\todo[size=\tiny,inline,color=green!30]{#1 \\ \hfill --- P.}}
-
 \date{\today}
 
-\author[G.~Codenotti]{Giulia Codenotti}
-\author[F.~Santos]{Francisco Santos}
+%\author[G.~Codenotti]{Giulia Codenotti}
+%\author[F.~Santos]{Francisco Santos}
 
-\address[G.~Codenotti]
-{
-Institut f\"ur Mathematik, Freie Universit\"at Berlin, Germany
-}
-\email{giulia.codenotti@fu-berlin.de}
+%\address[G.~Codenotti]
+%{
+%Institut f\"ur Mathematik, Freie Universit\"at Berlin, Germany
+%}
+%\email{codenotti@math.uni-frankfurt.de}
 
-\address[F.~Santos]
-{
-Department of Mathematics, Statistics and Computer Science, University of Cantabria, Spain
-}
-\email{francisco.santos@unican.es}
+%\address[F.~Santos]
+%{
+%Department of Mathematics, Statistics and Computer Science, University of Cantabria, Spain
+%}
+%\email{francisco.santos@unican.es}
 
-\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
-Work of F.
+%\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
+%Work of F.
 Santos is also supported by grants PID2019-106188GB-I00 funded by MCIN/AEI/10.13039/501100011033 and by FPU19/04163 of the Spanish Government and by project CLaPPo (21.SI03.64658) of Universidad de Cantabria and Banco Santander}
@@
 \title{Unimodular covers of 3-dimensional parallelepipeds and Cayley sums}
+\author{}
 
@@
 \maketitle
-
-%\tableofcontents
 
@@
 When the normal fan of a polytope $Q$ refines that of another polytope $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is easily seen to be equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$.
-%\paco{did not find this in the google books version of \cite{Grunbaum}; found it in Postnikov-Reiner-Williams, Tsuchiya, a survey by Rekha Thomas, and other places.
-Perhaps we can leave it without reference}
 This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor defined by $P$ on  $X_Q$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that what we here call  ``weak Minkowski summand'' is simply called ``Minkowski summand'' there).
@@
 Here a  \emph{prismatoid} is a polytope whose vertices all lie in two parallel facets.
-%A polytope has width $1$ if its vertices lie in two \emph{consecutive} parallel lattice hyperplanes. Observe that this is the same as being ($SL(\Z,d)$-equivalent to) a Cayley sum.
 
@@
 
-%\paco{deleted former corollary}
-%\Paco{
-%I am a bit puzzled.
-After the next corollary we say that \cite[Cor.~4.2]{SantosZiegler} proves a case of this, but in fact it proves the whole of it, in a stronger form: it does not need width one, and it gives a "dissection" and not a mere cover (dissection= cover by simplices with disjoint interiors).
-%I propose to delete it:
-%}
-%{\color{red}
-%\begin{corollary}
-%%\label{coro:width1}
-%Every integer dilation $kP$, $k\ge 2$, of a lattice $3$-polytope $P$ of width $1$ has a unimodular cover.
-%\end{corollary}
-%
-%A special case of the latter are integer dilations of empty tetrahedra. That their dilations have
-%unimodular covers is \cite[Cor.~4.2]{SantosZiegler} (and is also implicit in \cite{KantorSarkaria}).
-%}
-
-\paco{added ref to Gubeladze}
-\giulia{changed it}
-\paco{changed again, adding more references}
-
 Let us mention that recent work of Gubeladze~\cite{Gubeladze} shows another class of 3-polytopes admitting unimodular covers: the convex hulls of all lattice points inside an ellipsoid; these had previously been shown to be IDP by Bruns, Gubeladze and Michalek~\cite{BGM}.
 To date there are no known examples of IDP $3$-polytopes without a unimodular cover, although such polytopes exist in higher dimension~\cite{BG}.
-%Let us also mention the following recent results of Gubeladze~\cite{Gubeladze}.
-%In part (2), a polytope $P$ is called \emph{very ample} if the vertex cone $\cone(P-v)$ is contained in the translation $P-v$, for every vertex $v$ of $P$.
 
-%\begin{theorem}[Gubeladze~\cite{Gubeladze}]
-%Let $P$ be a lattice $3$-polytope. 
-%\begin{enumerate}
-%\item If $P=E\cap \Z^3$ for some (solid) ellipsoid $E$, then $P$ has a unimodular cover.
-%\item If $P$ is \emph{very ample} then the boundary of $P$ has a unimodular cover. That is, there are unimodular simplices contained in $P$ whose union contains $\partial P$.
-%\end{enumerate}
-%\end{theorem}
-
-%\paco{removed closing sentence}
-%\medskip
-%We believe that the $3$-polytopes in all these statements have unimodular triangulations, but this remains an open question.
-
-\subsection*{Acknowledgements:} We would like to thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and Christian Haase for helpful discussions.
+%\subsection*{Acknowledgements:} We would like to thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and Christian Haase for helpful discussions.
 
@@
 By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
-%
-
-%\begin{figure}[htb]
-%\includegraphics[scale=.28]{Whites}
-%\caption{The parallelograms $\conv(p_1p_2q_3q_4)$ and $\conv(q_2q_1p_4p_3)$. The large red dots are lattice points.}
-%\label{fig:corner}
-%\end{figure}
 
@@
 This finishes the proof of the claim, and of the lemma.
-%
-%Let $f\in (\R^3)^*$ be any functional. Fix $i \in [4]$; observe that if $f(q_i) > f( p_j)$ for all $j\neq i$, then 
-%\begin{gather*}
-%f(q_i) \overset{(i)}{>} f(q_j) \overset{(ii)}{>} f(p_i) \\
-%f(q_i) \overset{(iii)}{>} f(p_j) \overset{(iv)}{>} f(p_i) \\
-%f(p_j)\overset{(v)}{>}f(q_k)
-%\end{gather*}
-%for all $j \neq i$, $k \notin \{j, i\}$. 
-%
-%Indeed, for any $\{i,j,k,l\}=[4]$ we have 
-%
-%\begin{align*}
-%f(q_i) > f( p_j) \Leftrightarrow \frac12 f(p_l+&p_k -p_j-p_i) > 0  \Leftrightarrow f(q_j) > f(p_i)\\
-%& \Updownarrow \\
-% f(q_l) < f(p_k) \Leftrightarrow \frac12 f(-p_l-&p_k +p_j+p_i) < 0 \Leftrightarrow f(q_k) < f(p_l),
-%\end{align*}
-%which shows the equivalence of $(i)$, $(iii)$ and $(v)$; note that $(iii)$ is our assumption.
-Inequalities $(ii)$ and $(iv)$ follow by taking the sum of two inequalities of the form $\frac12 f(p_l+p_k -p_j-p_i) > 0$.
-%
-%Using this observation, we can show that each of pairs of parallel facets of $P$ cuts out at most one of the $q_i$s:
-%if $q_i$ lies outside of $P$, it must violate a facet-defining inequality $f(x) \leq a$.Thus $f(q_i) > a \geq f(p_j)$, and by our observation, it follows that all remaining $q_j$ satisfy the inequality $f(x) \leq a$.
-Since $P$ is a parallelogram, there is a facet with inequality $f(x) \geq b$; by what is said above, the minimum value of $f(x)$ is attained at $p_i$, which is in $P$.
-Thus $f(q_j) > f(p_i) \geq b$, that is, the inequality is satisfied by all the $q_j$.
-%
-%Since at most one of the $q_i$ does not satisfy the inequalities corresponding to a pairs of parallel facets of $P$, and there are only three pairs of facets, one of $q_i$s must satisfy all the inequalities and therefore be inside of $P$.
+
 \end{proof}
-%\giulia{here it may also be nice to add a picture}
 
@@
 By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point other than the $p_i$s, which we denote by $u$.
-%Observe that from the proof of \Cref{lemma:corner} we know more: $u$ lies in the face $\conv(p_1, p_2, q_4)$.\giulia{we need this for the volume computation below.
-perhaps it should be made explicit in the statemente of the lemma...}
 %
@@
 The $S_i$ may not be empty, but we can triangulate them into empty tetrahedra, which by inductive hypothesis they can be covered unimodularly.
-%
-%Observe that  $T'_4$ and/or one of the $S_i$ can degenerate to a triangle, if $u$ lies in $\conv(p_1,p_2,p_3)$, but this does not affect the argument.
-%
-%: indeed, $\Vol(S_3)=\Vol(T'_4)$, since these tetrahedra have a common face $\conv(p_1, p_2, l)$ and the remaining vertex is at distance one from this face in both.
-Now $T'_4$ is contained in $T_4$, and we know that each corner tetrahedron $T_i$ has volume half that of $T$.
-The remaining two tetrahedra $S_1$ and $S_2$ must therefore have volume summing to that of $T$, and therefore each has volume strictly less than $T$.
 \end{proof}
@@
 Arbitrarily triangulate the parallelepiped into empty lattice tetrahedra and apply \Cref{coro:coverpara} to these tetrahedra.
-%Let $P \subseteq \R^3$ be a lattice parallelepiped.
-It can be triangulated into empty lattice tetrahedra, simply by taking a triangulation which uses all lattice points as vertices.
-If all the tetrahedra in the triangulation are unimodular we are done.
-Otherwise, suppose $T$ is an empty tetrahedron in the triangulation which is not unimodular.
-\Cref{lemma:3<4} guarantees that one of the corner tetrahedra $T_i$ of the parallelepiped $C(T)$ circumscribed to $T$ is contained in $P$.
-Without loss of generality, suppose $T_4 = \conv(p_1, p_2, p_3,q_4)$ is in $P$.
-By \Cref{lemma:corner}, we know that $T_4$ contains a lattice point, which we denote by $u$.
-%
-%
-%This allows us to cover $T$ with lattice simplices of smaller volume recursively, until it is covered by unimodular ones.
 \end{proof}
 
-%\begin{remark}
-%\label{rem:octahedra_and_tri-prisms}
-
-%We finish with a question which, if answered in the positive, would imply \Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three.
 Let us say that a lattice $3$-polytope $P$ \emph{has the circumscribed parallelepiped property} if it satisfies the conclusion of \Cref{lemma:3<4}:  ``for every empty tetrahedron $T$ contained in $P$ at least one of the four corner tetrahedra in $C(T)$ is
@@
 
-%The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ being $(d+1)$-dimensional.
-
 The Cayley Trick also provides the following canonical bijections:
@@
 contains a lattice point. See \Cref{fig:strips}.
-%\[
-%(\int(p) + \vecline q) \cap P
-%\qquad \text{or} \qquad
-%\int(q + \vecline p) \cap Q
-%\]
-%contains a lattice point.
 \end{lemma}
@@
 In particular, we have covered $T$ by the three tetrahedra in $\mathcal T^-$, which are of smaller volume and hence have unimodular covers by inductive assumption.
-%
-%A unimodular triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained simply by unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$).
-This leaves simplices of type $(2,2)$.
-In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.
-%
-%Let $S=\{p_1, p_2\} + \{q_1, q_2\}$ be such a parallelogram.
-By \Cref{lemma:cayley} we know that there is a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $S$, w.l.o.g.
-suppose it is $p+q_1$.
-Then we can cover $S$ by the triangle $\{p_1, p_2, p\} + q_1$ and parallelograms $\{p_1, p\}  + \{q_1, q_2\}$ and $\{p_2, p\} + \{q_1, q_2\}$.
-As above, the triangle can be unimodularly triangulated, while the volume of the two parallelograms sums up to the volume of $S$, and therefore each has volume strictly smaller than $S$.
-By iterating this procedure, we can cover $S$ by unimodular cells.
 \end{proof}
@@
 the smoothness of the prismatoid implies that every edge of the form $[u\times \{0\}, v\times \{k\}]$ has lattice points in all slices. 
-%In \Cref{coro:width1}, since $P$ has width one, $P\cong \cayley(Q_1, Q_2)$ for some $Q_1$ and $Q_2$. 
 Hence,
@@
 
-%\begin{claim}
-%$b_2$ and $t_2$ are non-zero and have the same sign. W.l.o.g.~we assume they are positive.
-%\end{claim}
-%
-%\begin{proof}
-%If, say, $b_2 \le 0 \le t_2$.
-Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$).
-This implies the width w.r.t to $f_q$, the functional constant on $q_1q_2$, of any segment with vertices in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
-%\end{proof}
-
 \begin{claim}
@@
 \end{claim}
-
-%\begin{claim}
-%The slopes of $b$ and $t$ must either be both strictly larger than the slope of $q$ or both be strictly smaller. W.l.o.g.~we assume they are larger.
-%\end{claim}
-%
-%\begin{proof}
-%If, say, $b$ has positive slope and $t$ negative. Then since $t$ intersects the vertical line $x=0$ spanned by $p$ below $p_2$, and $b$ intersects it above $p_1$, the slopes of $b$ and $t$ imply that
 
@@
 \end{figure}
-
-%\Giulia{fix it to look more unimodular at the bottom}
-%\begin{figure}[htb]
-%  \centering
-%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-%  1]
-%%coordinate axes
-%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%%        \tkzLabelY[orig=false,label options={font=\tiny}]
-%%        \tkzGrid
-%%        \tkzDrawX
-%%        \tkzDrawY
-%
-%%edges b, t
-%  \node[
-%%label={[shift={(0,-0.67)}]$b_1$}, 
-%mynode] (b1) at (-3,-2) {}; 
-%  \node[
-%%label={[shift={(0,-0.67)}]$b_2$}, 
-%mynode] (b2) at (2, 0.2) {};  
-%  \node[
-%%label={[shift={(0,0)}]$t_1$}, 
-%mynode] (t1) at (-2.7, 1) {};  
-%  \node[
-%%label={[shift={(0,0)}]$t_2$}, 
-%mynode] (t2) at (1.6,0.7) {};  
-%
-%%shaded region where r can be  
-%  \draw[fill=gray!80, nearly transparent] (2.93,0.61) -- (0,0.81)  -- (0,-0.68) -- cycle;
-%
-%
-%%p
-%  \node[
-%%label={[shift={(0,-0.67)}]$p_1$}, 
-%mynode] (p1) at (0,-1) {}; 
-%  \node[
-%%label={[shift={(0,0)}]$p_2$}, 
-%mynode] (p2) at (0,1) {};  
-%%  \node[{p}]  at (-0.3,0.95)
-%
-%
-%%q
-%  \node[
-%%label={[shift={(0,-0.67)}]$q_1$}, 
-%mynode] (q1) at (-1.7,-0.4) {}; 
-%  \node[
-%%label={[shift={(0,-0.67)}]$q_2$}, 
-%mynode] (q2) at (0.3,-0.4) {};  
-%
-%%segments b, t, p, q
-%  \draw[ name path = b, black, thick] (b1) node[below, xshift=1.7cm, yshift=0.7cm] {b} -- (b2); 
-%  \draw[name path = t, black, thick] (t1) node[below, xshift=1.2cm, yshift=-0.1cm] {t} -- (t2);
-%  \draw[name path = p, black, thick] (p1) node[left, xshift=0 cm, yshift=1.3cm] {p} -- (p2);
-%  \draw[name path = q, black, thick] (q1) node[below, xshift=0.4cm, yshift=0cm] {q} -- (q2);
-%
-%
-%%prolonged lines through b and t
-% \tkzDrawLine[add=0.2 and 0.3, color=black, dashed](b1, b2);
-% \tkzDrawLine[ add=0.3 and 0.4, color=black, dashed](t1, t2);
-% 
-%%functional f_q
-%  \tkzDefLine[orthogonal=through p1](p1,p2)
-%  \tkzDrawLine[add = 2 and 1,color=black, dashed](p1,tkzPointResult)
-%  \node[label={\small $\{f_q(x)=-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab1) at (3.3, -1) {};
-%
-%
-%  \tkzDefLine[orthogonal=through p2](p1,p2) 
-%  \tkzDrawLine[add = 2 and 1,color=black, dashed](p2,tkzPointResult) 
-%    \node[label={\small $\{f_q(x)=w-1\}$}, xshift=0.4cm, yshift=-0.2cm] (lab2) at (3.3, 1) {};
-%
-%\end{tikzpicture}
-%\caption{Picture for Claim 4.2}
-%\label{fig:claim2}
-%\end{figure}
 
@@
 
-%\begin{figure}
-%\includegraphics[ width=.7\textwidth]{Claim43.jpg}
-%\caption{Figure for Claim 4.3}
-%\label{fig:claim3}
-%\end{figure}
 \begin{figure}[htb]
@@
 
-%Starting here we assume index 2.
-%
-%\begin{claim}
-%The slopes of $t$ and $b$ are strictly smaller than one and that of $r$ strictly greater than one.
-%\end{claim}
-%
-%\begin{proof}
-%Otherwise the points $(0,1/2)$, $(0,-1/2)$, $(1/2,0)$, respectively, would be in the corresponding band.
-%\end{proof}
-%
-%We are now ready to show a contradiction.
-Let $u$ and $v$ be the end-points of the translated copy of $r$ in $Q$, with $\vec r=v-u$.
-Let $d$ be the line of slope $1$ through $u$, which is a lattice line.
-Let $d'$ be the next parallel lattice line above $d$.
-By Claim A.3.1, the intersection of $d'$ with the vertical line through $u$ is above $Q$.
-Since the slope of $t$ is less than one, the same happens for every point of $d'$ to the right of $u$.
-In particular, $v$ must be below $d'$ and hence it must be on or below $d$.
-This contradicts the fact that the slope of $r$ is strictly greater than one.
-QED
-
 Let $g$ be the primitive lattice functional constant on $[p_1, q_2]$ (and therefore constant also on $[q_1, q_1+q_2-p_1]$).
@@
 $g(t_1) > g(t_2)$, $g(b_1) > g(b_2)$, and $g(r_1) < g(r_2)$.
-%The slopes of $t$ and $b$ are strictly smaller than the slope of  $[p_1, q_2]$ and that of $r=[r_1, r_2]$ (and $l=[l_1, l_2]$) is strictly greater than it.
 \end{claim}
@@
 
-%
-%\appendix
-%
-%\section{Old stuff}
-%
-%
-%\begin{proof}
-%%Suppose by contradiction that there is no lattice point as described in the lemma.
-Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
-The same holds for $Q$ and its corresponding strip.
-Thus $P$ has two edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $e_1$ and $e_2$.
-The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.
-%
-%Let us fix coordinates in which $p_1p_2$ and $q_1q_2$ are a vertical and horizontal segment of length one intersecting at the origin. That is, there are numbers $\alpha, \beta \in (0,1)$ such that
-%\[
-%p_1=(0,-\alpha), \quad
-%p_2=(0,1-\alpha), \quad
-%q_1=(-\beta,0), \quad
-%q_2=(1-\beta,0).
-%\] 
-%The ambient lattice is a certain lattice containing $p_1,p_2,q_1,q_2$ (the case of index two is when $\alpha=\beta=1/2$ and the lattice is spanned by these four points).
-%
-%Suppose by contradiction that there is no lattice point as described in the lemma.
-Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon.
-The same holds for $Q$ and its corresponding strip.
-Thus $P$ has two primitive edges which each have one vertex on each side of the strip $p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right).
-The same holds for $Q$ and the strip $q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle$, and we call the edges  $b, t$ (with $b$ below and $t$ above).
-To make things more explicit, we assume the following coordinates for the edge-vectors:
-%\[
-%\vec l = (l_1,l_2), \quad
-%\vec r= (r_1,r_2), \quad
-%\vec b= (b_1,b_2), \quad
-%\vec t= (t_1,t_2).
-%\]
-%Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).
-%
-%
-%
-%
-%
-%
-%Since the normal fan of $Q$ refines that of $P$, $Q$ will contain edges $e_1'$ and $e_2'$ parallel to $e_1$ and $e_2$ respectively.
-These edges of $Q$ are all distinct and appear alternating in $Q$, that is, they appear in the order $f_1, e_1', f_2, e_2'$.
-Indeed, consider the quadrilateral $S$ with vertices $q_1, p_1+l, q_2, p_2+l$, where $l$ is a lattice point chosen so that the segments $q_1q_2$ and $p_1+l,p_2+l$ intersect in their interior; we can further require that $\conv( q_1, p_1+l, q_2)$ is unimodular \giulia{not sure if we use it}.
-$S$ lies inside the $Q$-strip, while its translate $S-l$ lies inside the $P$-strip.
-Thus in order to have no lattice points in the strips, we must have that the edge $f_i$ separates the vertex $p_i +l$ of $S$ from the remaining three, and  the same for the edge $e_i +l$ and vertex $q_i$ of $S$.
-This is equivalent to saying that the normals to the edges $f_1, e_1, f_2, e_2$ lie in different open cones of the normal fan of $S$, in this cyclic order.
-%
-%- both $f_1$ or $f_2$ must meet the line through $q_1$ and $q_2$, else they would not cut out the lattice point they need to. same goes for $e_1$ and $e_2$ and the line through $p_1$ and $p_2$.
-%
-%claim: the segments $e_1 + l$ and $e_2 + l$ respectively intersect both $f_1$ and $f_2$. 
-%
-%claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both. 
-%
-%
-%
-%
-%
-%%Consider the segments on the lattice lines parallel to $p_1 p_2$ cut out by the edges $f_1$ and $f_2$ \giulia{this should be formulated more comprehensibly}.
-Since these segments do not contain lattice points, they must have lattice length stricly less than $1$ (or the width wrt $q$ is less than $k$, the index of $p,q$), which means that the smallest width in direction of $p_1 p_2$ between a "consecutive" vertices of $f_1$ and of $f_2$ must also be strictly less than $k$.
-%
-%%Further, we observe that both vertices of $f_1$ (or $f_2$) cannot lie strictly above the line through $q_1$ and $q_2$, else the strip would contain a lattice point. 
-%\end{proof}
-%
-%
-%
-%\begin{figure}[htb]
-%  \centering
-%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-%  .7]
-%%coordinate axes
-%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%%        \tkzLabelY[orig=false,label options={font=\tiny}]
-%        \tkzGrid
-%%       \tkzDrawX
-%%       \tkzDrawY
-%
-%%hexagon
-%  \node[mynode] (A) at (0,-1) {}; 
-%%  \node[mynode] (B) at (1,1) {};  
-%  \node[mynode] (C) at (2,3) {};  
-%  \node[mynode] (D) at (0,1) {};  
-%%  \node[mynode] (E) at (-1,-1) {};  
-%  \node[mynode] (F) at (-2,-3) {};  
-%
-%%other points
-%%  \node[mynode] () at (0,-1) {}; 
-%
-%
-%  \draw[black] (A) -- (C) -- (D) -- (F) -- (A);
-%  \draw[black, thick] (A) -- (D);
-%\end{tikzpicture}
-%%
-%\begin{tikzpicture}[mynode/.style={circle, inner sep=1.5pt, fill=red}, scale =
-%  .7]
-%%coordinate axes
-%   \tkzInit[xmax=3.5,ymax=3.5,xmin=-3.5,ymin=-3.5]
-%%        \tkzLabelX[orig=false,label options={font=\tiny}]
-%%        \tkzLabelY[orig=false,label options={font=\tiny}]
-%        \tkzGrid
-%%        \tkzDrawX
-%%        \tkzDrawY
-%
-%%hexagon
-%  \node[mynode] (A) at (1,0) {}; 
-%  \node[mynode] (B) at (2,1) {};  
-%  \node[mynode] (C) at (3,3) {};  
-%  \node[mynode] (D) at (-1,0) {};  
-%  \node[mynode] (E) at (-2,-1) {};  
-%  \node[mynode] (F) at (-3,-3) {};  
-%
-%  \node[mynode] (p1) at (-1,0) {}; 
-%  \node[mynode] (p2) at (1,0) {};  
-%
-%  \draw[black] (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- (A);
-%  \draw[black, thick] (p1) -- (p2);
-%\end{tikzpicture}
-%
-%\centerline{\hfill $P$ \hfill\hfill $Q$ \hfill}
-%
-%\caption{Partial counter-example to \Cref{lemma:cayley}.
-With respect to the standard lattice it is not a counter-example because (a) the edges $p_1p_2$ and $q_1q_2$ are not primitive and (b) the origin is a lattice point in the interior of both bands (but it is the only one).
-%%
-%With respect to the lattice of index two it is not a counter-example either, because two vertices of $Q$ are not in that lattice.
-%}
-%\label{fig:width3}
-%\end{figure}
-
 \end{document}

### 2022-12-12 unimodular-covers-CombTheory-reviewing.tex
worked on proof of 3.2

--- 2022-03-01 unimodular-covers-CombTheory-submitted.tex
@@
 \newcommand{\Image}{\ensuremath{\mathrm{Im}}\hspace{1pt}}
+
+\usepackage[colorinlistoftodos]{todonotes}
+
+\newcommand{\giulia}[1]{\todo[size=\tiny,color=blue!30]{#1 \\ \hfill --- G.}}
+\newcommand{\Giulia}[1]{\todo[size=\tiny,inline,color=blue!30]{#1 \\ \hfill --- G.}}
+
+\newcommand{\paco}[1]{\todo[size=\tiny,color=green!30]{#1 \\ \hfill --- P.}}
+\newcommand{\Paco}[1]{\todo[size=\tiny,inline,color=green!30]{#1 \\ \hfill --- P.}}
 
@@
 
+\giulia{unless b or t are equal to q}
+
 \begin{proof}
@@
 
-We assume w.l.o.g. that the maximum on $t$ (and hence on $b$) is achieved in $V_2$, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).
+We assume without loss of generality that the maximum on $t$ (and hence on $b$) is achieved in $V_2$, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).
+Otherwise the following considerations can be applied to $V_1$.
 
@@
 \label{claim:r}
-Assume w.l.o.g.~that $b$ and $t$ either are parallel or their affine spans cross in $V_2$. Then, 
+Assume without loss of generality that $b$ and $t$ either are parallel or their affine spans cross in $V_2$ (if they cross in $V_1$, the same claim can be reworded for the segment $l$). Then, 
 \begin{enumerate}
@@
 \[
-g([p_1, q_2])= g( [q_1, q_1+q_2-p_1]) +1. 
+g([p_1, q_2])= g( [q_1, q_1+q_2-p_1]) -1. 
 \]
+\giulia{changed the +1 to a -1, so that the following claim is true}
 

### 2022-12-16 unimodular-covers-CombTheory-reviewing.tex
Add files via upload
Rewiewing process for Comb Theory
finished a pass with review Y

@@
 We abbreviate this by saying that ``$P$ is IDP''.
-Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (normality of toric varieties).
+Being IDP is interesting in the context of  both enumerative combinatorics (Ehrhart theory) and algebraic geometry (projective normality of toric
+varieties).
 It falls into a hierarchy of several properties each stronger than the previous one; see, e.g., \cite[Section 2.D]{BGbook}, \cite[Sect.
@@
 When the normal fan of a polytope $Q$ refines that of another polytope $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is easily seen to be equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$.
-This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor defined by $P$ on  $X_Q$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Prop.~6.3.12]{CLS}, but observe that what we here call  ``weak Minkowski summand'' is simply called ``Minkowski summand'' there).
+\Giulia{RefY says to add the definition of Minkowski sum here, rather than have the definition in section 3; I think it would break up the flow of the introduction.
+Perhaps a compromise is to add a comment like, "where + denotes the Minkowski sum defined in Section 3".
+Then again, the Cayley sum is defined both later in the intro and in section 3.
+What do you think?}
+This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor defined by $P$ on  $X_Q$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Thm.~6.3.12]{CLS}, but observe that what we here call  ``weak Minkowski summand'' is called ``$\N$-Minkowski summand'' there).
 
@@
 More precisely, they show that any such polytope can be covered by lattice 
-parallelepipeds and unimodular simplices, both of which are trivially IDP.
+parallelepipeds (affine images of $3$-cubes) and unimodular simplices, both of which are trivially IDP.
 %
@@
 \[
-\cayley(P,Q) := \conv(P\times\{0\} \cup Q \times \{1\}) \subset \R^3.
+\cayley(P,Q) := \conv(P\times\{0\} \cup Q \times \{1\}) \subset \R^{d+1}.
 \]
-We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but this does not need $P$ and $Q$ to be full-dimensional.
-It only requires the linear subspaces parallel to them to span $\R^d$.
+We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but $P$ or $Q$ do not necessarily need to be fulldimensional.
+We only require the linear subspaces parallel to them to span $\R^d$.
 
@@
 \label{def:circunpara}
-Let $T$ be a tetrahedron with vertices $p_1$, $p_2$, $p_3$, and $p_4$. Consider the points $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$, $i\in [4]$, and let
+Let $T$ be a tetrahedron with vertices $p_1$, $p_2$, $p_3$, and $p_4$. Consider the points $q_i= \frac12 (p_1+p_2+p_3+p_4) - p_i$ for each $i\in \{1,2,3,4\}$, and let
 \[
-C(T)=\conv(p_i,q_i: i\in[4]).
+C(T)=\conv(p_i,q_i: i\in\{1,2,3,4\}).
 \] 
-$C(T)$ is a parallelepiped with facets $\conv(p_i, p_j, q_k, q_l)$ for all choices of $\{i,j,k,l\}=[4]$. We call it the \emph{parallelepiped circumscribed} to $T$.
+$C(T)$ is a parallelepiped with facets $\conv(p_i, p_j, q_k, q_l)$ for all choices of $\{i,j,k,l\}=\{1,2,3,4\}$. We call it the \emph{parallelepiped circumscribed} to $T$.
 
-For each $i \in [4]$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=[4]$; we call these $T_i$ the \emph{corner tetrahedra} of $C(T)$. Together with $T$ they triangulate $C(T)$.
+For each $i \in \{1,2,3,4\}$, let $T_i=\conv(q_i, p_j, p_k, p_l)$, with $\{i,j,k,l\}=\{1,2,3,4\}$; we call the $T_i$ \emph{corner tetrahedra} of $C(T)$. Together with $T$ they triangulate $C(T)$.
 \end{definition}
@@
 
+Observe that the points $q_i$ need not be lattice points. However, the following lemma shows that we can find lattice points in each corner tetrahedron.
 \begin{lemma}
@@
 \label{lemma:3<4}
-Let $P$ be a lattice parallelepiped and let $T\subset P$ be a tetrahedron. Then, at least one of the four corner tetrahedra $T_i$ of the circumscribed parallelogram $C(T)$ is fully contained in $P$.
+Let $P$ be a lattice parallelepiped and let $T\subset P$ be a tetrahedron. Then, at least one of the four corner tetrahedra $T_i$ of the circumscribed parallelepiped $C(T)$ is fully contained in $P$.
 \end{lemma}
@@
 
-We call \emph{band} any region of the form $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*$ and closed interval $[\alpha,\beta]\subset \R$.
+We call \emph{band} any region of the form $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*\setminus \{0\}$ and closed interval $[\alpha,\beta]\subset \R$.
 We claim that any band containing $T$ must contain at least three of the $q_i$s. 
@@
 \label{eq:second}
-f(q_2) = f(q_2+p_2-p_1) = f(q_2)+(f(p_2)-f(p_1)) < \alpha + (\beta-\alpha) = \beta,
+f(q_2) = f(q_1+p_1-p_2) = f(q_1)+(f(p_1)-f(p_2)) < \alpha + (\beta-\alpha) = \beta,
 \end{gather}
@@
 
-The translation of vector $\frac12 (p_1+p_2-p_3-p_4)$ sends $q_1,q_2,p_3,p_4$ to $p_2,p_1,q_4,q_3$ (in this order). By applying this to inequality \eqref{eq:third}, we obtain
+The translation by the vector $\frac12 (p_1+p_2-p_3-p_4)$ sends $q_1,q_2,p_3,p_4$ to $p_2,p_1,q_4,q_3$ (in this order). By applying this to inequality \eqref{eq:third}, we obtain
 \[
@@
 %
-This finishes the proof of the claim, and of the lemma.
+This concludes the proof of the claim, and of the lemma.
 
@@
 \end{corollary}
-
+\giulia{I would ignore comment Y.16, where if I understand correctly a different proof idea is suggested...}
 \begin{proof}
@@
 Thus, $S_1$, $S_2$ and $S_3$ cover $T$ and have volume strictly smaller than $T$.
-The $S_i$ may not be empty, but we can triangulate them into empty tetrahedra, which by inductive hypothesis they can be covered unimodularly.
+The $S_i$ may not be empty, but we can triangulate them into empty tetrahedra, which by inductive hypothesis can be covered unimodularly.
 \end{proof}
@@
 In turn, our proof that parallelepipeds have the property (\Cref{lemma:3<4}) is based on the fact that they have only three (pairs of) normal vectors. 
-In the following two examples we show a smooth $3$-polytope and two $3$-polytope with four normal vectors that do not have the property. The latter are not IDP:
+In the following two examples we show a smooth $3$-polytope and two $3$-polytopes with four normal vectors that do not have the property. The latter are not IDP:
 
@@
 In particular, none of the corner tetrahedra of $T$ is contained in $P$.
-\end{example}
+Another way to see this is to write explicitly the remaining vertices $q_i$ of the circumscribed parallelepiped, which are $(\frac{a+1}{2}, \frac{b+1}{2}, 1), (\frac{a+1}{2}, \frac{b+1}{2}, 0), (\frac{1-a}{2}, \frac{b-1}{2}, 1), (\frac{a-1}{2}, \frac{1-b}{2}, 1)$.
+None of these points are contained in $P$, and therefore none of the corner tetrahedra are either.
+\end{example}\giulia{Added the last part to answer comment Y.19}
 
@@
 
-An affirmative answer to the following question (weaker than the circumscribed parallelepiped property) would still  imply that smooth $3$-polytopes can be unimodularly covered, and hence\Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three is:
+An affirmative answer to the following question (weaker than the circumscribed parallelepiped property) would still  imply that smooth $3$-polytopes can be unimodularly covered, and hence~\Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three is:
 
@@
 If $\cayley(P,Q)$ is IDP then the pair $(P,Q)$ is mixed IDP.
-\end{proposition}
+\end{proposition} \giulia{We just called the pair IDP in the introduction... Which notation do we prefer?}
 
@@
 we call \emph{prod-simplices} in $P+Q$ the Minkowski sums $T_1+T_2$ where $T_1\subset P$ and $T_2\subset Q$ are simplices with complementary affine spans.
-A prod-simplex is \emph{unimodular} if the edge vectors from a vertex of $T_1$ and from a vertex of $T_2$ form a unimodular basis.
+A prod-simplex is \emph{unimodular} if the edge vectors from any vertex of $T_1$ and from any vertex of $T_2$ form a unimodular basis.
 
@@
 The case that we need to study are therefore tetrahedra of type $(2,2)$, which are Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
+These correspond to prod-simplices of two segments in $P+Q$, which are parallelograms.
 The following lemma, whose proof we postpone to \Cref{sec:the_lemma}, is crucial to understand how to unimodularly cover these tetrahedra. 
-We use the following conventions: if $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$.
+We use the following conventions: if $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open line segments with endpoints $a,b$.
 Given a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and  the line spanned by $\vec s$ by $\vecline s$.
@@
 Let us denote $\tilde u$, $\tilde p_1$, $\tilde p_2$, $\tilde q_1$, $\tilde q_2$ the points corresponding to $u, p_1, p_2, q_1, q_2$ in $\cayley(P,Q)$.
-That is, $\tilde p_i = p\times\{0\}$, $\tilde q_i = p\times\{1\}$, and $\tilde u = u\times\{0\}$.
-Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that of the segments $[u,q_i]$ crosses the triangle $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$, see \Cref{fig:flip}.
+That is, $\tilde p_i = p_i\times\{1\}$, $\tilde q_i = 	q_i\times\{0\}$, and $\tilde u = u\times\{1\}$.\giulia{to match the figure, P is now at height 1 instead of 0}
+Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that one of the segments $[u,q_i]$ crosses one of the triangles $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$, see \Cref{fig:flip}.
 
@@
 \includegraphics[scale=.3]{flip}
-\caption{$[u,q_2]$ intersects $\conv(p_1,p_2,q_1)$}
+\caption{$[\tilde{u},\tilde{q}_2]$ intersects $\conv(\tilde{p}_1,\tilde{p}_2,\tilde{q}_1)$}
 \label{fig:flip}
-\end{figure}
+\end{figure}\giulia{added tildes to the points}
 
@@
 \begin{gather*}
-\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_i\}) \right\},
+\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_j\}) \right\},
 \\
-\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_j\}) \}.
+\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_i\}) \}.
 \end{gather*}
 The tetrahedra $\cayley(t, \{q_j\})$ and $\cayley(t, \{q_i\})$ are unimodular, which implies that $T=\cayley(p,q)$ has volume equal to the sum of the volumes of $\cayley([p_1,u],q)$ and $\cayley([p_2,u],q)$.
-In particular, we have covered $T$ by the three tetrahedra in $\mathcal T^-$, which are of smaller volume and hence have unimodular covers by inductive assumption.
+In particular, we have covered $T$ by the three tetrahedra in $\mathcal T^-$, which are of smaller volume and hence have unimodular covers by induction hypothesis.
 \end{proof}
@@
 
+We can assume that $q_1$ is the origin, and thus $f_q(q)=0$.
 Observe that in the strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=-1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
-Since translating the polytopes by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
+Since translating the polytopes independently by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
 That is, $f_q(p_1)=-1$, or equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
@@
 
-We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq 0\}$ and $V_2=\{f_p(x) \geq 0\}$. 
-In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants. 
-See \Cref{fig:setup}.
+We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq f_p(p)\}$ and $V_2=\{f_p(x) \geq f_p(p)\}$. 
+In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants (see \Cref{fig:setup}).
+This is always possible via an affine transformation.
+\Giulia{I'm not sure this last sentence addition is necessary, but I thought it would help the ref clear up confusion (Y.31) about p being a vertical segment-sure, this isn't possible with a unimod transformation, but we don't need to fix the lattice...}
 %
@@
 
-\begin{claim}
+\begin{claim}\label{claim:width}
 The following inequalities hold, 
@@
 Thus the triangle $T=\conv(p_2, p_1, l_1)$ is contained in $P$ and its edge $[p_1, l_1]$ is an integer dilation of $q$.
-Since $\width_{f_q}(T) =w \geq 2$, $T$ must contain a lattice point in the interior of the strip.
+
+Since $T$ contains $p$ and a copy of $q$, its area (normalized to a fundamental domain) is $w/2 \geq 1$, and by Pick's theorem it must contain a lattice point other than its vertices.
+Since $p$ and $q$ are primitive, this lattice point must lie in the interior of the strip.\giulia{added an explanation here, according to the suggestion Y.32}
+%Since $\width_{f_q}(T) =w \geq 2$, $T$ must contain a lattice point in the interior of the strip.
 \end{proof}
@@
 
-\giulia{unless b or t are equal to q}
+\begin{proof}
+Were $f_q(b_2-b_1)$ or $f_q(t_2 - t_1)$ zero, that is, $b$ or $t$ horizontal, it would have to contain a translate of $q$ whose endpoints are necessarily on the boundary of the strip.
+The left (or right) endpoint of this translate and of $q$ are lattice points lying on a line parallel to $p$, and since both this translate of $q$ and $q$ cross the segment $p$, they force $p$ to contain a lattice point in the interior, a contradiction to $p$ being primitive.
 
-\begin{proof}
 Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $V_1$ and that the maximum on $b$ lies in $V_2$.
-
+Observe that both $t$ and $b$ must intersect $p$, otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.
+%
 Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$.
 This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2}.
-This is a contradiction, since $P$ is a Minkowski summand of $Q$ and therefore $Q$ must have an edge parallel to $t$.
+This is a contradiction, since $P$ is a weak Minkowski summand of $Q$ and therefore $Q$ must have an edge parallel to $r$.
 \end{proof}
@@
 \label{claim:r}
-Assume without loss of generality that $b$ and $t$ either are parallel or their affine spans cross in $V_2$ (if they cross in $V_1$, the same claim can be reworded for the segment $l$). Then, 
+Assume without loss of generality that $b$ and $t$ either are parallel or their affine spans cross in $V_2$ (if they cross in $V_1$, the same claim can be reworded for $V_1$ and $l$). Then, 
 \begin{enumerate}
-\item The intersection of $Q$ with any line parallel to $p$ in $V_2$ has width w.r.t. $f_q$ strictly smaller than $w$.
+\item The intersection of $Q$ with any line parallel to $p$ in $V_2$ has width with respect to $f_q$ strictly smaller than $w$.
 \item $f_p(r_2) > f_p(r_1)$, that is, $f_p$ achieves its maximum over $r$ in $H_2$.
@@
 \begin{proof}
-Both $t$ and $b$ must intersect $p$, otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.
-Their intersections with $p$ are thus endpoints of a segment of width w.r.t $f_q$ less than $w$, the width of $p$.
+Both $t$ and $b$ must intersect $p$, as seen in the proof of \Cref{claim:b_and_t}.
+Their intersections with $p$ are thus endpoints of a segment of width with respect to $f_q$ less than $w$, the width of $p$.
 Since $t$ and $b$ cross in $V_2$, the same is true for any segment parallel to $p$ contained in $Q \cap V_2$.
@@
 
-For part (b), If $f_p(r_2) \leq f_p(r_1)$, it would be impossible to fit a translated copy $r'$ of $r$ in the correct side of $Q$: $r$ would need to lie inside the triangle delimited by the affine line $\langle t \rangle$ and the inequalities $f_q(x) \geq f_q(r_1)$, $f_p(x) \leq f_p(r_1)$.
-However, this triangular region has width less than $w$  w.r.t.
-$f_q$, by combining part (a) with the fact that $f_p$ and $f_q$ increase in the same direction along $t$, see \Cref{fig:claim3}.
+For part (2), recall that by \Cref{claim:width}, $\width_{f_q}(r) \geq w$.
+If $f_p(r_2) \leq f_p(r_1)$, it would be impossible to fit a translated copy $r'$ of $r$ in the correct side of $Q$: since $f_q$ increases along $t$, $r'$ has width in direction $f_q$ smaller than the segment parallel to $p$ with endpoint $r'_1$.
+This segment by part (1) has width less than $w$ in direction $q$, a contradiction.
+\giulia{simplified the proof as suggested in Y.38}
+
+%$r'$ would need to lie inside the triangle delimited by the affine line $\langle t \rangle$ and the inequalities $f_q(x) \geq f_q(r_1)$, $f_p(x) \leq f_p(r_1)$.
+However, this triangular region has width less than $w$  with respect to $f_q$, by combining part (1) with the fact that $f_p$ and $f_q$ increase in the same direction along $t$, see \Cref{fig:claim3}.
 \end{proof}
@@
 \]
-\giulia{changed the +1 to a -1, so that the following claim is true}
+\giulia{changed the +1 to a -1, so that the inequalities in the following claim are the right way around--we had flipped everything}
 
@@
 Since $b$ and $t$ must respectively separate $p_1$ and $q_1+q_2-p_1$ from the other two vertices of the parallelogram $\conv(q_1, p_1, q_2, q_1+q_2-p_1)$, they must respectively intersect its (parallel) edges $[p_1, q_2]$ and $[q_1, q_1+q_2-p_1]$, which implies the stated inequalities for $b$ and $t$.
-The same argument  applied to the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-p_2)$, yields the inequalities for $\ell$ and $r$.
+The same argument  applied to the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-q_2)$, yields the inequalities for $r$.
 \end{proof}
@@
 
-We are now ready to show a contradiction.
-Since the normal fan of $Q$ refines that of $P$, $Q$ must have an edge $r'$ which is a translated copy of $r$.
-Let $r_1'$ and $r_2'$ be its endpoints.
-Now consider the lattice line $d$ through $r_1'$ parallel to  $[p_1, q_2]$, that is, $g$ is constant on $d$.
-Let $d'$ be the parallel line defined by $g(d')=g(d)+1$.
+We are now ready to show a contradiction. Since the normal fan of $Q$ refines that of $P$, $Q$ must have an edge $r'$ which is a translated copy of $r$. Let $r_1'$ and $r_2'$ be its endpoints.
 
-Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.  
-Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,
-and in particular for $r_2'$. Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
+\Giulia{Changed the proof from here}
+Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on the line spanned by $t$.
+The width of $s$ with respect to $g$ is $g(s_2)-g(s_1) <1$, because $s$ is shorter than $Q \cap p$, which is strictly contained between the consecutive parallel lattice lines through $p_1$ and $q_2$ and through $q_1$ and $q_1+q_2-p_1$ where $g$ is constant.
+
+We now observe that $g(r_2) < g(s_2)$, because $r_2$ lies to the right of $s_2$ ($r$ has positive slope) and below the line spanned by $t$, and $g$ decreases moving to the right along $t$ (\Cref{claim:g}).
+Thus $g(r_2)-g(r_1) = g(r'_2)-g(r'_1) < g(s_2)-g(s_1) <1$, that is, $g(r_2)=g(r_1)$.
+This contradicts \Cref{claim:g}.
+%Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$, and in particular for $r_2'$.
+Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
 \end{proof}

### 2023-01-07 unimodular-covers-CombTheory-reviewing.tex
dealt with referee X

@@
 \label{eq:mixedIDP}
-(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
+(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d,
 \end{align*}
-%
+where $A+B :=\{a+b: a\in A, b\in B\}$ denotes the \emph{Minkowski sum} of two sets $A,B\subset \R^d$.
+
 A lattice polytope $Q$ is called \emph{smooth} if it is simple and the primitive edge directions at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth.
@@
 When the normal fan of a polytope $Q$ refines that of another polytope $P$, as in the second conjecture, we say that $P$ \emph{is a weak Minkowski summand of $Q$}, since this is easily seen to be equivalent to the existence of a polytope $P'$ such that $P+P' = k Q$ for some dilation constant $k>0$.
-\Giulia{RefY says to add the definition of Minkowski sum here, rather than have the definition in section 3; I think it would break up the flow of the introduction.
-Perhaps a compromise is to add a comment like, "where + denotes the Minkowski sum defined in Section 3".
-Then again, the Cayley sum is defined both later in the intro and in section 3.
-What do you think?}
 This property has the following algebraic implication for the projective toric variety $X_Q$: $P$ is a weak Minkowski summand of $Q$ if and only if the Cartier divisor defined by $P$ on  $X_Q$ is \emph{numerically effective}, or ``nef'' (see~\cite[Cor.~6.2.15, Thm.~6.3.12]{CLS}, but observe that what we here call  ``weak Minkowski summand'' is called ``$\N$-Minkowski summand'' there).
@@
 Motivated by these and other questions, several authors have studied the IDP property for different classes of lattice polytopes,
-with special attention to dimension $3$ (in dimension 2 is it straightforward that ever lattice polygon has unimodular triangulations). 
+with special attention to dimension $3$ (in dimension 2 it is straightforward that every lattice polygon has unimodular triangulations). 
 %
@@
 \[
-u:=(1,1,0)\in \conv(p_1p_2q_3) \subset T_4, \quad
-v:=(0,-1,0)\in \conv(p_1p_2q_4) \subset T_3.
+u:=(1,1,0)\in \conv(p_1,p_2,q_3) \subset T_3, \quad
+v:=(0,-1,0)\in \conv(p_1,p_2,q_4) \subset T_4.
 \]
@@
 Observe that $u+v=p_1+p_2=q_3+q_4$.
-Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.
-Hence, its translate $\conv(q_2p_3q_1p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$.
-By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
+Now, this implies that the quadrilateral $\conv(p_1,q_4,p_2,q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$.
+Hence, its translate $\conv(q_2,p_3,q_1,p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$.
+By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2,p_3,q_1,p_4)$ must contain lattice points in both triangles $\conv(q_2,p_3,p_4)\subset T_1$ and $\conv(q_1,p_3,p_4)\subset T_2$.
 
@@
 
-We call \emph{band} any region of the form $f^{-1}([\alpha,\beta])$ for some functional $f\in (\R^3)^*\setminus \{0\}$ and closed interval $[\alpha,\beta]\subset \R$.
+We call \emph{band} any region of the form $f^{-1}([\alpha,\beta])$ for some linear functional $f\in (\R^3)^*\setminus \{0\}$ and closed interval $[\alpha,\beta]\subset \R$.
 We claim that any band containing $T$ must contain at least three of the $q_i$s. 
@@
 To prove the claim, suppose that $q_1\not\in B:= f^{-1}([\alpha,\beta])$ for a certain band $B \supset T$. 
-Without loss of generality, say $f(q_1)<\alpha$. Then the equalities $q_1+q_2=p_3+p_4$ and $q_1+p_1=q_2+p_2$ respectively give:
+Without loss of generality, say $f(q_1)<\alpha$. Then the equalities $q_1+q_i=p_j+p_k$ and $q_1+p_1=q_i+p_i$, where $\{i,j,k\}=\{2,3,4\}$, respectively give:
+\paco{used $i$ instead of $2$, to shorten proof}
 \begin{gather}
 \label{eq:first}
-f(q_2) = f(p_3+p_4-q_1) = f(p_3)+f(p_4)-f(q_1) > 2\alpha-\alpha=\alpha,\\
+f(q_i) = f(p_j+p_k-q_1) = f(p_j)+f(p_k)-f(q_1) > 2\alpha-\alpha=\alpha,\\
 \label{eq:second}
-f(q_2) = f(q_1+p_1-p_2) = f(q_1)+(f(p_1)-f(p_2)) < \alpha + (\beta-\alpha) = \beta,
+f(q_i) = f(q_1+p_1-p_2) = f(q_1)+f(p_1)-f(p_i) < \alpha + \beta-\alpha = \beta,
 \end{gather}
-so that $q_2 \in B$.
-
-Inequality \eqref{eq:first} also implies
-\begin{equation}
-\label{eq:third}
-f(q_1) <  f(p_i) < f(q_2), \quad\text{ for $i=3,4$}.
-\end{equation}
-
-The translation by the vector $\frac12 (p_1+p_2-p_3-p_4)$ sends $q_1,q_2,p_3,p_4$ to $p_2,p_1,q_4,q_3$ (in this order). By applying this to inequality \eqref{eq:third}, we obtain
-\[
-\alpha \le f(p_2) < f(q_i) < f(p_1) \le \beta, \quad\text{ for $i=3,4$},
-\]
-so that $q_3,q_4 \in B$.
+so that $q_i \in B$ for $i\in \{2,3,4\}$.
+%
+%Inequality \eqref{eq:first} also implies
+%\begin{equation}
+%\label{eq:third}
+%f(q_1) <  f(p_i) < f(q_2), \quad\text{ for $i=3,4$}.
+%\end{equation}
+%
+%The translation by the vector $\frac12 (p_1+p_2-p_3-p_4)$ sends $q_1,q_2,p_3,p_4$ to $p_2,p_1,q_4,q_3$ (in this order). By applying this to inequality \eqref{eq:third}, we obtain
+%\[
+%\alpha \le f(p_2) < f(q_i) < f(p_1) \le \beta, \quad\text{ for $i=3,4$},
+%\]
+%so that $q_3,q_4 \in B$.
 %
@@
 \end{corollary}
-\giulia{I would ignore comment Y.16, where if I understand correctly a different proof idea is suggested...}
+%\giulia{I would ignore comment Y.16, where if I understand correctly a different proof idea is suggested...}
 \begin{proof}
@@
 $P=\conv([0,a]\times [0,1] \times\{0\} \cup [0,1]\times [0,b] \times\{1\})$, for big $a$ and $b$.
-This is smooth and contains the big empty tetrahedron $T$ with vertices (0,0,0), (a,1,0), (0,0,1), (1,b,1) which occupies most of its volume.
+This is smooth and contains a big empty tetrahedron $T$ with vertices $(0,0,0)$, $(a,1,0)$, $(0,0,1)$, $(1,b,1)$ which occupies most of its volume.
 In particular, none of the corner tetrahedra of $T$ is contained in $P$.
-Another way to see this is to write explicitly the remaining vertices $q_i$ of the circumscribed parallelepiped, which are $(\frac{a+1}{2}, \frac{b+1}{2}, 1), (\frac{a+1}{2}, \frac{b+1}{2}, 0), (\frac{1-a}{2}, \frac{b-1}{2}, 1), (\frac{a-1}{2}, \frac{1-b}{2}, 1)$.
+
+More explicitly, the the remaining vertices $q_i$ of the circumscribed parallelepiped are $(\frac{a+1}{2}, \frac{b+1}{2}, 1), (\frac{a+1}{2}, \frac{b+1}{2}, 0), (\frac{1-a}{2}, \frac{b-1}{2}, 1), (\frac{a-1}{2}, \frac{1-b}{2}, 1)$.
 None of these points are contained in $P$, and therefore none of the corner tetrahedra are either.
-\end{example}\giulia{Added the last part to answer comment Y.19}
+\end{example}
+%\giulia{Added the last part to answer comment Y.19}
 
-\begin{example}[Non-IDP octahedron and triangular prism]
+\begin{example}[Non-IDP polytopes with four facet directions]
 \label{ex:non-IDP}
-The following lattice octahedron $Q$ and triangular prism $P$ are not IDP:
+The following triangular prism $P$ and centrally symmetric octahedron $Q$ are not IDP:
 \begin{gather}
-Q= \conv((0,1,1),(1,0,1),(1,1,0),(0,-1,-1),(-1,0,-1),(-1,-1,0)),\\
-P=\conv((0,1,1),(1,0,1),(1,1,0),(-1,0,0),(0,-1,0),(0,0,-1)).
+P=\conv((0,1,1),(1,0,1),(1,1,0),(-1,0,0),(0,-1,0),(0,0,-1)),\\
+Q= \conv((0,1,1),(1,0,1),(1,1,0),(0,-1,-1),(-1,0,-1),(-1,-1,0)).
 \end{gather}
 
-Indeed, in both polytopes the only lattice points are the six vertices and the origin.
-The point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope.
-Hence, they are not IDP, which implies they do not admit unimodular covers.
+Indeed, in both cases the 
+%only lattice points are the six vertices and the origin. The 
+point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope. 
+%Hence, they are not IDP, which implies they do not admit unimodular covers.
 \end{example}
 
-An affirmative answer to the following question (weaker than the circumscribed parallelepiped property) would still  imply that smooth $3$-polytopes can be unimodularly covered, and hence~\Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three is:
+An affirmative answer to the following question (weaker than the circumscribed parallelepiped property) would still  imply that smooth $3$-polytopes can be unimodularly covered and, hence,~\Cref{conj:Oda}\eqref{itm:smoothIDP} in dimension three:
 
@@
 \[
-2\cayley(P, Q) \cap (\R^d\times \{1\}) \cong P+Q,
+2\cayley(P, Q) \cap (\R^d\times \{1\}) = (P+Q) \times \{1\} \cong P+Q,
 \]
@@
 \label{prop:mixedIDP}
-If $\cayley(P,Q)$ is IDP then the pair $(P,Q)$ is mixed IDP.
-\end{proposition} \giulia{We just called the pair IDP in the introduction... Which notation do we prefer?}
+If $\cayley(P,Q)$ is IDP then the pair $(P,Q)$ is IDP.
+\end{proposition} 
+%\giulia{We just called the pair IDP in the introduction... Which notation do we prefer?}
 
 The Cayley Trick also provides the following canonical bijections:
-
 \[
@@
 A triangulation of $\cayley(P,Q)\subset \R^3$ consists of tetrahedra of types $(1,3)$, $(2,2)$ and $(3,1)$, where the type denotes how many vertices they have in $P$ and in $Q$.
-Empty tetrahedra of types $(1,3)$ or $(3,1)$, which are Cayley sums of a triangle in $P$ and a point in $Q$, or viceversa, are automatically unimodular.
+Empty tetrahedra of types $(1,3)$ or $(3,1)$, which are Cayley sums of an empty (hence unimodular) triangle in $P$ and a point in $Q$, or viceversa, are automatically unimodular.
 The case that we need to study are therefore tetrahedra of type $(2,2)$, which are Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
-These correspond to prod-simplices of two segments in $P+Q$, which are parallelograms.
+These correspond to prod-simplices of two segments in $P+Q$, which are \emph{parallelograms}.
 The following lemma, whose proof we postpone to \Cref{sec:the_lemma}, is crucial to understand how to unimodularly cover these tetrahedra.
+
 We use the following conventions: if $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open line segments with endpoints $a,b$.
-Given a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and  the line spanned by $\vec s$ by $\vecline s$.
+Given a segment $s=[a,b]$, we denote $\vec s$ the vector $b-a$ and denote $\vecline s$ the line spanned by $\vec s$.
+A lattice parallelogram $p+q$ is called \emph{unimodular} if it is a fundamental domain for the lattice.
+Equivalently, if $\cayley(p,q)$ is a unimodular tetrahedron.
+\paco{added def of unimodular paralellogram}
 
@@
 \label{lemma:cayley}
-%\paco{$Q$ is 2-dim, but $P$ may not}
-Let $Q$ be a two-dimensional lattice polytope and $P$ a weak Minkowski summand of it. 
+\paco{referee X suggests to give an ``example'' of the lemma, but I am not sure what could be helpful}
+Let $Q\subset \R^2$ be a two-dimensional lattice polytope and $P\subset \R^2$ a weak Minkowski summand of it. 
 Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the lines spanned by them.
@@
 \]
-contains a lattice point. See \Cref{fig:strips}.
+contains a lattice point. 
 \end{lemma}
+
+See \Cref{fig:strips} for an illustration of the two regions in the statement, which we call \emph{strips}. In this figure and the forthcoming ones in \Cref{sec:the_lemma} 
+we draw $p$ as a vertical segment and $q$ as a horizontal one for convenience.
+This is always possible via a linear transformation (which of course changes the lattice; in the proof we do not assume the lattice to be $\Z^2$).
+\paco{moved this remark here, since it already affects this proof}
+%\Giulia{I'm not sure this last sentence addition is necessary, but I thought it would help the ref clear up confusion (Y.31) about p being a vertical segment-sure, this isn't possible with a unimod transformation, but we don't need to fix the lattice...}
 
@@
 That is, $\tilde p_i = p_i\times\{1\}$, $\tilde q_i = 	q_i\times\{0\}$, and $\tilde u = u\times\{1\}$.\giulia{to match the figure, P is now at height 1 instead of 0}
-Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that one of the segments $[u,q_i]$ crosses one of the triangles $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$, see \Cref{fig:flip}.
+Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that one of the segments $[u,q_i]$ crosses the interior of one of the triangles $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$.
+Without loss of generality assume that $[\tilde{u},\tilde{q}_2]$ crosses $\conv(\tilde{p}_1,\tilde{p}_2,\tilde{q}_1)$, as in \Cref{fig:flip}.
 
@@
 \begin{gather*}
-\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_j\}) \right\},
+\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_2\}) \right\},
 \\
-\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_i\}) \}.
+\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_1\}) \}.
 \end{gather*}
-The tetrahedra $\cayley(t, \{q_j\})$ and $\cayley(t, \{q_i\})$ are unimodular, which implies that $T=\cayley(p,q)$ has volume equal to the sum of the volumes of $\cayley([p_1,u],q)$ and $\cayley([p_2,u],q)$.
+The tetrahedra $\cayley(t, \{q_1\})$ and $\cayley(t, \{q_2\})$ are unimodular, which implies that $T=\cayley(p,q)$ has volume equal to the sum of the volumes of $\cayley([p_1,u],q)$ and $\cayley([p_2,u],q)$.
 In particular, we have covered $T$ by the three tetrahedra in $\mathcal T^-$, which are of smaller volume and hence have unimodular covers by induction hypothesis.
@@
 
-We can assume that $q_1$ is the origin, and thus $f_q(q)=0$.
-Observe that in the strip $q +\vecline p$, there is a unique lattice point on the line $f_q(x)=-1$;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
-Since translating the polytopes independently by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
+\paco{rphrased to address X.10}
+Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that $q_1$ is the origin and that $f_q(p_1)=-1$. 
+Observe that then $f_q(p_2)$ must be strictly positive, since $f_q(p_2)=0$ would imply that $p+q$ is a unimodular paralellogram.
+Moreover, since $p$ is primitive $p_1$ and $p_2$ cannot be in the boundary of $q +\vecline p$, which implies that $p_1$ is the unique lattice point with $f_q(x)=-1$ in $q +\vecline p$.
+Similarly, the unique lattice point in the strip with $f_q(x)=1$ is  $q_1+q_2 -p_1$.
+
+%
+%;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
+%Since translating the polytopes independently by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
 That is, $f_q(p_1)=-1$, or equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
@@
 
-We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq f_p(p)\}$ and $V_2=\{f_p(x) \geq f_p(p)\}$. 
-In the figures, we draw $p$ as a vertical segment and $q$ as a horizontal one, so that $H_i \cap V_j$ are the four quadrants (see \Cref{fig:setup}).
-This is always possible via an affine transformation.
-\Giulia{I'm not sure this last sentence addition is necessary, but I thought it would help the ref clear up confusion (Y.31) about p being a vertical segment-sure, this isn't possible with a unimod transformation, but we don't need to fix the lattice...}
+We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq f_p(p)\}$ and $V_2=\{f_p(x) \geq f_p(p)\}$.%
+\paco{added footnote}
+\footnote{Since in our figures $p$ and $q$ are vertical and horizontal, we use the letters $V$ and $H$ for the half-planes they defined.
+Similarly, later in the proof we use the letters $b$, $t$, $r$, and $l$ for certain  points and segments meaning ``bottom'', ``top'', ``right'' and ``left''.}
 %
@@
 Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain.
-In what follows, the width of a functional $f$ on a set $S$, denoted $\width_f(S)$ is defined as the difference $\sup_{x\in S} f(x)-\inf_{x\in S} f(x)$
+In what follows, the width of a functional $f$ on a set $S$, denoted $\width_f(S)$ is defined as the difference $\sup_{x\in S} f(x)-\inf_{x\in S} f(x)$.
 Then:
@@
 
-The same holds for $P$ and the strip $p+\vecline q$, and we call the segments $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively.
+The same holds for $P$ and the strip $p+\vecline q$, and we call the segments 
+\paco{not sure why we use $\ell$ for the segment and $l$ for the endpoints...}
+$\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively.
 The only difference is that in the case that $P$ is one dimensional we have $\ell=r=p$.
@@
 
-Observe that a priori one of $l$ and $r$ can coincide with $p$, if this is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if this is on the boundary of $Q$.
+Observe that a priori one of $\ell$ and $r$ can coincide with $p$, if $p$ is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if $q$ is on the boundary of $Q$.
 
@@
 \begin{proof}
-Were $f_q(b_2-b_1)$ or $f_q(t_2 - t_1)$ zero, that is, $b$ or $t$ horizontal, it would have to contain a translate of $q$ whose endpoints are necessarily on the boundary of the strip.
-The left (or right) endpoint of this translate and of $q$ are lattice points lying on a line parallel to $p$, and since both this translate of $q$ and $q$ cross the segment $p$, they force $p$ to contain a lattice point in the interior, a contradiction to $p$ being primitive.
-
-Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $V_1$ and that the maximum on $b$ lies in $V_2$.
-Observe that both $t$ and $b$ must intersect $p$, otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.
+Both $t$ and $b$ must cross the interior of $p$, or otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.
+\paco{simplified proof}
+To seek a contradiction assume, as in \Cref{fig:claim2}, that 
+\[
+f_q(t_2 - t_1) \le 0 \le f_q(b_2-b_1).
+\]
+That is, $f_q$ decreases (perhaps weakly) along $t$ and increases along $b$, as $f_p$ increases on both.
+This implies that $Q \cap V_2$ is contained in the open strip $\{f_q(p_1) <f_q(x)< f_q(p_2)\}$, of width $w$.
 %
-Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$.
-This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2}.
-This is a contradiction, since $P$ is a weak Minkowski summand of $Q$ and therefore $Q$ must have an edge parallel to $r$.
+%
+%Then, $f_q(t_2) \le  f_q (t_1)$ implies $f_q(t_2)$ equals the minimum of $f_q$ on $t$, which is smaller than $f_q(p_2)=w-1$. Similarly, 
+%
+% f_q(t_2)
+%
+%This implies 
+%
+%
+%
+%Let us first show that $f_q(b_2-b_1)\ne 0 \ne f_q(t_2 - t_1)$.
+If, say, $f_q(b_2-b_1)= 0$, then $b$ is parallel to $q$ and its endpoints $b_1$ and $b_2$ must be on the boundary of the strip.
+By \Cref{claim:width} this implies $b=q$.
+%
+%Then, $[b_i,q_i]$ is parallel to $p$, for $i=1,2$.
+%
+%The left (or right) endpoint of this translate and of $q$ are lattice points lying on a line parallel to $p$, and since both this translate of $q$ and $q$ cross the segment $p$, they force $p$ to contain a lattice point in the interior, a contradiction to $p$ being primitive.
+%	
+%Suppose by contradiction that the maximum of $f_q$ on $t$  lies in $V_1$ and that the maximum on $b$ lies in $V_2$. %
+%Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. 
+This strip cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, which gives a contradiction: since $P$ is a weak Minkowski summand of $Q$,  $Q$ must have edges parallel to $r$ both in $V_1$ and $V_2$.
 \end{proof}
@@
 \begin{proof}
-Both $t$ and $b$ must intersect $p$, as seen in the proof of \Cref{claim:b_and_t}.
+Both $t$ and $b$ must intersect $p$, as said in the proof of \Cref{claim:b_and_t}.
 Their intersections with $p$ are thus endpoints of a segment of width with respect to $f_q$ less than $w$, the width of $p$.
@@
 Let $g$ be the primitive lattice functional constant on $[p_1, q_2]$ (and therefore constant also on $[q_1, q_1+q_2-p_1]$).
-By the assumption on $p_1$, the values of $g$ on these segments differer by $1$.
+By the assumption on $p_1$, the values of $g$ on these segments differ by $1$.
 We choose the sign of $g$ so that

### 2023-01-08 JCOMB-reports/response to referees.txt
minor edits. changed name of response file

We have carefully gone through all the (very helpful and detailed) suggestions of both referees and we have implemented almost all of them. Below is a response explaining what we did:

Response to referee X
---------------------
1-7: agreed and corrected.
8: we are aware that the proof of Lemma 3.2 is complicated. This is why we put it in its own section 4, and in section 3 we use Lemma 3.2 as a black-box to prove Theorem 1.6. We are not sure what kind of example we could/should put.
9: agreed and corrected.
10: we have rephrased in a hopefully clearer way.
11: we have hopefully solved this too. In particular, the proofs of Claims 4.2, 4.3 and the conclusion after 4.4 have been edited and clarified.

Response to referee Y
---------------------
1: agreed & changed
2: added definition of Minkowski sum where it is first used (in the definition of IDP-ness).
3-9: agreed & changed
10: here the referee seems not to notice that the points q_i are not necessarily lattice points. We have added a line to this effect.
11-15: agreed & changed
16: this would be another way to go, but we have decided to keep the original formulation
17-23: agreed & changed
24: added a line in the explanation above lemma 3.2 explaining that the prod-simplices we are interested in are sums of segments, ie. parallelograms
25-30: agreed & changed
31: in the text we never make the assumption that the segments p and q are vertical and horizontal. We just draw them that way and use the letters $V$ and $H$ for the half-planes that they define for convenience. And, anyway, p and q can be made vertical and horizontal via a linear transformation (which of course changes the lattice, but is not a problem here. The proof never assumes the lattice to be the integers). We have added a sentence to this effect before Figure 3.2, which already shows $p$ and $q$ as vertical and horizontal.
32-38: agreed & changed
39: see our comment to 31. What may actually be misleading is that in some of our pictures the lattice still looks orthogonal (that is, having a square fundamental domain), but that fact is never used in the proofs, which are totally "affine".
40: thank you! There was a sign mistake, with a "-1" instead of "+1" in the equation of g the inequalities are now correct.
41: agreed & changed
42-43: the conclusion of the proof of lemma 3.2 has been edited for clarity and to correct the mistake from comment 40.

### 2023-01-08 unimodular-covers-CombTheory-reviewing.tex
minor edits. changed name of response file

@@
 where $A+B :=\{a+b: a\in A, b\in B\}$ denotes the \emph{Minkowski sum} of two sets $A,B\subset \R^d$.
+\paco{added def of Minkowski sum}
 
@@
 None of these points are contained in $P$, and therefore none of the corner tetrahedra are either.
+\giulia{Added the last part to answer comment Y.19}
+\paco{edited it}
 \end{example}
-%\giulia{Added the last part to answer comment Y.19}
 
@@
 Indeed, in both cases the 
-%only lattice points are the six vertices and the origin. The 
+\paco{shortened}
 point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope. 
-%Hence, they are not IDP, which implies they do not admit unimodular covers.
 \end{example}
@@
 \end{proposition} 
-%\giulia{We just called the pair IDP in the introduction... Which notation do we prefer?}
+\giulia{We just called the pair IDP in the introduction... Which notation do we prefer?}
+\paco{we stick to just IDP}
 

### 2023-05-15 JCOMB-reports/response to referees.txt
uploaded last referee comments to git, and implemented some of them

@@
-We have carefully gone through all the (very helpful and detailed) suggestions of both referees and we have implemented almost all of them. Below is a response explaining what we did:
+We have carefully gone through all the (very helpful and detailed) suggestions of the referees and we have implemented almost all of them. Below is a response explaining what we did:
 
@@
 This is why we put it in its own section 4, and in section 3 we use Lemma 3.2 as a black-box to prove Theorem 1.6.
-We are not sure what kind of example we could/should put.
+We are not sure what kind of example we could/should add to clarify the statement, but have streamlined some of the steps in the proof.
 9: agreed and corrected.
@@
 We just draw them that way and use the letters $V$ and $H$ for the half-planes that they define for convenience.
-And, anyway, p and q can be made vertical and horizontal via a linear transformation (which of course changes the lattice, but is not a problem here.
+And, anyway, p and q could be made vertical and horizontal via a linear transformation (which of course changes the lattice, but is not a problem here.
 The proof never assumes the lattice to be the integers).
-We have added a sentence to this effect before Figure 3.2, which already shows $p$ and $q$ as vertical and horizontal.
+We have added a sentence to this effect before Figure 3.2.
 32-38: agreed & changed
@@
 What may actually be misleading is that in some of our pictures the lattice still looks orthogonal (that is, having a square fundamental domain), but that fact is never used in the proofs, which are totally "affine".
-40: thank you! There was a sign mistake, with a "-1" instead of "+1" in the equation of g the inequalities are now correct.
+40: thank you for this correction! There was a sign mistake, with a "-1" instead of "+1" in the equation of g the inequalities are now correct.
 41: agreed & changed

### 2023-05-15 unimodular-covers-CombTheory-final.tex
uploaded last referee comments to git, and implemented some of them

--- 2023-01-08 unimodular-covers-CombTheory-reviewing.tex
@@
 where $A+B :=\{a+b: a\in A, b\in B\}$ denotes the \emph{Minkowski sum} of two sets $A,B\subset \R^d$.
-\paco{added def of Minkowski sum}
 
@@
 \]
-We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but $P$ or $Q$ do not necessarily need to be fulldimensional.
+We normally require $\cayley(P,Q)$ to be full-dimensional (otherwise we can delete coordinates) but $P$ or $Q$ do not necessarily need to be full-dimensional.
 We only require the linear subspaces parallel to them to span $\R^d$.
@@
 
-Let us mention that recent work of Gubeladze~\cite{Gubeladze} shows another class of 3-polytopes admitting unimodular covers: the convex hulls of all lattice points inside an ellipsoid; these had previously been shown to be IDP by Bruns, Gubeladze and Michalek~\cite{BGM}.
+Let us mention that recent work of Gubeladze~\cite{Gubeladze} shows another class of 3-polytopes admitting unimodular covers: the convex hulls of all lattice points inside an ellipsoid; these had previously been shown to be IDP by Bruns, Gubeladze and Micha{\l}ek~\cite{BGM}.
 To date there are no known examples of IDP $3$-polytopes without a unimodular cover, although such polytopes exist in higher dimension~\cite{BG}.
 
-%\subsection*{Acknowledgements:} We would like to thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and Christian Haase for helpful discussions.
+\subsection*{Acknowledgements:} We  thank two anonymous referees for helpful comments on a previous version of this paper.
 
@@
 Hence, its translate $\conv(q_2,p_3,q_1,p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$.
-By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2,p_3,q_1,p_4)$ must contain lattice points in both triangles $\conv(q_2,p_3,p_4)\subset T_1$ and $\conv(q_1,p_3,p_4)\subset T_2$.
+By central symmetry around its center $\left(\frac{a}2,\frac{b}2,1\right)$, $\conv(q_2,p_3,q_1,p_4)$ must contain lattice points in both triangles $\conv(q_2,p_3,p_4)\subset T_1$ and $\conv(q_1,p_3,p_4)\subset T_2$.
 
@@
 Without loss of generality, say $f(q_1)<\alpha$. Then the equalities $q_1+q_i=p_j+p_k$ and $q_1+p_1=q_i+p_i$, where $\{i,j,k\}=\{2,3,4\}$, respectively give:
-\paco{used $i$ instead of $2$, to shorten proof}
 \begin{gather}
@@
 \end{corollary}
-%\giulia{I would ignore comment Y.16, where if I understand correctly a different proof idea is suggested...}
 \begin{proof}
@@
 
-More explicitly, the the remaining vertices $q_i$ of the circumscribed parallelepiped are $(\frac{a+1}{2}, \frac{b+1}{2}, 1), (\frac{a+1}{2}, \frac{b+1}{2}, 0), (\frac{1-a}{2}, \frac{b-1}{2}, 1), (\frac{a-1}{2}, \frac{1-b}{2}, 1)$.
+More explicitly, the remaining vertices $q_i$ of the circumscribed parallelepiped are $(\frac{a+1}{2}, \frac{b+1}{2}, 1), (\frac{a+1}{2}, \frac{b+1}{2}, 0), (\frac{1-a}{2}, \frac{b-1}{2}, 1), (\frac{a-1}{2}, \frac{1-b}{2}, 1)$.
 None of these points are contained in $P$, and therefore none of the corner tetrahedra are either.
-\giulia{Added the last part to answer comment Y.19}
-\paco{edited it}
 \end{example}
@@
 Indeed, in both cases the 
-\paco{shortened}
 point $(1,1,1)$ lies in the second dilation but is not the sum of two lattice points in the polytope. 
@@
 \end{proposition}
-\giulia{We just called the pair IDP in the introduction... Which notation do we prefer?}
-\paco{we stick to just IDP}
 
@@
 we call \emph{prod-simplices} in $P+Q$ the Minkowski sums $T_1+T_2$ where $T_1\subset P$ and $T_2\subset Q$ are simplices with complementary affine spans.
-A prod-simplex is \emph{unimodular} if the edge vectors from any vertex of $T_1$ and from any vertex of $T_2$ form a unimodular basis.
+A prod-simplex is \emph{unimodular} if the union of edge-vectors from any vertex of $T_1$ and any vertex of $T_2$ form a unimodular basis.
 
@@
 Equivalently, if $\cayley(p,q)$ is a unimodular tetrahedron.
-\paco{added def of unimodular paralellogram}
 
@@
 \label{lemma:cayley}
-\paco{referee X suggests to give an ``example'' of the lemma, but I am not sure what could be helpful}
 Let $Q\subset \R^2$ be a two-dimensional lattice polytope and $P\subset \R^2$ a weak Minkowski summand of it. 
@@
 This is always possible via a linear transformation (which of course changes the lattice; in the proof we do not assume the lattice to be $\Z^2$).
-\paco{moved this remark here, since it already affects this proof}
-%\Giulia{I'm not sure this last sentence addition is necessary, but I thought it would help the ref clear up confusion (Y.31) about p being a vertical segment-sure, this isn't possible with a unimod transformation, but we don't need to fix the lattice...}
 
@@
 Let us denote $\tilde u$, $\tilde p_1$, $\tilde p_2$, $\tilde q_1$, $\tilde q_2$ the points corresponding to $u, p_1, p_2, q_1, q_2$ in $\cayley(P,Q)$.
-That is, $\tilde p_i = p_i\times\{1\}$, $\tilde q_i = 	q_i\times\{0\}$, and $\tilde u = u\times\{1\}$.\giulia{to match the figure, P is now at height 1 instead of 0}
-Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that one of the segments $[u,q_i]$ crosses the interior of one of the triangles $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$.
+That is, $\tilde p_i = p_i\times\{1\}$, $\tilde q_i = 	q_i\times\{0\}$, and $\tilde u = u\times\{1\}$.
+Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that one of the segments $[\tilde u,\tilde q_i]$ crosses the interior of one of the triangles $\conv(\tilde p_1,\tilde p_2,\tilde q_j)$, where $\{i,j\}=\{1,2\}$.
 Without loss of generality assume that $[\tilde{u},\tilde{q}_2]$ crosses $\conv(\tilde{p}_1,\tilde{p}_2,\tilde{q}_1)$, as in \Cref{fig:flip}.
@@
 \label{fig:flip}
-\end{figure}\giulia{added tildes to the points}
+\end{figure}
 
@@
 \begin{gather*}
-\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_2\}) \right\},
+\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_1\}) \right\},
 \\
-\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_1\}) \}.
+\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_2\}) \}.
 \end{gather*}
@@
 \[
-P:= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\},
+P:= \conv(Q_1\times\{0\} \cup Q_2 \times \{k\}),
 \]
@@
 
-\paco{rphrased to address X.10}
 Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that $q_1$ is the origin and that $f_q(p_1)=-1$. 
@@
 We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq f_p(p)\}$ and $V_2=\{f_p(x) \geq f_p(p)\}$.%
-\paco{added footnote}
 \footnote{Since in our figures $p$ and $q$ are vertical and horizontal, we use the letters $V$ and $H$ for the half-planes they defined.
@@
 The same holds for $P$ and the strip $p+\vecline q$, and we call the segments 
-\paco{not sure why we use $\ell$ for the segment and $l$ for the endpoints...}
-$\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively.
-The only difference is that in the case that $P$ is one dimensional we have $\ell=r=p$.
+$l=[l_1, l_2]$ and $r=[r_1, r_2]$, with $l$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively.
+The only difference is that in the case that $P$ is one dimensional we have $l=r=p$.
 Again we have
@@
 
-Observe that a priori one of $\ell$ and $r$ can coincide with $p$, if $p$ is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if $q$ is on the boundary of $Q$.
+Observe that a priori one of $l$ and $r$ can coincide with $p$, if $p$ is on the boundary of $P$, and similarly one of $t,b$ might be $q$, if $q$ is on the boundary of $Q$.
 
@@
 \begin{align*}
-\width_{f_q}(\ell) , 
+\width_{f_q}(l) , 
 \width_{f_q}(r) ,
@@
 
-If one of the inequalities, say the one for $\ell$, is not strict, then $\ell$ has one endpoint on each of the boundary lines of $(p + \vecline q)$.
-Unless $\ell = p$, one of the endpoints of $\ell$ is not an endpoint of $p$, say $l_1 \neq p_1$.
+If one of the inequalities, say the one for $l$, is not strict, then $l$ has one endpoint on each of the boundary lines of $(p + \vecline q)$.
+Unless $l = p$, one of the endpoints of $l$ is not an endpoint of $p$, say $l_1 \neq p_1$.
 Thus the triangle $T=\conv(p_2, p_1, l_1)$ is contained in $P$ and its edge $[p_1, l_1]$ is an integer dilation of $q$.
@@
 Since $T$ contains $p$ and a copy of $q$, its area (normalized to a fundamental domain) is $w/2 \geq 1$, and by Pick's theorem it must contain a lattice point other than its vertices.
-Since $p$ and $q$ are primitive, this lattice point must lie in the interior of the strip.\giulia{added an explanation here, according to the suggestion Y.32}
+Since $p$ and $q$ are primitive, this lattice point must lie in the interior of the strip.
 %Since $\width_{f_q}(T) =w \geq 2$, $T$ must contain a lattice point in the interior of the strip.
@@
 Both $t$ and $b$ must cross the interior of $p$, or otherwise $p_1$ or $p_2$ are the lattice points we are looking for in $Q$.
-\paco{simplified proof}
 To seek a contradiction assume, as in \Cref{fig:claim2}, that 
@@
 This segment by part (1) has width less than $w$ in direction $q$, a contradiction.
-\giulia{simplified the proof as suggested in Y.38}
 
@@
 \]
-\giulia{changed the +1 to a -1, so that the inequalities in the following claim are the right way around--we had flipped everything}
 
@@
 
-\Giulia{Changed the proof from here}
 Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on the line spanned by $t$.
@@
 \bibitem{BGM}
-  Winfried Bruns, Joseph Gubeladze, and Mateusz Michalek. Quantum Jumps of Normal Polytopes. \emph{Discrete Comput. Geom.}, 56(1) (2016), 181--215. \doi{10.1007/s00454-016-9773-7}
+  Winfried Bruns, Joseph Gubeladze, and Mateusz Micha{\l}ek. Quantum Jumps of Normal Polytopes. \emph{Discrete Comput. Geom.}, 56(1) (2016), 181--215. \doi{10.1007/s00454-016-9773-7}
 

### 2023-06-19 unimodular-covers-CombTheory-final.tex

@@
 Let $f_q$ be the primitive lattice functional constant on $q$ and $f_p$ the one constant on $p$. We assume that $f_q(p_1) < f_q(p_2)$ and $f_p(q_1) < f_p(q_2)$.
+Observe that since the segments $p_1p_2$ and $q_1q_2$ are primitive, the fact the $f_p$ and $f_q$ are also primitive implies that the area of the parallelogram $p+q$ equals both $f_p(q_2) - f_p(q_1)$ and $f_q(p_2) - f_q(p_1)$.
+%
+Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that $q_1$ is the origin, that $p$ belongs to the strip $q+\vecline p$  and that $f_q(p_1)=-1$.
 
-Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that $q_1$ is the origin and that $f_q(p_1)=-1$. 
-Observe that then $f_q(p_2)$ must be strictly positive, since $f_q(p_2)=0$ would imply that $p+q$ is a unimodular paralellogram.
-Moreover, since $p$ is primitive $p_1$ and $p_2$ cannot be in the boundary of $q +\vecline p$, which implies that $p_1$ is the unique lattice point with $f_q(x)=-1$ in $q +\vecline p$.
-Similarly, the unique lattice point in the strip with $f_q(x)=1$ is  $q_1+q_2 -p_1$.
+Then $f_q(p_2)$ must be strictly positive, since $f_q(p_2)=0$ would imply that $p+q$ is a unimodular paralellogram.
+Moreover, since $p$ is primitive  $p$ must be in the interior of $q +\vecline p$ or otherwise we would have $p_2$ equal to one of $q_1$ or $q_2$, and then $f_q(p_2)=0$.
+This implies that $p_1$ is the unique lattice point with $f_q(x)=-1$ in $q +\vecline p$. Similarly, the unique lattice point in the strip with $f_q(x)=1$ is  $q_1+q_2 -p_1$.
 
@@
 %Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. 
-This strip cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, which gives a contradiction: since $P$ is a weak Minkowski summand of $Q$,  $Q$ must have edges parallel to $r$ both in $V_1$ and $V_2$.
+This strip cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, which gives a contradiction: since $P$ is a weak Minkowski summand of $Q$, $Q$ must have an edge parallel to $r$ and with exterior normal pointing to the right.
+This edge must lie in $V_2$, which is impossible.
 \end{proof}
@@
 
-We are now ready to show a contradiction. Since the normal fan of $Q$ refines that of $P$, $Q$ must have an edge $r'$ which is a translated copy of $r$. Let $r_1'$ and $r_2'$ be its endpoints.
+We are now ready to show a contradiction.
+Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right.
+Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
+ Let $r_1'$ and $r_2'$ be its endpoints.
 
-Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on the line spanned by $t$.
+Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ and $s_2$ on the line spanned by $t$.
 The width of $s$ with respect to $g$ is $g(s_2)-g(s_1) <1$, because $s$ is shorter than $Q \cap p$, which is strictly contained between the consecutive parallel lattice lines through $p_1$ and $q_2$ and through $q_1$ and $q_1+q_2-p_1$ where $g$ is constant.
 
-We now observe that $g(r_2) < g(s_2)$, because $r_2$ lies to the right of $s_2$ ($r$ has positive slope) and below the line spanned by $t$, and $g$ decreases moving to the right along $t$ (\Cref{claim:g}).
+We now observe that $g(r'_2) < g(s_2)$, because $r'_2$ lies to the right of $s_2$ ($r$ has positive slope) and below the line spanned by $t$, and $g$ decreases moving to the right along $t$ (\Cref{claim:g}).
 Thus $g(r_2)-g(r_1) = g(r'_2)-g(r'_1) < g(s_2)-g(s_1) <1$, that is, $g(r_2)=g(r_1)$.

### 2023-06-19 unimodular-covers-CombTheory-final.tex
edits to setup in Section 4

@@
 Let $Q\subset \R^2$ be a two-dimensional lattice polytope and $P\subset \R^2$ a weak Minkowski summand of it. 
-Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the lines spanned by them.
+Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the vector lines spanned by them.
 If the parallelogram $p + q$ is not unimodular, then at least one of the regions
@@
 \label{sec:the_lemma}
-Let $f_q$ be the primitive lattice functional constant on $q$ and $f_p$ the one constant on $p$. We assume that $f_q(p_1) < f_q(p_2)$ and $f_p(q_1) < f_p(q_2)$.
-Observe that since the segments $p_1p_2$ and $q_1q_2$ are primitive, the fact the $f_p$ and $f_q$ are also primitive implies that the area of the parallelogram $p+q$ equals both $f_p(q_2) - f_p(q_1)$ and $f_q(p_2) - f_q(p_1)$.
-%
-Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that $q_1$ is the origin, that $p$ belongs to the strip $q+\vecline p$  and that $f_q(p_1)=-1$.
 
-Then $f_q(p_2)$ must be strictly positive, since $f_q(p_2)=0$ would imply that $p+q$ is a unimodular paralellogram.
-Moreover, since $p$ is primitive  $p$ must be in the interior of $q +\vecline p$ or otherwise we would have $p_2$ equal to one of $q_1$ or $q_2$, and then $f_q(p_2)=0$.
-This implies that $p_1$ is the unique lattice point with $f_q(x)=-1$ in $q +\vecline p$. Similarly, the unique lattice point in the strip with $f_q(x)=1$ is  $q_1+q_2 -p_1$.
-
-%
-%;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
-%Since translating the polytopes independently by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
-That is, $f_q(p_1)=-1$, or equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
-Similarly, the unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.
+Let $f_q$ be the primitive lattice (affine) functional vanishing on $q$ and $f_p$ the one vanishing on $p$. We assume that $f_q(p_1) < f_q(p_2)$ and $f_p(q_1) < f_p(q_2)$.
+Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain.
+In what follows, the width of a functional $f$ on a set $S$, denoted $\width_f(S)$ is defined as the difference $\sup_{x\in S} f(x)-\inf_{x\in S} f(x)$.
+Then primitiveness of $p$, $q$, $f_p$ and $f_q$ implies that:
+\begin{align*}
+w=&\width_{f_q}(p + \vecline q )=\width_{f_q}(p) = f_q(p_2) - f_q(p_1) =\\
+=&\width_{f_p}(q +\vecline p ) = \width_{f_p}(q)= f_p(q_2) - f_p(q_1).
+\end{align*}
 
@@
 
-Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain.
-In what follows, the width of a functional $f$ on a set $S$, denoted $\width_f(S)$ is defined as the difference $\sup_{x\in S} f(x)-\inf_{x\in S} f(x)$.
-Then:
+%
+Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that the segments $p$ and $q$ intersect in their relative interiors and that $f_q(p_1)=-1$.
+That is,
 \[
-w=\width_{f_q}(p + \vecline q )=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\vecline p ). 
+-1=f_q(p_1) < 0 < f_q(p_1), \quad
+-f_p(q_1) < 0 < f_p(q_2).
 \]
+In particular, $p$ must be in the interior of $q +\vecline p$ and $q$ in the interior of $p +\vecline q$
+Our assumption also implies that $p_1$ is the unique lattice point with $f_q(x)=-1$ in $q +\vecline p$. Similarly, the unique lattice point in the strip with $f_q(x)=1$ is  $q_1+q_2 -p_1$.
+See Figure~\ref{fig:setup} for an illustration of this setup.
+
+%
+%;  indeed, since $q$ is primitive, the only way that in the strip there could be two lattice points on $f_q(x)=-1$ is if they were on the boundary of the strip, which would however imply that $p+q$ is a unimodular paralellogram, against our assumptions.
+%Since translating the polytopes independently by lattice vectors will not result in any loss of generality, we can assume that $p_1$ is that unique lattice point.
+That is, $f_q(p_1)=-1$, or equivalently, the triangle $\conv(q_1, q_2, p_1)$ is unimodular.
+Similarly, the unique lattice point in the strip on the line $f_q(x)=1$ is then $q_1+q_2 -p_1$.
 
@@
 Let $g$ be the primitive lattice functional constant on $[p_1, q_2]$ (and therefore constant also on $[q_1, q_1+q_2-p_1]$).
-By the assumption on $p_1$, the values of $g$ on these segments differ by $1$.
+By the assumption that $f_q(p_1)=-1$ the quadrilateral spanned by these two parallel segments is unimodular, so  the values of $g$ on them differ by $1$.
 We choose the sign of $g$ so that

### 2023-06-19 unimodular-covers-CombTheory-final.tex
new claim

@@
 
-We let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq f_p(p)\}$ and $V_2=\{f_p(x) \geq f_p(p)\}$.%
-\footnote{Since in our figures $p$ and $q$ are vertical and horizontal, we use the letters $V$ and $H$ for the half-planes they defined.
-Similarly, later in the proof we use the letters $b$, $t$, $r$, and $l$ for certain  points and segments meaning ``bottom'', ``top'', ``right'' and ``left''.}
-%
-\begin{figure}[htb]
-\includegraphics[scale=.3]{setup.png}
-\caption{Setup for the proof of \Cref{lemma:cayley}}
-\label{fig:setup}
-\end{figure}
-
 %
@@
 Our assumption also implies that $p_1$ is the unique lattice point with $f_q(x)=-1$ in $q +\vecline p$. Similarly, the unique lattice point in the strip with $f_q(x)=1$ is  $q_1+q_2 -p_1$.
+
+Finally, we let $H_1=\{f_q(x) \leq 0\}$ and $H_2=\{f_q(x) \geq 0\}$; similarly let $V_1=\{f_p(x) \leq f_p(p)\}$ and $V_2=\{f_p(x) \geq f_p(p)\}$.%
+\footnote{Since in our figures $p$ and $q$ are vertical and horizontal, we use the letters $V$ and $H$ for the half-planes they defined.
+Similarly, later in the proof we use the letters $b$, $t$, $r$, and $l$ for certain  points and segments meaning ``bottom'', ``top'', ``right'' and ``left''.}
 See Figure~\ref{fig:setup} for an illustration of this setup.
+%
+\begin{figure}[htb]
+\includegraphics[scale=.3]{setup.png}
+\caption{Setup for the proof of \Cref{lemma:cayley}}
+\label{fig:setup}
+\end{figure}
 
@@
 
-We are now ready to show a contradiction.
-Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right.
-Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
+\begin{claim}
+\label{claim:g2}
+$g(r_2) - g(r_1) <1$.
+\end{claim}
+
+\begin{proof}
+UNFINISHED
+
+Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right. Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
  Let $r_1'$ and $r_2'$ be its endpoints.
@@
 Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
+\end{proof}
+
+We now have a contradiction. Claim~\ref{claim:g} says that $g(r_2) - g(r_1) >0$ and Claim~\ref{claim:g2}  that $g(r_2) - g(r_1) <1$. Since  $g(r_2) - g(r_1)$ is an integer, this is impossible.
 \end{proof}

### 2023-06-19 unimodular-covers-CombTheory-final.tex
solved the last issue

@@
 -1=f_q(p_1) < 0 < f_q(p_1), \quad
--f_p(q_1) < 0 < f_p(q_2).
+f_p(q_1) < 0 < f_p(q_2).
 \]
@@
 Let $g$ be the primitive lattice functional constant on $[p_1, q_2]$ (and therefore constant also on $[q_1, q_1+q_2-p_1]$).
-By the assumption that $f_q(p_1)=-1$ the quadrilateral spanned by these two parallel segments is unimodular, so  the values of $g$ on them differ by $1$.
-We choose the sign of $g$ so that
+By the assumption that $f_q(p_1)=-1$ the quadrilateral $:=\conv(q_1, p_1, q_2, q_1+q_2-p_1)$
+ is unimodular, so  the values of $g$ on the edges $[p_1, q_2]$ and $[q_1, q_1+q_2-p_1]$ differ by $1$. We choose the sign of $g$ so that 
 \[
@@
 Since $b$ and $t$ must respectively separate $p_1$ and $q_1+q_2-p_1$ from the other two vertices of the parallelogram $\conv(q_1, p_1, q_2, q_1+q_2-p_1)$, they must respectively intersect its (parallel) edges $[p_1, q_2]$ and $[q_1, q_1+q_2-p_1]$, which implies the stated inequalities for $b$ and $t$.
-The same argument  applied to the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-q_2)$, yields the inequalities for $r$.
+The same argument  applied to the parallelogram  $\conv(p_1, q_2, p_2, p_1+p_2-q_2)$, yields the inequality for $r$.
 \end{proof}
@@
 \scalebox{.75}{\input{claim4.pdf_t}}
-\caption{Illustration of the proof of \Cref{claim:g}}
+\caption{Illustration of the proof of \Cref{claim:g}. The functional $g$ is constant along the dashed diagonal lines}
 \label{fig:claim4}
@@
 
-\begin{claim}
-\label{claim:g2}
-$g(r_2) - g(r_1) <1$.
-\end{claim}
-
-\begin{proof}
-UNFINISHED
-
-Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right. Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
+We are now ready to show a contradiction.
+Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right.
+Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
  Let $r_1'$ and $r_2'$ be its endpoints.
 
+We consider two cases, depending on whether $q_1+q_2-p_1$ lies ``to the right'' or ``to the left'' of $p_1$.
+That is, whether $f_p(q_1+q_2-p_1)$ is positive or negative.
+%For illustration use the left part of Figure~\ref{fig:claim4} in which $f_p(q_1+q_2-p_1)=0$, and in which both arguments work:
+
+\begin{itemize}
+\item \underline{$f_p(q_1+q_2-p_1)> 0$}. In this case the segment $r$ in the boundary of $P$ separates $p$ from  $q_1+q_2-p_1$.
+Let $r''_1$ and $r''_2$ be the intersection points of the line containing $r$ with $b$ and $t$ respectively. We have that
+$r'=[r'_1,r'_2]$ is parallel to $[r''_1,r''_2]$ but shorter, so that
+\[
+w= \width_{f_q}(p) \le \width_{f_q}(r) = \width_{f_q}(r') \le  f_q(r''_2) -f_q(r''_1), 
+\]
+where the first inequality comes from \eqref{eq:widthp}.
+But on the other hand
+\[
+f_q(r''_2) < f_q(q_1+q_2-p_1) = 1, \quad f_q(r''_1) > f_q(p_1) =-1,
+\]
+which leads to the contradiction $w<2$.
+
+\item \underline{$f_p(q_1+q_2-p_1)\ge 0$}.
 Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ and $s_2$ on the line spanned by $t$.
@@
 We now observe that $g(r'_2) < g(s_2)$, because $r'_2$ lies to the right of $s_2$ ($r$ has positive slope) and below the line spanned by $t$, and $g$ decreases moving to the right along $t$ (\Cref{claim:g}).
-Thus $g(r_2)-g(r_1) = g(r'_2)-g(r'_1) < g(s_2)-g(s_1) <1$, that is, $g(r_2)=g(r_1)$.
-This contradicts \Cref{claim:g}.
+Thus $g(r_2)-g(r_1) = g(r'_2)-g(r'_1) < g(s_2)-g(s_1) <1$.
+Since $g(r_2)$ and $g(r_1)$ are integers this leads to $g(r_2)\le g(r_1)$, in contradiction with \Cref{claim:g}.
 %Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$, and in particular for $r_2'$.
 Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
-\end{proof}
-
-We now have a contradiction. Claim~\ref{claim:g} says that $g(r_2) - g(r_1) >0$ and Claim~\ref{claim:g2}  that $g(r_2) - g(r_1) <1$. Since  $g(r_2) - g(r_1)$ is an integer, this is impossible.
+\end{itemize}
 \end{proof}

### 2023-06-21 unimodular-covers-CombTheory-final.tex
typo
added comments to final correction to ref Y

@@
 \end{claim}
+\Giulia{Do we ever use the claim 3(1), other than to prove 3(2)? If not, we can remove it from the statement...}
 
@@
 We are now ready to show a contradiction.
-Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right.
+Since the normal fan of $Q$ refines that of $P$, the boundary of $Q$ must contain an edge parallel to $r$ on the right.
 Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
@@
 %For illustration use the left part of Figure~\ref{fig:claim4} in which $f_p(q_1+q_2-p_1)=0$, and in which both arguments work:
+\giulia{we put the origin in $q_1$, so the functional $f_p$ is not $0$ at $p$...
+so we could say $q_1 + q_2-p_1$ in $V_1$ or $V_2$...
+Though I'm not against changing our origin to lie in the intersection of $p$ and $q$}
 
@@
 Let $r''_1$ and $r''_2$ be the intersection points of the line containing $r$ with $b$ and $t$ respectively. We have that
-$r'=[r'_1,r'_2]$ is parallel to $[r''_1,r''_2]$ but shorter, so that
+$r'=[r'_1,r'_2]$ is parallel to $[r''_1,r''_2]$ but shorter, \giulia{I would say that $r'$ needs to lie between the lines along $b$ and $t$ to explain why it's shorter} so that
 \[
@@
 
-\item \underline{$f_p(q_1+q_2-p_1)\ge 0$}.
+\item \underline{$f_p(q_1+q_2-p_1)\le 0$}.
 Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ and $s_2$ on the line spanned by $t$.
-The width of $s$ with respect to $g$ is $g(s_2)-g(s_1) <1$, because $s$ is shorter than $Q \cap p$, which is strictly contained between the consecutive parallel lattice lines through $p_1$ and $q_2$ and through $q_1$ and $q_1+q_2-p_1$ where $g$ is constant.
+The width of $s$ with respect to $g$ is $g(s_2)-g(s_1) <1$, because $s$ is shorter than $Q \cap p$,\giulia{we still need to explain why this is true...} which is strictly contained between the consecutive parallel lattice lines through $p_1$ and $q_2$ and through $q_1$ and $q_1+q_2-p_1$ where $g$ is constant.
 

### 2023-06-22 unimodular-covers-CombTheory-final.tex
new claim for final argument

--- 2023-06-19 unimodular-covers-CombTheory-final.tex
@@
 
-Let $f_q$ be the primitive lattice (affine) functional vanishing on $q$ and $f_p$ the one vanishing on $p$. We assume that $f_q(p_1) < f_q(p_2)$ and $f_p(q_1) < f_p(q_2)$.
+Let $f_q$ be the primitive lattice (affine) functional vanishing on $q$ and $f_p$ the one vanishing on $p$, taking their signs so that $f_q(p_1) < f_q(p_2)$ and $f_p(q_1) < f_p(q_2)$.
 Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain.
@@
 %
-Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that the segments $p$ and $q$ intersect in their relative interiors and that $f_q(p_1)=-1$.
-That is,
+Since we can perform without loss of generality respective lattice translations to $P$ and to $Q$, we assume that 
+$p_1$ lies on the line $\{f_q=-1\}$ and in the strip $q +\vecline p$.
+It  must then lie in the interior of the strip, or otherwise we would have that $p_2$ equals one of  $q_1$ or $q_2$ and $w=1$.
+That is, we have that $p$ and $q$ intersect in their relative interiors and that
 \[
@@
 \]
-In particular, $p$ must be in the interior of $q +\vecline p$ and $q$ in the interior of $p +\vecline q$
+%In particular, $p$ must be in the interior of $q +\vecline p$ and $q$ in the interior of $p +\vecline q$
 Our assumption also implies that $p_1$ is the unique lattice point with $f_q(x)=-1$ in $q +\vecline p$. Similarly, the unique lattice point in the strip with $f_q(x)=1$ is  $q_1+q_2 -p_1$.
@@
 
-The last two claims can be summarized as saying that in the pictures $b$, $t$ and $r$ have positive slope.
+The last two claims can be summarized as saying that in the pictures $b$, $t$ and $r$ have positive slope and that the slope of $b$ is greater or equal than that of $t$.
 Observe that this implies that $q$ is not in the boundary of $Q$ and $p \neq r$, so both $P$ and $Q$ are full dimensional.
@@
 
+Let $r'=[r'_1,r'_2]$ be the segment collinear with $r$ and with endpoints $r'_1$ and $r'_2$ lying respectively on the lines containing $b$ and $t$.
+($r'_1$ must in fact lie in $b$, but $r'_2$ may be outside $t$).
+
+\begin{claim}
+%\paco{new claim. Note I changed the roles of $r'$ and $r''$, since (the new) $r''$ is only used inside the claim}
+\label{claim:r}
+$r'$ has at least the same length as $r$. 
+\end{claim}
+
+\begin{proof}
+Since the normal fan of $Q$ refines that of $P$, the boundary of $Q$ must contain an edge parallel to $r$ on the right.
+Let $r''$ be a primitive segment along this edge.
+Then $r''$ is a translated copy of $r$, and parallel to $r'$.
+
+Now, $r''$ is contained also between the lines  along $b$ and $t$, and is to the right of $r'$ (because $r$, hence $r'$, has $q_2$ on its right, and $r''$ has $q_2$ on its left).
+Our assumption that the lines along $b$ and $t$ are either parallel or meet in $v_2$ implies the inequality.
+\end{proof}
+
 We are now ready to show a contradiction.
-Since the normal fan of $Q$ refines that of $P$, $Q$ must contain an edge parallel to $r$ on the right.
-Let $r'$ be a primitive segment along this edge, so that $r'$ is a translated copy of $r$.
- Let $r_1'$ and $r_2'$ be its endpoints.
-
 We consider two cases, depending on whether $q_1+q_2-p_1$ lies ``to the right'' or ``to the left'' of $p_1$.
@@
 \begin{itemize}
-\item \underline{$f_p(q_1+q_2-p_1)> 0$}. In this case the segment $r$ in the boundary of $P$ separates $p$ from  $q_1+q_2-p_1$.
-Let $r''_1$ and $r''_2$ be the intersection points of the line containing $r$ with $b$ and $t$ respectively. We have that
-$r'=[r'_1,r'_2]$ is parallel to $[r''_1,r''_2]$ but shorter, so that
+\item \underline{$f_p(q_1+q_2-p_1)> 0$}. In this case the line containing $r$ and $r'$ separates $p$ from  $q_1+q_2-p_1$.
+This implies
 \[
-w= \width_{f_q}(p) \le \width_{f_q}(r) = \width_{f_q}(r') \le  f_q(r''_2) -f_q(r''_1), 
+f_q(r'_2) < f_q(q_1+q_2-p_1) = 1, \quad f_q(r'_1) > f_q(p_1) =-1,
 \]
-where the first inequality comes from \eqref{eq:widthp}.
-But on the other hand
+while Claims~\ref{claim:r} and~\ref{claim:width} imply that $\width_{f_q}(r')\ge \width_{f_q}(r)  \ge w\ge 2$.
+
+\item \underline{$f_p(q_1+q_2-p_1)\le 0$}.
+We observe that $g(r_2) - g(r_1)$ is positive (and integer) because $r$ has positive slope and separates $p_1$ from $q_2$. But Claim~\ref{claim:r} gives
 \[
-f_q(r''_2) < f_q(q_1+q_2-p_1) = 1, \quad f_q(r''_1) > f_q(p_1) =-1,
+g(r_2) - g(r_1) \le g(r'_2) - g(r'_1) 
 \]
-which leads to the contradiction $w<2$.
-
-\item \underline{$f_p(q_1+q_2-p_1)\ge 0$}.
-Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ and $s_2$ on the line spanned by $t$.
-The width of $s$ with respect to $g$ is $g(s_2)-g(s_1) <1$, because $s$ is shorter than $Q \cap p$, which is strictly contained between the consecutive parallel lattice lines through $p_1$ and $q_2$ and through $q_1$ and $q_1+q_2-p_1$ where $g$ is constant.
-
-We now observe that $g(r'_2) < g(s_2)$, because $r'_2$ lies to the right of $s_2$ ($r$ has positive slope) and below the line spanned by $t$, and $g$ decreases moving to the right along $t$ (\Cref{claim:g}).
-Thus $g(r_2)-g(r_1) = g(r'_2)-g(r'_1) < g(s_2)-g(s_1) <1$.
-Since $g(r_2)$ and $g(r_1)$ are integers this leads to $g(r_2)\le g(r_1)$, in contradiction with \Cref{claim:g}.
-%Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$, and in particular for $r_2'$.
-Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
+which is in this case strictly smaller than $g(q_1+q_2-p_1) - g(p_1) =1$.
+\qedhere
 \end{itemize}

### 2023-08-10 unimodular-covers-CombTheory-final.tex
corrected repeated label

@@
 %\paco{new claim. Note I changed the roles of $r'$ and $r''$, since (the new) $r''$ is only used inside the claim}
-\label{claim:r}
+\label{claim:r2}
 $r'$ has at least the same length as $r$. 
@@
 \]
-while Claims~\ref{claim:r} and~\ref{claim:width} imply that $\width_{f_q}(r')\ge \width_{f_q}(r)  \ge w\ge 2$.
+while Claims~\ref{claim:r2} and~\ref{claim:width} imply that $\width_{f_q}(r')\ge \width_{f_q}(r)  \ge w\ge 2$.
 
 \item \underline{$f_p(q_1+q_2-p_1)\le 0$}.
-We observe that $g(r_2) - g(r_1)$ is positive (and integer) because $r$ has positive slope and separates $p_1$ from $q_2$. But Claim~\ref{claim:r} gives
+We observe that $g(r_2) - g(r_1)$ is positive (and integer) because $r$ has positive slope and separates $p_1$ from $q_2$. But Claim~\ref{claim:r2} gives
 \[

### 2023-09-26 ct-final/unimodular-covers-ctstyle.tex
ctstyle version

--- 2023-08-10 unimodular-covers-CombTheory-final.tex
@@
-\documentclass{amsart}
+% SAMPLE ARTICLE FOR COMBINATORIAL THEORY
+%
+% CT papers *must* follow this sample file.
+% Please, do not erase comments in the head of this sample file as
+% they are used to process your file semi-automatically.
 
-\textheight=22.5cm
-\topmargin=0.5cm
+% COPYRIGHT NOTICE
+% We use the CC BY licence. It allows others to freely distribute 
+% your work while giving credit to you. In particular CC BY allows 
+% remixing, tweaking and building upon your work.  
+% See https://creativecommons.org/licenses/ for a full explanation
+% of the Creative Commons licenses.
 
-\usepackage{amssymb}
-\usepackage{graphicx,color}
-\usepackage{amsmath}
-\usepackage{url}
-\usepackage[colorlinks=true]{hyperref}
-\usepackage{doi}
-%\usepackage{pictex}
+% DOCUMENT CLASS AND CT.STY
+% All CT papers start with these two lines. Do not change make
+% any changes in the ct.sty file as we use our own copy.
+\documentclass[12pt]{article}
+\usepackage{ct}
 
-\usepackage{cleveref}
-\Crefname{counterexample}{counterexample}{counterexamples}
-\Crefname{counterexample}{Counterexample}{Counterexamples}
-\Crefname{conjecture}{conjecture}{conjectures}
-\Crefname{conjecture}{Conjecture}{Conjectures}
+% GEOMETRY AND SPACES
+% To keep the style homogeneous among all CT papers, please
+%   * remove all commands that change parameters such as margins or page sizes. 
+%   * avoid using any vspace, hspace, newpage, and other spacing commands.
 
-\usepackage{tikz-cd}
-\usepackage{tikz}
-\usetikzlibrary{matrix}
-\usepackage{tkz-euclide}
-\usetikzlibrary{intersections}
-%\usetkzobj{all}
+% PACKAGES
+% The following packages are already loaded in the style file:
+%   * amssymb, amsthm, amsmath for mathematics,
+%   * graphicx for images,
+%   * hyperref and cleveref for references.
+% Load your own packages below only if they are really needed:
+% \usepackage{...}
 
-\newcommand{\limites}{}
+% THEOREM ENVIRONMENTS
+% The following environments are already declared in the style file:
+%   * theorem-like environments: theorem, lemma, corollary, proposition, fact, observation, claim
+%   * definition-like environments: definition, example, conjecture, open, problem, question
+%   * remark-like environments: remark, note
+% Declare your own environments below only if needed:
+% \newtheorem{myenv}[theorem]{MyEnvironment}
 
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{claim}[theorem]{Claim}
+% SPECIFICATIONS
+% The editors will insert the specifications of your paper.
+\specs{n}{vol (issue)}{year}{}
 
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{notation}[theorem]{Notation}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{problem}[theorem]{Problem}
-\newtheorem{question}[theorem]{Question}
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{remark}[theorem]{Remark}
+% DATES
+% Give the submission and acceptance dates in the format shown.
+% The editors will insert the publication date in the third argument.
+\dateline{Jan 1, 2020}{Jan 2, 2020}{TBD}
 
-% Macros in general
+% KEYWORDS
+% Give one or more keywords separated by commas. No dot at the end.
+\keywords{lattice polytopes, unimodular covers, integer decomposition property}
 
-\newcommand{\F}{ \ensuremath{\mathbb{F}}}
+% SUBJECT CODES
+% Give one or more subject codes separated by commas. No dot at the end.
+% Codes are available from http://www.ams.org/mathscinet/freeTools.html
+\MSC{52B10, 52B20, 52C17}
+
+% TITLE
+% If needed, include a line break (\\) at an appropriate place in the title.
+\title{Unimodular covers of 3-dimensional\\ parallelepipeds and Cayley sums}
+
+% AUTHORS AND AFFILIATIONS
+% Input author, affiliation, address and support information as follows.
+% The address should include the country, but does not have to include the street address. 
+% Give at least one email address.
+% Grants and institutional support information should be included in the thanks command. Other
+% acknowledgements to colleagues or referees should appear in the Acknowledgements sections
+% at the end of the paper, before the bibliography.
+
+\author[1]{Giulia Codenotti\thanks{}}
+\author[2]{Francisco Santos\thanks{Supported by grants PID2019-106188GB-I00 funded by MCIN/AEI/10.13039/501100011033 and CLaPPo (21.SI03.64658) of Universidad de Cantabria and Banco Santander}}
+%\author[3]{Some Third Author}
+
+% Write your affiation
+\affil[1]{%
+% Name of Department, Name of Institution, City and / or State, Country
+Institute of Mathematics, Freie Universit\"at Berlin,
+
+% leave a blank line and put your email here 
+\email{giulia.codenotti@fu-berlin.de}%
+}
+
+% In case some authors share the same affiliation, use the following style.
+% Do not use 'merging' of emails from the same institution such as \email{\{ssa,sta\}@uwn.edu.au}%
+\affil[2]{%
+% Name of Department, Name of Institution, City and / or State, Country
+Department of Mathematics, Statistics and Computer Science, University of Cantabria, Spain
+
+% leave a blank line and write the emails in the author's order
+\email{francisco.santos@unican.es}%
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% AUTHORS COMMANDS
+% Enter here all commands you need (and no more) in this paper.
+%\newcommand{\F}{ \ensuremath{\mathbb{F}}}
 \newcommand{\N}{ \ensuremath{\mathbb{N}}}
 \newcommand{\Z}{ \ensuremath{\mathbb{Z}}}
-\newcommand{\C}{ \ensuremath{\mathbb{C}}}
+%\newcommand{\C}{ \ensuremath{\mathbb{C}}}
 \newcommand{\R}{ \ensuremath{\mathbb{R}}}
 
-\newcommand{\T}{ \ensuremath{\mathcal{T}}}
+\newcommand{\width}{\operatorname{width}}
+\newcommand{\area}{\operatorname{area}}
+\newcommand{\conv}{\ensuremath{\mathrm{conv}}\hspace{1pt}}
+\newcommand{\cayley}{\operatorname{Cay}}
 
-\newcommand{\bb}{{\mathbf{b}}}
-\newcommand{\cc}{{\mathbf{c}}}
-\newcommand{\ttt}{\mathbf{t}}
-\newcommand{\GL}{{GL}_r (K)}
-
-\renewcommand{\int}{\operatorname{int}}
-\newcommand{\width}{\operatorname{width}}
-\newcommand{\ind}{\operatorname{index}}
-\newcommand{\Vol}{\operatorname{Vol}}
-\newcommand{\area}{\operatorname{area}}
-\newcommand{\length}{\operatorname{length}}
-
-\renewcommand{\vec}[1]{\overrightarrow#1}
 \newcommand{\vecline}[1]{\langle \vec #1 \rangle}
 
-\newcommand{\cone}{\ensuremath{\mathrm{cone}}\hspace{1pt}}
-\newcommand{\conv}{\ensuremath{\mathrm{conv}}\hspace{1pt}}
-\newcommand{\cayley}{\operatorname{Cay}}
-\newcommand{\rank}{\operatorname{rank}}
-\newcommand{\reg}{\operatorname{reg}}
+%\newcommand{\Image}{\ensuremath{\mathrm{Im}}\hspace{1pt}}
 
-\newcommand{\Image}{\ensuremath{\mathrm{Im}}\hspace{1pt}}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\giulia}[1]{\todo[size=\tiny,color=blue!30]{#1 \\ \hfill --- G.}}
-\newcommand{\Giulia}[1]{\todo[size=\tiny,inline,color=blue!30]{#1 \\ \hfill --- G.}}
-
-\newcommand{\paco}[1]{\todo[size=\tiny,color=green!30]{#1 \\ \hfill --- P.}}
-\newcommand{\Paco}[1]{\todo[size=\tiny,inline,color=green!30]{#1 \\ \hfill --- P.}}
-
-\date{\today}
-
-%\author[G.~Codenotti]{Giulia Codenotti}
-%\author[F.~Santos]{Francisco Santos}
-
-%\address[G.~Codenotti]
-%{
-%Institut f\"ur Mathematik, Freie Universit\"at Berlin, Germany
-%}
-%\email{codenotti@math.uni-frankfurt.de}
-
-%\address[F.~Santos]
-%{
-%Department of Mathematics, Statistics and Computer Science, University of Cantabria, Spain
-%}
-%\email{francisco.santos@unican.es}
-
-%\thanks{The authors were supported by the Einstein Foundation Berlin under grant EVF-2015-230. 
-%Work of F.
-Santos is also supported by grants PID2019-106188GB-I00 funded by MCIN/AEI/10.13039/501100011033 and by FPU19/04163 of the Spanish Government and by project CLaPPo (21.SI03.64658) of Universidad de Cantabria and Banco Santander}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{Unimodular covers of 3-dimensional parallelepipeds and Cayley sums}
-\author{}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
+%\writedatatofile
+
+\maketitle
+
+% ABSTRACT
+% CT papers must include an abstract. The abstract should consist of a
+% succinct statement of background followed by a listing of the
+% principal new results that are to be found in the paper. The abstract
+% should be informative, clear, and as complete as possible. Phrases
+% like "we investigate..." or "we study..." should be kept to a minimum
+% in favor of "we prove that..."  or "we show that...".  Do not
+% include equation numbers, unexpanded citations (such as "[23]"), or
+% any other references to things in the paper that are not defined in
+% the abstract. The abstract may be distributed without the rest of the
+% paper so it must be entirely self-contained.  Try to include all words
+% and phrases that someone might search for when looking for your paper.
+% You can use some basic LaTeX commands in the abstract, but not any
+% user defined macros.
 
@@
 
-\maketitle
+% TABLE OF CONTENTS, LIST OF FIGURES, LIST OF TABLES
+% Please, do not include a table of contents, a list of figures, or a
+% list of tables. They will be removed by the editors (and the command
+% is actually redefined in the ct.sty file).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 To date there are no known examples of IDP $3$-polytopes without a unimodular cover, although such polytopes exist in higher dimension~\cite{BG}.
-
-\subsection*{Acknowledgements:} We  thank two anonymous referees for helpful comments on a previous version of this paper.
 
@@
 
-\begin{thebibliography}{99}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\bibitem{BHHHJKM2019}
-Matthias Beck, Christian Haase, Akihiro Higashitani, Johannes Hofscheier, Katharina Jochemko, Lukas Katth\"an, Mateusz Micha{\l}ek,
-Smooth Centrally Symmetric Polytopes in Dimension 3 are IDP,
-\emph{Ann.~Combin.}, 23 (2019), 255--262.
-\doi{10.1007/s00026-019-00418-x}
+% ACKNOWLEDGEMENTS
+% Include acknowledgements to colleagues and referee here.
+% Funding and grant support should appear in footnotes on the front page, using the 
+% thanks command in the authors command (see above).
 
-\bibitem{BG}
-Winfried Bruns and Joseph Gubeladze, Normality and covering properties of affine semigroups, \emph{J. Reine Angew. Math.} 510 (1999), 151--178. \doi{10.1515/crll.1999.044}
+\section*{Acknowledgements}
+We  thank two anonymous referees for helpful comments on a previous version of this paper.
 
-\bibitem{BGbook}
-Winfried Bruns and Joseph Gubeladze, \emph{Polytopes, rings, and k-theory}, Monographs in Mathematics,
-  Springer-Verlag, 2009, XIV, 461 p. 52 illus.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\bibitem{BGM}
-  Winfried Bruns, Joseph Gubeladze, and Mateusz Micha{\l}ek. Quantum Jumps of Normal Polytopes. \emph{Discrete Comput. Geom.}, 56(1) (2016), 181--215. \doi{10.1007/s00454-016-9773-7}
+% BIBLIOGRAPHY
+% Pease provide us with a bibtex file for your bibliography
+% You can find examples in ct-sample.bib. 
+% Please use the correct entrytype 
+%     article: any article published in a periodical like a journal article or magazine article
+%     book: a book
+%     booklet: like a book but without a designated publisher
+%     conference: a conference paper
+%     inbook: a section or chapter in a book
+%     incollection: an article in a collection
+%     inproceedings: a conference paper (same as the conference entry type)
+%     manual: a technical manual
+%     masterthesis: a Masters thesis
+%     misc: used if nothing else fits
+%     phdthesis: a PhD thesis
+%     proceedings: the whole conference proceedings
+%     techreport: a technical report, government report or white paper
+%     unpublished: a work that has not yet been officially published
+%
+% When it exists, please add a DOI to your reference using the doi field, this will be shown as a link in the final pdf
+% Similarly for preprints, please put the arXiv reference into the eprint field 
+% When necessary, you can provide a link through the URL field. Please do not give both the URL and the DOI.
+%
+% We encourage you to use MathSciNet https://mathscinet.ams.org/mathscinet
+% or other equivalent bibliography databases to get full and correct bibliographic entries.
 
-\bibitem{CLS}
-David A.~Cox, John B.~Little, and Hal K.~Schenck.
-{\em {T}oric {V}arieties}.
-AMS, Providence, 2011.
-
- \bibitem{DLRS2010}
-J. A. De Loera, J. Rambau, F. Santos
-\emph{Triangulations: Structures for Algorithms and Applications}, 539 pp.
-Algorithms and Computation in Mathematics, Vol. 25, Springer-Verlag. 
-ISBN: 978-3-642-12970-4
-
-\bibitem{Fakhruddin}
-Najmuddin Fakhruddin.
-\newblock Multiplication maps of linear systems on smooth projective toric surfaces.
-  Preprint, 
-  \href{https://arxiv.org/abs/math/0208178}{arXiv:math/0208178}, August 2002.
-
-%\bibitem{Grunbaum}
-%{Branko Gr\"unbaum}.
-%{\em {C}onvex {P}olytopes}.
-%Wiley, London, 1967.
-
-\bibitem{Gubeladze}
-Joseph Gubeladze,
-Normal polytopes and ellipsoids,
-\emph{Elect. J. Combin.}, 28:4 (2021), P4.7.
-\doi{10.37236/10338}
-
-\bibitem{HaaseHof}
-Christian Haase and Jan Hofmann, 
-Convex-normal (Pairs of) Polytopes, 
-\emph{Canad. Math. Bull.} 60:3 (2017), 510--521. \doi{10.4153/CMB-2016-057-0}
-
-  \bibitem{HPPS-survey}
-Christian Haase, Andreas Paffenholz, Lindsay C. Piechnik, Francisco Santos. Existence of unimodular triangulations - positive results. 
- \emph{Mem. Amer. Math. Soc.},  Vol.~270 1N.~321, American Math. Society, 2021. DOI: \doi{10.1090/memo/1321}
-%ISBNs: 978-1-4704-4716-8 (print); 978-1-4704-6530-8 (online)
-
-\bibitem{HNPS2008}
-Christian Haase, Benjamin Nill, Andreas Paffenholz, and Francisco Santos, Lattice points in Minkowski sums, 
-\emph{Electron.~J.~Combin.}, 15:1 (2008), Note 11, 5 pp. \doi{10.37236/886}
-
-\bibitem{KantorSarkaria}
-Jean-Michel Kantor and Karanbir~S.\ Sarkaria.
-On primitive subdivisions of an elementary tetrahedron, 
-\emph{Pacific J.~Math.} \textbf{211} (2003), no.~1,  123--155. \doi{10.2140/pjm.2003.211.123}
-
-\bibitem{mfo2004}
-Mini-workshop: Ehrhart Quasipolynomials: Algebra, Combinatorics, and Geometry, Oberwolfach Rep.
-1 (2004), no.
-3, 2071-2101, Abstracts from the mini-workshop held August 15-21, 2004, Organized by Jes\'us A.~De Loera and Christian Haase, Oberwolfach Reports.
-Vol.
-1, no.
-3.
-\doi{10.14760/OWR-2004-39}
-
-\bibitem{mfo2007}
-Mini-workshop: Projective normality of smooth toric varieties, Oberwolfach Rep.
-4 (2007), no.
-39/2007, Abstracts from the mini-workshop held August 12-18, 2007.
-Organized by Christian Haase, Takayuki Hibi and Diane MacLagan.
-\doi{10.14760/OWR-2007-39}
-
-\bibitem{Oda1997}
-Tadao Oda. Problems on Minkowski sums of convex lattice polytopes. 
-Abstract submitted at the Oberwolfach Conference ''Combinatorial Convexity and Algebraic Geometry'' 26.10--01.11, 1997.
-Available as arXiv preprint \href{https://arxiv.org/abs/0812.1418}{arXiv:0812.1418}, 2008.
-
-\bibitem{Ogata}
-Shoetsu Ogata.
-\newblock {Multiplication maps of complete linear systems on projective toric surfaces.}
-\newblock {\em Interdiscip. Inf. Sci.}, 12(2):93--107, 2006. \doi{10.4036/iis.2006.93}
-
-\bibitem{Santos-hirsch}
-Francisco Santos.
-A counter-example to the Hirsch Conjecture.
-\emph{Ann.~Math.~(2)}, 176 (July 2012), 383--412. 
-\doi{10.4007/annals.2012.176.1.7}
-
-\bibitem{SantosZiegler}
-Francisco Santos and G{\"u}nter~M. Ziegler.
-\newblock {Unimodular triangulations of dilated 3-polytopes}, 
-\newblock {\em Trans. Moscow Math. Soc.} (2013), 293--311. \doi{10.1007/s00493-014-2959-9}
-
-\bibitem{Tsuchiya}
-Akiyoshi Tsuchiya.
-Cayley sums and Minkowski sums of 2-convex-normal lattice polytopes.
-Preprint,  \href{https://arxiv.org/abs/1804.10538}{arXiv:1804.10538}, April 2018.
-
-\bibitem{White1964}
-G.~K.~White.
-Lattice tetrahedra.
-\emph{Canadian J.~Math.} 16 (1964), 389--396. \doi{10.4153/CJM-1964-040-2}
-
-\end{thebibliography}
+\bibliographystyle{alphaurl}
+\bibliography{unimodular-covers}
 
